### 2023-10-25 paper.tex
first commit

\documentclass{article}

\title{Dissertation Proposal}
\author{Kuang-Chen Lu}
\date{\today}

\bibliography{bibfile}

\usepackage{geometry} % Adjust page margins
\usepackage{amsmath, amssymb} % For math symbols and equations
\usepackage{graphicx} % For including images
\usepackage{titlesec} % For customizing section titles
\usepackage{biblatex} % For managing references
\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\miscon}[1]{\textbf{#1}}
\newcommand{\surmisedMiscon}[1]{\miscon{#1}$^\ddag$}
\newcommand{\pValue}[0]{\textit{p}\nobreakdash-value}

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{tabularx}   %% For tables with growable column
\usepackage{cleveref}   %% For smart ref
\usepackage{graphicx}   %% For inserting images
\graphicspath{ {./images/} }  %% configure the graphicx package
\usepackage{multirow}
\usepackage{multicol}
\usepackage{ragged2e}  %% For \raggedright
\usepackage{listings}  %% For typesetting code
\usepackage{enumitem}  %% For noitemsep and topsep
\lstdefinelanguage{smol} {
  alsoletter={ABCDEFGHIJKLMNOPQRSTUVWXYZabcdefghijklmnopqrstuvwxyz-!?*},
  morekeywords={defvar,deffun,lambda,set!,let,let*,letrec,begin,if,cond,else,and,or},
  morecomment=[l]{;},
  morestring=[b]",
}
\lstset{
  basicstyle=\small\ttfamily,
  breaklines=true,
  language=smol,
  % aboveskip=-0.3 \baselineskip,   % remove extra space before
  % belowskip=-0.8 \baselineskip,     % and after code blocks
}
%% Custom commands:
%% new terms are introduced with a boldface
%% (the common style is italic, but KC thinks bf is easier to locate)
\newcommand{\newterm}[1]{\textbf{#1}}

%% abbrev. for JavaScript etc.
\newcommand{\js}{JavaScript}
\newcommand{\py}{Python}

%% abbrev. for quizzes
\newcommand{\QuizFun}{Quiz 1}
\newcommand{\QuizState}{Quiz 2}
\newcommand{\QuizHof}{Quiz 3}
\newcommand{\quizFun}{Quiz 1}
\newcommand{\quizState}{Quiz 2}
\newcommand{\quizHof}{Quiz 3}

%% abbrev. for tutorials
\newcommand{\TutDef}{The \textbf{def} tutorial}
\newcommand{\TutSet}{The \textbf{set} tutorial}
\newcommand{\TutVec}{The \textbf{vec} tutorial}
\newcommand{\TutLam}{The \textbf{lam} tutorial}
\newcommand{\TutLet}{The \textbf{let} tutorial}
\newcommand{\tutDef}{the \textbf{def} tutorial}
\newcommand{\tutSet}{the \textbf{set} tutorial}
\newcommand{\tutVec}{the \textbf{vec} tutorial}
\newcommand{\tutLam}{the \textbf{lam} tutorial}
\newcommand{\tutLet}{the \textbf{let} tutorial}

%% abbrev. for referring to supplementary materials
\newcommand{\supplRef}[1]{\texttt{#1} in the supplementary materials}

\begin{document}

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Misconceptions about core linguistic concepts like mutable
  variables, mutable compound data, and their interaction with scope
  and higher-order functions seem to be widespread.
  But
  how do we detect
  them, given that experts have blind spots and may not realize the
  myriad ways in which students can misunderstand programs?
  Furthermore, once identified, what can we do to correct them?

  In this paper, we present a curated list of misconceptions, and an
  instrument to detect them. These are distilled from student work
  over several years and match and extend prior research.  We also
  present an automated, self-guided tutoring system. The tutor builds
  on strategies in the education literature and is explicitly designed
  around identifying and correcting misconceptions.

  We have tested the tutor in multiple settings.
  Our data consistently show that (a) the misconceptions we tackle are
  widespread, and (b) the tutor appears to improve understanding.
\end{abstract}

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}
\label{s:intro}

A large number of widely used modern programming languages share a
common semantic basis:
\begin{itemize}

  \item lexical scope
  \item nested scope
  \item eager evaluation
  \item sequential evaluation (per ``thread'')
  \item mutable first-\emph{order} variables
  \item mutable first-\emph{class} structures (objects, vectors, etc.)
  \item higher-order functions that close over bindings
  \item automated memory management (e.g., garbage collection)

\end{itemize}
This semantic core can be seen in languages from ``object-oriented''
languages like C\# and Java, to ``scripting'' languages like
JavaScript, Python, and Ruby, to ``functional'' languages like the ML
and Lisp families. Of course, there are sometimes restrictions
(e.g., Java has restrictions on closures) and extensions (such as the
documented semantic oddities of JavaScript and Python \cite{bernhardtWat2012, guhaEssenceJavascript2010, politzPythonFullMonty2013, politzTestedSemanticsGetters2012}). Still, this semantic core bridges many syntaxes, and
understanding it helps when transferring knowledge from old languages
to new ones. In recognition of this deep commonality, in this paper we
choose to call this the \emph{Standard Model of Languages} (SMoL).

Unfortunately, this combination of features appears to also be non-trivial for
programmers to understand. Consider the following scenario: to create a
calculator, we have to construct a callback that can be attached to
each button. The following Python program seems to achieve the
construction of these callbacks and the act of pushing the buttons in
order:
\newpage
\begin{lstlisting}[language=Python]
button_list = []

for i in range(10):
    button = lambda: print(i)
    button_list.append(button)

for button in button_list:
    button()
\end{lstlisting}
That is, a user would expect to see \verb|0| through \verb|9|. In
fact, however, it prints \verb|9| ten times.

As the related work section (\Cref{s:rel-work}) describes,
multiple researchers, in different
countries and different kinds of post-secondary educational contexts,
have studied how students fare with scope and state. They consistently
find that even advanced students have difficulty with such programs
and even programs simpler than this.

In fact, these problems are not at all limited to students. This
specific looping problem even trips up industrial programmers. The
C\# language \emph{changed} to produce \verb|0| through \verb|9| \cite{lippertClosingLoopVariable2009}. It is now also the focus of a language design
change in Go \cite{chaseFixingLoopsGo2023} for a new kind of looping
construct \cite{coxSpecLessErrorprone2023}.
It continues to trip up programmers
(e.g., \cite{sharveyCreatingFunctionsLambdas2022})
despite being
documented as a ``gotcha'' in Python \cite{reitzHitchhikerGuidePython2016}.
% SK: But not everyone has read the document. I don't think the document is
% a Python official document. Also, this sentence sounds like stating that
% the cited Python programmers are stupid.

Most of the time, however, these misunderstandings do not lead to
language design changes. Nor are changes necessarily desirable: the
SMoL feature set has presumably evolved because it is convenient
for writing non-trivial programs (e.g., mutable structures) without
being too unwieldy (e.g., no dynamic scope). Furthermore, having a
good mental model of these features
is essential to understand ownership \cite{clarkeOwnershipTypesFlexible1998},
manage parallelism, and more. Thus, we
still need to train students and other programmers on the semantic
features \emph{and their interactions} in SMoL.

In response, we have (a) run a multi-year, multi-phase study to
identify these misunderstandings while trying to work around the blind
spots of experts (and also drawing on prior work),
and (b) constructed the SMoL Tutor, an interactive tutor to address
these misunderstandings. The SMoL Tutor is not a \emph{programming} tutor;
rather, it assumes the user has some basic facility with programming
(i.e., that they are already familiar with concepts like vectors,
mutation, and higher-order functions). It is built around SMoL's
features and their interactions, and draws on cognitive and
educational psychology concepts to expressly identify
and correct misconceptions.

Concretely, we make the following contributions:
\begin{enumerate}

  \item We provide a collection of brief programs for detecting
        misconceptions about the features of SMoL and their
        interaction.
        These programs are short and
        most are readily translatable to a wide variety of
        programming languages that share the SMoL characteristics.

      \item We show that multiple populations have problems with these
        programs.

      \item We present a list of curated misconceptions generated
        after multiple rounds of analysis.

      \item We implement these misconceptions as interpreters,
        which find weaknesses in our manual analysis and hold
        potential for generative use in the future.

  \item We present a tutoring system that measurably corrects these
        misconceptions.

\end{enumerate}

\paragraph{A Note on Terminology} We use the term
``behavior'' to refer to the meaning of programs in terms of the
answers they produce. A more standard term for this would, of course, be
``semantics''. However, the term ``semantics tutor'' might
mislead some readers into thinking it
teaches people to read or write a formal semantics, e.g., an
introduction to ``Greek'' notation. Because that is not the kind of
tutor we are describing, to avoid confusion, we use the term
``behavior'' instead.

\section{Background: Tutoring Systems and Pedagogic Techniques}
\label{s:background}

There is an extensive body of literature on tutoring systems (\cite{vanlehnBehaviorTutoringSystems2006a} is
a quality survey), and indeed whole conferences are dedicated to
them. We draw on this literature. In particular, it is common in the
literature to talk about a ``two-loop'' architecture \cite{vanlehnBehaviorTutoringSystems2006a} where the outer
loop iterates through ``tasks'' (i.e., educational activities) and the inner loop iterates through UI events within a task. We follow the same
structure in our tutor (\Cref{s:tutor-uiux}).

Many tutoring systems focus on teaching programming
(such as the well-known and heavily studied LISP Tutor \cite{andersonLISPTutor1985}), and in
the process undoubtedly address some program behavior
misconceptions. Our SMoL Tutor differs in a notable way: it does not
try to teach programming per se. Instead, it assumes a basic
programming background and focuses entirely on program behavior misconceptions
and correcting them. We are not aware of a tutoring system (in computer science) that has
this specific design.

The SMoL Tutor is firmly grounded in one technique from
cognitive and educational psychology. The fundamental problem is: how
do you fix a misconception? One approach is to only ``present the
right answer'', for fear that discussing the wrong conception might
actually reinforce it. Instead, there is a body of literature starting
from \cite{posnerTheoryConceptualChange1982} that presents a theory of conceptual
change, at whose heart is the \emph{refutation text}. A refutation
text tackles the misconception directly, discussing and providing a
refutation for the incorrect idea. Several studies \cite{schroederRefutationTextFacilitates2022} have shown
their effectiveness in multiple other domains.

The SMoL Tutor's content structure is also influenced by work on \emph{case comparisons}
(which draws analogies between examples).
\cite{alfieriLearningCaseComparisons2013a} suggests that asking
(rather than not asking) students to find similarities between cases,
and providing principles \emph{after} the comparisons (rather than
before or not at all), are associated with better learning
outcomes.

\section{The SMoL Language}\label{s:smol}

\begin{figure}[t]
  \centering
  \begin{lstlisting}
    t ::= d
        | e
    d ::= (defvar x e)
        | (deffun (f x ...) body)
    e ::= c
        | x
        | (lambda (x ...) body)
        | (let ([x e] ...) body)
        | (begin e ... e)
        | (set! x e)
        | (if e e e)
        | (cond [e body] ... [else body])
        | (cond [e body] ...)
        | (e e ...)
    body    ::= t ... e
    program ::= t ...
  \end{lstlisting}
  \caption{The syntax of SMoL.}
  \label{f:smol-syntax}
\end{figure}

\begin{table}[t]
  \centering\begin{tabularx}{\linewidth}{r X}
    \textbf{Operators}            & \textbf{Meaning}                                       \\\hline
    \texttt{+ - * /}              & Arithmetic Operators                                   \\
    \texttt{< > <= >=}            & Number comparison                                      \\
    \texttt{mvec}                 & Create a (mutable) vector (a.k.a. array)               \\
    \texttt{vec-ref}              & Look up a vector element                               \\
    \texttt{vec-set!}             & Replace a vector element                               \\
    \texttt{vec-len}              & Get the length of a vector                             \\
    \texttt{mpair}                & Create a 2-element (mutable) vector                              \\
    \texttt{left right}           & Look up the first/second element of a 2-element vector \\
    \texttt{set-left! set-right!} & Replace the first/second element of a 2-element vector \\
    \texttt{eq?}                    & Equality                                               \\
  \end{tabularx}
  \caption{Primitive operators of SMoL.}
  \label{t:smol-primitives}
\end{table}

SMoL is designed to capture common features of many modern
languages.
The syntax of SMoL is presented in~\Cref{f:smol-syntax}, where
\lstinline|t| stands for terms, \lstinline|d| stands for definitions, \lstinline|e|
stands for expressions, \lstinline|c| stands for constants (i.e., number,
boolean, and string), and \lstinline|x| and \lstinline|f| are identifiers
(variables). The last kind of expression (i.e., \lstinline|(e e ...)|) is
function application.

SMoL's Lispy syntax is
an artifact of the course (\Cref{s:univ-1-course}) using Racket
\cite{friedmanEssentialsProgrammingLanguages2001, krishnamurthiProgrammingLanguagesApplication2007};
however, it also proved to be pedagogically valuable.
We have found the parentheses useful when discussing
scope in conjunction with local-binding features like \lstinline|let|.
This avoids the various confusing ``variable lifting'' semantics found
in languages like Python (\cite{politzPythonFullMonty2013}; see also
posts like \cite{froadieWhatScopeVariable2022}), where the actual defined range of a
variable is not apparent from the source code.

Nevertheless, most SMoL programs are easy to translate to other
languages. We present machine-translated Python and JavaScript
versions of our programs in \supplRef{Translated\_Programs.html}. Furthermore, we intend to
make a multi-lingual tutor in the future (\Cref{s:discussion}).

The semantics of SMoL is as described in \Cref{s:intro}.
SMoL includes limited primitive operators (\Cref{t:smol-primitives}) to work with numbers, strings, and
vectors. It provides only one equality operator, which tests for exact
equality between atomic values and for pointer equality between other
values.

Students are given a working implementation of SMoL, built
as a \verb|#lang| language inside Racket \cite{felleisenProgrammableProgrammingLanguage2018}. This language
provides only the defined syntax and semantics of SMoL, with no
(other) Racket features present.
(As in Racket, arguments evaluate left-to-right, to give stateful programs an
unambiguous semantics.) The implementation is available online
[URL anonymized].

\section{Paper Roadmap}
\label{s:roadmap}
\label{s:univ-1-course}

This paper describes a four-year effort to obtain a quality instrument
to measure misconceptions and to build a tutor to address them.

The first step was a two-year process to generate (a) programs that
tend to trip up students and (b) incorrect responses to those programs
(\Cref{s:gen-prob-quizius}), using a tool for the purpose
(\Cref{s:quizius}). These were then curated (\Cref{s:smol-quizzes}) to
produce an instrument that was tested for one more year
(\Cref{s:misconceptions-quiz}). This underwent one more round of
curation to produce the final instrument~(\Cref{s:tutor-refine-program-set}).

We then present the tutor (\Cref{s:system-tutor}). In particular, we
present the list of misconceptions---grounded in student data---that
are the focus of this project (\Cref{s:misconceptions-tutor}). We also
evaluate the tutor's effectiveness at correcting them
(\Cref{s:effect-tutor}).

All this work is done with students in a ``Principles of Programming
Languages'' class at a selective, private US university. The class has
about 70--75 students per year. It is not required, so students take
it by choice. Virtually all are computer science majors. Most are in
their third or fourth year of tertiary education; about 10\% are
graduate students. All have had at least one semester of imperative
programming, and most have significantly more experience with it. Most
have had close to a semester of functional programming. The student
work described here was required, but students were graded on effort,
not correctness.

Naturally, we should wonder to what extent the demographic affects the
results: we may simply be studying weaknesses at this institution! We
therefore work with two other populations
(\Cref{s:generality-tutor}). In both, we find similar results.

\section{Generating and Collating Problems}
\label{s:gen-prob-quizius}

In \Cref{s:rel-work}, we discuss several papers that have provided
reports of student misconceptions with different fragments of
SMoL. However, it is difficult to know how comprehensive these
are. While some are unclear on the origin of their programs, they
generally seem to be expert-generated.

The problem with expert-generated lists is that they can be quite
incomplete. Education researchers have documented the
phenomenon of the \emph{expert blind spot} \cite{nathanExpertBlindSpot2001}: experts
simply do not conceive of many learner difficulties.
Thus, we need
methods to identify problems beyond what experts conceive.

Additionally, in this paper, we intentionally use the word
mis\emph{conceptions} rather than mis\emph{take}. A mistake can happen
for any reason (e.g., selecting the wrong answer from a menu). A
misconception, however, implies a conceptual problem: the student has
formed an incorrect concept in their head. For instance, they may
think that mutable structures are copied on function calls, or that
scope is resolved dynamically. This requires probing what they are
thinking.

Finally, we are inspired by the significant body of education research
on \emph{concept inventories} \cite{hestenesForceConceptInventory1992} (with a growing
number for computer science, as a survey lists \cite{taylorComputerScienceConcept2014}). In
terms of mechanics, a concept inventory is just an instrument
consisting of multiple-choice questions (MCQs), where each question has
one correct answer and several wrong ones. However, the wrong ones are
chosen with great care. Each one has been validated so that if a
student picks it, we can quite unambiguously determine \emph{what
  misconception the student has}. For instance, if the question is
``What is \verb|sqrt(4)|?'', then \verb|37| is probably an
uninteresting wrong answer, but if people appear to confuse
square-roots with squares, then \verb|16| would be present as an
answer.\footnote{Concept inventories are thus useful in many
  settings. For instance, an educator can use them with clickers to
  get quick feedback from a class. If several students pick a
  specific wrong answer, the educator not only knows they are wrong,
  but also has a strong inkling of \emph{precisely what} misconception
  that group has and can address it directly. We expect our
  instruments to be useful in the same way.}

All these, however, add up to a somewhat challenging demand. We want
to produce a list of questions (each one an MCQ) such that
\begin{enumerate}

\item we can get past the expert blind spot,

\item we have a sense of what misconceptions students have, and

\item we can generally associate wrong answers with specific
  misconceptions, approaching a concept inventory.

\end{enumerate}

\subsection{Generating Problems Using Quizius}
\label{s:quizius}

Our main solution to the expert blind spot is to use the Quizius system
\cite{saarinenHarnessingWisdomClasses2019}. In contrast to the very
heavyweight process (involving a lot of expert time)
that is generally
used to create a concept inventory, Quizius uses a lightweight,
interactive approach to obtain fairly comparable data, which an expert
can then shape into a quality instrument.

In Quizius, experts create a prompt; in our case, we asked students to
create small but ``interesting'' programs using the SMoL
language. Quizius shows this prompt to students and gathers their
answers. Each student is then shown a set of programs created by other
students and asked to predict (without running it) the value produced
by the program.\footnote{In the course (\Cref{s:univ-1-course}),
  students were given credit for using Quizius but not penalized for wrong
  answers, reducing their incentive to ``cheat'' by running
  programs. They were also told that doing so would diminish the value of
  their answers. Some students seemed to do so anyway, but most
  honored the directive.} Students are also asked to provide a rationale for why
they think it will produce that output.

Quizius runs interactively during an assignment period.
At each point, it needs to determine which
previously authored program to show a student.
It can
either ``exploit'' a given program that already has responses
or ``explore'' a new
one. Quizius thus treats this as a multi-armed bandit problem
\cite{katehakisMultiArmedBanditProblem1987} and uses that to choose a program.

The output from Quizius is (a) a collection of programs; (b) for
each program, a collection of predicted answers; and (c) for each
answer, a rationale. Clustering the answers is easy (after
ignoring some small syntactic differences). Thus, for
each cluster, we obtain a set of rationales.

After running Quizius in the course (\Cref{s:univ-1-course}),
we took over as experts. Determining which is the right
answer is easy. Where expert knowledge is useful is in
\emph{clustering the rationales}. If all the rationales for a wrong
answer are fairly similar, this is strong evidence that there is a
common misconception that generates it. If, however, there are
multiple rationale clusters, that means the program
is not discriminative enough to distinguish the misconceptions, and it
needs to be further refined to tell them apart. Interestingly, even
the correct answer needs to be analyzed, because sometimes correct
answers do have incorrect rationales (again, suggesting the program
needs refinement to discriminate correct conceptions
from misconceptions).

Prior work using Quizius \cite{saarinenHarnessingWisdomClasses2019} finds that students do author programs
that the experts did not imagine. In our case, we seeded Quizius with
programs from prior papers (\Cref{s:rel-work}), which gives the first
few students programs to respond to. However, we found
that Quizius significantly expanded the scope of our problems and
misconceptions. In our final instrument, most programs were directly
or indirectly inspired by the output of Quizius.

\subsection{Collating Problems}
\label{s:smol-quizzes}

While Quizius is very useful in principle, it also produced data that
needed significant curation for the following reasons:
\begin{itemize}

\item A problem may have produced diverse outputs simply because it
  was written in a very confusing way. Such programs do not reveal
  any useful \emph{behavior} misconceptions, and must therefore be
  filtered out. For instance:

  \begin{lstlisting}
  (defvar x 1)
  (defvar y 2)
  (defvar z 3)
  (deffun (sum a ...) (+ a ...))
  (sum x y z)
  \end{lstlisting}

  A reader might think that \verb|sum| takes variable arguments (so
  the program produces \verb|6|), but in fact \verb|...| is a single
  variable, so this produces an arity error.

\item Some programs relied on (or stumbled upon) underspecified aspects
  of unimportant (in particular, non-standard) parts of SMoL, such as floating-point versus
  rational arithmetic.

\item A problem may have produced diverse outputs simply because it is
  hard to parse or to (mentally) trace its execution. One example was a
  17-line program with 6 similar-looking and -named functions. As
  another example:

\begin{lstlisting}
  (defvar a (or (/ 1 (- 0.25 0.25)) (/ 1 0.0)))
  (defvar b (and (/ 1 (- 0.25 0.25)) (/ 1 0.0)))
  (defvar c (and (/ 1 0.0) (/ 1 (- 0.25 -0.25))))
  (defvar d (or (/ 1 0) (/ 1 (- 0.25 -0.25))))
  (and (or a c) (or b d))
\end{lstlisting}

  This program is not only confusing, it \emph{also} tests interpretations of (a) exact
  versus inexact numbers and (b) truthy/falsiness, leading to
  significant (but not very useful) answer diversity.

\item As noted above, a program's wrong (or even correct) answers may
  correspond to multiple (mis)conceptions. In these cases, the program
  must be refined to be more discriminative.

\end{itemize}
and so on.
We therefore manually curated the Quizius output
to address these issues.

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\textwidth]{images/smol-quizzes-example.png}
  \caption{Screenshot of a SMoL Quiz question.}
  \label{f:smol-quizzes-example}
\end{figure}

\subsection{The SMoL Quizzes}%
\label{s:misconceptions-quiz}

Having curated the output, we had to confirm that these programs were
still effective! That is, they needed to actually find student errors.

We therefore turned the programs into a set of quizzes (in the US
sense: namely, a brief test of knowledge) that we call the SMoL
Quizzes. There were three quizzes, ordered by linguistic
complexity. The first consisted of only basic operators and
first-order functions. The second added variable and structure
mutation. The third added \lstinline|lambda| and higher-order functions.

The goal of the SMoL Quizzes was to confirm that the aforementioned
processes of cleansing and enriching the problems was successful. The
quizzes were therefore administered in the third year of this project in the
same course.
\Cref{f:smol-quizzes-example} shows a
sample question. Every question got an ``Other'' option. If chosen,
the quiz gave the user a text box with the caption ``Please
specify''. The goal here was to record any other answers, which in
turn might lead to fresh misconceptions.

Question orders were partially randomized. We wanted students to get
some easy, warm-up questions initially, so those were kept at the
beginning. Similarly, we wanted programs that are syntactically
similar to stay close to each other in the quiz. This is so that, when
students got a second such program, they would not have to look far to
find the first one and confirm that they are indeed (slightly)
different, rather than wonder if they were seeing the same program
again.

Students only received feedback after having completed a whole
quiz. At the end of each quiz, they received both summative feedback
\emph{and} a refutation text that explained every program that
appeared in the quiz. Students were also encouraged to run the
programs, but we have little reason to believe that they did (and
certainly they asked few questions on the class forum about them).

Due to space limitations, we present the entire instrument in \supplRef{SMoL Quizzes/instrument}.
Here we focus on a few programs where student choices correspond to
misconceptions identified earlier, thereby also showing
that the curated programs are
still effective.
Syntactically, \verb|#t| and \verb|#f| are true and false,
while \verb|#(...)| is a vector. We use a * to indicate the correct
answer (which also matches the implementation's output).

\subsubsection{Quiz 1: Basic Operators and First-Order Functions}%
\label{s:misconceptions-quiz:fun}

% This quiz starts with two warm-up and not very interesting questions
% and then other questions. The order of the other questions are fully
% randomized.

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{l l}
    \textbf{Questions}                       &
    \textbf{Frequency Table of Answers}          \\\hline
    \lstinputlisting{programs/def_twice.txt} & {
        \begin{tabularx}{1in}{r l}
          $^\dag$59\%  & \verb|2 0|             \\
          *32\% & \verb|Error|           \\
          4\%   & \verb|Other: 2 2|      \\
          3\%   & ``Nothing is printed'' \\
          2\%   & \verb|0 0|             \\
        \end{tabularx}
    }                                            \\\hline
    \lstinputlisting{programs/dyn_scope.txt} & {
        \begin{tabularx}{1in}{r l}
          *65\% & \verb|Error| \\
          $^\dag$29\%  & \verb|42 #t| \\
          6\%   & Other        \\
        \end{tabularx}
    }                                            \\\hline
    \lstinputlisting{programs/reassign.txt}  & {
        \begin{tabularx}{1in}{r l}
          *53\% & \verb|#(6 5) 5|                 \\
          *20\% & \verb|Error|                    \\
          $^\dag$11\%  & \verb|#(6 6) 5|                 \\
          $^\dag$11\%  & \verb|#(6 6) 6|                 \\
          5\%   & ``Nothing is printed'' or Other \\
        \end{tabularx}
    }                                            \\\hline
  \end{tabularx}
  \caption{Some questions and student answers from \quizFun. Each
    table row is a program and the (relative) frequencies of
    student answers. Answers that can be considered correct are marked
    with \texttt{*}. Wrong answers that we discuss are
    marked with $^\dag$.}
  \label{t:quizzes-data-fun}
\end{table}
\Cref{t:quizzes-data-fun} lists programs in \quizFun{} that we
consider the most interesting. These data confirm the presence of
scope-related misconceptions:
\begin{enumerate}
  \item At least 59\% of students incorrectly believe that a
        variable can be defined twice in one block.
  \item 29\% of students believe that a variable defined in a function
        will be available in the top-level (or global) environment
        after the function is called. That is, these students may have a
        dynamic scope misconception.
  \item 22\% (11\% + 11\%) of students believe variables themselves
        can be passed as arguments and redefined inside the function.
        They disagree on whether the redefinition persists after the
        function call.
\end{enumerate}

\subsubsection{Quiz 2: Adding Variable and Structure Mutation}%
\label{s:misconceptions-quiz:state}

% This quiz randomizes the order of all questions.

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{l l}
    \textbf{Question}                           &
    \textbf{Frequency Table of Answers}             \\\hline
    \lstinputlisting{programs/circularity.txt}  & {
        \begin{tabularx}{1in}{r p{2.5in}}
          $^\dag$50\%  & \verb|#(#(2 #(2 3)) #(2 3))| \\
          *29\% & \verb|x=#(x x)| or
          something similar. Both \verb|(left x)| and \verb|(right x)|
          are \verb|x| itself.                 \\
          $^\dag$16\%  & \verb|Error|                 \\
          5\%   & Other                        \\
        \end{tabularx}
    }                                               \\\hline
    \lstinputlisting{programs/mvec_in_mvec.txt} & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *62\% & \verb|#(#(100 2 3 4) #(100 2 3 4))| \\
          $^\dag$26\%  & \verb|#(#(1 2 3 4) #(100 2 3 4))|   \\
          6\%   & \texttt{Error}                      \\
          4\%   & \verb|#(#(1 2 3 4) #(1 2 3 4))|     \\
          2\%   & Other                               \\
        \end{tabularx}
    }                                               \\\hline
    \lstinputlisting{programs/vset_in_let.txt}  & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *59\% & \verb|#(10)|  \\
          $^\dag$36\%  & \verb|#(123)| \\
          5\%   & Other         \\
        \end{tabularx}
    }                                               \\\hline
    \lstinputlisting{programs/var_as_arg.txt}   & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *65\% & \texttt{12} \\
          $^\dag$31\%  & \texttt{0}  \\
          4\%   & Other       \\
        \end{tabularx}
    }                                               \\\hline
    \lstinputlisting{programs/setset.txt}       & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *59\% & Other: \texttt{5 7}   \\
          $^\dag$27\% & \texttt{6 7}          \\
          11\%  & \texttt{5 5}          \\
          2\%  & Other: \texttt{Error} \\
          1\%  & Other: \texttt{5 6}   \\
        \end{tabularx}
    }                                               \\\hline
  \end{tabularx}
  \caption{Some questions and student answers from \quizState. Each
    table row is a program and the (relative) frequencies of
    student answers. Answers that can be considered correct are marked
    with \texttt{*}. Wrong answers that we discuss are
    marked with $^\dag$.}
  \label{t:quizzes-data-state}
\end{table}
\Cref{t:quizzes-data-state} lists interesting programs after the
addition of state. These data suggest the students have aliasing-related misconceptions:
\begin{itemize}
  \item Up to 50\% of students think vectors are copied rather than
  aliased.
  \item 16\% of students think trying to construct and print a
  self-referring vector would error. (This program is ambiguous:
  \emph{constructing} works in most SMoL languages, but \emph{printing}
  may well cause a problem. These identified ambiguities are
  addressed in \Cref{s:tutor-refine-program-set}.)
  \item 31\% of students think a variable is aliased by a parameter if
        the two variables have the same name. Perhaps interestingly,
        fewer students (27\%) think the variable aliasing would
        happen if the two variables had different names.
\end{itemize}

\subsubsection{Quiz 3: Adding Closures and Higher-Order Functions}%
\label{s:misconceptions-quiz:hof}

% This quiz includes two groups of questions. Each group contains
% programs that are really similar to each other. The order of the two
% groups and the other questions are randomized. Question order within
% each group is also randomized. The first two programs in the
% \Cref{t:quizzes-data-hof} belongs to one group, which is about
% changing a function's behavior by mutating a variable defined outside
% the function. The other group is about equality between functions. The
% latter group is not included in the main text because we think it is
% not very related to the theme of this paper.

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{l l}
    \textbf{Question}                                 &
    \textbf{Frequency Table of Answers}                   \\\hline
    \lstinputlisting{programs/fun_returns_lambda.txt} & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *82\% & \verb|3|     \\
          $^\dag$17\%  & \verb|Error| \\
          1\%   & Other        \\
        \end{tabularx}
    }                                                     \\\hline
    \lstinputlisting{programs/fun_and_state_1.txt}    & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *74\% & \verb|4|            \\
          $^\dag$21\%  & \verb|3|            \\
          6\%   & Other: \verb|Error| \\
        \end{tabularx}
    }                                                     \\\hline
    \lstinputlisting{programs/fun_and_state_2.txt}    & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *88\% & \verb|4|            \\
          $^\dag$10\%  & \verb|3|            \\
          3\%   & Other: \verb|Error| \\
        \end{tabularx}
    }                                                     \\\hline
    \lstinputlisting{programs/counter.txt}            & {
        \begin{tabularx}{1in}{r p{2.5in}}
          *62\% & \verb|1 1 2 3 2| \\
          $^\dag$34\%  & \verb|1 1 1 1 1| \\
          4\%   & Other            \\
          1\%   & \verb|1 1 2 3 4| \\
        \end{tabularx}
    }                                                     \\\hline
  \end{tabularx}
  \caption{Some questions and student answers from \quizHof. Each
    table row is a program and the (relative) frequencies of
    student answers. Answers that can be considered correct are marked
    with \texttt{*}. Wrong answers that we discuss are
    marked with $^\dag$.}
  \label{t:quizzes-data-hof}
\end{table}
\Cref{t:quizzes-data-hof} lists interesting programs
after the addition of closures and higher-order functions.
It extends what we saw with loops in
\Cref{s:intro}: that students have misconceptions about their interaction with mutable variables:
\begin{itemize}
  \item 17\% of students think \texttt{lambda} functions can't refer
        to free variables.
  \item 21\% of the students think mutating a variable defined outside
        a lambda can't possibly change the behavior of the lambda.
        Perhaps interestingly, this
        misconception seems to depend on how the lambda is constructed
        (compare the middle two programs).
  \item Student understanding of the
        \lstinline|let|-over-\lstinline|lambda| pattern is weak.
\end{itemize}

\section{The SMoL Tutor}
\label{s:system-tutor}

So far, we have focused on identifying problems. As noted earlier
(\Cref{s:smol-quizzes}), we provided students with refutation texts after the quizzes,
but it is unclear to what extent students read, understood, or
internalized these. We also wanted to determine whether other
populations run into these issues, and it was unclear whether we would
be able to administer passive quizzes to them.

Furthermore, while the quizzes may be a useful diagnostic, our goal is
not only to find faults but to improve the understanding of basic
programming language behavior. We believe (and presumably so does
anyone else who writes a formal semantics!)\ that an understanding of
these basic program behaviors is important, and even more so when it
comes to understanding concurrency, ownership, and other advanced
features. Even more simply, misunderstanding these clearly impacts
development and debugging time.

In response, we decided to create a \emph{tutor}: the SMoL Tutor. It is built
around our quiz instruments and the detected misconceptions. We
describe the Tutor from a user's perspective in \Cref{s:tutor-uiux}, and
explain how it was populated from the
SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
We then discuss what
we learned from its data (\Cref{s:misconceptions-tutor}), and finally
evaluate its effectiveness as a \emph{tutor} (\Cref{s:effect-tutor}).

\subsection{The User Experience}
\label{s:tutor-uiux}

The Tutor covers five major topics, shown on the left in
\Cref{t:tutor-goal-sentences}. The larger topics are further broken
down into 2--3 modules. The goal was for students to spend at most
20--30 minutes per module. Our data show that in practice, students
spent about 9.8 (median) minutes.

Each tutorial consists of a sequence of
questions. Most questions in the Tutor are \newterm{interpreting
questions}.\footnote{The Tutor also asked students to perform some
other activities, like showing programs and asking for their heap
content, which we do not cover in this paper.}
 These questions (illustrated in
\Cref{f:smol-tutor-example}) are versions of the SMoL
Quizzes (obtained by the process described in \Cref{s:tutor-refine-program-set}).
In each of these questions, students are shown a program
and asked to predict the program's running result(s) by answering an
MCQ (with an ``Other'' option). After making a choice, students receive
feedback. If a student chooses incorrectly, they are (a) given an
explanation that is \emph{based on the misconception associated with
  that wrong answer} (or a generic one, if there is not a specific
misconception),
and then (b) asked to answer a
second question:
\begin{enumerate}

\item The second question is semantically the same as the first, but
  with superficial changes (e.g., variable names, constants, and operators are
  changed) so that the student cannot immediately
  guess the answer.

\item Instead of multiple-choice, students must \emph{type} the answer
  into a text box. (The Tutor normalizes text to accommodate variations.)
  This is
  intentional. First, we wanted to force reflection on the
  explanation just given, whereas with an
  MCQ, students could just guess. Second, we felt that students would
  find typing more onerous than clicking. In case students
  had just guessed on a question, our hope is that the penalty of having to
  type means, on \emph{subsequent} tasks, they
  would be more likely to pause and reflect before choosing.

\end{enumerate}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\textwidth]{images/smol-tutor-example.png}
  \caption{Screenshots of an interpreting question in SMoL Tutor. The
    top-left shows the initial state, where the question is presented
    as an MCQ. If a student chooses a wrong answer, they will receive
    feedback (bottom-left) and will be asked a similar question
    (right). The similar question must be answered by typing.}
  \label{f:smol-tutor-example}
\end{figure}

In addition to asking students questions, the Tutor along the way
introduces terminology and states the true conceptions. These are the
teaching goals of the Tutor. We therefore refer to these as
\newterm{goal sentences}. \Cref{t:tutor-goal-sentences} lists
the (abbreviated) goal sentences for each tutorial.
Readers can find the full Tutor in \supplRef{SMoL Tutor/instrument}.

\begin{table}[t]
  % \begin{tabularx}{250pt}{|c|X|c|X|} \end{tabularx}
  \centering\begin{tabularx}{\linewidth}{
      >{\hsize=.5\hsize\linewidth=\hsize}X
      >{\hsize=1.5\hsize\linewidth=\hsize}X}
    \textbf{Tutorial}                             & \textbf{Goal Sentences}
    \\\hline
    \newterm{scope}: Variable definitions and function definitions &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Referring to an unbound variable leads to an error.
      \item Define "blocks".
      \item SMoL is lexically scoped rather than dynamically scoped.
      \item SMoL disallows defining a variable twice in one block.
      \item SMoL is eager (and evaluates from left to right) rather
            than lazy, reactive, or relational.
    \end{enumerate}
    \\\hline
    \newterm{mut-vars}: Variable updates &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Variable assignment respects the hierarchical structure of blocks.
      \item Variables do not alias.
    \end{enumerate}
    \\\hline
    \newterm{begin}: Sequencing expressions &
    A sequencing expression evaluates its sub-expressions from left to right and
    returns the value of the last sub-expression.
    \\\hline
    \newterm{vectors}: Vectors and vector updates &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Define heap and memory addresses.
      \item Vectors can be referred to by vectors, including themselves.
      \item Vectors are not copied but aliased by bindings.
      \item Constructing vectors doesn't alter environments.
      \item Introducing bindings doesn't alter the (value part of) heap.
    \end{enumerate}
    \\\hline
    \newterm{lambda}: Lambda expressions &
    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
      \item Functions are (first-class) values.
      \item Functions remember the environment where they were created.
      \item Functions can be created with lambda expressions.
      \item A function definition can be viewed as a variable definition plus a lambda expression.
    \end{enumerate}
    \\\hline
    \newterm{local}: Local binding forms &
    Introduce \texttt{let}, \texttt{letrec}, and \texttt{let*}.
    % \\\hline
  \end{tabularx}
  \caption{SMoL Tutor tutorials and their goal sentences.}
  \label{t:tutor-goal-sentences}
\end{table}

Some later tutorials include questions about earlier tutorials so
that we can check whether students remember the concepts across
modules. In particular, the \textbf{mut-vars} tutorial starts with questions
about \textbf{scope}, and \textbf{lambda} starts with questions about
\textbf{mut-vars} and \textbf{vectors}.

Of the modules, \textbf{local} is the least portable across
languages. While local binding is of course present in other
languages, this is mostly covered using nested \lstinline|defvar|s in
earlier modules, starting with \textbf{scope}. The distinction central
to this module (between three local-binding constructs with slightly
different scopes) is primarily a focus of Lispy languages. Therefore,
we exclude this module from our analysis and instruments, and indeed
plan to deprecate the module in future versions.

\Cref{s:smol-quizzes} discusses the partial randomization employed by
the SMoL Quizzes. In contrast, the SMoL Tutor does \emph{not}
randomize question order. This is because the Tutor consists of more
than just questions: it also has explanatory text. This text is based
on the preceding problems. Authoring it is therefore somewhat like
writing a textbook, with complex dependencies that can easily be
broken.

\subsection{Collating Problems for the Tutor}%
\label{s:tutor-refine-program-set}

The SMoL Tutor's interpreting questions are the final instrument of
this paper.
We bootstrap it from the SMoL
Quizzes, but made the following alterations:
\begin{enumerate}

  % fill gaps in goal sentences
  \item In the process of developing the goal sentences, we felt the
        existing programs were insufficient to cover the ideas we wanted
        students to work through. For instance, SMoL Quizzes did not
        alias vectors using \lstinline|defvar|, so we added this program:

\begin{lstlisting}
(defvar x (mvec 100))
(defvar y x)
(vec-set! x 0 200)
y
\end{lstlisting}

  \item We renamed vector operators to avoid confusion between
        \lstinline|vset!|, which replaces a vector element, and \lstinline|set!|,
        which mutates variables. We rename all vector operators from, for example,
        \lstinline|vset!| to \lstinline|vec-set!|.

  \item We deferred the local binding forms to the end, since we felt
        they were less essential. We therefore rewrote programs that use
        it to not depend on it.

        % \item We removed some questions that we deemed ``Lispy'' (e.g.,
        %   involving the difference between parenthetical and infix syntax).

  \item We removed some programs that relied on underspecified aspects or
        unimportant aspects of SMoL. For example, some depended on
        whether operations on pairs could be applied to arbitrary vectors:

\begin{lstlisting}
  (pair? (mpair 1 2))
  (pair? (mvec 1 2))
  (pair? '#(1 2))
  (pair? '(1 2))
\end{lstlisting}

        As another example, one hinged on
        whether or not a function's formal parameter could have the same
        name as the function itself:

\begin{lstlisting}
  (deffun (f f) f)
  (f 5)
\end{lstlisting}

  \item We resolved ambiguities in some programs, either adding
    answer choices or even adding other questions to tease
    out different interpretations.

  \item To reduce the number of concepts, we removed programs that relied
        upon \emph{im}mutable vectors and lists, because they did not
        seem to create problems. (For brevity, we leave these out of
        the presentation of SMoL in \Cref{s:smol}, though they are in the implementation.)

  \item We removed questions related to function equality, which was
        not a focus of the Tutor.\footnote{However, we believe there is
          a good deal of confusion about notions of equality. We
          intend to
          add that as a major tutorial component in future versions.}

  \item We removed programs of a ``Lispy'' nature, such as one where
    the answer depended on whether the reader correctly understood
    this inequality:

\begin{lstlisting}
      (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
\end{lstlisting}

    This checks whether $\verb|n| < \verb|3|$, but some presumably
    vocalized it as ``greater than \verb|3|, \verb|n|?''.

\end{enumerate}
Most of these steps either modify or elide programs. Two cases,
filling gaps in light of the goal sentences and resolving ambiguities,
introduce
programs. Starting with the 37 programs in the SMoL Quizzes, we added
52 more programs to arrive at a total of 89.

In addition to generating this set of programs, we also
modified the answer options. In addition to retaining the correct and
incorrect answers from the SMoL Quizzes (and constructing our best
guess of analogous incorrect answers for the new problems), we added
more wrong answers.

The reason is as follows. The SMoL Quizzes often have very few wrong
choices: of the 37 tasks, 26 have only three choices (including the
correct answer and the ``Other'' option\footnote{In a few cases, we
  removed the \emph{correct} answer from the choices. A student would
  therefore have to click on ``Other'' and type it in. The idea was to
  make the ``Other'' option more salient and plausible.}),
seven have 4, and only four have 5 or 6.
Thus, in most
cases, students have a 25\% or even 33\% chance of just guessing the
right answer or successfully using a process of elimination. By
increasing the number of options, we hoped to
greatly reduce the odds of getting the right answer by chance or
by elimination.

It was important to add wrong answers that are not utterly
implausible, because those would become easy to eliminate. Therefore,
we added numeric constants mentioned in the problem, permuted some of
the values in case of multiple outputs, and so on. In general, we
ended up increasing the number of choices substantially:
only 8 out of the 89 questions have three or four choices; 70
questions have 5--8 choices; and 11 questions have 9--14 choices. We hope
this reduced both guessing and elimination, and forced students to
actually think through the program. Of course, these new answers do
not have a clear associated misconception, so their mistakes are given
the generic explanation.

\section{Misconceptions Detected by the SMoL Tutor}%
\label{s:misconceptions-tutor}

We now examine what we learned from the interpreting
questions in terms of program behavior
misconceptions. But first, we explain how we associate incorrect
answers with misconceptions.

We do not present the results of \textbf{begin} and
\textbf{local} (\Cref{t:tutor-goal-sentences}), because they focus on constructs not usually found in
non-Lispy languages. They were included in the Tutor to help students
write programs for the course, but are not interesting from a SMoL
perspective.

\subsection{Misinterpreters}
\label{s:misinterp-intro}

In principle, an expert can identify what misconceptions a particular
incorrect answer might correspond to. In practice, we found this
rather difficult for three reasons. First, with 89 programs, each with
several answers, it's easy to make mistakes. Second, our own expert blind
spot may prevent us from seeing an interpretation that would
lead to an additional association. Finally, and specific to our case,
since we had either curated or written all the programs and answers,
we were likely to miss some associations we had not intended.

To address this problem, we formalized misconceptions as
interpreters. That is, for each misconception we created a
corresponding \emph{misinterpreter}. This is an interpreter for the
SMoL syntax that intentionally has a semantic error corresponding to
that misconception. Put differently, a misinterpreter is a like a
``definitional interpreter for a misconception''.

By running programs through the misinterpreters, we can more uniformly
and rigorously identify \emph{all} the misconceptions associated with
a wrong answer. Furthermore, if we identify a new misconception (or
alter a misinterpreter), it is easy to automatically re-classify all
the answers. By using misinterpreters, we indeed found new
interpretations for existing program-answer pairs.
We provide all the misinterpreters in the artifact.

\subsection{A Catalog of Misconceptions}

We iteratively created
our final catalog of misconceptions
(from the perspective of the data in this paper).
We started
with misinterpreters representing the misconceptions described in
\Cref{s:misconceptions-quiz}. These are misconceptions for which we
have reasonable validation (due to the prose in Quizius), so we call
them \emph{grounded} misconceptions. We then looked at wrong answers
not covered by these but chosen by students, and did our best to
distill these into misconceptions. These are \emph{surmised}
misconceptions (which we
identify with a $^\ddag$), which need to be validated in the future.\footnote{To keep
  the time spent in each module reasonable, the SMoL Tutor does not
  ask students for rationales for their programs. However, we could
  easily modify it to do so a small number of times per user, to make
  progress towards this need.} We then re-ran the misinterpreters
against the chosen answers. We terminated when all the remaining wrong
answers were either (a) found in very few students (we found a gap
between 23\% and 13\%, and hence took 14\% as the threshold),
(b) difficult for
us to attribute to a misconception, or (c) appeared to us to be
``Lispy'' and hence not of broad interest.

\Cref{t:tutor-grounded-misconceptions-1,t:tutor-grounded-misconceptions-2,t:tutor-surmised-misconceptions}
list the final catalog.
For each, we also
present a Tutor question
for which the marked wrong answer can be explained by \emph{only} the
named misconception.
That is, that program-answer pair is a representative example of that
misconception.\footnote{An astute reader may well imagine another
  misinterpretation; but we presumably did not find evidence of it in our
  data. We do welcome other misinterpretations, for which we
  can add more misinterpreters!}

  \begin{table}[t]
    \centering
    \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}
      \textbf{Misconception}                                                     &
      \textbf{Question}                                                          &
      \textbf{Table of Answers}
      \\\hline
      \miscon{CallByRef} Function calls alias variables.                         &
       % mut-vars.not_aliased_by_funarg_2 (35)
       \lstinputlisting{programs/call-by-ref.txt}                                                                    & {
        \begin{tabularx}{1in}{r l}
          78\%  & \verb|12|      \\
          11\% & \verb|error|      \\
          **10\%   & \verb|0|  \\
          1\%   & \verb|23|  \\
        \end{tabularx}
    }
      \\\hline
      \miscon{CallsCopyStructs} Function calls copy data structures.      &
      % vectors2.alias_with_funcall (48)
      \lstinputlisting{programs/calls-copy-structs.txt}
      & {
        \begin{tabularx}{1in}{r l}
          90\%  & \verb|#(173 0)|      \\
          **10\%   & \verb|#(1 0)|  \\
        \end{tabularx}
    } \\\hline
      \miscon{DeepClosure} Closures copy the \textit{values} of free variables.  &
      % lambda3.lambda_remembers_env (70)
      \lstinputlisting{programs/deep-closure.txt}                                & {
          \begin{tabularx}{1in}{r l}
            86\%  & \verb|4|      \\
            **10\% & \verb|3|      \\
            3\%   & \verb|error|  \\
            1\%   & \verb|lambda| \\
          \end{tabularx}
      }                                                                                   \\\hline
      \miscon{DefByRef} Variable definitions alias variables.                    &
      % mut-vars.not_aliased_by_defvar_1 (29)
      \lstinputlisting{programs/def-by-ref.txt}                                  & {
          \begin{tabularx}{1in}{r p{0.2\textwidth}}
            85\%   & \verb|0 12|                \\
            **12\% & \verb|0 0|                 \\
            2\%    & \verb|error|               \\
            1\%    & depends on implementation. \\
          \end{tabularx}
        }
      \\\hline
      \miscon{DefOrSet}  Both definitions and variable assignments are
      interpreted as follows: if a variable is not defined in the
      current environment, it is defined. Otherwise, it is mutated to
      the new value.
      &
      % program: mut-vars.update_undefined (28)
      \lstinputlisting{programs/def-or-set.txt}                                  & {
          \begin{tabularx}{1in}{r l}
            85\%   & \verb|error| \\
            **15\% & \verb|2|     \\
          \end{tabularx}
        }
      \\\hline
      \miscon{DefsCopyStructs} Variable definitions copy structures recursively. &
      % vectors2.alias_with_defvar (47)
      \lstinputlisting{programs/defs-copy-structs.txt}                                & {
          \begin{tabularx}{1in}{r l}
            **67\% & \verb|#(100)| \\
            30\%   & \verb|#(200)| \\
            1\%    & \verb|#(300)| \\
            1\%    & \verb|error|  \\
          \end{tabularx}
      }                                                                                   \\\hline
    \end{tabularx}
    \caption{Ground misconceptions identified by the SMoL~Tutor. Answers marked
      with ``**'' represent the misconception. (Part I)}
    \label{t:tutor-grounded-misconceptions-1}
  \end{table}

  \begin{table}[t]
    \centering
    \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}
      \textbf{Misconception}                                                     &
      \textbf{Question}                                                          &
      \textbf{Table of Answers}
      \\\hline
      \miscon{FlatEnv} There is only one environment, the global
      environment. (This misconception is a kind of dynamic scope.)              &
      % scope.error_when_refer_to_undefined (9)
      \lstinputlisting{programs/flat-env.txt}                                    & {
          \begin{tabularx}{1in}{r l}
            96\%  & \verb|error| \\
            **4\% & \verb|402|   \\
          \end{tabularx}
        }
      \\\hline
      \miscon{FunNotVal} Functions are \textit{not} considered
      first-class values. They can't be bound to other variables, passed as arguments, or referred to by data structures.
                                                                                 &
      % lambda1.fun_as_parameter (62)
      \lstinputlisting{programs/fun-not-val.txt}                                 & {
          \begin{tabularx}{1in}{r l}
            83\%   & \verb|4|     \\
            **11\% & \verb|error| \\
            4\%    & \verb|2|     \\
            1\%    & \verb|8|     \\
          \end{tabularx}
      }                                                                                   \\\hline
      \miscon{IsolatedFun} Functions can't refer to free variables
      except for the built-in ones.
                                                                                 &
      % scope.refer_global_is_possible (5)
      \lstinputlisting{programs/isolated-fun.txt}                                & {
          \begin{tabularx}{1in}{r l}
            77\%   & \verb|3|     \\
            **23\% & \verb|error| \\
          \end{tabularx}
        }
      \\\hline
      \miscon{NoCircularity} Data structures can't (possibly indirectly)
      refer to themselves.                                                       &
      % vector2.warmup_circularity (53)
      \lstinputlisting{programs/no-circularity.txt}                              & {
          \begin{tabularx}{1in}{r p{0.2\textwidth}}
            76\%   & \verb|3|                   \\
            **14\% & \verb|error|               \\
            9\%    & Run out of memory or time. \\
            1\%    & \verb|+inf|                \\
          \end{tabularx}
      }                                                                                   \\\hline
      \miscon{StructByRef} Data structures might refer to variables by their references.                                                       &
      % vector2.alias_var_in_mvec (49)
      \lstinputlisting{programs/struct-by-ref.txt}                              & {
          \begin{tabularx}{1in}{r p{0.2\textwidth}}
            67\%   & \verb|#(1 2 3)|                   \\
            **24\% & \verb|#(1 2 4)|               \\
            9\%    & error \\
          \end{tabularx}
      }                                                                                   \\\hline
      \miscon{StructsCopyStructs} Storing data structures into data structures makes copies.                                                       &
      % lambda1.smol_quiz_circularity (58)
      \lstinputlisting{programs/structs-copy-struct.txt}                              & {
          \begin{tabularx}{1in}{r p{0.2\textwidth}}
            65\%   & \verb|#0=#(#0# #0#)| (Racket circular object notation) \\
            24\%   & \verb|#(#(2 3) #(2 3))|               \\
            **6\%   & \verb|#(#(2 #(2 3)) #(2 3))|                   \\
            6\%    & error \\
          \end{tabularx}
      }                                                                                   \\\hline
    \end{tabularx}
    \caption{Ground misconceptions identified by the SMoL~Tutor. Answers marked
      with ``**'' represent the misconception. (Part II)}
    \label{t:tutor-grounded-misconceptions-2}
  \end{table}

\begin{table}[t]
  \centering
  \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}
    \textbf{Misconception}                                                             &
    \textbf{Question}                                                                  &
    \textbf{Table of Answers}                                                              \\\hline
    \surmisedMiscon{NestedDef} Sequences of definitions are
    interpreted as if they are written in nested blocks. A definition
    is not in the scope of later definitions.
    &
    % program: scope.what_is_x_4 (13)
    \lstinputlisting{programs/nested-def.txt}                                          & {
        \begin{tabularx}{1in}{r l}
          92\%  & \verb|2| \\
          **8\% & \verb|1| \\
        \end{tabularx}
    }                                                                                      \\\hline
    \surmisedMiscon{Lazy} Expressions are only evaluated when their values are needed. &
    % order.bad_order (20)
    \lstinputlisting{programs/lazy.txt}                                                & {
        \begin{tabularx}{1in}{r l}
          **57\% & \verb|1 3|   \\
          43\%   & \verb|error| \\
        \end{tabularx}
    }                                                                                      \\\hline
  \end{tabularx}
  \caption{Surmised misconceptions identified by the SMoL~Tutor. Answers marked
    with ``**'' represent the misconception.}
  \label{t:tutor-surmised-misconceptions}
\end{table}

\paragraph{Confirmed Misconceptions}

The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.

\paragraph{Added Misconceptions}

The misinterpreters helped us find misinterpretations that we had
overlooked. For instance, consider this program from \quizState:
\begin{lstlisting}
(defvar x 12)
(deffun (f x)
  (set! x 0))
(f x)
x
\end{lstlisting}
We had assumed the wrong answer
\texttt{0} is caused by \miscon{CallByRef}.
However, our misinterpreters made us realize it can also be explained
by \miscon{FlatEnv}. This could also explain why we see a difference in
error rate when the formal and actual parameter have the same
name (as mentioned in \Cref{s:misconceptions-quiz:state}).

% Tutor's refined program sets and misinterpreters also helped us to
% identify programs that detect one of the misconceptions but not the
% other. Consider this new program
% % mut-vars.not_aliased_by_defvar_1 (29)
% \begin{lstlisting}
%   (defvar x 12)
%   (defvar y x)
%   (set! x 0)
%   x
%   y
% \end{lstlisting}
% 12\% of students asserted that the following program produces
% \texttt{0 0}, which suggest that they have the \textbf{BindByRef}
% misconception. Students with a \miscon{FlatEnv} misconception would
% have chosen \texttt{0 12}, which happens to be the correct answer.
% (1\% wrote \textit{depends on
% implementation}; 2\% asserted \texttt{error}; the remaining 85\%
% asserted \texttt{0 12}).

% The second program in \Cref{t:quizzes-data-fun} is already a good
% detector for \miscon{FlatEnv}. SMoL Tutor includes an alternative
% program that does not depend on equality checks.
% % scope.error_when_refer_to_undefined (9)
% \begin{lstlisting}
%   (deffun (addy x)
%     (defvar y 200)
%     (+ x y))
%   (+ (addy 2) y)
% \end{lstlisting}
% 4\% of students asserted that the following program produces
% \texttt{402}, which suggest that they have the \miscon{FlatEnv}
% misconception. Students with a \textbf{BindByRef} misconception would
% have chosen \texttt{error}, which happens to be the correct answer.
% (the remaining 96\% asserted \texttt{error}).

\paragraph{A New Potential Misconception}

For the following program, added in the Tutor:
\begin{lstlisting}
  (defvar y (+ x 2))
  (defvar x 1)
  x
  y
\end{lstlisting}
56\% of students asserted that it produces
\texttt{1 3}. Based on this, we surmise that students might have
another misconception, which we define as \surmisedMiscon{Lazy}. (Recall that
SMoL is eager, but even in many lazy languages, this would be an error.)

\paragraph{Another Potential Misconception, and Its Effect on
  Interpreting Descriptions.}%
\label{ss:misinterpreters:submisconception}

Consider this program from the SMoL Tutor:
\begin{lstlisting}
(defvar x 1)
(deffun (main)
  (deffun (get-x) x)
  (defvar x 2)
  (get-x))

(main)
\end{lstlisting}
This program, suitably translated, would produce \verb|2| in a wide
variety of languages (Python, JavaScript, Racket, Java, etc.), because
\verb|get-x| and the inner \verb|x| are in the same scope block. The
answer \verb|1| cannot be explained by any of our existing
misconceptions. Based on this, we surmise a new misconception,
\surmisedMiscon{NestedDef}. (Though its frequency falls below our
threshold, our reading of answers suggests this may be more
widespread, and we feel it needs to be investigated more.)

Once we turned this into a misinterpreter, we found that it
unexpectedly captured the program-answer pair for the following
program from \quizFun---which should be an error, due to the
double-binding of \verb|x| in the same scope block---and the answer
\verb|2 0|:
\begin{lstlisting}
  (defvar x 0)
  (defvar y x)
  (defvar x 2)
  x
  y
\end{lstlisting}
Previously, we had interpreted this only as \miscon{DefOrSet}, because
students had stated that the second \lstinline|(defvar x ....)|
``mutates'' or ``redefines'' \verb|x|. This is reminiscent of the
behavior of languages like Python, which use the same syntax both for
binding new variables and for mutating existing ones.

The problem here is that the word ``redefine'' underspecifies how the
second definition is interpreted. We had interpreted it as
\textit{mutating} the binding established by the first definition,
which fits \miscon{DefOrSet}. However, another possibility is that
it \emph{shadows} the binding (i.e., establishes a new scope block). We
did not recognize that the original misconception is underspecified
until we uncovered the new (surmised) misconception.

\paragraph{Summary}

To summarize, we used the SMoL Tutor, with an expanded set of
programs, to investigate student misconceptions. We implemented the
idea of misinterpreters to help us properly classify student
performance. We were able to reconfirm all the previously identified
misconceptions, refine some of them, and also identify potential new
ones (that need further investigation).

\section{Is The Tutor Effective?}
\label{s:effect-tutor}

\begin{figure}[t]

  \includegraphics[width=1\textwidth]{images/Misconceptions_Decay.png}

  \caption{How many students chose a wrong answer that (uniquely)
    represents a misconception? (Downward tendency suggests improvement
    over time.)}
  \label{f:tutor-trend-plots}

\end{figure}

Recall that the SMoL Tutor is not only a collection of MCQs: it is
also a \emph{tutor}! So far, we have investigated the value of the
MCQs. Now we examine its tutoring aspect. Concretely, we ask:
\begin{description}

\item[RQ] How effective is the Tutor at correcting each misconception?

\end{description}
To study this, we perform the following analysis per
misconception. Using misinterpreters, we identify those questions
whose wrong answers fit only one misconception (i.e., only one output
matches that produced by that misinterpreter, and it matches only that
misinterpreter). We then examine student performance over time across
those questions. This would let us examine how they do on just that
topic, in isolation, over time.

We present the result of this analysis in two forms. Graphically, we
show plots in \Cref{f:tutor-trend-plots}. Each figure shows the percentage of students who
chose the answer corresponding to that misconception. Ideally, we
would like to see these percentages diminish.

Indeed, that is what we see in most of the graphs. The exceptions are
\miscon{CallsCopyStructs}, \surmisedMiscon{NestedDef}, and
\miscon{StructsCopyStructs}, which have only one problem (and hence no
trend), and \miscon{DefOrSet} and \miscon{FunNotVal}, which show an
increase. The lack of improvement for \miscon{FunNotVal} is
unsurprising because the Tutor does not explicitly address this issue,
focusing on closures created by \lstinline|lambda|s rather than named
functions. (However, this does not explain the increase!)

We also perform a logistic regression to see whether these
improvements are significant (at a $p < 0.05$ threshold); details are
in \supplRef{Paper.html}. Of the 11:
\begin{itemize}

\item Of the nine seemingly improved (i.e., decreased)
  misconceptions:
  \begin{itemize}
  \item Two are \emph{not} significant: \miscon{CallByRef}
      (\pValue\ = .074); \miscon{NoCircularity} (\pValue\ = .075).
    \item The other seven \emph{are} significant.
    \end{itemize}
  \item However, the two with an increasing trend (\miscon{DefOrSet}
    and \miscon{FunNotVal}) are \emph{also} significant.
\end{itemize}

These data broadly suggest that the Tutor is a net positive.
Ultimately, however, what these data really show is a need for
improvement in the Tutor. When designing the Tutor questions, many
more were intended to be representative of single
misconceptions. However, as noted in \Cref{s:misinterp-intro}, it is
easy to be incomplete (or incorrect) in ascribing
misconceptions. Furthermore, as their set grows, it is difficult to
reassess all the problems manually. Once evaluated using
misinterpreters, we found far fewer problems than we would have liked.
Concretely, only 40 of the 71 eligible problems (after removing the
non-SMoL modules) were useful in the above
analysis.

We therefore view the above data as
purely formative: they suggest that the Tutor \emph{most likely did
  not do harm} and perhaps even \emph{may have done some
  good}. However, it would be improper to read too much into the
analysis. It is quite possible that some of the other problems would
have found issues. Rather, what we really see is value in the
misinterpreter concept. It is not only useful for analysis, it is also
valuable for \emph{problem design}: in the next iteration of the
Tutor, we will use misinterpreters actively to shape the incorrect answers and,
as necessary, update the programs as well. We therefore hope to have
much more thorough analyses in future work.

\section{Performance on Other Populations}
\label{s:generality-tutor}

There is, of course, a significant danger that the data above have
been overfitted to only one institution (\Cref{s:univ-1-course},
henceforth University 1),
thereby actually reflecting the state of its curriculum rather than
some greater truth about program behavior understanding.  We already
have some reason to believe this is not the case: the related work
discussed in \Cref{s:rel-work} is drawn from many institutions in
multiple countries with different educational preparations, levels,
and demographics. Nevertheless, that gives us only
limited information about the specific questions and misconceptions
described above.

Fortunately, we were able to deploy the Tutor on two other
populations:
\begin{itemize}

\item University 2 is a primarily public university in the US. It is
  one of the largest Hispanic-serving institutions in the country. As
  such, its demographic is extremely different from those whose data
  were used above. The Tutor was used in one course in Spring 2023,
  taken by 12 students.
  The course is a third-year, programming
  language course. The students are required to have taken two
  introductory programming courses (C++ focused).

\item A separate instance of the Tutor was published on the
  website of a programming languages textbook [ANON]. Over the course
  of 8 months, 597 people started with the first module and 103
  users made it to the last one. To protect privacy, we intentionally
  do not record demographic information, but we conjecture that the
  population is largely self-learners  (who are known to use the
  accompanying book),
  including some professional
  programmers. It is extremely
  unlikely to be the students from either university, because they
  would not get credit for their work on the public instance; they
  needed to use the university-specific instance. Furthermore, since
  they were not penalized for wrong answers, it would make little
  sense to do a ``test run'' on the public instance. Finally, we note
  that there is no overlap between the dates of submission on the
  public instance and the semester at University 1.

\end{itemize}
These two populations are therefore at least somewhat different from
the original population, and help us assess whether the problems we
identify are merely an artifact of the first institution.

To evaluate, we computed a Spearman's rank correlation $\rho$, ranking
questions by what percentage of students got the question
right. Between the original university and University 2,
we obtain a \pValue~=~2.013e-07. Between the original university and
the online population, we obtain a \pValue~<~2.2e-16. These show that
the other two populations performed similarly to the original one.
While
further validation on other populations remains essential, these
suggest that the questions are \emph{finding misconceptions that may
  be universal}.

\section{Related Work}
\label{s:rel-work}

\begin{table}[t]
  \centering
  \begin{tabularx}{\textwidth}{>{\hsize=.6\hsize\RaggedRight}X >{\hsize=.8\hsize\RaggedRight}X >{\hsize=.5\hsize\RaggedRight}X >{\hsize=1.4\hsize\RaggedRight}X}
    \textbf{Publication}                       &
    \textbf{Population}                       &
    \textbf{Languages}                       &
    \textbf{Misconceptions}
    \\\hline
    \cite{fleuryParameterPassingRules1991} &
    Unclear, likely CS2 students &
    Pascal &
    \surmisedMiscon{CallerEnv}; \miscon{IsolatedFun}; \miscon{DefOrSet}
    (See their RULEs 2, 3a, and 3b in TABLE 2;
    RULE 1 doesn't apply to us.)
    \\\hline
    \cite{goldmanIdentifyingImportantDifficult2008,goldmanSettingScopeConcept2010} &
    CS1 students &
    Java, Python, Scheme &
    Scope and memory model (See their Figures 4 and 5)
    \\\hline
    \cite{fislerAssessingTeachingScope2017} &
    Third- and fourth-year undergrads &
    Java and Scheme &
    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{CallsCopyStruct}; \miscon{DefByRef}
    (See their Section 4)
    \\\hline
    \cite{saarinenHarnessingWisdomClasses2019} &
    CS2 students &
    Java &
    \miscon{StructByRef} (Their G2); \miscon{DefByRef} or \miscon{DefCopyStructs} (Their G3); \miscon{CallsCopyStruct} (Their G4).
    \\\hline
    \cite{strombackProgressionStudentsAbility2023} &
    CS masters &
    Python &
    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
    (See their section 4.2)
    \\\hline
    \cite{strombackProgressionStudentsAbility2023} &
    CS undergrads &
    C++ &
    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
    (See their section 4.2)
    \\\hline
  \end{tabularx}
  \caption{Similar misconceptions found in prior research.}
  \label{t:rel-work}
\end{table}

\paragraph{Misconceptions}

Misconceptions related to scope, mutation, and higher-order functions have been
widely identified in varying populations
(from CS1 students to graduate students to users of online forums)
over many years (since at last 1991)
in varying programming languages (from Java to Racket) and in
different countries (such as the USA and Sweden).
\Cref{t:rel-work} lists works that seem most relevant to us.

There are some small differences.
\cite{fleuryParameterPassingRules1991} identifies a dynamic scope misconception
that is different from \miscon{FlatEnv}:
\begin{description}
  \item[\surmisedMiscon{CallerEnv}] Function values don't remember their environments.
  When a function is called, the function body is evaluated in an
  environment that extends from the \textit{caller's} environment.
\end{description}
We don't include \surmisedMiscon{CallerEnv} in our analysis because in our
data, all wrong answers that can be explained by \surmisedMiscon{CallerEnv}
are also explainable by \miscon{FlatEnv}.

Appendix A of \cite{sorvaVisualProgramSimulation2012} provides an
extensive survey of misconceptions reported
in research up to 2012.
There are overlaps between our survey and
theirs. For instance, our \surmisedMiscon{CallerEnv} is their No.~47.
However, because their descriptions are brief, it is difficult to tell
whether a misconception in their survey matches our misconceptions.
Because
they provide neither misinterpreters nor representative program-output
pairs, it is difficult to determine the overlaps precisely (showing
the value of providing these two machine-runnable descriptions).
At any rate,
we certainly find
no equivalent of \miscon{FlatEnv}, \miscon{DeepClosure}, and
\miscon{DefByRef} in their survey.

\paragraph{Goal Sentences}

Our idea of ``goal sentences'' is not substantially different from the
\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
We only used these sentences (or rules) to guide the
design of the Tutor and as text in the Tutor itself.
\cite{duranRulesProgramBehavior2021} argue for other uses of such
sentences.

\paragraph{Misinterpreters}

Our idea of misinterpreters is related to mystery languages
\cite{diwanPLdetectiveSystemTeaching2004,
pombrioTeachingProgrammingLanguages2017}. Both approaches use
evaluators that represent alternative semantics to the same syntax.
However, the two are complementary.
In mystery languages, instructors design the space of semantics with
pedagogic intent, and
students must create programs to explore that space. Misinterpreters,
in contrast, are driven by student input, while the programs are provided
by instructors. The two approaches also have different goals:
mystery languages focus on encouraging students to experiment with
languages; misinterpreters aim at capturing students' misconceptions.

\section{Threats to Validity}
\label{s:threats}

\subsection{Construct Validity}

Did we measure the right thing?
While our goal is to study student understanding of language
behavior, what we actually measure is performance on MCQs on select
programs. MCQs as a mechanism introduce various clear biases, though
we add the ``Other'' options to somewhat alleviate them. The small
size, syntactic details (such as intentionally meaningless variable
names), and other aspects of the programs may also impact our ability
to measure understanding. (This could go both ways: students may have
no trouble understanding behavior in a small program but may struggle
to do so in a larger one.)

A particularly notable threat is the use of Lispy
syntax. We chose it for two reasons: both because it helps clarify
scope (\Cref{s:smol}) and because the rest of the course used
Racket. However, students may well perform differently with more
familiar syntaxes. It seems unlikely their performance would be \emph{too}
different given the many languages that have produced similar
misconceptions (\Cref{t:rel-work}). Nevertheless, we intend to use a
variety of syntaxes to examine this issue further.

Finally, we curated programs by hand, so they may well reflect our own
biases about misconceptions. It may be possible to mitigate this
problem by synthesizing MCQ programs using the misinterpreters.

\subsection{Internal Validity}

Is our reasoning valid? We have applied standard techniques and
measurements for evaluating student responses. However, our
instruments still lack the validity of a proper concept
inventory. Their creation requires heavyweight processes (such as
Delphi methods \cite{goldmanSettingScopeConcept2010}) that require many hours of expert attention as
well as conversations with learners, and can hence be prohibitive in
cost. The Quizius method (\Cref{s:quizius}) was created precisely to
be an inexpensive method that provides a good proxy.

Ultimately, our goal is to provide a reasonable instrument for
widespread use. While the set of all misconceptions could be
unbounded, we believe our tasks, especially as embedded in the Tutor,
provide a good starting point for others. In particular, if students
select a wrong answer, that is still of \emph{some} use to an
educator, even if the precise misconception cannot be pinned down with
the highest accuracy. We therefore believe our instruments, and our
misinterpreter technique, are of general value.

A failure in our Tutor's logging infrastructure led us to miss some
responses. These are unlikely to be task-specific because the Tutor
has a generic framework that should perform the same across
tasks. Moreover, on average only $0.4\%\ (\textrm{sd} = 0.6\%)$ of values are missing. Therefore, we do not believe
this had a noticeable impact. (In addition, every wrong answer is
still wrong! We may just not have exactly the right proportions of
them.)

\subsection{External Validity}

How well do our results generalize? Certainly there is reason to
question whether our results would apply to other
populations. \Cref{s:generality-tutor} provides preliminary evidence
that the results are not specific to one institution, and
\Cref{s:rel-work} suggests these issues are widespread. Nevertheless,
much more broad testing is needed to confirm our specific instruments.

The other major concern is the tie to Lispy syntax. Learners in other
settings may do worse or even better with it. Building a Tutor that
supports multiple, and more traditional, syntaxes should help address
this issue. We defer this to future work.

\section{Discussion}%
\label{s:discussion}

The paper has already identified several areas for future work:
\begin{itemize}
\item testing on more varied populations;
\item using other syntaxes;
\item having more questions that uniquely identify minsconceptions;
  and,
\item enabling textual responses in the Tutor so we can better
  characterize wrong answers.
\end{itemize}
These can help get us even closer to a good approximation of a true concept
inventory. We would also like the Tutor to make more use of the education
theories discussed in \Cref{s:background}.

We focus here on some issues that we think warrant broader
discussion.

\paragraph{The State-Aliasing-Function Triangle}

Our problematic programs hardly include any ``advanced'' programming
features: there are no threads, asynchrony, sophisticated type systems, ownership,
inversion of control, etc. Indeed, many of those features typically
build on an understanding of these basics (e.g., it is hard to make
sense of ownership \cite{clarkeOwnershipTypesFlexible1998} without a good understanding of this
triangle). But many populations seem to struggle even with this, which
may explain why concepts like ownership are considered hard \cite{crichtonGroundedConceptualModel2023}.

It is worth noting that we find problems even without all three
components, as \Cref{s:smol-quizzes} shows! Nevertheless, we think it
would be useful for curricula to focus on achieving mastery of this
triangle. This may require a deep revision of widely accepted
pedagogy. For instance, it is common to explain variables as
``boxes''. But if taken seriously by a learner, this metaphor may
cause more harm than good. As prior research \cite{putnamSummaryMisconceptionsHigh1986,groverMeasuringStudentLearning2017,hermansThinkingOutBox2018} shows, students expect that a variable can then contain more
than one value, removing one makes the other accessible, etc.  That
research has not explored aliasing, but the metaphor may affect that
too. A box is a closed object, so a (mutable) value put ``into'' it
clearly can't be modified by another ``box'' (variable)---i.e., it not
only doesn't explain but is antithetical to aliasing.

\paragraph{Terminology}

It is common to teach programming using the terms ``call-by-value''
and ``call-by-reference''. The reader will note that we instead have
\emph{three} \miscon{ByRef} misconceptions.

The emphasis on ``calling'' suggests that the semantics is associated
\emph{only} with calls. That leaves open what happens when one simply
binds a variable. In a reasonable language (and definitionally, in
SMoL), the behavior is exactly the same: the formal parameter can be
viewed as a binding to the value of the actual. But students often
form inconsistent views because the ``call'' terminology breaks this
deep similarity. We therefore recommend that languages in general use
the term \emph{bind-by-}, to emphasize the underlying semantic unity
of these syntactically different mechanisms.

We feel that further confusion is caused by terms like ``pass'' and
``return''. We often vocalize the call \verb|(f x)| as ``passing
\verb|x| to \verb|f|''; saying ``passing \emph{the value of} \verb|x|
to \verb|f|'' is a mouthful. But is it then surprising that students
think \verb|x| is being aliased? Similarly, consider a statement like
\verb|return y| (in Python syntax). Of course, semanticists understand
that it is the \emph{value} of \verb|y|, not \verb|y| itself, that is
being returned. Nevertheless, it is not surprising if this also
results in assumptions that \verb|y| is either aliased or that it has
escaped from the function (leading to an interpretation of dynamic
scope). We believe there is a need for significant research to
investigate these kinds of effects, which are similar to but not
strictly the same as ``vernacular misconceptions'' \cite{nationalresearchcouncilScienceTeachingReconsidered1997}.

\end{document}

### 2023-10-25 paper.tex
fix references
clean up tables

@@
 
-\title{Dissertation Proposal}
-\author{Kuang-Chen Lu}
-\date{\today}
-
-\bibliography{bibfile}
-
+\usepackage[utf8]{inputenc} % Use UTF-8 encoding
+\usepackage[T1]{fontenc} % Use the T1 font encoding
 \usepackage{geometry} % Adjust page margins
@@
 \usepackage{titlesec} % For customizing section titles
-\usepackage{biblatex} % For managing references
+\usepackage[style=authoryear]{biblatex}  % For managing references
+\addbibresource{bibfile.bib}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 \usepackage{tabularx}   %% For tables with growable column
+\newcolumntype{Y}{>{\centering\arraybackslash}X}
 \usepackage{cleveref}   %% For smart ref
@@
 
+\title{Dissertation Proposal}
+\author{Kuang-Chen Lu}
+\date{\today}
+
 \begin{document}
+\maketitle
 
@@
   \centering
-  \begin{tabularx}{\linewidth}{l l}
+  \begin{tabularx}{\linewidth}{l X}
     \textbf{Questions}                       &
-    \textbf{Frequency Table of Answers}          \\\hline
+    \textbf{Table of Answers}          \\\hline
     \lstinputlisting{programs/def_twice.txt} & {
-        \begin{tabularx}{1in}{r l}
+        \begin{tabularx}{\linewidth}{r X}
           $^\dag$59\%  & \verb|2 0|             \\
@@
     \lstinputlisting{programs/dyn_scope.txt} & {
-        \begin{tabularx}{1in}{r l}
+        \begin{tabularx}{\linewidth}{r X}
           *65\% & \verb|Error| \\
@@
     \lstinputlisting{programs/reassign.txt}  & {
-        \begin{tabularx}{1in}{r l}
+        \begin{tabularx}{\linewidth}{r X}
           *53\% & \verb|#(6 5) 5|                 \\
@@
   \centering
-  \begin{tabularx}{\linewidth}{l l}
+  \begin{tabularx}{\linewidth}{l X}
     \textbf{Question}                           &
-    \textbf{Frequency Table of Answers}             \\\hline
+    \textbf{Table of Answers}             \\\hline
     \lstinputlisting{programs/circularity.txt}  & {
-        \begin{tabularx}{1in}{r p{2.5in}}
+        \begin{tabularx}{\linewidth}{r X}
           $^\dag$50\%  & \verb|#(#(2 #(2 3)) #(2 3))| \\
@@
     \lstinputlisting{programs/mvec_in_mvec.txt} & {
-        \begin{tabularx}{1in}{r p{2.5in}}
+        \begin{tabularx}{\linewidth}{r X}
           *62\% & \verb|#(#(100 2 3 4) #(100 2 3 4))| \\
@@
     \lstinputlisting{programs/vset_in_let.txt}  & {
-        \begin{tabularx}{1in}{r p{2.5in}}
+        \begin{tabularx}{\linewidth}{r X}
           *59\% & \verb|#(10)|  \\
@@
     \lstinputlisting{programs/var_as_arg.txt}   & {
-        \begin{tabularx}{1in}{r p{2.5in}}
+        \begin{tabularx}{\linewidth}{r X}
           *65\% & \texttt{12} \\
@@
     \lstinputlisting{programs/setset.txt}       & {
-        \begin{tabularx}{1in}{r p{2.5in}}
+        \begin{tabularx}{\linewidth}{r X}
           *59\% & Other: \texttt{5 7}   \\
@@
   \centering
-  \begin{tabularx}{\linewidth}{l l}
+  \begin{tabularx}{\linewidth}{l X}
     \textbf{Question}                                 &
-    \textbf{Frequency Table of Answers}                   \\\hline
+    \textbf{Table of Answers}                   \\\hline
     \lstinputlisting{programs/fun_returns_lambda.txt} & {
-        \begin{tabularx}{1in}{r p{2.5in}}
+        \begin{tabularx}{\linewidth}{r X}
           *82\% & \verb|3|     \\
@@
     \lstinputlisting{programs/fun_and_state_1.txt}    & {
-        \begin{tabularx}{1in}{r p{2.5in}}
+        \begin{tabularx}{\linewidth}{r X}
           *74\% & \verb|4|            \\
@@
     \lstinputlisting{programs/fun_and_state_2.txt}    & {
-        \begin{tabularx}{1in}{r p{2.5in}}
+        \begin{tabularx}{\linewidth}{r X}
           *88\% & \verb|4|            \\
@@
     \lstinputlisting{programs/counter.txt}            & {
-        \begin{tabularx}{1in}{r p{2.5in}}
+        \begin{tabularx}{\linewidth}{r X}
           *62\% & \verb|1 1 2 3 2| \\
@@
     \centering
-    \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}
+    \begin{tabularx}{\textwidth}{m{0.25\textwidth} l Y}
       \textbf{Misconception}                                                     &
@@
       \lstinputlisting{programs/def-by-ref.txt}                                  & {
-          \begin{tabularx}{1in}{r p{0.2\textwidth}}
+          \begin{tabularx}{1in}{r Y}
             85\%   & \verb|0 12|                \\
@@
 
+\section{References}
+
+\printbibliography
+
 \end{document}

### 2023-10-25 paper.tex
clean up some tables
move figures to the paper body

@@
   \centering
-  \includegraphics[width=0.8\textwidth]{images/smol-quizzes-example.png}
+  \includegraphics[width=\linewidth]{images/smol-quizzes-example.png}
   \caption{Screenshot of a SMoL Quiz question.}
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{images/smol-tutor-example.png}
+  \includegraphics[width=\textwidth]{images/smol-tutor-example.png}
   \caption{Screenshots of an interpreting question in SMoL Tutor. The
@@
 \begin{table}[t]
-  % \begin{tabularx}{250pt}{|c|X|c|X|} \end{tabularx}
-  \centering\begin{tabularx}{\linewidth}{
-      >{\hsize=.5\hsize\linewidth=\hsize}X
-      >{\hsize=1.5\hsize\linewidth=\hsize}X}
+  \centering\begin{tabularx}{\linewidth}{p{0.2\linewidth} X}
     \textbf{Tutorial}                             & \textbf{Goal Sentences}
@@
 
-  \begin{table}[t]
-    \centering
-    \begin{tabularx}{\textwidth}{m{0.25\textwidth} l Y}
+  \begin{table}[!h]
+    \centering\begin{tabularx}{\linewidth}{X l X}
       \textbf{Misconception}                                                     &
@@
        \lstinputlisting{programs/call-by-ref.txt}                                                                    & {
-        \begin{tabularx}{1in}{r l}
+        \begin{tabularx}{\linewidth}{r X}
           78\%  & \verb|12|      \\
@@
       & {
-        \begin{tabularx}{1in}{r l}
+        \begin{tabularx}{\linewidth}{r X}
           90\%  & \verb|#(173 0)|      \\
@@
       \lstinputlisting{programs/deep-closure.txt}                                & {
-          \begin{tabularx}{1in}{r l}
+          \begin{tabularx}{\linewidth}{r X}
             86\%  & \verb|4|      \\
@@
       \lstinputlisting{programs/def-by-ref.txt}                                  & {
-          \begin{tabularx}{1in}{r Y}
+          \begin{tabularx}{\linewidth}{r X}
             85\%   & \verb|0 12|                \\
@@
       \lstinputlisting{programs/def-or-set.txt}                                  & {
-          \begin{tabularx}{1in}{r l}
+          \begin{tabularx}{\linewidth}{r X}
             85\%   & \verb|error| \\
@@
       \lstinputlisting{programs/defs-copy-structs.txt}                                & {
-          \begin{tabularx}{1in}{r l}
+          \begin{tabularx}{\linewidth}{r X}
             **67\% & \verb|#(100)| \\
@@
 
-  \begin{table}[t]
+  \begin{table}[!h]
     \centering
-    \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}
+    \begin{tabularx}{\linewidth}{X l X}
       \textbf{Misconception}                                                     &
@@
       \lstinputlisting{programs/flat-env.txt}                                    & {
-          \begin{tabularx}{1in}{r l}
+          \begin{tabularx}{\linewidth}{r X}
             96\%  & \verb|error| \\
@@
       \lstinputlisting{programs/fun-not-val.txt}                                 & {
-          \begin{tabularx}{1in}{r l}
+          \begin{tabularx}{\linewidth}{r X}
             83\%   & \verb|4|     \\
@@
       \lstinputlisting{programs/isolated-fun.txt}                                & {
-          \begin{tabularx}{1in}{r l}
+          \begin{tabularx}{\linewidth}{r X}
             77\%   & \verb|3|     \\
@@
       \lstinputlisting{programs/no-circularity.txt}                              & {
-          \begin{tabularx}{1in}{r p{0.2\textwidth}}
+          \begin{tabularx}{\linewidth}{r X}
             76\%   & \verb|3|                   \\
@@
       \lstinputlisting{programs/struct-by-ref.txt}                              & {
-          \begin{tabularx}{1in}{r p{0.2\textwidth}}
+          \begin{tabularx}{\linewidth}{r X}
             67\%   & \verb|#(1 2 3)|                   \\
@@
       \lstinputlisting{programs/structs-copy-struct.txt}                              & {
-          \begin{tabularx}{1in}{r p{0.2\textwidth}}
+          \begin{tabularx}{\linewidth}{r X}
             65\%   & \verb|#0=#(#0# #0#)| (Racket circular object notation) \\
@@
 
-\begin{table}[t]
+\begin{table}[!h]
   \centering
-  \begin{tabularx}{\linewidth}{m{0.3\textwidth} l l}
+  \begin{tabularx}{\linewidth}{X l X}
     \textbf{Misconception}                                                             &
@@
     \lstinputlisting{programs/nested-def.txt}                                          & {
-        \begin{tabularx}{1in}{r l}
+        \begin{tabularx}{\linewidth}{r X}
           92\%  & \verb|2| \\
@@
     \lstinputlisting{programs/lazy.txt}                                                & {
-        \begin{tabularx}{1in}{r l}
+        \begin{tabularx}{\linewidth}{r X}
           **57\% & \verb|1 3|   \\
@@
 
-\begin{figure}[t]
-
-  \includegraphics[width=1\textwidth]{images/Misconceptions_Decay.png}
+\begin{figure}[!h]
+
+  \includegraphics[width=\linewidth]{images/Misconceptions_Decay.png}
 

### 2023-10-26 paper.tex
line numbers
rqs and future plan

@@
 \documentclass{article}
+
+\usepackage[switch]{lineno}
+\usepackage{xcolor}
+\renewcommand\thelinenumber{\textcolor{red}{\arabic{linenumber}}}
 
@@
 \maketitle
+\linenumbers
 
@@
 order:
-\newpage
 \begin{lstlisting}[language=Python]
@@
 
-In response, we have (a) run a multi-year, multi-phase study to
+This lead to my \textbf{research questions}:
+
+\begin{description}
+  \item[RQ1] What are the misconceptions that are about program
+  behavior, apply to most modern programming languages, and present
+  (even) in students with prior programming backgrounds?
+  \item[RQ2] What would be a tutoring system (the SMoL Tutor) that
+  draws on cognitive and educational psychology concepts and that aim
+  at correcting the misconceptions?
+  \item[RQ3] How effective is SMoL Tutor at correcting the
+  misconceptions?
+\end{description}
+
+To answer my research questions, I have (a) run a multi-year, multi-phase study to
 identify these misunderstandings while trying to work around the blind
@@
 
-Concretely, we make the following contributions:
+Concretely, I have made the following contributions:
 \begin{enumerate}
@@
         misconceptions about the features of SMoL and their
-        interaction.
-        These programs are short and
-        most are readily translatable to a wide variety of
-        programming languages that share the SMoL characteristics.
+        interaction. These programs are short and most are readily
+        translatable to a wide variety of programming languages that
+        share the SMoL characteristics.
 
@@
 
-      \item We implement these misconceptions as interpreters,
-        which find weaknesses in our manual analysis and hold
-        potential for generative use in the future.
-
-  \item We present a tutoring system that measurably corrects these
-        misconceptions.
+  \item We implement these misconceptions as interpreters, which
+        find weaknesses in our manual analysis and hold potential for
+        generative use in the future.
+
+  \item We present a tutoring system, the SMoL Tutor, that measurably
+        corrects these misconceptions.
 
 \end{enumerate}
+
+In the near future, I would like to
 
@@
 versions of our programs in \supplRef{Translated\_Programs.html}. Furthermore, we intend to
-make a multi-lingual tutor in the future (\Cref{s:discussion}).
+make a multi-lingual tutor in the future (\Cref{s:future-work}).
 
@@
 
+Because we don't randomize Tutor question order, it is hard to tell
+whether the general decreasing trend in \cref{f:tutor-trend-plots}
+shows misconceptions being corrected, or shows properties of the
+questions (e.g., maybe later questions in the plots are easier because
+their programs happen to be shorter).
+
 \subsection{External Validity}
@@
 supports multiple, and more traditional, syntaxes should help address
-this issue. We defer this to future work.
-
-\section{Discussion}%
-\label{s:discussion}
-
-The paper has already identified several areas for future work:
-\begin{itemize}
-\item testing on more varied populations;
-\item using other syntaxes;
-\item having more questions that uniquely identify minsconceptions;
-  and,
-\item enabling textual responses in the Tutor so we can better
-  characterize wrong answers.
-\end{itemize}
-These can help get us even closer to a good approximation of a true concept
-inventory. We would also like the Tutor to make more use of the education
-theories discussed in \Cref{s:background}.
-
-We focus here on some issues that we think warrant broader
-discussion.
-
-\paragraph{The State-Aliasing-Function Triangle}
-
-Our problematic programs hardly include any ``advanced'' programming
-features: there are no threads, asynchrony, sophisticated type systems, ownership,
-inversion of control, etc. Indeed, many of those features typically
-build on an understanding of these basics (e.g., it is hard to make
-sense of ownership \cite{clarkeOwnershipTypesFlexible1998} without a good understanding of this
-triangle). But many populations seem to struggle even with this, which
-may explain why concepts like ownership are considered hard \cite{crichtonGroundedConceptualModel2023}.
-
-It is worth noting that we find problems even without all three
-components, as \Cref{s:smol-quizzes} shows! Nevertheless, we think it
-would be useful for curricula to focus on achieving mastery of this
-triangle. This may require a deep revision of widely accepted
-pedagogy. For instance, it is common to explain variables as
-``boxes''. But if taken seriously by a learner, this metaphor may
-cause more harm than good.
-As prior research \cite{putnamSummaryMisconceptionsHigh1986,groverMeasuringStudentLearning2017,hermansThinkingOutBox2018} shows, students expect that a variable can then contain more
-than one value, removing one makes the other accessible, etc.  That
-research has not explored aliasing, but the metaphor may affect that
-too. A box is a closed object, so a (mutable) value put ``into'' it
-clearly can't be modified by another ``box'' (variable)---i.e., it not
-only doesn't explain but is antithetical to aliasing.
-
-\paragraph{Terminology}
-
-It is common to teach programming using the terms ``call-by-value''
-and ``call-by-reference''. The reader will note that we instead have
-\emph{three} \miscon{ByRef} misconceptions.
-
-The emphasis on ``calling'' suggests that the semantics is associated
-\emph{only} with calls. That leaves open what happens when one simply
-binds a variable. In a reasonable language (and definitionally, in
-SMoL), the behavior is exactly the same: the formal parameter can be
-viewed as a binding to the value of the actual. But students often
-form inconsistent views because the ``call'' terminology breaks this
-deep similarity. We therefore recommend that languages in general use
-the term \emph{bind-by-}, to emphasize the underlying semantic unity
-of these syntactically different mechanisms.
-
-We feel that further confusion is caused by terms like ``pass'' and
-``return''. We often vocalize the call \verb|(f x)| as ``passing
-\verb|x| to \verb|f|''; saying ``passing \emph{the value of} \verb|x|
-to \verb|f|'' is a mouthful. But is it then surprising that students
-think \verb|x| is being aliased? Similarly, consider a statement like
-\verb|return y| (in Python syntax). Of course, semanticists understand
-that it is the \emph{value} of \verb|y|, not \verb|y| itself, that is
-being returned. Nevertheless, it is not surprising if this also
-results in assumptions that \verb|y| is either aliased or that it has
-escaped from the function (leading to an interpretation of dynamic
-scope). We believe there is a need for significant research to
-investigate these kinds of effects, which are similar to but not
-strictly the same as ``vernacular misconceptions'' \cite{nationalresearchcouncilScienceTeachingReconsidered1997}.
+this issue. We defer this to future work (\cref{s:future-work}).
+
+\section{Future Work \& Timeline}%
+\label{s:future-work}
+
+This semester (Fall 2023) I am already deploying a newer SMoL Tutor at
+Brown. This new tutor is able to display programs in JavaScript and
+Python, in addition to the Lispy syntax. The goal of this semester is
+to compare the new data with the current results, i.e., a comparison
+between uni-syntax tutor and multi-syntax tutor.
+
+By the end of Fall 2023, I would like to improve the SMoL Tutor:
+\begin{enumerate}
+  \item Ask students to explained when they gave a wrong answer in a
+  interepreting question. With student explanations, I check whether
+  choosing the wrong answer indicates holding the corresponding misconception(s).
+  \item Adjust the collection of interpreting questions such that
+  every misconception has at least two questions that aims at
+  correcting the misconception.
+  \item Partially randomize the order of questions.
+  \item Try to reduce the chance of missing values.
+\end{enumerate}
+
+Several faculties from other institutions have expressed interest in
+using the Tutor in Spring 2024. My goal is to collect data from at
+least two other institutions. I am officially taking a personal leave
+in Spring 2024, so I do not plan to make more progress other than
+data collection.
+
+During Fall 2024, I plan to collect a new round of data from Brown. I
+expect this datasets and the datasets from Spring 2024 would give a
+more rigorously evaluation of the effectiveness of SMoL Tutor. The
+improvements to be made in Fall 2023 should resolve many of the
+threats to validity. In this semester I also want to write a new paper
+about the SMoL Tutor.
+
+I plan to defense in Spring 2025.
 

### 2023-10-26 proposal.tex
complete
clean up

--- 2023-10-26 paper.tex
@@
 \addbibresource{bibfile.bib}
+\usepackage{hyperref}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 
-  In this paper, we present a curated list of misconceptions, and an
+  In this proposal, I present a curated list of misconceptions, and an
   instrument to detect them. These are distilled from student work
-  over several years and match and extend prior research.  We also
+  over several years and match and extend prior research.  I also
   present an automated, self-guided tutoring system. The tutor builds
@@
 
-  We have tested the tutor in multiple settings.
-  Our data consistently show that (a) the misconceptions we tackle are
+  I have tested the tutor in multiple settings.
+  Our data consistently show that (a) the misconceptions I tackle are
   widespread, and (b) the tutor appears to improve understanding.
+
+  [FILL future work]
 \end{abstract}
@@
   \item[RQ2] What would be a tutoring system (the SMoL Tutor) that
-  draws on cognitive and educational psychology concepts and that aim
-  at correcting the misconceptions?
+  draws on cognitive and educational psychology concepts, utilizes a
+  notional machine (the Stacker), and aims at correcting the
+  misconceptions?
   \item[RQ3] How effective is SMoL Tutor at correcting the
@@
 
-  \item We provide a collection of brief programs for detecting
+  \item I provide a collection of brief programs for detecting
         misconceptions about the features of SMoL and their
@@
 
-  \item We show that multiple populations have problems with these
+  \item I show that multiple populations have problems with these
         programs.
 
-  \item We present a list of curated misconceptions generated
+  \item I present a list of curated misconceptions generated
         after multiple rounds of analysis.
 
-  \item We implement these misconceptions as interpreters, which
+  \item I implement these misconceptions as interpreters, which
         find weaknesses in our manual analysis and hold potential for
@@
 
-  \item We present a tutoring system, the SMoL Tutor, that measurably
+  \item I present a tutoring system, the SMoL Tutor, that measurably
         corrects these misconceptions.
@@
 
-In the near future, I would like to
-
-\paragraph{A Note on Terminology} We use the term
+This document first presents the current results and ends with future
+work \& timeline (\cref{s:future-work}).
+
+\paragraph{A Note on Terminology} I use the term
 ``behavior'' to refer to the meaning of programs in terms of the
@@
 introduction to ``Greek'' notation. Because that is not the kind of
-tutor we are describing, to avoid confusion, we use the term
+tutor I are describing, to avoid confusion, I use the term
 ``behavior'' instead.
@@
 a quality survey), and indeed whole conferences are dedicated to
-them. We draw on this literature. In particular, it is common in the
+them. I draw on this literature. In particular, it is common in the
 literature to talk about a ``two-loop'' architecture \cite{vanlehnBehaviorTutoringSystems2006a} where the outer
-loop iterates through ``tasks'' (i.e., educational activities) and the inner loop iterates through UI events within a task. We follow the same
+loop iterates through ``tasks'' (i.e., educational activities) and the inner loop iterates through UI events within a task. I follow the same
 structure in our tutor (\Cref{s:tutor-uiux}).
@@
 programming background and focuses entirely on program behavior misconceptions
-and correcting them. We are not aware of a tutoring system (in computer science) that has
+and correcting them. I are not aware of a tutoring system (in computer science) that has
 this specific design.
@@
 however, it also proved to be pedagogically valuable.
-We have found the parentheses useful when discussing
+I have found the parentheses useful when discussing
 scope in conjunction with local-binding features like \lstinline|let|.
@@
 Nevertheless, most SMoL programs are easy to translate to other
-languages. We present machine-translated Python and JavaScript
+languages. I present machine-translated Python and JavaScript
 versions of our programs in \supplRef{Translated\_Programs.html}. Furthermore, we intend to
@@
 (As in Racket, arguments evaluate left-to-right, to give stateful programs an
-unambiguous semantics.) The implementation is available online
-[URL anonymized].
-
-\section{Paper Roadmap}
+unambiguous semantics.) The implementation is available online:
+\href{https://github.com/shriram/smol}{https://github.com/shriram/smol}.
+
+\section{Proposal Roadmap}
 \label{s:roadmap}
@@
 
-This paper describes a four-year effort to obtain a quality instrument
+This proposal describes a four-year effort to obtain a quality instrument
 to measure misconceptions and to build a tutor to address them.
@@
 
-We then present the tutor (\Cref{s:system-tutor}). In particular, we
+I then present the tutor (\Cref{s:system-tutor}). In particular, we
 present the list of misconceptions---grounded in student data---that
-are the focus of this project (\Cref{s:misconceptions-tutor}). We also
+are the focus of this project (\Cref{s:misconceptions-tutor}). I also
 evaluate the tutor's effectiveness at correcting them
@@
   but also has a strong inkling of \emph{precisely what} misconception
-  that group has and can address it directly. We expect our
+  that group has and can address it directly. I expect our
   instruments to be useful in the same way.}
 
-All these, however, add up to a somewhat challenging demand. We want
+All these, however, add up to a somewhat challenging demand. I want
 to produce a list of questions (each one an MCQ) such that
@@
 and so on.
-We therefore manually curated the Quizius output
+I therefore manually curated the Quizius output
 to address these issues.
@@
 
-We therefore turned the programs into a set of quizzes (in the US
+I therefore turned the programs into a set of quizzes (in the US
 sense: namely, a brief test of knowledge) that we call the SMoL
@@
 
-Question orders were partially randomized. We wanted students to get
+Question orders were partially randomized. I wanted students to get
 some easy, warm-up questions initially, so those were kept at the
@@
 Syntactically, \verb|#t| and \verb|#f| are true and false,
-while \verb|#(...)| is a vector. We use a * to indicate the correct
+while \verb|#(...)| is a vector. I use a * to indicate the correct
 answer (which also matches the implementation's output).
@@
 but it is unclear to what extent students read, understood, or
-internalized these. We also wanted to determine whether other
+internalized these. I also wanted to determine whether other
 populations run into these issues, and it was unclear whether we would
@@
 not only to find faults but to improve the understanding of basic
-programming language behavior. We believe (and presumably so does
+programming language behavior. I believe (and presumably so does
 anyone else who writes a formal semantics!)\ that an understanding of
@@
 SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
-We then discuss what
+I then discuss what
 we learned from its data (\Cref{s:misconceptions-tutor}), and finally
@@
 introduces terminology and states the true conceptions. These are the
-teaching goals of the Tutor. We therefore refer to these as
+teaching goals of the Tutor. I therefore refer to these as
 \newterm{goal sentences}. \Cref{t:tutor-goal-sentences} lists
@@
 this paper.
-We bootstrap it from the SMoL
+I bootstrap it from the SMoL
 Quizzes, but made the following alterations:
@@
 
-  \item We renamed vector operators to avoid confusion between
+  \item I renamed vector operators to avoid confusion between
         \lstinline|vset!|, which replaces a vector element, and \lstinline|set!|,
-        which mutates variables. We rename all vector operators from, for example,
+        which mutates variables. I rename all vector operators from, for example,
         \lstinline|vset!| to \lstinline|vec-set!|.
 
-  \item We deferred the local binding forms to the end, since we felt
-        they were less essential. We therefore rewrote programs that use
+  \item I deferred the local binding forms to the end, since we felt
+        they were less essential. I therefore rewrote programs that use
         it to not depend on it.
 
-        % \item We removed some questions that we deemed ``Lispy'' (e.g.,
+        % \item I removed some questions that we deemed ``Lispy'' (e.g.,
         %   involving the difference between parenthetical and infix syntax).
 
-  \item We removed some programs that relied on underspecified aspects or
+  \item I removed some programs that relied on underspecified aspects or
         unimportant aspects of SMoL. For example, some depended on
@@
 
-  \item We resolved ambiguities in some programs, either adding
+  \item I resolved ambiguities in some programs, either adding
     answer choices or even adding other questions to tease
@@
 
-  \item We removed questions related to function equality, which was
+  \item I removed questions related to function equality, which was
         not a focus of the Tutor.\footnote{However, we believe there is
@@
 
-  \item We removed programs of a ``Lispy'' nature, such as one where
+  \item I removed programs of a ``Lispy'' nature, such as one where
     the answer depended on whether the reader correctly understood
@@
 only 8 out of the 89 questions have three or four choices; 70
-questions have 5--8 choices; and 11 questions have 9--14 choices. We hope
+questions have 5--8 choices; and 11 questions have 9--14 choices. I hope
 this reduced both guessing and elimination, and forced students to
@@
 
-We now examine what we learned from the interpreting
+I now examine what we learned from the interpreting
 questions in terms of program behavior
@@
 
-We do not present the results of \textbf{begin} and
+I do not present the results of \textbf{begin} and
 \textbf{local} (\Cref{t:tutor-goal-sentences}), because they focus on constructs not usually found in
@@
 interpretations for existing program-answer pairs.
-We provide all the misinterpreters in the artifact.
+I provide all the misinterpreters in the artifact.
 
@@
 
-We iteratively created
+I iteratively created
 our final catalog of misconceptions
 (from the perspective of the data in this paper).
-We started
+I started
 with misinterpreters representing the misconceptions described in
@@
 have reasonable validation (due to the prose in Quizius), so we call
-them \emph{grounded} misconceptions. We then looked at wrong answers
+them \emph{grounded} misconceptions. I then looked at wrong answers
 not covered by these but chosen by students, and did our best to
@@
   easily modify it to do so a small number of times per user, to make
-  progress towards this need.} We then re-ran the misinterpreters
-against the chosen answers. We terminated when all the remaining wrong
+  progress towards this need.} I then re-ran the misinterpreters
+against the chosen answers. I terminated when all the remaining wrong
 answers were either (a) found in very few students (we found a gap
@@
   misinterpretation; but we presumably did not find evidence of it in our
-  data. We do welcome other misinterpretations, for which we
+  data. I do welcome other misinterpretations, for which we
   can add more misinterpreters!}
@@
 \end{lstlisting}
-We had assumed the wrong answer
+I had assumed the wrong answer
 \texttt{0} is caused by \miscon{CallByRef}.
@@
 The problem here is that the word ``redefine'' underspecifies how the
-second definition is interpreted. We had interpreted it as
+second definition is interpreted. I had interpreted it as
 \textit{mutating} the binding established by the first definition,
@@
 To summarize, we used the SMoL Tutor, with an expanded set of
-programs, to investigate student misconceptions. We implemented the
+programs, to investigate student misconceptions. I implemented the
 idea of misinterpreters to help us properly classify student
-performance. We were able to reconfirm all the previously identified
+performance. I were able to reconfirm all the previously identified
 misconceptions, refine some of them, and also identify potential new
@@
 matches that produced by that misinterpreter, and it matches only that
-misinterpreter). We then examine student performance over time across
+misinterpreter). I then examine student performance over time across
 those questions. This would let us examine how they do on just that
@@
 
-We present the result of this analysis in two forms. Graphically, we
+I present the result of this analysis in two forms. Graphically, we
 show plots in \Cref{f:tutor-trend-plots}. Each figure shows the percentage of students who
@@
 
-We also perform a logistic regression to see whether these
+I also perform a logistic regression to see whether these
 improvements are significant (at a $p < 0.05$ threshold); details are
@@
 
-We therefore view the above data as
+I therefore view the above data as
 purely formative: they suggest that the Tutor \emph{most likely did
@@
 Tutor, we will use misinterpreters actively to shape the incorrect answers and,
-as necessary, update the programs as well. We therefore hope to have
+as necessary, update the programs as well. I therefore hope to have
 much more thorough analyses in future work.
@@
 thereby actually reflecting the state of its curriculum rather than
-some greater truth about program behavior understanding.  We already
+some greater truth about program behavior understanding.  I already
 have some reason to believe this is not the case: the related work
@@
 \end{description}
-We don't include \surmisedMiscon{CallerEnv} in our analysis because in our
+I don't include \surmisedMiscon{CallerEnv} in our analysis because in our
 data, all wrong answers that can be explained by \surmisedMiscon{CallerEnv}
@@
 \textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
-We only used these sentences (or rules) to guide the
+I only used these sentences (or rules) to guide the
 design of the Tutor and as text in the Tutor itself.
@@
 A particularly notable threat is the use of Lispy
-syntax. We chose it for two reasons: both because it helps clarify
+syntax. I chose it for two reasons: both because it helps clarify
 scope (\Cref{s:smol}) and because the rest of the course used
@@
 
-Is our reasoning valid? We have applied standard techniques and
+Is our reasoning valid? I have applied standard techniques and
 measurements for evaluating student responses. However, our
@@
 educator, even if the precise misconception cannot be pinned down with
-the highest accuracy. We therefore believe our instruments, and our
+the highest accuracy. I therefore believe our instruments, and our
 misinterpreter technique, are of general value.
@@
 this had a noticeable impact. (In addition, every wrong answer is
-still wrong! We may just not have exactly the right proportions of
+still wrong! I may just not have exactly the right proportions of
 them.)
@@
 supports multiple, and more traditional, syntaxes should help address
-this issue. We defer this to future work (\cref{s:future-work}).
+this issue. I defer this to future work (\cref{s:future-work}).
 

### 2023-10-28 proposal.tex
lift lit review

@@
 outcomes.
+
+\section{Background: Misconceptions}
+
+\begin{table}[t]
+  \centering
+  \begin{tabularx}{\textwidth}{>{\hsize=.6\hsize\RaggedRight}X >{\hsize=.8\hsize\RaggedRight}X >{\hsize=.5\hsize\RaggedRight}X >{\hsize=1.4\hsize\RaggedRight}X}
+    \textbf{Publication}                       &
+    \textbf{Population}                       &
+    \textbf{Languages}                       &
+    \textbf{Misconceptions}
+    \\\hline
+    \cite{fleuryParameterPassingRules1991} &
+    Unclear, likely CS2 students &
+    Pascal &
+    \surmisedMiscon{CallerEnv}; \miscon{IsolatedFun}; \miscon{DefOrSet}
+    (See their RULEs 2, 3a, and 3b in TABLE 2;
+    RULE 1 doesn't apply to us.)
+    \\\hline
+    \cite{goldmanIdentifyingImportantDifficult2008,goldmanSettingScopeConcept2010} &
+    CS1 students &
+    Java, Python, Scheme &
+    Scope and memory model (See their Figures 4 and 5)
+    \\\hline
+    \cite{fislerAssessingTeachingScope2017} &
+    Third- and fourth-year undergrads &
+    Java and Scheme &
+    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{CallsCopyStruct}; \miscon{DefByRef}
+    (See their Section 4)
+    \\\hline
+    \cite{saarinenHarnessingWisdomClasses2019} &
+    CS2 students &
+    Java &
+    \miscon{StructByRef} (Their G2); \miscon{DefByRef} or \miscon{DefCopyStructs} (Their G3); \miscon{CallsCopyStruct} (Their G4).
+    \\\hline
+    \cite{strombackProgressionStudentsAbility2023} &
+    CS masters &
+    Python &
+    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
+    (See their section 4.2)
+    \\\hline
+    \cite{strombackProgressionStudentsAbility2023} &
+    CS undergrads &
+    C++ &
+    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
+    (See their section 4.2)
+    \\\hline
+  \end{tabularx}
+  \caption{Similar misconceptions found in prior research.}
+  \label{t:rel-work}
+\end{table}
+
+Misconceptions related to scope, mutation, and higher-order functions have been
+widely identified in varying populations
+(from CS1 students to graduate students to users of online forums)
+over many years (since at last 1991)
+in varying programming languages (from Java to Racket) and in
+different countries (such as the USA and Sweden).
+\Cref{t:rel-work} lists works that seem most relevant to us.
+
+There are some small differences.
+\cite{fleuryParameterPassingRules1991} identifies a dynamic scope misconception
+that is different from \miscon{FlatEnv}:
+\begin{description}
+  \item[\surmisedMiscon{CallerEnv}] Function values don't remember their environments.
+  When a function is called, the function body is evaluated in an
+  environment that extends from the \textit{caller's} environment.
+\end{description}
+I don't include \surmisedMiscon{CallerEnv} in our analysis because in our
+data, all wrong answers that can be explained by \surmisedMiscon{CallerEnv}
+are also explainable by \miscon{FlatEnv}.
+
+Appendix A of \cite{sorvaVisualProgramSimulation2012} provides an
+extensive survey of misconceptions reported
+in research up to 2012.
+There are overlaps between our survey and
+theirs. For instance, our \surmisedMiscon{CallerEnv} is their No.~47.
+However, because their descriptions are brief, it is difficult to tell
+whether a misconception in their survey matches our misconceptions.
+Because
+they provide neither misinterpreters nor representative program-output
+pairs, it is difficult to determine the overlaps precisely (showing
+the value of providing these two machine-runnable descriptions).
+At any rate,
+we certainly find
+no equivalent of \miscon{FlatEnv}, \miscon{DeepClosure}, and
+\miscon{DefByRef} in their survey.
+
+\section{Background: Program Behavior}
+
+Our idea of ``goal sentences'' is not substantially different from the
+\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
+I only used these sentences (or rules) to guide the
+design of the Tutor and as text in the Tutor itself.
+\cite{duranRulesProgramBehavior2021} argue for other uses of such
+sentences.
+
+\section{Research Methodology}%
+\label{s:research-methodology}
+
+\subsection{\textbf{RQ1:} What are the misconceptions that are about program behavior, apply to most modern programming languages, and present (even) in students with prior programming backgrounds?}
+
+This research question imposes three predicates on misconceptions:
+\begin{enumerate}
+  \item are about program behavior,
+  \item apply to most modern programming languages, and
+  \item present (even) in students with prior programming background
+\end{enumerate}
+The first predicate is the probably the most subtle, so I will talk
+about it last. I approach the second predicate by modeling the
+``consensus'' of modern programming language with SMoL. Because SMoL
+represents the semantic core, misconceptions about SMoL presumably
+also apply to most modern PLs. I approach the third and last predicate
+by collecting data in population that very likely have prior experience.
+For example, my priliminary results are build upon data collected in
+a upper-year programming language course.
+
+Let me come back to the first predicate. I would like to measure misconceptions
+with tasks like the following:
+
+[FILL]
+
+I think tasks like the following
+is best suited for measuring program-behavior misconceptions, in comparison
+to tasks commonly found in literature.
+
+% To find misconceptions about
+% program behavior, I need to
+% \begin{enumerate}
+%   \item define what counts as a misconception about program behavior, and
+%   \item describe how to find such a misconception
+% \end{enumerate}
+% Only conception that causes a student to predict a wrong result of a program
+% counts as a misconception about program behavior. Here are a few example
+% of misconceptions that I do \textit{not} consider about program behavior:
+% \begin{enumerate}
+%   \item \texttt{[1, 2, 3]} allocates 3 bytes. [CITE]
+%   \item Variables reside on heap [CITE]
+% \end{enumerate}
+% I do not consider them because they go too deep into implementation
+% details [FILL]. By limiting the scope to program behavior, I suspect
+% the misconceptions I found would be more relevant to most reading,
+% writing, and debugging tasks.
+
+% Misconceptions about program behavior should not be specific to
+% certain variable names, constants, or indentation. [FILL]
 
@@
 
-\section{Related Work}
-\label{s:rel-work}
-
-\begin{table}[t]
-  \centering
-  \begin{tabularx}{\textwidth}{>{\hsize=.6\hsize\RaggedRight}X >{\hsize=.8\hsize\RaggedRight}X >{\hsize=.5\hsize\RaggedRight}X >{\hsize=1.4\hsize\RaggedRight}X}
-    \textbf{Publication}                       &
-    \textbf{Population}                       &
-    \textbf{Languages}                       &
-    \textbf{Misconceptions}
-    \\\hline
-    \cite{fleuryParameterPassingRules1991} &
-    Unclear, likely CS2 students &
-    Pascal &
-    \surmisedMiscon{CallerEnv}; \miscon{IsolatedFun}; \miscon{DefOrSet}
-    (See their RULEs 2, 3a, and 3b in TABLE 2;
-    RULE 1 doesn't apply to us.)
-    \\\hline
-    \cite{goldmanIdentifyingImportantDifficult2008,goldmanSettingScopeConcept2010} &
-    CS1 students &
-    Java, Python, Scheme &
-    Scope and memory model (See their Figures 4 and 5)
-    \\\hline
-    \cite{fislerAssessingTeachingScope2017} &
-    Third- and fourth-year undergrads &
-    Java and Scheme &
-    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{CallsCopyStruct}; \miscon{DefByRef}
-    (See their Section 4)
-    \\\hline
-    \cite{saarinenHarnessingWisdomClasses2019} &
-    CS2 students &
-    Java &
-    \miscon{StructByRef} (Their G2); \miscon{DefByRef} or \miscon{DefCopyStructs} (Their G3); \miscon{CallsCopyStruct} (Their G4).
-    \\\hline
-    \cite{strombackProgressionStudentsAbility2023} &
-    CS masters &
-    Python &
-    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
-    (See their section 4.2)
-    \\\hline
-    \cite{strombackProgressionStudentsAbility2023} &
-    CS undergrads &
-    C++ &
-    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
-    (See their section 4.2)
-    \\\hline
-  \end{tabularx}
-  \caption{Similar misconceptions found in prior research.}
-  \label{t:rel-work}
-\end{table}
-
-\paragraph{Misconceptions}
-
-Misconceptions related to scope, mutation, and higher-order functions have been
-widely identified in varying populations
-(from CS1 students to graduate students to users of online forums)
-over many years (since at last 1991)
-in varying programming languages (from Java to Racket) and in
-different countries (such as the USA and Sweden).
-\Cref{t:rel-work} lists works that seem most relevant to us.
-
-There are some small differences.
-\cite{fleuryParameterPassingRules1991} identifies a dynamic scope misconception
-that is different from \miscon{FlatEnv}:
-\begin{description}
-  \item[\surmisedMiscon{CallerEnv}] Function values don't remember their environments.
-  When a function is called, the function body is evaluated in an
-  environment that extends from the \textit{caller's} environment.
-\end{description}
-I don't include \surmisedMiscon{CallerEnv} in our analysis because in our
-data, all wrong answers that can be explained by \surmisedMiscon{CallerEnv}
-are also explainable by \miscon{FlatEnv}.
-
-Appendix A of \cite{sorvaVisualProgramSimulation2012} provides an
-extensive survey of misconceptions reported
-in research up to 2012.
-There are overlaps between our survey and
-theirs. For instance, our \surmisedMiscon{CallerEnv} is their No.~47.
-However, because their descriptions are brief, it is difficult to tell
-whether a misconception in their survey matches our misconceptions.
-Because
-they provide neither misinterpreters nor representative program-output
-pairs, it is difficult to determine the overlaps precisely (showing
-the value of providing these two machine-runnable descriptions).
-At any rate,
-we certainly find
-no equivalent of \miscon{FlatEnv}, \miscon{DeepClosure}, and
-\miscon{DefByRef} in their survey.
-
-\paragraph{Goal Sentences}
-
-Our idea of ``goal sentences'' is not substantially different from the
-\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
-I only used these sentences (or rules) to guide the
-design of the Tutor and as text in the Tutor itself.
-\cite{duranRulesProgramBehavior2021} argue for other uses of such
-sentences.
-
-\paragraph{Misinterpreters}
+\section{Threats to Validity}
+\label{s:threats}
+
+\subsection{Construct Validity}
+
+Did we measure the right thing?
+While our goal is to study student understanding of language
+behavior, what we actually measure is performance on MCQs on select
+programs. MCQs as a mechanism introduce various clear biases, though
+we add the ``Other'' options to somewhat alleviate them. The small
+size, syntactic details (such as intentionally meaningless variable
+names), and other aspects of the programs may also impact our ability
+to measure understanding. (This could go both ways: students may have
+no trouble understanding behavior in a small program but may struggle
+to do so in a larger one.)
+
+A particularly notable threat is the use of Lispy
+syntax. I chose it for two reasons: both because it helps clarify
+scope (\Cref{s:smol}) and because the rest of the course used
+Racket. However, students may well perform differently with more
+familiar syntaxes. It seems unlikely their performance would be \emph{too}
+different given the many languages that have produced similar
+misconceptions (\Cref{t:rel-work}). Nevertheless, we intend to use a
+variety of syntaxes to examine this issue further.
+
+Finally, we curated programs by hand, so they may well reflect our own
+biases about misconceptions. It may be possible to mitigate this
+problem by synthesizing MCQ programs using the misinterpreters.
+
+\subsection{Internal Validity}
+
+Is our reasoning valid? I have applied standard techniques and
+measurements for evaluating student responses. However, our
+instruments still lack the validity of a proper concept
+inventory. Their creation requires heavyweight processes (such as
+Delphi methods \cite{goldmanSettingScopeConcept2010}) that require many hours of expert attention as
+well as conversations with learners, and can hence be prohibitive in
+cost. The Quizius method (\Cref{s:quizius}) was created precisely to
+be an inexpensive method that provides a good proxy.
+
+Ultimately, our goal is to provide a reasonable instrument for
+widespread use. While the set of all misconceptions could be
+unbounded, we believe our tasks, especially as embedded in the Tutor,
+provide a good starting point for others. In particular, if students
+select a wrong answer, that is still of \emph{some} use to an
+educator, even if the precise misconception cannot be pinned down with
+the highest accuracy. I therefore believe our instruments, and our
+misinterpreter technique, are of general value.
+
+A failure in our Tutor's logging infrastructure led us to miss some
+responses. These are unlikely to be task-specific because the Tutor
+has a generic framework that should perform the same across
+tasks. Moreover, on average only $0.4\%\ (\textrm{sd} = 0.6\%)$ of values are missing. Therefore, we do not believe
+this had a noticeable impact. (In addition, every wrong answer is
+still wrong! I may just not have exactly the right proportions of
+them.)
+
+Because we don't randomize Tutor question order, it is hard to tell
+whether the general decreasing trend in \cref{f:tutor-trend-plots}
+shows misconceptions being corrected, or shows properties of the
+questions (e.g., maybe later questions in the plots are easier because
+their programs happen to be shorter).
+
+\subsection{External Validity}
+
+How well do our results generalize? Certainly there is reason to
+question whether our results would apply to other
+populations. \Cref{s:generality-tutor} provides preliminary evidence
+that the results are not specific to one institution, and
+\Cref{s:rel-work} suggests these issues are widespread. Nevertheless,
+much more broad testing is needed to confirm our specific instruments.
+
+The other major concern is the tie to Lispy syntax. Learners in other
+settings may do worse or even better with it. Building a Tutor that
+supports multiple, and more traditional, syntaxes should help address
+this issue. I defer this to future work (\cref{s:future-work}).
+
+\section{Related Work}%
+\label{s:related-work}
+
+\subsection{Misinterpreters}
 
@@
 
-\section{Threats to Validity}
-\label{s:threats}
-
-\subsection{Construct Validity}
-
-Did we measure the right thing?
-While our goal is to study student understanding of language
-behavior, what we actually measure is performance on MCQs on select
-programs. MCQs as a mechanism introduce various clear biases, though
-we add the ``Other'' options to somewhat alleviate them. The small
-size, syntactic details (such as intentionally meaningless variable
-names), and other aspects of the programs may also impact our ability
-to measure understanding. (This could go both ways: students may have
-no trouble understanding behavior in a small program but may struggle
-to do so in a larger one.)
-
-A particularly notable threat is the use of Lispy
-syntax. I chose it for two reasons: both because it helps clarify
-scope (\Cref{s:smol}) and because the rest of the course used
-Racket. However, students may well perform differently with more
-familiar syntaxes. It seems unlikely their performance would be \emph{too}
-different given the many languages that have produced similar
-misconceptions (\Cref{t:rel-work}). Nevertheless, we intend to use a
-variety of syntaxes to examine this issue further.
-
-Finally, we curated programs by hand, so they may well reflect our own
-biases about misconceptions. It may be possible to mitigate this
-problem by synthesizing MCQ programs using the misinterpreters.
-
-\subsection{Internal Validity}
-
-Is our reasoning valid? I have applied standard techniques and
-measurements for evaluating student responses. However, our
-instruments still lack the validity of a proper concept
-inventory. Their creation requires heavyweight processes (such as
-Delphi methods \cite{goldmanSettingScopeConcept2010}) that require many hours of expert attention as
-well as conversations with learners, and can hence be prohibitive in
-cost. The Quizius method (\Cref{s:quizius}) was created precisely to
-be an inexpensive method that provides a good proxy.
-
-Ultimately, our goal is to provide a reasonable instrument for
-widespread use. While the set of all misconceptions could be
-unbounded, we believe our tasks, especially as embedded in the Tutor,
-provide a good starting point for others. In particular, if students
-select a wrong answer, that is still of \emph{some} use to an
-educator, even if the precise misconception cannot be pinned down with
-the highest accuracy. I therefore believe our instruments, and our
-misinterpreter technique, are of general value.
-
-A failure in our Tutor's logging infrastructure led us to miss some
-responses. These are unlikely to be task-specific because the Tutor
-has a generic framework that should perform the same across
-tasks. Moreover, on average only $0.4\%\ (\textrm{sd} = 0.6\%)$ of values are missing. Therefore, we do not believe
-this had a noticeable impact. (In addition, every wrong answer is
-still wrong! I may just not have exactly the right proportions of
-them.)
-
-Because we don't randomize Tutor question order, it is hard to tell
-whether the general decreasing trend in \cref{f:tutor-trend-plots}
-shows misconceptions being corrected, or shows properties of the
-questions (e.g., maybe later questions in the plots are easier because
-their programs happen to be shorter).
-
-\subsection{External Validity}
-
-How well do our results generalize? Certainly there is reason to
-question whether our results would apply to other
-populations. \Cref{s:generality-tutor} provides preliminary evidence
-that the results are not specific to one institution, and
-\Cref{s:rel-work} suggests these issues are widespread. Nevertheless,
-much more broad testing is needed to confirm our specific instruments.
-
-The other major concern is the tie to Lispy syntax. Learners in other
-settings may do worse or even better with it. Building a Tutor that
-supports multiple, and more traditional, syntaxes should help address
-this issue. I defer this to future work (\cref{s:future-work}).
-
 \section{Future Work \& Timeline}%

### 2023-10-31 proposal.tex
plan for RQ1

--- 2023-10-26 proposal.tex
@@
 \newcommand{\pValue}[0]{\textit{p}\nobreakdash-value}
+\newcommand{\rqI}{What are the misconceptions that are about program behavior, apply to most modern programming languages, and present (even) in students with prior programming backgrounds?}
+\newcommand{\rqII}{What would be a tutoring system that aims at correcting the misconceptions, draws on cognitive and educational psychology concepts, utilizes a notional machine, and seems to be effective?}
 
@@
 \begin{description}
-  \item[RQ1] What are the misconceptions that are about program
-  behavior, apply to most modern programming languages, and present
-  (even) in students with prior programming backgrounds?
-  \item[RQ2] What would be a tutoring system (the SMoL Tutor) that
-  draws on cognitive and educational psychology concepts, utilizes a
-  notional machine (the Stacker), and aims at correcting the
-  misconceptions?
-  \item[RQ3] How effective is SMoL Tutor at correcting the
-  misconceptions?
+  \item[RQ1] \rqI
+  \item[RQ2] \rqII
 \end{description}
 
-To answer my research questions, I have (a) run a multi-year, multi-phase study to
-identify these misunderstandings while trying to work around the blind
-spots of experts (and also drawing on prior work),
-and (b) constructed the SMoL Tutor, an interactive tutor to address
-these misunderstandings. The SMoL Tutor is not a \emph{programming} tutor;
-rather, it assumes the user has some basic facility with programming
-(i.e., that they are already familiar with concepts like vectors,
-mutation, and higher-order functions). It is built around SMoL's
-features and their interactions, and draws on cognitive and
-educational psychology concepts to expressly identify
+To answer my research questions, I have (a) run a multi-year,
+multi-phase study to identify these misunderstandings while trying to
+work around the blind spots of experts (and also drawing on prior
+work), and (b) constructed the SMoL Tutor, an interactive tutor to
+address these misunderstandings. The SMoL Tutor is not a
+\emph{programming} tutor; rather, it assumes the user has some basic
+facility with programming (i.e., that they are already familiar with
+concepts like vectors, mutation, and higher-order functions). It is
+built around SMoL's features and their interactions, and draws on
+cognitive and educational psychology concepts to expressly identify
 and correct misconceptions.
@@
         corrects these misconceptions.
+
+  \item I present a notional machine, the Stacker, that is used by the
+  SMoL Tutor to explain programs and might be useful in itsown.
 
@@
 outcomes.
+
+\section{Background: Misconceptions}%
+\label{s:background-misconceptions}
+
+\begin{table}[t]
+  \centering
+  \begin{tabularx}{\textwidth}{>{\hsize=.6\hsize\RaggedRight}X >{\hsize=.8\hsize\RaggedRight}X >{\hsize=.5\hsize\RaggedRight}X >{\hsize=1.4\hsize\RaggedRight}X}
+    \textbf{Publication}                       &
+    \textbf{Population}                       &
+    \textbf{Languages}                       &
+    \textbf{Misconceptions}
+    \\\hline
+    \cite{fleuryParameterPassingRules1991} &
+    Unclear, likely CS2 students &
+    Pascal &
+    \surmisedMiscon{CallerEnv}; \miscon{IsolatedFun}; \miscon{DefOrSet}
+    (See their RULEs 2, 3a, and 3b in TABLE 2;
+    RULE 1 doesn't apply to us.)
+    \\\hline
+    \cite{goldmanIdentifyingImportantDifficult2008,goldmanSettingScopeConcept2010} &
+    CS1 students &
+    Java, Python, Scheme &
+    Scope and memory model (See their Figures 4 and 5)
+    \\\hline
+    \cite{fislerAssessingTeachingScope2017} &
+    Third- and fourth-year undergrads &
+    Java and Scheme &
+    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{CallsCopyStruct}; \miscon{DefByRef}
+    (See their Section 4)
+    \\\hline
+    \cite{saarinenHarnessingWisdomClasses2019} &
+    CS2 students &
+    Java &
+    \miscon{StructByRef} (Their G2); \miscon{DefByRef} or \miscon{DefCopyStructs} (Their G3); \miscon{CallsCopyStruct} (Their G4).
+    \\\hline
+    \cite{strombackProgressionStudentsAbility2023} &
+    CS masters &
+    Python &
+    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
+    (See their section 4.2)
+    \\\hline
+    \cite{strombackProgressionStudentsAbility2023} &
+    CS undergrads &
+    C++ &
+    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
+    (See their section 4.2)
+    \\\hline
+  \end{tabularx}
+  \caption{Similar misconceptions found in prior research.}
+  \label{t:rel-work}
+\end{table}
+
+Misconceptions related to scope, mutation, and higher-order functions have been
+widely identified in varying populations
+(from CS1 students to graduate students to users of online forums)
+over many years (since at last 1991)
+in varying programming languages (from Java to Racket) and in
+different countries (such as the USA and Sweden).
+\Cref{t:rel-work} lists works that seem most relevant to us.
+
+There are some small differences.
+\cite{fleuryParameterPassingRules1991} identifies a dynamic scope misconception
+that is different from \miscon{FlatEnv}:
+\begin{description}
+  \item[\surmisedMiscon{CallerEnv}] Function values don't remember their environments.
+  When a function is called, the function body is evaluated in an
+  environment that extends from the \textit{caller's} environment.
+\end{description}
+I don't include \surmisedMiscon{CallerEnv} in our analysis because in our
+data, all wrong answers that can be explained by \surmisedMiscon{CallerEnv}
+are also explainable by \miscon{FlatEnv}.
+
+Appendix A of \cite{sorvaVisualProgramSimulation2012} provides an
+extensive survey of misconceptions reported
+in research up to 2012.
+There are overlaps between our survey and
+theirs. For instance, our \surmisedMiscon{CallerEnv} is their No.~47.
+However, because their descriptions are brief, it is difficult to tell
+whether a misconception in their survey matches our misconceptions.
+Because
+they provide neither misinterpreters nor representative program-output
+pairs, it is difficult to determine the overlaps precisely (showing
+the value of providing these two machine-runnable descriptions).
+At any rate,
+we certainly find
+no equivalent of \miscon{FlatEnv}, \miscon{DeepClosure}, and
+\miscon{DefByRef} in their survey.
+
+\section{Background: Program Behavior}
+
+Our idea of ``goal sentences'' is not substantially different from the
+\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
+I only used these sentences (or rules) to guide the
+design of the Tutor and as text in the Tutor itself.
+\cite{duranRulesProgramBehavior2021} argue for other uses of such
+sentences.
+
+\section{Plan for \textbf{RQ1}}%
+\label{s:methodology-rq1}
+
+\begin{description}
+  \item[RQ1] \rqI
+\end{description}
+
+To ensure that I am studying program behavior rather than some
+implementation details, I plan to focus on the results of running
+programs. I approach the second predicate by modeling the
+``consensus'' of modern programming language with SMoL. Because SMoL
+represents the semantic core, misconceptions about SMoL presumably
+also apply to most modern PLs. I approach the third and last predicate
+by collecting data in population that very likely have prior
+experience. For example, my priliminary results are build upon data
+collected in a upper-year programming language course.
+
+In order to find a reasonably comprehensive collection of
+misconceptions, I seek for inputs from students. Most prior works
+(\Cref{s:background-misconceptions}) have experts coming up with a
+list of potential misconceptions and then filter the potential
+misconceptions with student inputs. This approach can't possibly find
+misconceptions not considered by experts and hence is necessarily
+affected by experts' blind spots. I plan to overcome experts'
+blind spots by avoid presuming misconceptions. Concretely, my first
+step was a two-year process to find (a) student-written programs that
+tend to trip up students and (b) incorrect responses to those programs
+(\Cref{s:gen-prob-quizius}), using a tool for the purpose
+(\Cref{s:quizius}).
+
+Another source of misconceptions are student responses in the tutoring
+system. The tutoring system, which I will explain in details when I
+talk about RQ2 (\Cref{s:methodology-rq2}), includes many
+multiple-choice questions (MCQs) that asks students to predict the
+result(s) of running a program. MCQ is a commonly used task format for
+detecting misconceptions [CITE]. MCQs in their own only detect
+\emph{patterns of errors}. To detect \emph{misconceptions} -- student
+conceptions that produces pattern of errors, I need student-written
+explanations for their wrong answers. To this end, I plan to ask a
+follow up questions after the MCQs: when students make a wrong choice,
+the follow-up asks \emph{why} they made the choice.
+
+I expect to find multiple misconceptions. Each misconception is
+represented by a collection of program-result-explanation triples.
+There should be a small number of programs. For each of these
+programs, a wrong result is commonly predicted by students from
+various population. Furthermore, these students give similar explanations
+for all the wrong answers. I also would like to provide a definitional
+interpreter for the misconception.
+
+For example, I might find a misconception called \textbf{DefByRef}:
+when students are asked to predict the result of
+\begin{lstlisting}
+  (defvar x 12)
+  (defvar y x)
+  (set! y 0)
+  x
+  y
+\end{lstlisting}
+many of them responded \lstinline|12 0|; and when asked to predict
+\begin{lstlisting}
+  (defvar x 12)
+  (defvar y x)
+  (set! x 0)
+  x
+  y
+\end{lstlisting}
+many of them responded \lstinline|0 12|; and these students all
+explained something like ``\lstinline|y| is \lstinline|x|''. To
+represent this misconception, I provide a interpreter that almost
+works like the referential interperter of SMoL, but \lstinline|defvar|
+aliases variables.
+
+\section{Plan for \textbf{RQ2}}%
+\label{s:methodology-rq2}
+
+\begin{description}
+  \item[RQ2] \rqII
+\end{description}
+
+Prior works suggest that presenting refutation text and encouraging
+students to compare cases are helpful for learning, and that notional
+machines are useful for explaning program behavior.
+
+I have built a tutor that incorporate these ideas
+(\Cref{s:system-tutor}).
+
+This tutor uses a lot of MCQs illustrated in
+\Cref{f:smol-tutor-example}.
 
@@
   \item Partially randomize the order of questions.
-  \item Try to reduce the chance of missing values.
 \end{enumerate}

### 2023-11-01 proposal.tex
clean up RQ1

--- 2023-10-26 proposal.tex
@@
 \newcommand{\pValue}[0]{\textit{p}\nobreakdash-value}
+\newcommand{\rqI}{What are the program-behavior misconceptions that apply to most modern programming languages and present (even) in students with prior programming backgrounds?}
+\newcommand{\rqII}{What would be a tutoring system that aims at correcting the misconceptions, draws on cognitive and educational psychology concepts, utilizes a notional machine, and seems to be effective?}
 
@@
 outcomes.
+
+\section{Background: Misconceptions}%
+\label{s:background-misconceptions}
+
+\begin{table}[t]
+  \centering
+  \begin{tabularx}{\textwidth}{>{\hsize=.6\hsize\RaggedRight}X >{\hsize=.8\hsize\RaggedRight}X >{\hsize=.5\hsize\RaggedRight}X >{\hsize=1.4\hsize\RaggedRight}X}
+    \textbf{Publication}                       &
+    \textbf{Population}                       &
+    \textbf{Languages}                       &
+    \textbf{Misconceptions}
+    \\\hline
+    \cite{fleuryParameterPassingRules1991} &
+    Unclear, likely CS2 students &
+    Pascal &
+    \surmisedMiscon{CallerEnv}; \miscon{IsolatedFun}; \miscon{DefOrSet}
+    (See their RULEs 2, 3a, and 3b in TABLE 2;
+    RULE 1 doesn't apply to us.)
+    \\\hline
+    \cite{goldmanIdentifyingImportantDifficult2008,goldmanSettingScopeConcept2010} &
+    CS1 students &
+    Java, Python, Scheme &
+    Scope and memory model (See their Figures 4 and 5)
+    \\\hline
+    \cite{fislerAssessingTeachingScope2017} &
+    Third- and fourth-year undergrads &
+    Java and Scheme &
+    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{CallsCopyStruct}; \miscon{DefByRef}
+    (See their Section 4)
+    \\\hline
+    \cite{saarinenHarnessingWisdomClasses2019} &
+    CS2 students &
+    Java &
+    \miscon{StructByRef} (Their G2); \miscon{DefByRef} or \miscon{DefCopyStructs} (Their G3); \miscon{CallsCopyStruct} (Their G4).
+    \\\hline
+    \cite{strombackProgressionStudentsAbility2023} &
+    CS masters &
+    Python &
+    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
+    (See their section 4.2)
+    \\\hline
+    \cite{strombackProgressionStudentsAbility2023} &
+    CS undergrads &
+    C++ &
+    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
+    (See their section 4.2)
+    \\\hline
+  \end{tabularx}
+  \caption{Similar misconceptions found in prior research.}
+  \label{t:rel-work}
+\end{table}
+
+Misconceptions related to scope, mutation, and higher-order functions have been
+widely identified in varying populations
+(from CS1 students to graduate students to users of online forums)
+over many years (since at last 1991)
+in varying programming languages (from Java to Racket) and in
+different countries (such as the USA and Sweden).
+\Cref{t:rel-work} lists works that seem most relevant to us.
+
+There are some small differences.
+\cite{fleuryParameterPassingRules1991} identifies a dynamic scope misconception
+that is different from \miscon{FlatEnv}:
+\begin{description}
+  \item[\surmisedMiscon{CallerEnv}] Function values don't remember their environments.
+  When a function is called, the function body is evaluated in an
+  environment that extends from the \textit{caller's} environment.
+\end{description}
+I don't include \surmisedMiscon{CallerEnv} in our analysis because in our
+data, all wrong answers that can be explained by \surmisedMiscon{CallerEnv}
+are also explainable by \miscon{FlatEnv}.
+
+Appendix A of \cite{sorvaVisualProgramSimulation2012} provides an
+extensive survey of misconceptions reported
+in research up to 2012.
+There are overlaps between our survey and
+theirs. For instance, our \surmisedMiscon{CallerEnv} is their No.~47.
+However, because their descriptions are brief, it is difficult to tell
+whether a misconception in their survey matches our misconceptions.
+Because
+they provide neither misinterpreters nor representative program-output
+pairs, it is difficult to determine the overlaps precisely (showing
+the value of providing these two machine-runnable descriptions).
+At any rate,
+we certainly find
+no equivalent of \miscon{FlatEnv}, \miscon{DeepClosure}, and
+\miscon{DefByRef} in their survey.
+
+\section{Background: Program Behavior}
+
+Our idea of ``goal sentences'' is not substantially different from the
+\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
+I only used these sentences (or rules) to guide the
+design of the Tutor and as text in the Tutor itself.
+\cite{duranRulesProgramBehavior2021} argue for other uses of such
+sentences.
+
+\section{Plan for \textbf{RQ1}}%
+\label{s:plan-rq1}
+
+\begin{description}
+  \item[RQ1] \rqI
+\end{description}
+
+To answer this RQ, I need to first define what I mean by ``apply to
+most modern programming languages''. As I described in the
+Introduction (\Cref{s:intro}), most modern languages share a semantic
+core. I have modelled this core as SMoL (\Cref{s:smol}). Because SMoL
+represents the semantic core, misconceptions about SMoL presumably
+also apply to most modern PLs.
+
+To find misconceptions that ``present (even) in students with
+prior programming backgrounds?'', I have been collecting data from
+multiple populations that very likely have prior experience. For
+example, my priliminary results are build upon data collected in a
+upper-year programming language course at Brown.
+
+To find a reasonably comprehensive collection of misconceptions, I
+seek for inputs from students. Most prior works
+(\Cref{s:background-misconceptions}) have experts coming up with a
+list of potential misconceptions and then filter the potential
+misconceptions with student inputs. This approach can't possibly find
+misconceptions not considered by experts and hence is necessarily
+affected by experts' blind spots. To overcome experts' blind spots, I
+analyze a dataset produced by a two-year process to find (a)
+student-written programs that tend to trip up students and (b)
+incorrect responses to those programs (\Cref{s:gen-prob-quizius}),
+using a tool for the purpose (\Cref{s:quizius}). I have been cleaning
+up those programs (\Cref{s:smol-quizzes,s:tutor-refine-program-set})
+and confirming that these programs are still tricky in the same way
+(\Cref{s:misconceptions-quiz,s:misconceptions-tutor}).
+
+I have been creating a tutoring system to answer my RQ2. The tutoring
+system collects more student data, which becomes another source of
+misconceptions. The tutoring system, which I explain in details when I
+talk about RQ2 (\Cref{s:plan-rq2}), includes many tasks that asks
+students to predict the result(s) of running a program. I plan to
+change the tutor such that it asks students' for their explanations
+when they give a wrong answer. I plan to eyeball the explanations to
+see if there are any new misconceptions.
+
+To be reasonably specific about misconceptions, I have created a
+definitional interpreter for each misconception that I have found. I
+plan to do the same things for potentially newly found/confirmed
+misconceptions.
+
+Eventually, I expect to build a table of misconceptions. Each
+misconception comes with two or more programs. For each of these
+programs, the wrong result that correspond to the misconception should
+be commonly predicted by students from various population.
+Furthermore, these students give similar explanations for all the
+wrong answers. For example, I might find a misconception called
+\miscon{DefByRef}: when students are asked to predict the result of
+\begin{lstlisting}
+  (defvar x 12)
+  (defvar y x)
+  (set! y 0)
+  x
+  y
+\end{lstlisting}
+X\% of them responded \lstinline|12 0|; and when asked to predict
+\begin{lstlisting}
+  (defvar x 12)
+  (defvar y x)
+  (set! x 0)
+  x
+  y
+\end{lstlisting}
+Y\% of them responded \lstinline|0 12|; and these students all
+explained something like ``\lstinline|y| is \lstinline|x|''.
+Preliminary results can be found in \Cref{s:misconceptions-tutor}.
+
+\section{Plan for \textbf{RQ2}}%
+\label{s:plan-rq2}
+
+\begin{description}
+  \item[RQ2] \rqII
+\end{description}
+
+Prior works suggest that presenting refutation text and encouraging
+students to compare cases are helpful for learning, and that notional
+machines are useful for explaning program behavior.
+
+I would first explain the major task format
+(\Cref{s:interpreting-task}) in the tutor and then explain how I would
+let the tutor incorporate ideas of refutation text
+(\Cref{s:refutationt-text}), case comparsion
+(\Cref{s:case-comparison}), and notional machines
+(\Cref{s:notional-machine}). Finally, I explain how I plan to evaluate
+the effectiveness of the tutor (\Cref{s:evaluation}).
+
+\subsection{Interpreting Task}%
+\label{s:interpreting-task}
+
+To build a tutor that incorporate these ideas, I plan to includes lots
+of multiple-choice questions (MCQs) in the tutor. I call these
+interpreting tasks because they ask students to interpret programs:
+each MCQ presents a program, for example,
+\begin{lstlisting}
+  (defvar x 12)
+  (defvar y x)
+  (set! y 0)
+  x
+  y
+\end{lstlisting}
+and then asks students to choose the results of the program.
+
+\paragraph{The MCQ Options} The options include the correct answers and
+wrong answers that correspond to known misconceptions (in this case,
+only \lstinline|0 0|, which correspond to \miscon{DefByRef}), among
+other options that I will explain shortly. Because students might give
+wrong answers that I do not anticipate, I plan to include an ``Other''
+option that, once choosen, allows students to enter arbitrary answer.
+If an MCQs only have a small number of options, students have a great
+chance (in this case, 50\%) of just guessing the right answer or
+successfully using a process of elimination. So I plan to add
+additional options that are not obviously wrong: options generated by
+moving and/or copying numbers in the current options. In this case, I
+will add \lstinline|12 12| and \lstinline|0 12|. In summary, each MCQ
+has the following options:
+\begin{itemize}
+  \item the correct answer,
+  \item wrong answers that correspond to known misconceptions,
+  \item options generated by moving and/or copying numbers in the
+  current options
+  \item the ``Other'' option (which, once choosen, allows students to
+  enter arbitrary answer)
+\end{itemize}
+
+\paragraph{Panelty to Wrong Choices} To discourage students from
+making a random choice, the tutor should gives panelty. Deciding the
+panelty is a non-trivial question. A common strategy is to grade
+students by their correctness. However, grading by correctness is not
+suitable for a tutoring system because we do not mind students doing
+badly on earlier tasks as long as they eventually master the concepts.
+Actually, if students generally do perfectly on all tasks, that is a
+failure of the tutoring system -- it is likely not teaching anything
+to the students. If grading by correctness is not a suitable option, a
+common alternative is to grade by completion. However, grading by
+completion alone encourage students to rush through the tutor and
+hence might encourage students to make random choices. In short, I
+would like to make students feel as follows:
+\begin{itemize}
+  \item They won't lose grade as long as they finish the tutor.
+  \item They should avoid giving wrong answers.
+\end{itemize}
+To encourage this feeling, I do the following: when I deployed the
+tutor in my institution, I graded by completion\footnote{However, I
+have no control over how other institutions use the tutor.}; when a
+student fails a task, the tutor gives an \emph{extra} task:
+\begin{enumerate}
+
+  \item The program is semantically the same as the first, but with
+    superficial changes (e.g., variable names, constants, and
+    operators are changed) so that the student cannot immediately
+    guess the answer.
+
+  \item Instead of multiple-choice, students must \emph{type} the
+    answer into a text box. (The tutor normalizes text to accommodate
+    variations.) This is intentional. First, I want to force
+    reflection on the explanation just given, whereas with an MCQ,
+    students could just guess. Second, I feel that students would
+    find typing more onerous than clicking. In case students had just
+    guessed on a question, my hope is that the penalty of having to
+    type means, on \emph{subsequent} tasks, they would be more likely
+    to pause and reflect before choosing.
+
+\end{enumerate}
+
+\subsection{Design the Tutor with Refutation Text and Other Explanations}%
+\label{s:refutationt-text}
+
+The tutor presents a refutation text when a student seems to hold a
+misconception. Let's reconsider the program in
+\Cref{s:interpreting-task}. If a student answers \lstinline|0 0|,
+which suggests they hold the \miscon{DefByRef} misconception, the
+tutor responds
+\begin{quote}
+  The correct answer is \lstinline|12 0|. \lstinline|y| was bound to
+  \lstinline|12| (i.e., the value of \lstinline|x|) rather than to
+  \lstinline|x|. So changing the value of \lstinline|y| does not
+  changes the value of \lstinline|x|.
+\end{quote}
+If a students gave a wrong answer that does not correspond to a
+misconception (either by choosing one of the generated option or by
+choosing the ``Other'' option), the tutor presents a generic
+explanation. For example, if a student answers \lstinline|12 12|, the
+tutor responds
+\begin{quote}
+  The correct answer is \lstinline|12 0|. The first definition binds
+  \lstinline|x| to \lstinline|12|. The second definition binds
+  \lstinline|y| to the value of \lstinline|x|, which is
+  \lstinline|12|. The \lstinline|set!| mutates the binding of
+  \lstinline|y|, so \lstinline|y| is now bound to \lstinline|0|.
+\end{quote}
+
+\subsection{Design the Tutor with Case Comparison}%
+\label{s:case-comparison}
+
+The tutor provides considerable opportunities for making case
+comparison: each interpreting task presents a SMoL program and the
+program's result. These program-result pairs are concrete cases of the
+semantics of SMoL. To incorporate the idea of case comparsion, the
+tutor prompts students to compare these cases and then to summarize
+the rules of program behavior.
+
+It is impractical to ask students to compare tens or even hundreads of
+programs. So
+
+\subsection{Design the Tutor with Case Comparison}%
+\label{s:notional-machine}
+
+\subsection{Evaluate the Tutor}%
+\label{s:evaluation}
+
+\section{Proposal Roadmap}
+\label{s:roadmap}
+\label{s:univ-1-course}
+
+This proposal describes a four-year effort to obtain a quality instrument
+to measure misconceptions and to build a tutor to address them.
+
+The first step was a two-year process to generate (a) programs that
+tend to trip up students and (b) incorrect responses to those programs
+(\Cref{s:gen-prob-quizius}), using a tool for the purpose
+(\Cref{s:quizius}). These were then curated (\Cref{s:smol-quizzes}) to
+produce an instrument that was tested for one more year
+(\Cref{s:misconceptions-quiz}). This underwent one more round of
+curation to produce the final instrument~(\Cref{s:tutor-refine-program-set}).
+
+I then present the tutor (\Cref{s:system-tutor}). In particular, we
+present the list of misconceptions---grounded in student data---that
+are the focus of this project (\Cref{s:misconceptions-tutor}). I also
+evaluate the tutor's effectiveness at correcting them
+(\Cref{s:effect-tutor}).
+
+All this work is done with students in a ``Principles of Programming
+Languages'' class at a selective, private US university. The class has
+about 70--75 students per year. It is not required, so students take
+it by choice. Virtually all are computer science majors. Most are in
+their third or fourth year of tertiary education; about 10\% are
+graduate students. All have had at least one semester of imperative
+programming, and most have significantly more experience with it. Most
+have had close to a semester of functional programming. The student
+work described here was required, but students were graded on effort,
+not correctness.
+
+Naturally, we should wonder to what extent the demographic affects the
+results: we may simply be studying weaknesses at this institution! We
+therefore work with two other populations
+(\Cref{s:generality-tutor}). In both, we find similar results.
 
@@
 \href{https://github.com/shriram/smol}{https://github.com/shriram/smol}.
-
-\section{Proposal Roadmap}
-\label{s:roadmap}
-\label{s:univ-1-course}
-
-This proposal describes a four-year effort to obtain a quality instrument
-to measure misconceptions and to build a tutor to address them.
-
-The first step was a two-year process to generate (a) programs that
-tend to trip up students and (b) incorrect responses to those programs
-(\Cref{s:gen-prob-quizius}), using a tool for the purpose
-(\Cref{s:quizius}). These were then curated (\Cref{s:smol-quizzes}) to
-produce an instrument that was tested for one more year
-(\Cref{s:misconceptions-quiz}). This underwent one more round of
-curation to produce the final instrument~(\Cref{s:tutor-refine-program-set}).
-
-I then present the tutor (\Cref{s:system-tutor}). In particular, we
-present the list of misconceptions---grounded in student data---that
-are the focus of this project (\Cref{s:misconceptions-tutor}). I also
-evaluate the tutor's effectiveness at correcting them
-(\Cref{s:effect-tutor}).
-
-All this work is done with students in a ``Principles of Programming
-Languages'' class at a selective, private US university. The class has
-about 70--75 students per year. It is not required, so students take
-it by choice. Virtually all are computer science majors. Most are in
-their third or fourth year of tertiary education; about 10\% are
-graduate students. All have had at least one semester of imperative
-programming, and most have significantly more experience with it. Most
-have had close to a semester of functional programming. The student
-work described here was required, but students were graded on effort,
-not correctness.
-
-Naturally, we should wonder to what extent the demographic affects the
-results: we may simply be studying weaknesses at this institution! We
-therefore work with two other populations
-(\Cref{s:generality-tutor}). In both, we find similar results.
 

### 2023-11-02 proposal.tex
tons of updates

@@
 \usepackage{hyperref}
-\usepackage[colorinlistoftodos]{todonotes}
 \newcommand{\miscon}[1]{\textbf{#1}}
@@
 \newcommand{\rqI}{What are the program-behavior misconceptions that apply to most modern programming languages and present (even) in students with prior programming backgrounds?}
-\newcommand{\rqII}{What would be a tutoring system that aims at correcting the misconceptions, draws on cognitive and educational psychology concepts, utilizes a notional machine, and seems to be effective?}
+\newcommand{\rqII}{What would be a tutoring system that aims at correcting the misconceptions, draws on cognitive, educational, or psychological concepts, and seems to be effective?}
 
@@
   breaklines=true,
-  language=smol,
-  % aboveskip=-0.3 \baselineskip,   % remove extra space before
-  % belowskip=-0.8 \baselineskip,     % and after code blocks
+  language=smol
 }
@@
 outcomes.
+
+[FILL Notional Machine]
+
+[FILL Language Levels] Prior research argues for the benefits of \emph{language levels}
+[CITE]: a langauge has better be taught as a sequence of languages,
+where each later language includes more langauge constructs than the
+previous one.
 
@@
 
-\section{Background: Program Behavior}
-
-Our idea of ``goal sentences'' is not substantially different from the
-\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
-I only used these sentences (or rules) to guide the
-design of the Tutor and as text in the Tutor itself.
-\cite{duranRulesProgramBehavior2021} argue for other uses of such
-sentences.
-
 \section{Plan for \textbf{RQ1}}%
@@
 
-To find misconceptions that ``present (even) in students with
-prior programming backgrounds?'', I have been collecting data from
-multiple populations that very likely have prior experience. For
-example, my priliminary results are build upon data collected in a
-upper-year programming language course at Brown.
+To find misconceptions that ``present (even) in students with prior
+programming backgrounds'', I have been collecting data from multiple
+populations that very likely have prior experience
+(\Cref{s:generality-tutor}). For example, my priliminary results are
+build upon data collected in a upper-year programming language course
+at Brown. [FILL list the data sources]
 
@@
 (\Cref{s:misconceptions-quiz,s:misconceptions-tutor}).
+\Cref{s:misconceptions-tutor} list misconceptions found with this
+process.
 
@@
 talk about RQ2 (\Cref{s:plan-rq2}), includes many tasks that asks
-students to predict the result(s) of running a program. I plan to
-change the tutor such that it asks students' for their explanations
-when they give a wrong answer. I plan to eyeball the explanations to
-see if there are any new misconceptions.
+students to predict the result(s) of running a program. I have found
+several unanticipated patterns of errors
+(\Cref{s:misconceptions-tutor}), which suggest the existence of
+unknown misconceptions. The current system do not collect students'
+rationale for their answers. So I do not know what are the underlying
+misconceptions (although some of the wrong answers are somewhat
+self-explanatory). I plan to change the system such that it asks
+students' for their explanations when they give a wrong answer. I plan
+to eyeball the explanations to find new misconceptions.
 
@@
 definitional interpreter for each misconception that I have found. I
-plan to do the same things for potentially newly found/confirmed
-misconceptions.
+plan to do the same things for newly found/confirmed misconceptions.
 
@@
 be commonly predicted by students from various population.
-Furthermore, these students give similar explanations for all the
-wrong answers. For example, I might find a misconception called
+Furthermore, these students should give similar explanations for all
+the wrong answers. For example, I might find a misconception called
 \miscon{DefByRef}: when students are asked to predict the result of
@@
 explained something like ``\lstinline|y| is \lstinline|x|''.
-Preliminary results can be found in \Cref{s:misconceptions-tutor}.
 
@@
 
-Prior works suggest that presenting refutation text and encouraging
-students to compare cases are helpful for learning, and that notional
-machines are useful for explaning program behavior.
-
-I would first explain the major task format
-(\Cref{s:interpreting-task}) in the tutor and then explain how I would
-let the tutor incorporate ideas of refutation text
-(\Cref{s:refutationt-text}), case comparsion
-(\Cref{s:case-comparison}), and notional machines
-(\Cref{s:notional-machine}). Finally, I explain how I plan to evaluate
-the effectiveness of the tutor (\Cref{s:evaluation}).
+To answer this RQ, I have done some literature review on what
+pedagogic techniques are helpful (\Cref{s:background}). My current
+takeaway is that refutation text, case comparison, and notional
+machines are helpful. And I have created a tutoring system (hereafter,
+the tutor) (\Cref{s:system-tutor}) that incorporates refutation text
+and case comparison (\Cref{s:tutor-uiux}). I am adding notional
+machine to the tutor and plan to explore more literature and to
+improve the tutor if I find new helpful techniques.
+
+I would like to first explain key aspects of the tutor
+(\Cref{s:tutorials,s:interpreting-task}) and then how the tutor
+incorporates ideas of refutation text (\Cref{s:refutationt-text}),
+case comparsion (\Cref{s:case-comparison}), and notional machines
+(\Cref{s:notional-machine}). Finally, I present my plan to evaluate
+the effectiveness of the tutor and how evaluation impacts the tutor
+design (\Cref{s:evaluation}).
+
+\subsection{Tutorials}%
+\label{s:tutorials}
+
+There are many misconceptions to correct. To make the tutoring process
+managable, I have design the tutor as a sequence of tutorials.
+
+Inspired by prior research on language levels (\Cref{s:background}),
+my sequence of tutorials is ordered, and each later tutorials cover
+one or more language constructs (\Cref{t:tutorials}).
+
+\begin{table}[t]
+  \centering\begin{tabularx}{\linewidth}{p{0.4\linewidth} X}
+    \textbf{Tutorial}                             & \textbf{Language Constructs}
+    \\\hline
+    \newterm{scope}: Lexical scope &
+    Primitives, variables, \lstinline|defvar|, and \lstinline|deffun|.
+    \\\hline
+    \newterm{mut-vars}: Mutable variables &
+    Adding \lstinline|set!|.
+    \\\hline
+    \newterm{vectors}: Vectors and vector updates &
+    Adding vector operators.
+    \\\hline
+    \newterm{lambda}: Lambda expressions &
+    Adding \lstinline|lambda|.
+    \\\hline
+    \newterm{local}: Local binding forms &
+    Adding \texttt{let}.
+    % \\\hline
+  \end{tabularx}
+  \caption{Tutorials and Their Language Constructs}
+  \label{t:tutorials}
+\end{table}
 
@@
 
-To build a tutor that incorporate these ideas, I plan to includes lots
-of multiple-choice questions (MCQs) in the tutor. I call these
-interpreting tasks because they ask students to interpret programs:
-each MCQ presents a program, for example,
+The tutor is mostly consist of \emph{interpreting tasks}:
+multiple-choice questions (MCQs) that asks students to interprete
+program.
+
+\paragraph{The MCQ Options} The options include (among other options
+that I will explain shortly) the correct answer and wrong answers that
+correspond to known misconceptions. Because students might conceive
+wrong answers that I do not anticipate, the tutor has been providing
+an ``Other'' option that, once choosen, allows students to enter
+arbitrary answer.
+
+Let's reconsider this program from \Cref{s:plan-rq1}
 \begin{lstlisting}
@@
 \end{lstlisting}
-and then asks students to choose the results of the program.
-
-\paragraph{The MCQ Options} The options include the correct answers and
-wrong answers that correspond to known misconceptions (in this case,
-only \lstinline|0 0|, which correspond to \miscon{DefByRef}), among
-other options that I will explain shortly. Because students might give
-wrong answers that I do not anticipate, I plan to include an ``Other''
-option that, once choosen, allows students to enter arbitrary answer.
-If an MCQs only have a small number of options, students have a great
-chance (in this case, 50\%) of just guessing the right answer or
-successfully using a process of elimination. So I plan to add
-additional options that are not obviously wrong: options generated by
-moving and/or copying numbers in the current options. In this case, I
-will add \lstinline|12 12| and \lstinline|0 12|. In summary, each MCQ
+In this case, the options should include the correct answer
+\lstinline|12 0| and a wrong answer \lstinline|0 0|, which correspond
+to the \miscon{DefByRef} misconception. If an MCQ only has a small
+number of options like this example, students have a great chance (in
+this case, 50\%) of just guessing the correct answer or successfully
+using a process of elimination. So I have been adding additional
+options that are not obviously wrong. The tutor is adding options with
+an ad hoc rules. Eventually, I plan to generate options by moving
+and/or copying numbers in the current options. In this case, my plan
+would add \lstinline|12 12| and \lstinline|0 12|. In summary, each MCQ
 has the following options:
@@
 \paragraph{Panelty to Wrong Choices} To discourage students from
-making a random choice, the tutor should gives panelty. Deciding the
-panelty is a non-trivial question. A common strategy is to grade
-students by their correctness. However, grading by correctness is not
-suitable for a tutoring system because we do not mind students doing
-badly on earlier tasks as long as they eventually master the concepts.
-Actually, if students generally do perfectly on all tasks, that is a
-failure of the tutoring system -- it is likely not teaching anything
-to the students. If grading by correctness is not a suitable option, a
-common alternative is to grade by completion. However, grading by
-completion alone encourage students to rush through the tutor and
-hence might encourage students to make random choices. In short, I
-would like to make students feel as follows:
+making a random choice, the tutor has been giving panelty. Designing
+the strategy of assigning panelty is non-trivial. A common strategy is
+to grade students by their correctness. However, grading by
+correctness is not suitable for a tutoring system because we do not
+mind students doing badly on earlier tasks as long as they eventually
+master the concepts. Actually, if students generally succeed in all
+tasks, the tutoring system is failing -- it does not teach anything to
+the students. A alternative common strategy is to grade by completion.
+However, grading by completion alone encourage students to rush
+through the tutorials and hence might encourage students to make
+random choices. In short, I want to make students feel as follows:
 \begin{itemize}
-  \item They won't lose grade as long as they finish the tutor.
+  \item They won't lose grade as long as they finish the tutorials.
   \item They should avoid giving wrong answers.
 \end{itemize}
-To encourage this feeling, I do the following: when I deployed the
-tutor in my institution, I graded by completion\footnote{However, I
-have no control over how other institutions use the tutor.}; when a
-student fails a task, the tutor gives an \emph{extra} task:
+To encourage this feeling, I have been doing the following: when I
+deployed the tutor in my institution, I graded by
+completion\footnote{I have no control over how other institutions use
+the tutor.}; when a student fails a task, the tutor gives an
+\emph{extra} task:
 \begin{enumerate}
@@
 
-\subsection{Design the Tutor with Refutation Text and Other Explanations}%
+\subsection{Refutation Text and Other Explanations}%
 \label{s:refutationt-text}
 
-The tutor presents a refutation text when a student seems to hold a
-misconception. Let's reconsider the program in
-\Cref{s:interpreting-task}. If a student answers \lstinline|0 0|,
-which suggests they hold the \miscon{DefByRef} misconception, the
-tutor responds
+The tutor has been presenting a refutation text when a student chooses
+a wrong answer that correspond to known misconception(s). Let's
+reconsider the program in \Cref{s:interpreting-task}. If a student
+answers \lstinline|0 0|, which suggests they hold the
+\miscon{DefByRef} misconception, the tutor responds
 \begin{quote}
-  The correct answer is \lstinline|12 0|. \lstinline|y| was bound to
-  \lstinline|12| (i.e., the value of \lstinline|x|) rather than to
-  \lstinline|x|. So changing the value of \lstinline|y| does not
-  changes the value of \lstinline|x|.
+  \lstinline|y| was bound to \lstinline|12| (i.e., the value of
+  \lstinline|x|) rather than to \lstinline|x|. So changing the value
+  of \lstinline|y| does not changes the value of \lstinline|x|.
 \end{quote}
@@
 \begin{quote}
-  The correct answer is \lstinline|12 0|. The first definition binds
-  \lstinline|x| to \lstinline|12|. The second definition binds
-  \lstinline|y| to the value of \lstinline|x|, which is
-  \lstinline|12|. The \lstinline|set!| mutates the binding of
+  The first definition binds \lstinline|x| to \lstinline|12|. The
+  second definition binds \lstinline|y| to the value of \lstinline|x|,
+  which is \lstinline|12|. The \lstinline|set!| mutates the binding of
   \lstinline|y|, so \lstinline|y| is now bound to \lstinline|0|.
 \end{quote}
-
-\subsection{Design the Tutor with Case Comparison}%
+If a student gave the correct answer, the tutor also presents the
+generic explanation.
+
+\subsection{Case Comparison}%
 \label{s:case-comparison}
@@
 semantics of SMoL. To incorporate the idea of case comparsion, the
-tutor prompts students to compare these cases and then to summarize
-the rules of program behavior.
-
-It is impractical to ask students to compare tens or even hundreads of
-programs. So
-
-\subsection{Design the Tutor with Case Comparison}%
+tutor currently prompts students to compare these cases and then to
+summarize the rules of program behavior at the end of \emph{some}
+tutorials. I plan to expand this to \emph{all} tutorials and use a
+consistent prompt. The prompt should asks about the behavior of new
+language constructs and their interaction with all previously
+introduced constructs. For example, at the end of \textbf{mut-vars}, I
+would want the tutor to ask
+\begin{quote}
+  What does \lstinline|set!| do? How does it interact with variables,
+  \lstinline|defvar|, and \lstinline|deffun|?
+\end{quote}
+
+\subsection{Multi-syntax}%
+\label{s:multi-syntax}
+
+[FILL]
+
+\subsection{Notional Machine}%
 \label{s:notional-machine}
+
+Prior research argues that illustrating how programs runs with
+notional machines (NMs) is benefitial for students [CITE]. I have created a
+web-based notional machine
+\href{https://lukuangchen.github.io/stacker-2023}{https://lukuangchen.github.io/stacker-2023}.
+
+The notional machine can be configurated with URL parameters. For example,
+\href{https://lukuangchen.github.io/stacker-2023/?program=\%28defvar+x+12\%29\%0A\%28defvar+y+x\%29\%0A\%28set\%21+y+0\%29\%0Ax\%0Ay}{this link}
+opens the NM with a program embeded.
+
+My plan is to change the tutor such that it provides a
+program-specific link after each interpreting task and instructs
+students to click the link for more explanation.
 
@@
 \label{s:evaluation}
+
+To fully answer my research questions, I should also try to prove that
+my tutor is effective. In this secton, I describe how I want to refine
+the design of the tutor to help me make the prove, what data I want to
+collect, and what (statistical) analysis do I want to perform on the
+data.
+
+I would like to evaluate the effectiveness from two perspectives:
+\begin{itemize}
+  \item How effective is the tutor at stopping students from holding
+  misconceptions?
+  \item How close are students' final conception to the true
+  conception?
+\end{itemize}
+
+I can infer whether students hold a misconception by analysing their
+responses to interpreting tasks. There are three possible responses to
+an interpreting task:
+\begin{enumerate}
+  \item They give the correct answer.
+  \item They give a wrong answer that correspond to some
+  misconceptions.
+  \item They give a wrong answer that do not correspond to any (known)
+  misconceptions.
+\end{enumerate}
+If a student gives a wrong answer corresponds to exactly one
+misconception \emph{and} an explanation that matches the
+misconception, I deduce that the student holds the misconception when
+doing the task.
+
+Given this observation, I plan to set up the tutor such that
+\begin{itemize}
+  \item Every wrong answer corresponds to up to one misconception. In
+  this case, I say the wrong answer \emph{represents} the
+  misconception.
+  \item Every misconception is represented by some questions.
+  \item In each tutorial, a misconception is either not represented by
+  any interpreting tasks or represented by exactly two tasks.
+  \item Within each tutorial, the order of interpreting tasks is
+  randomized.
+\end{itemize}
+
+For each tutorail and each misconception appear in the tutorial, I
+compare student performance on the \emph{first} question that involve
+the misconceptions with performance on the \emph{second} question. If
+the difference is clear (i.e., ``statistically significant''), I
+deduce that the tutor is effective at correcting the misconception.
 
@@
 
+\subsection{Program Behavior}
+
+Our idea of ``goal sentences'' is not substantially different from the
+\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
+I only used these sentences (or rules) to guide the
+design of the Tutor and as text in the Tutor itself.
+\cite{duranRulesProgramBehavior2021} argue for other uses of such
+sentences.
+
 \subsection{Misinterpreters}

### 2023-11-03 proposal.tex
format file
save new tutorials

@@
 \documentclass{article}
+
+% TODO
+% - clean up the background sections
 
@@
 \renewcommand\thelinenumber{\textcolor{red}{\arabic{linenumber}}}
-
 \usepackage[utf8]{inputenc} % Use UTF-8 encoding
@@
 \usepackage{hyperref}
-\newcommand{\miscon}[1]{\textbf{#1}}
-\newcommand{\surmisedMiscon}[1]{\miscon{#1}$^\ddag$}
-\newcommand{\pValue}[0]{\textit{p}\nobreakdash-value}
-\newcommand{\rqI}{What are the program-behavior misconceptions that apply to most modern programming languages and present (even) in students with prior programming backgrounds?}
-\newcommand{\rqII}{What would be a tutoring system that aims at correcting the misconceptions, draws on cognitive, educational, or psychological concepts, and seems to be effective?}
-
-%% Some recommended packages.
 \usepackage{booktabs}   %% For formal tables:
@@
 }
-%% Custom commands:
-%% new terms are introduced with a boldface
-%% (the common style is italic, but KC thinks bf is easier to locate)
+
+\newcommand{\miscon}[1]{\textbf{#1}}
+\newcommand{\surmisedMiscon}[1]{\miscon{#1}$^\ddag$}
 \newcommand{\newterm}[1]{\textbf{#1}}
+
+%% abbrev. for p-value
+\newcommand{\pValue}[0]{\textit{p}\nobreakdash-value}
+
+%% abbrev. for RQs
+\newcommand{\rqI}{What are the program-behavior misconceptions that apply to SMoL and present (even) in students with prior programming backgrounds?}
+\newcommand{\rqII}{What would be a tutoring system that aims at correcting the misconceptions, draws on cognitive, educational, or psychological concepts, and seems to be effective?}
 
@@
 
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
 \begin{abstract}
@@
   variables, mutable compound data, and their interaction with scope
-  and higher-order functions seem to be widespread.
-  But
-  how do we detect
-  them, given that experts have blind spots and may not realize the
-  myriad ways in which students can misunderstand programs?
+  and higher-order functions seem to be widespread. But how do we
+  detect them, given that experts have blind spots and may not realize
+  the myriad ways in which students can misunderstand programs?
   Furthermore, once identified, what can we do to correct them?
 
-  In this proposal, I present a curated list of misconceptions, and an
-  instrument to detect them. These are distilled from student work
-  over several years and match and extend prior research.  I also
-  present an automated, self-guided tutoring system. The tutor builds
-  on strategies in the education literature and is explicitly designed
-  around identifying and correcting misconceptions.
-
-  I have tested the tutor in multiple settings.
-  Our data consistently show that (a) the misconceptions I tackle are
-  widespread, and (b) the tutor appears to improve understanding.
-
-  [FILL future work]
+  I propose a plan for finding misconceptions possibly not anticipated
+  by experts, and the design of an automated, self-guided tutoring
+  system. The tutor builds on strategies in the cognitive and
+  educational literature and is explicitly designed around identifying
+  and correcting misconceptions. Preliminary results suggest (a) the
+  misconceptions I found are widespread, and (b) the tutor appears to
+  improve understanding.
 \end{abstract}
 
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
 \maketitle
@@
 
-As the related work section (\Cref{s:rel-work}) describes,
+As a background section (\Cref{s:background-misconceptions}) describes,
 multiple researchers, in different
@@
 
-To answer my research questions, I have (a) run a multi-year,
-multi-phase study to identify these misunderstandings while trying to
-work around the blind spots of experts (and also drawing on prior
-work), and (b) constructed the SMoL Tutor, an interactive tutor to
-address these misunderstandings. The SMoL Tutor is not a
-\emph{programming} tutor; rather, it assumes the user has some basic
-facility with programming (i.e., that they are already familiar with
-concepts like vectors, mutation, and higher-order functions). It is
-built around SMoL's features and their interactions, and draws on
-cognitive and educational psychology concepts to expressly identify
-and correct misconceptions.
-
-Concretely, I have made the following contributions:
-\begin{enumerate}
-
-  \item I provide a collection of brief programs for detecting
-        misconceptions about the features of SMoL and their
-        interaction. These programs are short and most are readily
-        translatable to a wide variety of programming languages that
-        share the SMoL characteristics.
-
-  \item I show that multiple populations have problems with these
-        programs.
-
-  \item I present a list of curated misconceptions generated
-        after multiple rounds of analysis.
-
-  \item I implement these misconceptions as interpreters, which
-        find weaknesses in our manual analysis and hold potential for
-        generative use in the future.
-
-  \item I present a tutoring system, the SMoL Tutor, that measurably
-        corrects these misconceptions.
-
-  \item I present a notional machine, the Stacker, that is used by the
-  SMoL Tutor to explain programs and might be useful in itsown.
-
-\end{enumerate}
-
-This document first presents the current results and ends with future
-work \& timeline (\cref{s:future-work}).
-
-\paragraph{A Note on Terminology} I use the term
-``behavior'' to refer to the meaning of programs in terms of the
-answers they produce. A more standard term for this would, of course, be
-``semantics''. However, the term ``semantics tutor'' might
-mislead some readers into thinking it
-teaches people to read or write a formal semantics, e.g., an
-introduction to ``Greek'' notation. Because that is not the kind of
-tutor I are describing, to avoid confusion, I use the term
-``behavior'' instead.
-
-\section{Background: Tutoring Systems and Pedagogic Techniques}
-\label{s:background}
-
-There is an extensive body of literature on tutoring systems (\cite{vanlehnBehaviorTutoringSystems2006a} is
-a quality survey), and indeed whole conferences are dedicated to
-them. I draw on this literature. In particular, it is common in the
-literature to talk about a ``two-loop'' architecture \cite{vanlehnBehaviorTutoringSystems2006a} where the outer
-loop iterates through ``tasks'' (i.e., educational activities) and the inner loop iterates through UI events within a task. I follow the same
-structure in our tutor (\Cref{s:tutor-uiux}).
-
-Many tutoring systems focus on teaching programming
-(such as the well-known and heavily studied LISP Tutor \cite{andersonLISPTutor1985}), and in
-the process undoubtedly address some program behavior
-misconceptions. Our SMoL Tutor differs in a notable way: it does not
-try to teach programming per se. Instead, it assumes a basic
-programming background and focuses entirely on program behavior misconceptions
-and correcting them. I are not aware of a tutoring system (in computer science) that has
-this specific design.
-
-The SMoL Tutor is firmly grounded in one technique from
-cognitive and educational psychology. The fundamental problem is: how
-do you fix a misconception? One approach is to only ``present the
-right answer'', for fear that discussing the wrong conception might
-actually reinforce it. Instead, there is a body of literature starting
-from \cite{posnerTheoryConceptualChange1982} that presents a theory of conceptual
-change, at whose heart is the \emph{refutation text}. A refutation
-text tackles the misconception directly, discussing and providing a
-refutation for the incorrect idea. Several studies \cite{schroederRefutationTextFacilitates2022} have shown
-their effectiveness in multiple other domains.
-
-The SMoL Tutor's content structure is also influenced by work on \emph{case comparisons}
-(which draws analogies between examples).
-\cite{alfieriLearningCaseComparisons2013a} suggests that asking
-(rather than not asking) students to find similarities between cases,
-and providing principles \emph{after} the comparisons (rather than
-before or not at all), are associated with better learning
-outcomes.
-
-[FILL Notional Machine]
-
-[FILL Language Levels] Prior research argues for the benefits of \emph{language levels}
-[CITE]: a langauge has better be taught as a sequence of languages,
-where each later language includes more langauge constructs than the
-previous one.
+In the rest of this document, I first define SMoL (\Cref{s:smol}),
+then presents literature review of misconceptions
+(\Cref{s:background-misconceptions}) and tutoring systems
+(\Cref{s:background-tutoring-systems}), then outline my research plans
+for the RQs (\Cref{s:plan-rq1,s:plan-rq2}). The remaining sections
+detail my current progress.
+
+% To answer my research questions, I have (a) run a multi-year,
+% multi-phase study to identify these misunderstandings while trying to
+% work around the blind spots of experts (and also drawing on prior
+% work), and (b) constructed the SMoL Tutor, an interactive tutor to
+% address these misunderstandings. The SMoL Tutor is not a
+% \emph{programming} tutor; rather, it assumes the user has some basic
+% facility with programming (i.e., that they are already familiar with
+% concepts like vectors, mutation, and higher-order functions). It is
+% built around SMoL's features and their interactions, and draws on
+% cognitive and educational psychology concepts to expressly identify
+% and correct misconceptions.
+
+% Concretely, I have made the following contributions:
+% \begin{enumerate}
+
+%   \item I provide a collection of brief programs for detecting
+%         misconceptions about the features of SMoL and their
+%         interaction. These programs are short and most are readily
+%         translatable to a wide variety of programming languages that
+%         share the SMoL characteristics.
+
+%   \item I show that multiple populations have problems with these
+%         programs.
+
+%   \item I present a list of curated misconceptions generated
+%         after multiple rounds of analysis.
+
+%   \item I implement these misconceptions as interpreters, which
+%         find weaknesses in our manual analysis and hold potential for
+%         generative use in the future.
+
+%   \item I present a tutoring system, the SMoL Tutor, that measurably
+%         corrects these misconceptions.
+
+%   \item I present a notional machine, the Stacker, that is used by the
+%         SMoL Tutor to explain programs and might be useful in itsown.
+
+% \end{enumerate}
+
+% This document first presents the current results and ends with future
+% work \& timeline (\cref{s:future-work}).
+
+% \paragraph{A Note on Terminology} I use the term
+% ``behavior'' to refer to the meaning of programs in terms of the
+% answers they produce. A more standard term for this would, of course, be
+% ``semantics''. However, the term ``semantics tutor'' might
+% mislead some readers into thinking it
+% teaches people to read or write a formal semantics, e.g., an
+% introduction to ``Greek'' notation. Because that is not the kind of
+% tutor I are describing, to avoid confusion, I use the term
+% ``behavior'' instead.
+
+\section{The SMoL Language}\label{s:smol}
+
+\begin{figure}[t]
+  \centering
+  \begin{lstlisting}
+    t ::= d
+        | e
+    d ::= (defvar x e)
+        | (deffun (f x ...) body)
+    e ::= c
+        | x
+        | (lambda (x ...) body)
+        | (let ([x e] ...) body)
+        | (begin e ... e)
+        | (set! x e)
+        | (if e e e)
+        | (cond [e body] ... [else body])
+        | (cond [e body] ...)
+        | (e e ...)
+    body    ::= t ... e
+    program ::= t ...
+  \end{lstlisting}
+  \caption{The syntax of SMoL.}
+  \label{f:smol-syntax}
+\end{figure}
+
+\begin{table}[t]
+  \centering\begin{tabularx}{\linewidth}{r X}
+    \textbf{Operators}            & \textbf{Meaning}                                       \\\hline
+    \texttt{+ - * /}              & Arithmetic Operators                                   \\
+    \texttt{< > <= >=}            & Number comparison                                      \\
+    \texttt{mvec}                 & Create a (mutable) vector (a.k.a. array)               \\
+    \texttt{vec-ref}              & Look up a vector element                               \\
+    \texttt{vec-set!}             & Replace a vector element                               \\
+    \texttt{vec-len}              & Get the length of a vector                             \\
+    \texttt{mpair}                & Create a 2-element (mutable) vector                    \\
+    \texttt{left right}           & Look up the first/second element of a 2-element vector \\
+    \texttt{set-left! set-right!} & Replace the first/second element of a 2-element vector \\
+    \texttt{eq?}                  & Equality                                               \\
+  \end{tabularx}
+  \caption{Primitive operators of SMoL.}
+  \label{t:smol-primitives}
+\end{table}
+
+SMoL is designed to capture common features of many modern
+languages.
+The syntax of SMoL is presented in~\Cref{f:smol-syntax}, where
+\lstinline|t| stands for terms, \lstinline|d| stands for definitions, \lstinline|e|
+stands for expressions, \lstinline|c| stands for constants (i.e., number,
+boolean, and string), and \lstinline|x| and \lstinline|f| are identifiers
+(variables). The last kind of expression (i.e., \lstinline|(e e ...)|) is
+function application.
+
+SMoL's Lispy syntax is
+an artifact of the course (\Cref{s:univ-1-course}) using Racket
+\cite{friedmanEssentialsProgrammingLanguages2001, krishnamurthiProgrammingLanguagesApplication2007};
+however, it also proved to be pedagogically valuable.
+I have found the parentheses useful when discussing
+scope in conjunction with local-binding features like \lstinline|let|.
+This avoids the various confusing ``variable lifting'' semantics found
+in languages like Python (\cite{politzPythonFullMonty2013}; see also
+posts like \cite{froadieWhatScopeVariable2022}), where the actual defined range of a
+variable is not apparent from the source code.
+
+Nevertheless, most SMoL programs are easy to translate to other
+languages. I present machine-translated Python and JavaScript
+versions of our programs in \supplRef{Translated\_Programs.html}. Furthermore, we intend to
+make a multi-lingual tutor in the future (\Cref{s:future-work}).
+
+The semantics of SMoL is as described in \Cref{s:intro}.
+SMoL includes limited primitive operators (\Cref{t:smol-primitives}) to work with numbers, strings, and
+vectors. It provides only one equality operator, which tests for exact
+equality between atomic values and for pointer equality between other
+values.
+
+Students are given a working implementation of SMoL, built
+as a \verb|#lang| language inside Racket \cite{felleisenProgrammableProgrammingLanguage2018}. This language
+provides only the defined syntax and semantics of SMoL, with no
+(other) Racket features present.
+(As in Racket, arguments evaluate left-to-right, to give stateful programs an
+unambiguous semantics.) The implementation is available online:
+\href{https://github.com/shriram/smol}{https://github.com/shriram/smol}.
 
@@
 
+\section{Background: Tutoring Systems}%
+\label{s:background-tutoring-systems}
+
+There is an extensive body of literature on tutoring systems
+(\cite{vanlehnBehaviorTutoringSystems2006a} is a quality survey), and
+indeed whole conferences are dedicated to them. I draw on this
+literature. In particular, it is common in the literature to talk
+about a ``two-loop'' architecture
+\cite{vanlehnBehaviorTutoringSystems2006a} where the outer loop
+iterates through ``tasks'' (i.e., educational activities) and the
+inner loop iterates through UI events within a task. I follow the same
+structure in my tutor (\Cref{s:tutor-uiux}).
+
+Many tutoring systems focus on teaching programming (such as the
+well-known and heavily studied LISP Tutor
+\cite{andersonLISPTutor1985}), and in the process undoubtedly address
+some program behavior misconceptions. My tutor differs in a notable
+way: it does not try to teach programming per se. Instead, it assumes
+a basic programming background and focuses entirely on program
+behavior misconceptions and correcting them. I are not aware of a
+tutoring system (in computer science) that has this specific design.
+
 \section{Plan for \textbf{RQ1}}%
@@
 
-To answer this RQ, I need to first define what I mean by ``apply to
-most modern programming languages''. As I described in the
-Introduction (\Cref{s:intro}), most modern languages share a semantic
-core. I have modelled this core as SMoL (\Cref{s:smol}). Because SMoL
-represents the semantic core, misconceptions about SMoL presumably
-also apply to most modern PLs.
-
 To find misconceptions that ``present (even) in students with prior
@@
 populations that very likely have prior experience
-(\Cref{s:generality-tutor}). For example, my priliminary results are
-build upon data collected in a upper-year programming language course
-at Brown. [FILL list the data sources]
+(\Cref{s:generality-tutor}). I have the following data sources so far:
+\begin{description}
+  \item[University 1] students in a ``Principles of Programming
+    Languages'' class at Brown University. The class has about 70--75
+    students per year. It is not required, so students take it by
+    choice. Virtually all are computer science majors. Most are in
+    their third or fourth year of tertiary education; about 10\% are
+    graduate students. All have had at least one semester of
+    imperative programming, and most have significantly more
+    experience with it. Most have had close to a semester of
+    functional programming. The student work described here was
+    required, but students were graded on effort, not correctness.
+  \item[University 2] a primarily public university in the US. It is
+    one of the largest Hispanic-serving institutions in the country. As
+    such, its demographic is extremely different from those whose data
+    were used above. The Tutor was used in one course in Spring 2023,
+    taken by 12 students. The course is a third-year, programming
+    language course. The students are required to have taken two
+    introductory programming courses (C++ focused).
+  \item[Textbook] an instrument was published on the website of a
+    programming languages textbook. Over the course of 8 months,
+    several hundreds of people submitted to the instrument. To protect
+    privacy, I intentionally do not record demographic information,
+    but I conjecture that the population is largely self-learners (who
+    are known to use the accompanying book), including some
+    professional programmers.
+\end{description}
 
@@
 list of potential misconceptions and then filter the potential
-misconceptions with student inputs. This approach can't possibly find
-misconceptions not considered by experts and hence is necessarily
-affected by experts' blind spots. To overcome experts' blind spots, I
-analyze a dataset produced by a two-year process to find (a)
-student-written programs that tend to trip up students and (b)
-incorrect responses to those programs (\Cref{s:gen-prob-quizius}),
-using a tool for the purpose (\Cref{s:quizius}). I have been cleaning
-up those programs (\Cref{s:smol-quizzes,s:tutor-refine-program-set})
-and confirming that these programs are still tricky in the same way
+misconceptions with student inputs. This approach can't find
+misconceptions not envisioned by experts and hence is hindered by
+experts' blind spots. To overcome such bias, I analyze a University~1
+dataset produced by a two-year process to find (a) student-written
+programs that tend to trip up students and (b) incorrect responses to
+those programs (\Cref{s:gen-prob-quizius}), using a tool for the
+purpose (\Cref{s:quizius}). I have been cleaning up those programs
+(\Cref{s:smol-quizzes,s:tutor-refine-program-set}) and confirming that
+these programs are still tricky in the same way
 (\Cref{s:misconceptions-quiz,s:misconceptions-tutor}).
-\Cref{s:misconceptions-tutor} list misconceptions found with this
-process.
-
-I have been creating a tutoring system to answer my RQ2. The tutoring
-system collects more student data, which becomes another source of
-misconceptions. The tutoring system, which I explain in details when I
-talk about RQ2 (\Cref{s:plan-rq2}), includes many tasks that asks
-students to predict the result(s) of running a program. I have found
-several unanticipated patterns of errors
-(\Cref{s:misconceptions-tutor}), which suggest the existence of
-unknown misconceptions. The current system do not collect students'
-rationale for their answers. So I do not know what are the underlying
-misconceptions (although some of the wrong answers are somewhat
-self-explanatory). I plan to change the system such that it asks
-students' for their explanations when they give a wrong answer. I plan
-to eyeball the explanations to find new misconceptions.
+\Cref{s:misconceptions-tutor} presents the final list of
+misconceptions found with this process.
+
+I have another source of misconceptions. I have been creating a tutor
+to answer my RQ2. The tutor, which I explain in details when I talk
+about RQ2 (\Cref{s:plan-rq2}), includes many tasks that asks students
+to predict the result(s) of running a program. I have found several
+unanticipated patterns of errors (\Cref{s:misconceptions-tutor}),
+which suggest the existence of unknown misconceptions. The tutor is
+not collecting students' rationale for their answers. So I do not know
+what are the underlying misconceptions (although some of the wrong
+answers are somewhat self-explanatory). I plan to change the tutor
+such that it asks students' for their explanations when they give a
+wrong answer. I plan to eyeball the explanations to find new
+misconceptions and to further confirm that anticipated wrong answers
+do represent the expected misconceptions.
 
@@
 definitional interpreter for each misconception that I have found. I
-plan to do the same things for newly found/confirmed misconceptions.
+plan to do the same things for any new misconceptions.
 
 Eventually, I expect to build a table of misconceptions. Each
-misconception comes with two or more programs. For each of these
+misconception comes with multiple programs. For each of these
 programs, the wrong result that correspond to the misconception should
@@
 To answer this RQ, I have done some literature review on what
-pedagogic techniques are helpful (\Cref{s:background}). My current
-takeaway is that refutation text, case comparison, and notional
-machines are helpful. And I have created a tutoring system (hereafter,
-the tutor) (\Cref{s:system-tutor}) that incorporates refutation text
-and case comparison (\Cref{s:tutor-uiux}). I am adding notional
-machine to the tutor and plan to explore more literature and to
-improve the tutor if I find new helpful techniques.
+pedagogic techniques are helpful (\Cref{s:pedagogic-techniques}). My
+current takeaway is that refutation text, case comparison, notional
+machines, and language levels are helpful.
+
+I have created a tutor (\Cref{s:system-tutor}) that incorporates all
+these ideas (\Cref{s:tutor-uiux}) and have collected data from all
+populations. I plan to explore more literature and to improve the
+tutor if I find new helpful techniques.
 
 I would like to first explain key aspects of the tutor
-(\Cref{s:tutorials,s:interpreting-task}) and then how the tutor
-incorporates ideas of refutation text (\Cref{s:refutationt-text}),
-case comparsion (\Cref{s:case-comparison}), and notional machines
+(\Cref{s:tutorials,s:interpreting-tasks}) and then my plan to
+incorporate ideas of refutation text (\Cref{s:refutationt-text}), case
+comparsion (\Cref{s:case-comparison}), and notional machines
 (\Cref{s:notional-machine}). Finally, I present my plan to evaluate
@@
 
-\subsection{Tutorials}%
+\subsection{Tutorials and Language Levels}%
 \label{s:tutorials}
@@
 managable, I have design the tutor as a sequence of tutorials.
-
-Inspired by prior research on language levels (\Cref{s:background}),
-my sequence of tutorials is ordered, and each later tutorials cover
-one or more language constructs (\Cref{t:tutorials}).
+Inspired by prior research on language levels
+(\Cref{s:pedagogic-techniques}), my sequence of tutorials is ordered,
+and each later tutorials cover one or more language constructs
+(\Cref{t:tutorials}).
 
@@
 
-\subsection{Interpreting Task}%
-\label{s:interpreting-task}
+\subsection{Interpreting Tasks}%
+\label{s:interpreting-tasks}
 
@@
 a wrong answer that correspond to known misconception(s). Let's
-reconsider the program in \Cref{s:interpreting-task}. If a student
+reconsider the program in \Cref{s:interpreting-tasks}. If a student
 answers \lstinline|0 0|, which suggests they hold the
@@
 
-\section{Proposal Roadmap}
-\label{s:roadmap}
-\label{s:univ-1-course}
-
-This proposal describes a four-year effort to obtain a quality instrument
-to measure misconceptions and to build a tutor to address them.
-
-The first step was a two-year process to generate (a) programs that
-tend to trip up students and (b) incorrect responses to those programs
-(\Cref{s:gen-prob-quizius}), using a tool for the purpose
-(\Cref{s:quizius}). These were then curated (\Cref{s:smol-quizzes}) to
-produce an instrument that was tested for one more year
-(\Cref{s:misconceptions-quiz}). This underwent one more round of
-curation to produce the final instrument~(\Cref{s:tutor-refine-program-set}).
-
-I then present the tutor (\Cref{s:system-tutor}). In particular, we
-present the list of misconceptions---grounded in student data---that
-are the focus of this project (\Cref{s:misconceptions-tutor}). I also
-evaluate the tutor's effectiveness at correcting them
-(\Cref{s:effect-tutor}).
-
-All this work is done with students in a ``Principles of Programming
-Languages'' class at a selective, private US university. The class has
-about 70--75 students per year. It is not required, so students take
-it by choice. Virtually all are computer science majors. Most are in
-their third or fourth year of tertiary education; about 10\% are
-graduate students. All have had at least one semester of imperative
-programming, and most have significantly more experience with it. Most
-have had close to a semester of functional programming. The student
-work described here was required, but students were graded on effort,
-not correctness.
-
-Naturally, we should wonder to what extent the demographic affects the
-results: we may simply be studying weaknesses at this institution! We
-therefore work with two other populations
-(\Cref{s:generality-tutor}). In both, we find similar results.
-
-\section{The SMoL Language}\label{s:smol}
-
-\begin{figure}[t]
-  \centering
-  \begin{lstlisting}
-    t ::= d
-        | e
-    d ::= (defvar x e)
-        | (deffun (f x ...) body)
-    e ::= c
-        | x
-        | (lambda (x ...) body)
-        | (let ([x e] ...) body)
-        | (begin e ... e)
-        | (set! x e)
-        | (if e e e)
-        | (cond [e body] ... [else body])
-        | (cond [e body] ...)
-        | (e e ...)
-    body    ::= t ... e
-    program ::= t ...
-  \end{lstlisting}
-  \caption{The syntax of SMoL.}
-  \label{f:smol-syntax}
-\end{figure}
-
-\begin{table}[t]
-  \centering\begin{tabularx}{\linewidth}{r X}
-    \textbf{Operators}            & \textbf{Meaning}                                       \\\hline
-    \texttt{+ - * /}              & Arithmetic Operators                                   \\
-    \texttt{< > <= >=}            & Number comparison                                      \\
-    \texttt{mvec}                 & Create a (mutable) vector (a.k.a. array)               \\
-    \texttt{vec-ref}              & Look up a vector element                               \\
-    \texttt{vec-set!}             & Replace a vector element                               \\
-    \texttt{vec-len}              & Get the length of a vector                             \\
-    \texttt{mpair}                & Create a 2-element (mutable) vector                              \\
-    \texttt{left right}           & Look up the first/second element of a 2-element vector \\
-    \texttt{set-left! set-right!} & Replace the first/second element of a 2-element vector \\
-    \texttt{eq?}                    & Equality                                               \\
-  \end{tabularx}
-  \caption{Primitive operators of SMoL.}
-  \label{t:smol-primitives}
-\end{table}
-
-SMoL is designed to capture common features of many modern
-languages.
-The syntax of SMoL is presented in~\Cref{f:smol-syntax}, where
-\lstinline|t| stands for terms, \lstinline|d| stands for definitions, \lstinline|e|
-stands for expressions, \lstinline|c| stands for constants (i.e., number,
-boolean, and string), and \lstinline|x| and \lstinline|f| are identifiers
-(variables). The last kind of expression (i.e., \lstinline|(e e ...)|) is
-function application.
-
-SMoL's Lispy syntax is
-an artifact of the course (\Cref{s:univ-1-course}) using Racket
-\cite{friedmanEssentialsProgrammingLanguages2001, krishnamurthiProgrammingLanguagesApplication2007};
-however, it also proved to be pedagogically valuable.
-I have found the parentheses useful when discussing
-scope in conjunction with local-binding features like \lstinline|let|.
-This avoids the various confusing ``variable lifting'' semantics found
-in languages like Python (\cite{politzPythonFullMonty2013}; see also
-posts like \cite{froadieWhatScopeVariable2022}), where the actual defined range of a
-variable is not apparent from the source code.
-
-Nevertheless, most SMoL programs are easy to translate to other
-languages. I present machine-translated Python and JavaScript
-versions of our programs in \supplRef{Translated\_Programs.html}. Furthermore, we intend to
-make a multi-lingual tutor in the future (\Cref{s:future-work}).
-
-The semantics of SMoL is as described in \Cref{s:intro}.
-SMoL includes limited primitive operators (\Cref{t:smol-primitives}) to work with numbers, strings, and
-vectors. It provides only one equality operator, which tests for exact
-equality between atomic values and for pointer equality between other
-values.
-
-Students are given a working implementation of SMoL, built
-as a \verb|#lang| language inside Racket \cite{felleisenProgrammableProgrammingLanguage2018}. This language
-provides only the defined syntax and semantics of SMoL, with no
-(other) Racket features present.
-(As in Racket, arguments evaluate left-to-right, to give stateful programs an
-unambiguous semantics.) The implementation is available online:
-\href{https://github.com/shriram/smol}{https://github.com/shriram/smol}.
+\section{What Pedagogic Techniques are Effective?}%
+\label{s:pedagogic-techniques}
+
+The SMoL Tutor is firmly grounded in one technique from
+cognitive and educational psychology. The fundamental problem is: how
+do you fix a misconception? One approach is to only ``present the
+right answer'', for fear that discussing the wrong conception might
+actually reinforce it. Instead, there is a body of literature starting
+from \cite{posnerTheoryConceptualChange1982} that presents a theory of conceptual
+change, at whose heart is the \emph{refutation text}. A refutation
+text tackles the misconception directly, discussing and providing a
+refutation for the incorrect idea. Several studies \cite{schroederRefutationTextFacilitates2022} have shown
+their effectiveness in multiple other domains.
+
+The SMoL Tutor's content structure is also influenced by work on \emph{case comparisons}
+(which draws analogies between examples).
+\cite{alfieriLearningCaseComparisons2013a} suggests that asking
+(rather than not asking) students to find similarities between cases,
+and providing principles \emph{after} the comparisons (rather than
+before or not at all), are associated with better learning
+outcomes.
+
+  [FILL Notional Machine]
+
+  [FILL Language Levels] Prior research argues for the benefits of \emph{language levels}
+[CITE]: a langauge has better be taught as a sequence of languages,
+where each later language includes more langauge constructs than the
+previous one.
 

### 2023-11-05 proposal.tex
save

@@
 
-[FILL]
+[FILL] Use a random default syntax per task? Log syntax switch.
 
@@
         misconceptions?
-  \item How close are students' final conception to the true
-        conception?
 \end{itemize}
@@
   \item In each tutorial, a misconception is either not represented by
-        any interpreting tasks or represented by exactly two tasks.
+        any interpreting tasks or represented by at least two tasks.
   \item Within each tutorial, the order of interpreting tasks is
@@
 \end{itemize}
-
-For each tutorail and each misconception appear in the tutorial, I
-compare student performance on the \emph{first} question that involve
-the misconceptions with performance on the \emph{second} question. If
-the difference is clear (i.e., ``statistically significant''), I
-deduce that the tutor is effective at correcting the misconception.
+For each tutorial and each misconception, I plan to do a model
+comparison (with the standard AIC and BIC criteria) on the following
+models/theories:
+\begin{description}
+  \item[TaskSpecific] Some students are more likely to choose the
+  representing wrong answers in certain tasks.
+  \item[PositionSpecific] Students are more likely to choose the
+  representing wrong answer in a task if the task appears earlier (or
+  later) in the tutorial.
+\end{description}
+
+If the \textbf{TaskSpecific} model best explains the data, I deduce
+that some programs are more confusing and that we need further
+investigation into the relationship between the tasks and misconceptions.
+
+If the \textbf{PositionSpecific} model best explains \emph{and} students
+perform better in later tasks, I deduce that the tutor seeems to be
+correcting misconceptions. If studnets perform worse in later taks, the
+tutor is potentially doing harm.
+
+To know students well, I would like to have as many tasks per
+misconception as possible. But in practice, the number of tasks is
+subject to the following limitations:
+\begin{enumerate}
+  \item I don't want the length of any tutorial to exceed 30 minutes.
+  Because the tutor is now asking for student explanation on choice,
+  students are likely to spend more time on each task then before.
+  \item Programs shouldn't be too long. Some misconceptions are really
+  similar and as a consequence have only a small number of short
+  programs to tell them apart.
+\end{enumerate}
+
+The current tutor is not in the final shape yet. In particular, many
+of the wrong answers correspond to more than one misconceptions; some
+misconceptions are represented by only one question; the order of
+interpreting tasks are not randomized. I plan to tackles the first two
+differences by adding and removing programs. Adding randomization
+requires me to carefully check the dependencies between current
+explanations and goal sentences and resolve those dependencies if any.
 

### 2023-11-07 proposal.tex
Clean up with helps from Overleaf

@@
 \end{lstlisting}
-That is, a user would expect to see \verb|0| through \verb|9|. In
-fact, however, it prints \verb|9| ten times.
+That is, a user would expect to see \lstinline|0| through \lstinline|9|. In
+fact, however, it prints \lstinline|9| ten times.
 
@@
 specific looping problem even trips up industrial programmers. The
-C\# language \emph{changed} to produce \verb|0| through \verb|9| \cite{lippertClosingLoopVariable2009}. It is now also the focus of a language design
+C\# language \emph{changed} to produce \lstinline|0| through \lstinline|9| \cite{lippertClosingLoopVariable2009}. It is now also the focus of a language design
 change in Go \cite{chaseFixingLoopsGo2023} for a new kind of looping
@@
 SMoL's Lispy syntax is
-an artifact of the course (\Cref{s:univ-1-course}) using Racket
+an artifact of the course (\Cref{s:populations}) using Racket
 \cite{friedmanEssentialsProgrammingLanguages2001, krishnamurthiProgrammingLanguagesApplication2007};
@@
 Students are given a working implementation of SMoL, built
-as a \verb|#lang| language inside Racket \cite{felleisenProgrammableProgrammingLanguage2018}. This language
+as a \lstinline|#lang| language inside Racket \cite{felleisenProgrammableProgrammingLanguage2018}. This language
 provides only the defined syntax and semantics of SMoL, with no
@@
 \end{description}
+
+\subsection{Multiple Populations}%
+\label{s:populations}
 
@@
 
+\subsection{Find Misconceptions in Student-written Programs}%
+\label{s:beyond-experts-blindsopts}
+
 To find a reasonably comprehensive collection of misconceptions, I
@@
 
+\subsection{Find Misconceptions with Expert-written Programs}
+
 I have another source of misconceptions. I have been creating a tutor
@@
 do represent the expected misconceptions.
+
+\subsection{Multiple-Syntax}
+
+[TODO]
+
+\subsection{Expected Results}
 
@@
 
-In \Cref{s:rel-work}, we discuss several papers that have provided
+In \Cref{s:background-misconceptions}, we discuss several papers that have provided
 reports of student misconceptions with different fragments of
@@
   misconception the student has}. For instance, if the question is
-``What is \verb|sqrt(4)|?'', then \verb|37| is probably an
+``What is \lstinline|sqrt(4)|?'', then \lstinline|37| is probably an
 uninteresting wrong answer, but if people appear to confuse
-square-roots with squares, then \verb|16| would be present as an
+square-roots with squares, then \lstinline|16| would be present as an
 answer.\footnote{Concept inventories are thus useful in many
@@
 students and asked to predict (without running it) the value produced
-by the program.\footnote{In the course (\Cref{s:univ-1-course}),
+by the program.\footnote{In the course (\Cref{s:populations}),
   students were given credit for using Quizius but not penalized for wrong
@@
 
-After running Quizius in the course (\Cref{s:univ-1-course}),
+After running Quizius in the course (\Cref{s:populations}),
 we took over as experts. Determining which is the right
@@
 that the experts did not imagine. In our case, we seeded Quizius with
-programs from prior papers (\Cref{s:rel-work}), which gives the first
+programs from prior papers (\Cref{s:background-misconceptions}), which gives the first
 few students programs to respond to. However, we found
@@
 
-        A reader might think that \verb|sum| takes variable arguments (so
-        the program produces \verb|6|), but in fact \verb|...| is a single
+        A reader might think that \lstinline|sum| takes variable arguments (so
+        the program produces \lstinline|6|), but in fact \lstinline|...| is a single
         variable, so this produces an arity error.
@@
 still effective.
-Syntactically, \verb|#t| and \verb|#f| are true and false,
-while \verb|#(...)| is a vector. I use a * to indicate the correct
+Syntactically, \lstinline|#t| and \lstinline|#f| are true and false,
+while \lstinline|#(...)| is a vector. I use a * to indicate the correct
 answer (which also matches the implementation's output).
@@
         \begin{tabularx}{\linewidth}{r X}
-          $^\dag$59\% & \verb|2 0|             \\
-          *32\%       & \verb|Error|           \\
-          4\%         & \verb|Other: 2 2|      \\
+          $^\dag$59\% & \lstinline|2 0|             \\
+          *32\%       & \lstinline|Error|           \\
+          4\%         & \lstinline|Other: 2 2|      \\
           3\%         & ``Nothing is printed'' \\
-          2\%         & \verb|0 0|             \\
+          2\%         & \lstinline|0 0|             \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          *65\%       & \verb|Error| \\
-          $^\dag$29\% & \verb|42 #t| \\
+          *65\%       & \lstinline|Error| \\
+          $^\dag$29\% & \lstinline|42 #t| \\
           6\%         & Other        \\
@@
         \begin{tabularx}{\linewidth}{r X}
-          *53\%       & \verb|#(6 5) 5|                 \\
-          *20\%       & \verb|Error|                    \\
-          $^\dag$11\% & \verb|#(6 6) 5|                 \\
-          $^\dag$11\% & \verb|#(6 6) 6|                 \\
+          *53\%       & \lstinline|#(6 5) 5|                 \\
+          *20\%       & \lstinline|Error|                    \\
+          $^\dag$11\% & \lstinline|#(6 6) 5|                 \\
+          $^\dag$11\% & \lstinline|#(6 6) 6|                 \\
           5\%         & ``Nothing is printed'' or Other \\
@@
         \begin{tabularx}{\linewidth}{r X}
-          $^\dag$50\% & \verb|#(#(2 #(2 3)) #(2 3))| \\
-          *29\%       & \verb|x=#(x x)| or
-          something similar. Both \verb|(left x)| and \verb|(right x)|
-          are \verb|x| itself.                       \\
-          $^\dag$16\% & \verb|Error|                 \\
+          $^\dag$50\% & \lstinline|#(#(2 #(2 3)) #(2 3))| \\
+          *29\%       & \lstinline|x=#(x x)| or
+          something similar. Both \lstinline|(left x)| and \lstinline|(right x)|
+          are \lstinline|x| itself.                       \\
+          $^\dag$16\% & \lstinline|Error|                 \\
           5\%         & Other                        \\
@@
         \begin{tabularx}{\linewidth}{r X}
-          *62\%       & \verb|#(#(100 2 3 4) #(100 2 3 4))| \\
-          $^\dag$26\% & \verb|#(#(1 2 3 4) #(100 2 3 4))|   \\
+          *62\%       & \lstinline|#(#(100 2 3 4) #(100 2 3 4))| \\
+          $^\dag$26\% & \lstinline|#(#(1 2 3 4) #(100 2 3 4))|   \\
           6\%         & \texttt{Error}                      \\
-          4\%         & \verb|#(#(1 2 3 4) #(1 2 3 4))|     \\
+          4\%         & \lstinline|#(#(1 2 3 4) #(1 2 3 4))|     \\
           2\%         & Other                               \\
@@
         \begin{tabularx}{\linewidth}{r X}
-          *59\%       & \verb|#(10)|  \\
-          $^\dag$36\% & \verb|#(123)| \\
+          *59\%       & \lstinline|#(10)|  \\
+          $^\dag$36\% & \lstinline|#(123)| \\
           5\%         & Other         \\
@@
         \begin{tabularx}{\linewidth}{r X}
-          *82\%       & \verb|3|     \\
-          $^\dag$17\% & \verb|Error| \\
+          *82\%       & \lstinline|3|     \\
+          $^\dag$17\% & \lstinline|Error| \\
           1\%         & Other        \\
@@
         \begin{tabularx}{\linewidth}{r X}
-          *74\%       & \verb|4|            \\
-          $^\dag$21\% & \verb|3|            \\
-          6\%         & Other: \verb|Error| \\
+          *74\%       & \lstinline|4|            \\
+          $^\dag$21\% & \lstinline|3|            \\
+          6\%         & Other: \lstinline|Error| \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          *88\%       & \verb|4|            \\
-          $^\dag$10\% & \verb|3|            \\
-          3\%         & Other: \verb|Error| \\
+          *88\%       & \lstinline|4|            \\
+          $^\dag$10\% & \lstinline|3|            \\
+          3\%         & Other: \lstinline|Error| \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          *62\%       & \verb|1 1 2 3 2| \\
-          $^\dag$34\% & \verb|1 1 1 1 1| \\
+          *62\%       & \lstinline|1 1 2 3 2| \\
+          $^\dag$34\% & \lstinline|1 1 1 1 1| \\
           4\%         & Other            \\
-          1\%         & \verb|1 1 2 3 4| \\
+          1\%         & \lstinline|1 1 2 3 4| \\
         \end{tabularx}
@@
         This checks whether $\verb|n| < \verb|3|$, but some presumably
-        vocalized it as ``greater than \verb|3|, \verb|n|?''.
+        vocalized it as ``greater than \lstinline|3|, \lstinline|n|?''.
 
@@
 
-\begin{table}[!h]
+\begin{table}[!ht]
   \centering\begin{tabularx}{\linewidth}{X l X}
@@
         \begin{tabularx}{\linewidth}{r X}
-          78\%   & \verb|12|    \\
-          11\%   & \verb|error| \\
-          **10\% & \verb|0|     \\
-          1\%    & \verb|23|    \\
+          78\%   & \lstinline|12|    \\
+          11\%   & \lstinline|error| \\
+          **10\% & \lstinline|0|     \\
+          1\%    & \lstinline|23|    \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          90\%   & \verb|#(173 0)| \\
-          **10\% & \verb|#(1 0)|   \\
+          90\%   & \lstinline|#(173 0)| \\
+          **10\% & \lstinline|#(1 0)|   \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          86\%   & \verb|4|      \\
-          **10\% & \verb|3|      \\
-          3\%    & \verb|error|  \\
-          1\%    & \verb|lambda| \\
+          86\%   & \lstinline|4|      \\
+          **10\% & \lstinline|3|      \\
+          3\%    & \lstinline|error|  \\
+          1\%    & \lstinline|lambda| \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          85\%   & \verb|0 12|                \\
-          **12\% & \verb|0 0|                 \\
-          2\%    & \verb|error|               \\
+          85\%   & \lstinline|0 12|                \\
+          **12\% & \lstinline|0 0|                 \\
+          2\%    & \lstinline|error|               \\
           1\%    & depends on implementation. \\
@@
         \begin{tabularx}{\linewidth}{r X}
-          85\%   & \verb|error| \\
-          **15\% & \verb|2|     \\
+          85\%   & \lstinline|error| \\
+          **15\% & \lstinline|2|     \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          **67\% & \verb|#(100)| \\
-          30\%   & \verb|#(200)| \\
-          1\%    & \verb|#(300)| \\
-          1\%    & \verb|error|  \\
+          **67\% & \lstinline|#(100)| \\
+          30\%   & \lstinline|#(200)| \\
+          1\%    & \lstinline|#(300)| \\
+          1\%    & \lstinline|error|  \\
         \end{tabularx}
@@
 
-\begin{table}[!h]
+\begin{table}[!ht]
   \centering
@@
         \begin{tabularx}{\linewidth}{r X}
-          96\%  & \verb|error| \\
-          **4\% & \verb|402|   \\
+          96\%  & \lstinline|error| \\
+          **4\% & \lstinline|402|   \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          83\%   & \verb|4|     \\
-          **11\% & \verb|error| \\
-          4\%    & \verb|2|     \\
-          1\%    & \verb|8|     \\
+          83\%   & \lstinline|4|     \\
+          **11\% & \lstinline|error| \\
+          4\%    & \lstinline|2|     \\
+          1\%    & \lstinline|8|     \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          77\%   & \verb|3|     \\
-          **23\% & \verb|error| \\
+          77\%   & \lstinline|3|     \\
+          **23\% & \lstinline|error| \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          76\%   & \verb|3|                   \\
-          **14\% & \verb|error|               \\
+          76\%   & \lstinline|3|                   \\
+          **14\% & \lstinline|error|               \\
           9\%    & Run out of memory or time. \\
-          1\%    & \verb|+inf|                \\
+          1\%    & \lstinline|+inf|                \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          67\%   & \verb|#(1 2 3)| \\
-          **24\% & \verb|#(1 2 4)| \\
+          67\%   & \lstinline|#(1 2 3)| \\
+          **24\% & \lstinline|#(1 2 4)| \\
           9\%    & error           \\
@@
         \begin{tabularx}{\linewidth}{r X}
-          65\%  & \verb|#0=#(#0# #0#)| (Racket circular object notation) \\
-          24\%  & \verb|#(#(2 3) #(2 3))|                                \\
-          **6\% & \verb|#(#(2 #(2 3)) #(2 3))|                           \\
+          65\%  & \lstinline|#0=#(#0# #0#)| (Racket circular object notation) \\
+          24\%  & \lstinline|#(#(2 3) #(2 3))|                                \\
+          **6\% & \lstinline|#(#(2 #(2 3)) #(2 3))|                           \\
           6\%   & error                                                  \\
@@
         \begin{tabularx}{\linewidth}{r X}
-          92\%  & \verb|2| \\
-          **8\% & \verb|1| \\
+          92\%  & \lstinline|2| \\
+          **8\% & \lstinline|1| \\
         \end{tabularx}
@@
         \begin{tabularx}{\linewidth}{r X}
-          **57\% & \verb|1 3|   \\
-          43\%   & \verb|error| \\
+          **57\% & \lstinline|1 3|   \\
+          43\%   & \lstinline|error| \\
         \end{tabularx}
@@
 \end{lstlisting}
-This program, suitably translated, would produce \verb|2| in a wide
+This program, suitably translated, would produce \lstinline|2| in a wide
 variety of languages (Python, JavaScript, Racket, Java, etc.), because
-\verb|get-x| and the inner \verb|x| are in the same scope block. The
-answer \verb|1| cannot be explained by any of our existing
+\lstinline|get-x| and the inner \lstinline|x| are in the same scope block. The
+answer \lstinline|1| cannot be explained by any of our existing
 misconceptions. Based on this, we surmise a new misconception,
@@
 program from \quizFun---which should be an error, due to the
-double-binding of \verb|x| in the same scope block---and the answer
-\verb|2 0|:
+double-binding of \lstinline|x| in the same scope block---and the answer
+\lstinline|2 0|:
 \begin{lstlisting}
@@
 students had stated that the second \lstinline|(defvar x ....)|
-``mutates'' or ``redefines'' \verb|x|. This is reminiscent of the
+``mutates'' or ``redefines'' \lstinline|x|. This is reminiscent of the
 behavior of languages like Python, which use the same syntax both for
@@
 
-\begin{figure}[!h]
+\begin{figure}[p]
 
@@
 There is, of course, a significant danger that the data above have
-been overfitted to only one institution (\Cref{s:univ-1-course},
+been overfitted to only one institution (\Cref{s:populations},
 henceforth University 1),
@@
 have some reason to believe this is not the case: the related work
-discussed in \Cref{s:rel-work} is drawn from many institutions in
+discussed in \Cref{s:background-misconceptions} is drawn from many institutions in
 multiple countries with different educational preparations, levels,
@@
 that the results are not specific to one institution, and
-\Cref{s:rel-work} suggests these issues are widespread. Nevertheless,
+\Cref{s:background-misconceptions} suggests these issues are widespread. Nevertheless,
 much more broad testing is needed to confirm our specific instruments.

### 2023-11-07 proposal.tex
try to switch to natbib
clean up citations

@@
 \usepackage{titlesec} % For customizing section titles
-\usepackage[style=authoryear]{biblatex}  % For managing references
-\addbibresource{bibfile.bib}
+% \usepackage[style=authoryear]{biblatex}  % For managing references
+\usepackage[round, square]{natbib}
+\bibliographystyle{plainnat}
 \usepackage{hyperref}
@@
 JavaScript, Python, and Ruby, to ``functional'' languages like the ML
-and Lisp families. Of course, there are sometimes restrictions
-(e.g., Java has restrictions on closures) and extensions (such as the
-documented semantic oddities of JavaScript and Python \cite{bernhardtWat2012, guhaEssenceJavascript2010, politzPythonFullMonty2013, politzTestedSemanticsGetters2012}).
-Still, this semantic core bridges many syntaxes, and
-understanding it helps when transferring knowledge from old languages
-to new ones. In recognition of this deep commonality, in this paper we
-choose to call this the \emph{Standard Model of Languages} (SMoL).
+and Lisp families. Of course, there are sometimes restrictions (e.g.,
+Java has restrictions on closures) and extensions (such as the
+documented semantic oddities of JavaScript and Python
+\citep{bernhardtWat2012, guhaEssenceJavascript2010,
+politzPythonFullMonty2013, politzTestedSemanticsGetters2012}). Still,
+this semantic core bridges many syntaxes, and understanding it helps
+when transferring knowledge from old languages to new ones. In
+recognition of this deep commonality, in this paper we choose to call
+this the \emph{Standard Model of Languages} (SMoL).
 
@@
 specific looping problem even trips up industrial programmers. The
-C\# language \emph{changed} to produce \lstinline|0| through \lstinline|9| \cite{lippertClosingLoopVariable2009}. It is now also the focus of a language design
-change in Go \cite{chaseFixingLoopsGo2023} for a new kind of looping
-construct \cite{coxSpecLessErrorprone2023}.
+C\# language \emph{changed} to produce \lstinline|0| through \lstinline|9| \citep{lippertClosingLoopVariable2009}. It is now also the focus of a language design
+change in Go \citep{chaseFixingLoopsGo2023} for a new kind of looping
+construct \citep{coxSpecLessErrorprone2023}.
 It continues to trip up programmers
@@
 despite being
-documented as a ``gotcha'' in Python \cite{reitzHitchhikerGuidePython2016}.
+documented as a ``gotcha'' in Python \citep{reitzHitchhikerGuidePython2016}.
 % SK: But not everyone has read the document. I don't think the document is
@@
 good mental model of these features
-is essential to understand ownership \cite{clarkeOwnershipTypesFlexible1998},
+is essential to understand ownership \citep{clarkeOwnershipTypesFlexible1998},
 manage parallelism, and more. Thus, we
@@
 an artifact of the course (\Cref{s:populations}) using Racket
-\cite{friedmanEssentialsProgrammingLanguages2001, krishnamurthiProgrammingLanguagesApplication2007};
+\citep{friedmanEssentialsProgrammingLanguages2001, krishnamurthiProgrammingLanguagesApplication2007};
 however, it also proved to be pedagogically valuable.
@@
 Students are given a working implementation of SMoL, built
-as a \lstinline|#lang| language inside Racket \cite{felleisenProgrammableProgrammingLanguage2018}. This language
+as a \lstinline|#lang| language inside Racket \citep{felleisenProgrammableProgrammingLanguage2018}. This language
 provides only the defined syntax and semantics of SMoL, with no
@@
 about a ``two-loop'' architecture
-\cite{vanlehnBehaviorTutoringSystems2006a} where the outer loop
+\citep{vanlehnBehaviorTutoringSystems2006a} where the outer loop
 iterates through ``tasks'' (i.e., educational activities) and the
@@
 well-known and heavily studied LISP Tutor
-\cite{andersonLISPTutor1985}), and in the process undoubtedly address
+\citep{andersonLISPTutor1985}), and in the process undoubtedly address
 some program behavior misconceptions. My tutor differs in a notable
@@
 text tackles the misconception directly, discussing and providing a
-refutation for the incorrect idea. Several studies \cite{schroederRefutationTextFacilitates2022} have shown
+refutation for the incorrect idea. Several studies \citep{schroederRefutationTextFacilitates2022} have shown
 their effectiveness in multiple other domains.
@@
 incomplete. Education researchers have documented the
-phenomenon of the \emph{expert blind spot} \cite{nathanExpertBlindSpot2001}: experts
+phenomenon of the \emph{expert blind spot} \citep{nathanExpertBlindSpot2001}: experts
 simply do not conceive of many learner difficulties.
@@
 Finally, we are inspired by the significant body of education research
-on \emph{concept inventories} \cite{hestenesForceConceptInventory1992} (with a growing
-number for computer science, as a survey lists \cite{taylorComputerScienceConcept2014}). In
+on \emph{concept inventories} \citep{hestenesForceConceptInventory1992} (with a growing
+number for computer science, as a survey lists \citep{taylorComputerScienceConcept2014}). In
 terms of mechanics, a concept inventory is just an instrument
@@
 Our main solution to the expert blind spot is to use the Quizius system
-\cite{saarinenHarnessingWisdomClasses2019}. In contrast to the very
+\citep{saarinenHarnessingWisdomClasses2019}. In contrast to the very
 heavyweight process (involving a lot of expert time)
@@
 one. Quizius thus treats this as a multi-armed bandit problem
-\cite{katehakisMultiArmedBanditProblem1987} and uses that to choose a program.
+\citep{katehakisMultiArmedBanditProblem1987} and uses that to choose a program.
 
@@
 
-Prior work using Quizius \cite{saarinenHarnessingWisdomClasses2019} finds that students do author programs
+Prior work using Quizius \citep{saarinenHarnessingWisdomClasses2019} finds that students do author programs
 that the experts did not imagine. In our case, we seeded Quizius with
@@
 inventory. Their creation requires heavyweight processes (such as
-Delphi methods \cite{goldmanSettingScopeConcept2010}) that require many hours of expert attention as
+Delphi methods \citep{goldmanSettingScopeConcept2010}) that require many hours of expert attention as
 well as conversations with learners, and can hence be prohibitive in
@@
 Our idea of misinterpreters is related to mystery languages
-\cite{diwanPLdetectiveSystemTeaching2004,
+\citep{diwanPLdetectiveSystemTeaching2004,
   pombrioTeachingProgrammingLanguages2017}. Both approaches use
@@
 
-\printbibliography
+\bibliography{bibfile}
 

### 2023-11-07 proposal.tex
the SMoL Lang

@@
 
-SMoL is designed to capture common features of many modern
-languages.
+SMoL is designed to capture common features of many modern languages.
 The syntax of SMoL is presented in~\Cref{f:smol-syntax}, where
-\lstinline|t| stands for terms, \lstinline|d| stands for definitions, \lstinline|e|
-stands for expressions, \lstinline|c| stands for constants (i.e., number,
-boolean, and string), and \lstinline|x| and \lstinline|f| are identifiers
-(variables). The last kind of expression (i.e., \lstinline|(e e ...)|) is
-function application.
-
-SMoL's Lispy syntax is
-an artifact of the course (\Cref{s:populations}) using Racket
-\citep{friedmanEssentialsProgrammingLanguages2001, krishnamurthiProgrammingLanguagesApplication2007};
-however, it also proved to be pedagogically valuable.
-I have found the parentheses useful when discussing
-scope in conjunction with local-binding features like \lstinline|let|.
-This avoids the various confusing ``variable lifting'' semantics found
-in languages like Python (\cite{politzPythonFullMonty2013}; see also
-posts like \cite{froadieWhatScopeVariable2022}), where the actual defined range of a
-variable is not apparent from the source code.
+\lstinline|t| stands for terms, \lstinline|d| stands for definitions,
+\lstinline|e| stands for expressions, \lstinline|c| stands for
+constants (i.e., number, boolean, and string), and \lstinline|x| and
+\lstinline|f| are identifiers (variables). The last kind of expression
+(i.e., \lstinline|(e e ...)|) is function application.
+
+SMoL defines a semantic but not the syntax. However, to repsent SMoL
+to students, I have to choose a syntax. I chose a Lispy syntax, which
+is an artifact of the major population where I conducted studies: a
+course (\Cref{s:populations}) using Racket
+\citep{friedmanEssentialsProgrammingLanguages2001,
+krishnamurthiProgrammingLanguagesApplication2007}. However, the Lispy
+syntax also proved to be pedagogically valuable. I have found the
+parentheses useful when discussing scope in conjunction with
+local-binding features like \lstinline|let|. This avoids the various
+confusing ``variable lifting'' semantics found in languages like
+Python (\cite{politzPythonFullMonty2013}; see also posts like
+\cite{froadieWhatScopeVariable2022}), where the actual defined range
+of a variable is not apparent from the source code.
 
 Nevertheless, most SMoL programs are easy to translate to other
-languages. I present machine-translated Python and JavaScript
-versions of our programs in \supplRef{Translated\_Programs.html}. Furthermore, we intend to
-make a multi-lingual tutor in the future (\Cref{s:future-work}).
-
-The semantics of SMoL is as described in \Cref{s:intro}.
-SMoL includes limited primitive operators (\Cref{t:smol-primitives}) to work with numbers, strings, and
-vectors. It provides only one equality operator, which tests for exact
-equality between atomic values and for pointer equality between other
-values.
-
-Students are given a working implementation of SMoL, built
-as a \lstinline|#lang| language inside Racket \citep{felleisenProgrammableProgrammingLanguage2018}. This language
+languages. I present machine-translated Python and JavaScript versions
+of my programs in \supplRef{Translated\_Programs.html}. Furthermore, I
+intend to make a multi-lingual tutor in the future
+(\Cref{s:future-work}).
+
+The semantics of SMoL is as described in \Cref{s:intro}. SMoL includes
+limited primitive operators (\Cref{t:smol-primitives}) to work with
+numbers, strings, and vectors. It provides only one equality operator,
+which tests for exact equality between atomic values and for pointer
+equality between other values.
+
+Students are given a working implementation of SMoL, built as a
+\lstinline|#lang| language inside Racket
+\citep{felleisenProgrammableProgrammingLanguage2018}. This language
 provides only the defined syntax and semantics of SMoL, with no
-(other) Racket features present.
-(As in Racket, arguments evaluate left-to-right, to give stateful programs an
-unambiguous semantics.) The implementation is available online:
+(other) Racket features present. (As in Racket, arguments evaluate
+left-to-right, to give stateful programs an unambiguous semantics.)
+The implementation is available online:
 \href{https://github.com/shriram/smol}{https://github.com/shriram/smol}.

### 2023-11-07 proposal.tex
background: misconceptions

@@
 The implementation is available online:
-\href{https://github.com/shriram/smol}{https://github.com/shriram/smol}.
+\begin{center}
+  \href{https://github.com/shriram/smol}{https://github.com/shriram/smol}
+\end{center}
 
@@
   \centering
-  \begin{tabularx}{\textwidth}{>{\hsize=.6\hsize\RaggedRight}X >{\hsize=.8\hsize\RaggedRight}X >{\hsize=.5\hsize\RaggedRight}X >{\hsize=1.4\hsize\RaggedRight}X}
+  \begin{tabularx}{\textwidth}{>{\hsize=.6\hsize\RaggedRight}X >{\hsize=.8\hsize\RaggedRight}X >{\hsize=.6\hsize\RaggedRight}X >{\hsize=1.3\hsize\RaggedRight}X}
     \textbf{Publication}                                                           &
@@
     \textbf{Languages}                                                             &
-    \textbf{Misconceptions}
+    \textbf{Topics}
     \\\hline
     \cite{fleuryParameterPassingRules1991}                                         &
-    Unclear, likely CS2 students                                                   &
+    Likely CS2 students                                                   &
     Pascal                                                                         &
-    \surmisedMiscon{CallerEnv}; \miscon{IsolatedFun}; \miscon{DefOrSet}
-    (See their RULEs 2, 3a, and 3b in TABLE 2;
-    RULE 1 doesn't apply to us.)
+    Scope
     \\\hline
@@
     Java and Scheme                                                                &
-    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{CallsCopyStruct}; \miscon{DefByRef}
-    (See their Section 4)
+    Scope, variable aliasing, and structure aliasing.
     \\\hline
@@
     Java                                                                           &
-    \miscon{StructByRef} (Their G2); \miscon{DefByRef} or \miscon{DefCopyStructs} (Their G3); \miscon{CallsCopyStruct} (Their G4).
+    Variable aliasing and structure aliasing.
     \\\hline
@@
     Python                                                                         &
-    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
-    (See their section 4.2)
+    Scope, variable aliasing, and structure aliasing.
     \\\hline
@@
     C++                                                                            &
-    \miscon{FlatEnv}; \miscon{CallByRef}; \miscon{DefsCopyStruct}
-    (See their section 4.2)
-    \\\hline
-  \end{tabularx}
-  \caption{Similar misconceptions found in prior research.}
+    Scope, variable aliasing, and structure aliasing.
+    \\\hline
+  \end{tabularx}
+  \caption{Topics of misconceptions found in prior research.}
   \label{t:rel-work}
@@
 \Cref{t:rel-work} lists works that seem most relevant to us.
-
-There are some small differences.
-\cite{fleuryParameterPassingRules1991} identifies a dynamic scope misconception
-that is different from \miscon{FlatEnv}:
-\begin{description}
-  \item[\surmisedMiscon{CallerEnv}] Function values don't remember their environments.
-    When a function is called, the function body is evaluated in an
-    environment that extends from the \textit{caller's} environment.
-\end{description}
-I don't include \surmisedMiscon{CallerEnv} in our analysis because in our
-data, all wrong answers that can be explained by \surmisedMiscon{CallerEnv}
-are also explainable by \miscon{FlatEnv}.
-
 Appendix A of \cite{sorvaVisualProgramSimulation2012} provides an
@@
 in research up to 2012.
-There are overlaps between our survey and
-theirs. For instance, our \surmisedMiscon{CallerEnv} is their No.~47.
-However, because their descriptions are brief, it is difficult to tell
-whether a misconception in their survey matches our misconceptions.
-Because
-they provide neither misinterpreters nor representative program-output
-pairs, it is difficult to determine the overlaps precisely (showing
-the value of providing these two machine-runnable descriptions).
-At any rate,
-we certainly find
-no equivalent of \miscon{FlatEnv}, \miscon{DeepClosure}, and
-\miscon{DefByRef} in their survey.
+
+% There are some small differences.
+% \cite{fleuryParameterPassingRules1991} identifies a dynamic scope misconception
+% that is different from \miscon{FlatEnv}:
+% \begin{description}
+%   \item[\surmisedMiscon{CallerEnv}] Function values don't remember their environments.
+%     When a function is called, the function body is evaluated in an
+%     environment that extends from the \textit{caller's} environment.
+% \end{description}
+% I don't include \surmisedMiscon{CallerEnv} in our analysis because in our
+% data, all wrong answers that can be explained by \surmisedMiscon{CallerEnv}
+% are also explainable by \miscon{FlatEnv}.
+
+% There are overlaps between our survey and
+% theirs. For instance, our \surmisedMiscon{CallerEnv} is their No.~47.
+% However, because their descriptions are brief, it is difficult to tell
+% whether a misconception in their survey matches our misconceptions.
+% Because
+% they provide neither misinterpreters nor representative program-output
+% pairs, it is difficult to determine the overlaps precisely (showing
+% the value of providing these two machine-runnable descriptions).
+% At any rate,
+% we certainly find
+% no equivalent of \miscon{FlatEnv}, \miscon{DeepClosure}, and
+% \miscon{DefByRef} in their survey.
 

### 2023-11-07 proposal.tex
RQ1

@@
 
-\subsection{Multiple Populations}%
+\subsection{Populations}%
 \label{s:populations}
@@
 
+I am contacting professors from various other universities. Several of
+them have expressed interests in using my instruments. So the final
+results will likely include more populations.
+
 \subsection{Find Misconceptions in Student-written Programs}%
-\label{s:beyond-experts-blindsopts}
+\label{s:find-misconceptions-with-quizius}
 
@@
 misconceptions not envisioned by experts and hence is hindered by
-experts' blind spots. To overcome such bias, I analyze a University~1
-dataset produced by a two-year process to find (a) student-written
-programs that tend to trip up students and (b) incorrect responses to
-those programs (\Cref{s:gen-prob-quizius}), using a tool for the
-purpose (\Cref{s:quizius}). I have been cleaning up those programs
+experts' blind spots.
+
+To overcome such bias, I analyze a University~1 dataset produced by a
+two-year process to find (a) student-written programs that tend to
+trip up students and (b) incorrect responses to those programs
+(\Cref{s:gen-prob-quizius}), using a tool for the purpose
+(\Cref{s:quizius}). I have been cleaning up those programs
 (\Cref{s:smol-quizzes,s:tutor-refine-program-set}) and confirming that
@@
 
-\subsection{Find Misconceptions with Expert-written Programs}
+\subsection{Find Misconceptions with Expert-written Programs}%
+\label{s:find-misconceptions-with-tutor}
 
 I have another source of misconceptions. I have been creating a tutor
-to answer my RQ2. The tutor, which I explain in details when I talk
-about RQ2 (\Cref{s:plan-rq2}), includes many tasks that asks students
-to predict the result(s) of running a program. I have found several
-unanticipated patterns of errors (\Cref{s:misconceptions-tutor}),
-which suggest the existence of unknown misconceptions. The tutor is
-not collecting students' rationale for their answers. So I do not know
-what are the underlying misconceptions (although some of the wrong
-answers are somewhat self-explanatory). I plan to change the tutor
-such that it asks students' for their explanations when they give a
-wrong answer. I plan to eyeball the explanations to find new
-misconceptions and to further confirm that anticipated wrong answers
-do represent the expected misconceptions.
-
-\subsection{Multiple-Syntax}
-
-[TODO]
+to answer my RQ2, which I explain in details in \Cref{s:plan-rq2}. The
+tutor includes many tasks that asks students to predict the result(s)
+of running a program. Although these programs originate from
+student-written programs (\Cref{s:find-misconceptions-with-quizius}), they
+have been modified to fit tutoring purpose. Perhaps because of the
+modification or because of the population variance, I have found
+several unanticipated patterns of errors
+(\Cref{s:misconceptions-tutor}), which suggest unknown misconceptions.
+
+I plan to change the tutor such that it asks students' to explain
+their prediction. I plan to eyeball the explanations to find new
+misconceptions. This piece of data also helps me to confirm that
+anticipated wrong answers do represent the expected misconceptions.
+
+\subsection{Reduce Syntactic Bias}
+
+The processes described in
+\Cref{s:find-misconceptions-with-quizius,s:find-misconceptions-with-tutor}
+have been conducted with SMoL's Lispy syntax (\Cref{s:smol}). This can
+potentially introduce biases to the found misconceptions. For example,
+current results find that some students think the following program
+produces \lstinline|'(4 5)| rather than \lstinline|'(1 2)|.
+\begin{lstlisting}
+  (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
+\end{lstlisting}
+This checks whether $\verb|3| > \verb|n|$, but those students
+presumably vocalized it as ``greater than \lstinline|3|,
+\lstinline|n|?''. Although I have tried to remove programs like this
+from my instruments, there is still a risk of having biase in student
+errors. I plan to adjust the process descrbed in
+\Cref{s:find-misconceptions-with-tutor} such that it presents programs
+in multiple languages: the SMoL version, a JavaScript translation, and
+a Python version.
 
 \subsection{Expected Results}
-
-To be reasonably specific about misconceptions, I have created a
-definitional interpreter for each misconception that I have found. I
-plan to do the same things for any new misconceptions.
 
@@
 explained something like ``\lstinline|y| is \lstinline|x|''.
+
+To be reasonably specific about misconceptions, I have created a
+definitional interpreter for each misconception that I have found. I
+plan to do the same things for any new misconceptions.
 

### 2023-11-07 proposal.tex
rq2

@@
 
-To find a reasonably comprehensive collection of misconceptions, I
-seek for inputs from students. Most prior works
-(\Cref{s:background-misconceptions}) have experts coming up with a
-list of potential misconceptions and then filter the potential
-misconceptions with student inputs. This approach can't find
-misconceptions not envisioned by experts and hence is hindered by
-experts' blind spots.
+In \Cref{s:background-misconceptions}, I discuss several papers that
+have provided reports of student misconceptions with different
+fragments of SMoL. However, it is difficult to know how comprehensive
+these are. While some are unclear on the origin of their programs,
+they generally seem to be expert-generated.
+
+The problem with expert-generated lists is that they can be quite
+incomplete. Education researchers have documented the phenomenon of
+the \emph{expert blind spot} \citep{nathanExpertBlindSpot2001}:
+experts simply do not conceive of many learner difficulties.
 
@@
 
-\subsection{Expected Results}
+\subsection{Expected Results}%
+\label{s:rq1-expected-results}
 
@@
 the wrong answers. For example, I might find a misconception called
-\miscon{DefByRef}: when students are asked to predict the result of
+\miscon{DefByRef} (See
+\Cref{t:tutor-grounded-misconceptions-1,t:tutor-grounded-misconceptions-2,t:tutor-surmised-misconceptions}
+in \Cref{s:misconceptions-tutor} for a list of misconceptions that I
+have found): when students are asked to predict the result of
 \begin{lstlisting}
@@
 current takeaway is that refutation text, case comparison, notional
-machines, and language levels are helpful.
-
-I have created a tutor (\Cref{s:system-tutor}) that incorporates all
-these ideas (\Cref{s:tutor-uiux}) and have collected data from all
-populations. I plan to explore more literature and to improve the
-tutor if I find new helpful techniques.
+machines, and language levels are helpful. I have created a tutor
+(\Cref{s:system-tutor}) that incorporates all these ideas
+(\Cref{s:tutor-uiux}) and have collected data from all populations. I
+plan to explore more literature and to improve the tutor if I find new
+helpful techniques.
 
@@
 
+Each tutorial might be delivered as multiple sections to avoid forcing
+students to continuesly work for more than 30 minutes.
+
 \begin{table}[t]
@@
 
-Let's reconsider this program from \Cref{s:plan-rq1}
+Let's reconsider this program from \Cref{s:rq1-expected-results}
 \begin{lstlisting}
@@
 making a random choice, the tutor has been giving panelty. Designing
-the strategy of assigning panelty is non-trivial. A common strategy is
-to grade students by their correctness. However, grading by
-correctness is not suitable for a tutoring system because we do not
-mind students doing badly on earlier tasks as long as they eventually
-master the concepts. Actually, if students generally succeed in all
-tasks, the tutoring system is failing -- it does not teach anything to
-the students. A alternative common strategy is to grade by completion.
+the strategy of assigning panelty is non-trivial:
+\begin{itemize}
+  \item A common strategy is to grade students by their correctness.
+  However, grading by correctness is not suitable for a tutoring
+  system because we do not mind students doing badly on earlier tasks
+  as long as they eventually master the concepts. Actually, if
+  students generally succeed in all tasks, the tutoring system is
+  failing -- it does not teach anything to the students.
+  \item A alternative common strategy is to grade by completion.
   However, grading by completion alone encourage students to rush
   through the tutorials and hence might encourage students to make
-random choices. In short, I want to make students feel as follows:
+  random choices.
+\end{itemize}
+In short, I want to make students feel as follows:
 \begin{itemize}
@@
 To encourage this feeling, I have been doing the following: when I
-deployed the tutor in my institution, I graded by
-completion\footnote{I have no control over how other institutions use
+deployed the tutor in my institution, University 1, I graded by
+completion\footnote{I encourage collaborators to grade by
+completeness, but I have no real control over how they actually use
 the tutor.}; when a student fails a task, the tutor gives an
@@
 
-The tutor provides considerable opportunities for making case
-comparison: each interpreting task presents a SMoL program and the
-program's result. These program-result pairs are concrete cases of the
-semantics of SMoL. To incorporate the idea of case comparsion, the
-tutor currently prompts students to compare these cases and then to
-summarize the rules of program behavior at the end of \emph{some}
-tutorials. I plan to expand this to \emph{all} tutorials and use a
-consistent prompt. The prompt should asks about the behavior of new
-language constructs and their interaction with all previously
-introduced constructs. For example, at the end of \textbf{mut-vars}, I
-would want the tutor to ask
+The tutor provides opportunities for making case comparison: each
+interpreting task presents a SMoL program and the program's result.
+These program-result pairs are concrete cases of the semantics of
+SMoL. To incorporate the idea of case comparsion, the tutor currently
+prompts students to compare these cases and then to summarize the
+rules of program behavior at the end of \emph{some} tutorials. I plan
+to expand this to \emph{all} tutorials and use a consistent prompt.
+The prompt should asks about the behavior of new language constructs
+and their interaction with all previously introduced constructs.
+
+The current scope tutorial uses the following prompt
 \begin{quote}
-  What does \lstinline|set!| do? How does it interact with variables,
-  \lstinline|defvar|, and \lstinline|deffun|?
+  When we see a variable reference, how do we find the corresponding
+  declaration, if any?
 \end{quote}
-
-\subsection{Multi-syntax}%
-\label{s:multi-syntax}
-
-[FILL] Use a random default syntax per task? Log syntax switch.
+This prompt has a few problems:
+\begin{itemize}
+  \item It is unclear to adapt it to other tutorials.
+  \item Students tend to give brief response possibly because the
+  questions is asking for a lot and because students are tired when
+  they reach the end of a tutorial.
+\end{itemize}.
+
+So I plan to change the question format into a notice-and-wonder style
+of questions, which is known to be less demanding
+(\Cref{s:pedagogic-techniques}). So I plan to replace the
+\textbf{scope} prompt with
+\begin{quote}
+  \emph{Think about} the following questions:
+  \begin{itemize}
+    \item How do \lstinline|defvar|s work?
+    \item How do \lstinline|deffun|s work?
+    \item How do variable referencess (e.g., \lstinline|x| in \lstinline|(+ x 1)|) work?
+    \item How do \lstinline|defvar|s, \lstinline|deffun|s, and variable references interact with each other?
+  \end{itemize}
+  If you believe you know the answers perfectly, please \emph{response with}
+  \begin{quote}
+    I know the answers with 100\% certainty.
+  \end{quote}
+  Otherwise, please \emph{response with} what you are uncertain about these language constructs.
+\end{quote}
+I plan to use a similar prompt in later tutorials. For example, in \textbf{mut-vars},
+I would want the tutor to ask
+\begin{quote}
+  \emph{Think about} the following questions:
+  \begin{itemize}
+    \item How do \lstinline|set!|s work?
+    \item How do \lstinline|set!|s interact with other language constructs (namely, \lstinline|defvar|s, \lstinline|deffun|s, and variable references)?
+  \end{itemize}
+  If you believe you know the answers perfectly, please \emph{response with}
+  \begin{quote}
+    I know the answers with 100\% certainty.
+  \end{quote}
+  Otherwise, please \emph{response with} what you are uncertain about these language constructs.
+\end{quote}
+I am hoping these new prompt are
 
@@
 
-Prior research argues that illustrating how programs runs with
-notional machines (NMs) is benefitial for students [CITE]. I have created a
-web-based notional machine
-\href{https://lukuangchen.github.io/stacker-2023}{https://lukuangchen.github.io/stacker-2023}.
-
-The notional machine can be configurated with URL parameters. For example,
-\href{https://lukuangchen.github.io/stacker-2023/?program=\%28defvar+x+12\%29\%0A\%28defvar+y+x\%29\%0A\%28set\%21+y+0\%29\%0Ax\%0Ay}{this link}
-opens the NM with a program embeded.
+Prior research argues (\Cref{s:pedagogic-techniques})that illustrating
+how programs runs with notional machines (NMs) is benefitial for
+students. I have created a web-based notional machine
+\begin{center}
+  \href{https://lukuangchen.github.io/stacker-2023}{https://lukuangchen.github.io/stacker-2023}
+\end{center}
+
+The notional machine can be configurated with URL parameters. For
+example,
+\href{https://lukuangchen.github.io/stacker-2023/?program=\%28defvar+x+12\%29\%0A\%28defvar+y+x\%29\%0A\%28set\%21+y+0\%29\%0Ax\%0Ay}{this
+link} opens the NM with a program embeded.
 
@@
 
-To fully answer my research questions, I should also try to prove that
-my tutor is effective. In this secton, I describe how I want to refine
-the design of the tutor to help me make the prove, what data I want to
-collect, and what (statistical) analysis do I want to perform on the
-data.
-
-I would like to evaluate the effectiveness from two perspectives:
-\begin{itemize}
-  \item How effective is the tutor at stopping students from holding
-        misconceptions?
-\end{itemize}
+To fully answer my research questions, I should also argue that my
+tutor is effective at correcting misconceptions. In this secton, I
+describe how I want to refine the design of the tutor to help me make
+the argument, what data I want to collect, and what (statistical)
+analysis do I want to perform on the data.
 
@@
   \item[PositionSpecific] Students are more likely to choose the
-  representing wrong answer in a task if the task appears earlier (or
-  later) in the tutorial.
+  representing wrong answer in task that appears earlier (or later) in
+  the tutorial.
 \end{description}
@@
 that some programs are more confusing and that we need further
-investigation into the relationship between the tasks and misconceptions.
-
-If the \textbf{PositionSpecific} model best explains \emph{and} students
-perform better in later tasks, I deduce that the tutor seeems to be
-correcting misconceptions. If studnets perform worse in later taks, the
-tutor is potentially doing harm.
+investigation into the relationship between the tasks and
+misconceptions. If the \textbf{PositionSpecific} model best explains
+\emph{and} students perform better in later tasks, I deduce that the
+tutor seeems to be correcting misconceptions. If studnets perform
+worse in later taks, the tutor is potentially doing harm.
 
@@
 \begin{enumerate}
-  \item I don't want the length of any tutorial to exceed 30 minutes.
-  Because the tutor is now asking for student explanation on choice,
-  students are likely to spend more time on each task then before.
+  \item I don't want the length of any tutorial section to exceed 30
+  minutes.
   \item Programs shouldn't be too long. Some misconceptions are really
@@
 
-\section{What Pedagogic Techniques are Effective?}%
-\label{s:pedagogic-techniques}
-
-The SMoL Tutor is firmly grounded in one technique from
-cognitive and educational psychology. The fundamental problem is: how
-do you fix a misconception? One approach is to only ``present the
-right answer'', for fear that discussing the wrong conception might
-actually reinforce it. Instead, there is a body of literature starting
-from \cite{posnerTheoryConceptualChange1982} that presents a theory of conceptual
-change, at whose heart is the \emph{refutation text}. A refutation
-text tackles the misconception directly, discussing and providing a
-refutation for the incorrect idea. Several studies \citep{schroederRefutationTextFacilitates2022} have shown
-their effectiveness in multiple other domains.
-
-The SMoL Tutor's content structure is also influenced by work on \emph{case comparisons}
-(which draws analogies between examples).
-\cite{alfieriLearningCaseComparisons2013a} suggests that asking
-(rather than not asking) students to find similarities between cases,
-and providing principles \emph{after} the comparisons (rather than
-before or not at all), are associated with better learning
-outcomes.
-
-  [FILL Notional Machine]
-
-  [FILL Language Levels] Prior research argues for the benefits of \emph{language levels}
-[CITE]: a langauge has better be taught as a sequence of languages,
-where each later language includes more langauge constructs than the
-previous one.
-
 \section{Generating and Collating Problems}
 \label{s:gen-prob-quizius}
-
-In \Cref{s:background-misconceptions}, we discuss several papers that have provided
-reports of student misconceptions with different fragments of
-SMoL. However, it is difficult to know how comprehensive these
-are. While some are unclear on the origin of their programs, they
-generally seem to be expert-generated.
-
-The problem with expert-generated lists is that they can be quite
-incomplete. Education researchers have documented the
-phenomenon of the \emph{expert blind spot} \citep{nathanExpertBlindSpot2001}: experts
-simply do not conceive of many learner difficulties.
-Thus, we need
-methods to identify problems beyond what experts conceive.
 
@@
 
+\section{What Pedagogic Techniques are Effective?}%
+\label{s:pedagogic-techniques}
+
+The SMoL Tutor is firmly grounded in one technique from
+cognitive and educational psychology. The fundamental problem is: how
+do you fix a misconception? One approach is to only ``present the
+right answer'', for fear that discussing the wrong conception might
+actually reinforce it. Instead, there is a body of literature starting
+from \cite{posnerTheoryConceptualChange1982} that presents a theory of conceptual
+change, at whose heart is the \emph{refutation text}. A refutation
+text tackles the misconception directly, discussing and providing a
+refutation for the incorrect idea. Several studies \citep{schroederRefutationTextFacilitates2022} have shown
+their effectiveness in multiple other domains.
+
+The SMoL Tutor's content structure is also influenced by work on \emph{case comparisons}
+(which draws analogies between examples).
+\cite{alfieriLearningCaseComparisons2013a} suggests that asking
+(rather than not asking) students to find similarities between cases,
+and providing principles \emph{after} the comparisons (rather than
+before or not at all), are associated with better learning
+outcomes.
+
+  [FILL Notional Machine]
+
+  [FILL Language Levels] Prior research argues for the benefits of \emph{language levels}
+[CITE]: a langauge has better be taught as a sequence of languages,
+where each later language includes more langauge constructs than the
+previous one.
+
 \subsection{The User Experience}

### 2023-11-07 proposal.tex
sec 7.1, 7.2

@@
 
+Known misconceptions are often detected with a kind of instruments
+called \emph{concept inventories}
+\citep{hestenesForceConceptInventory1992,taylorComputerScienceConcept2014}.
+In terms of mechanics, a concept inventory is just an instrument
+consisting of multiple-choice questions (MCQs), where each question
+has one correct answer and several wrong ones. However, the wrong ones
+are chosen with great care. Each one has been validated so that if a
+student picks it, we can quite unambiguously determine \emph{what
+misconception the student has}. For instance, if the question is
+``What is \lstinline|sqrt(4)|?'', then \lstinline|37| is probably an
+uninteresting wrong answer, but if people appear to confuse
+square-roots with squares, then \lstinline|16| would be present as an
+answer.\footnote{Concept inventories are thus useful in many settings.
+For instance, an educator can use them with clickers to get quick
+feedback from a class. If several students pick a specific wrong
+answer, the educator not only knows they are wrong, but also has a
+strong inkling of \emph{precisely what} misconception that group has
+and can address it directly.}
+
 \section{Background: Tutoring Systems}%
@@
 
-\section{Generating and Collating Problems}
+\section{Generating and Collating Problems}%
 \label{s:gen-prob-quizius}
 
-Additionally, in this paper, we intentionally use the word
-mis\emph{conceptions} rather than mis\emph{take}. A mistake can happen
-for any reason (e.g., selecting the wrong answer from a menu). A
-misconception, however, implies a conceptual problem: the student has
-formed an incorrect concept in their head. For instance, they may
-think that mutable structures are copied on function calls, or that
-scope is resolved dynamically. This requires probing what they are
-thinking.
-
-Finally, we are inspired by the significant body of education research
-on \emph{concept inventories} \citep{hestenesForceConceptInventory1992} (with a growing
-number for computer science, as a survey lists \citep{taylorComputerScienceConcept2014}). In
-terms of mechanics, a concept inventory is just an instrument
-consisting of multiple-choice questions (MCQs), where each question has
-one correct answer and several wrong ones. However, the wrong ones are
-chosen with great care. Each one has been validated so that if a
-student picks it, we can quite unambiguously determine \emph{what
-  misconception the student has}. For instance, if the question is
-``What is \lstinline|sqrt(4)|?'', then \lstinline|37| is probably an
-uninteresting wrong answer, but if people appear to confuse
-square-roots with squares, then \lstinline|16| would be present as an
-answer.\footnote{Concept inventories are thus useful in many
-  settings. For instance, an educator can use them with clickers to
-  get quick feedback from a class. If several students pick a
-  specific wrong answer, the educator not only knows they are wrong,
-  but also has a strong inkling of \emph{precisely what} misconception
-  that group has and can address it directly. I expect our
-  instruments to be useful in the same way.}
-
-All these, however, add up to a somewhat challenging demand. I want
-to produce a list of questions (each one an MCQ) such that
-\begin{enumerate}
-
-  \item we can get past the expert blind spot,
-
-  \item we have a sense of what misconceptions students have, and
-
-  \item we can generally associate wrong answers with specific
-        misconceptions, approaching a concept inventory.
-
-\end{enumerate}
+% Additionally, in this paper, we intentionally use the word
+% mis\emph{conceptions} rather than mis\emph{take}. A mistake can happen
+% for any reason (e.g., selecting the wrong answer from a menu). A
+% misconception, however, implies a conceptual problem: the student has
+% formed an incorrect concept in their head. For instance, they may
+% think that mutable structures are copied on function calls, or that
+% scope is resolved dynamically. This requires probing what they are
+% thinking.
+
+% All these, however, add up to a somewhat challenging demand. I want
+% to produce a list of questions (each one an MCQ) such that
+% \begin{enumerate}
+
+%   \item we can get past the expert blind spot,
+
+%   \item we have a sense of what misconceptions students have, and
+
+%   \item we can generally associate wrong answers with specific
+%         misconceptions, approaching a concept inventory.
+
+% \end{enumerate}
+
+This section describes preliminary results on finding misconceptions
+from student-written programs (\Cref{s:quizius}) and what I have done
+to distill misconception-detecting programs from student inputs
+(\Cref{s:smol-quizzes,s:tutor-refine-program-set}).
 
@@
 
-Our main solution to the expert blind spot is to use the Quizius system
+My main solution to the expert blind spot is to use the Quizius system
 \citep{saarinenHarnessingWisdomClasses2019}. In contrast to the very
-heavyweight process (involving a lot of expert time)
-that is generally
+heavyweight process (involving a lot of expert time) that is generally
 used to create a concept inventory, Quizius uses a lightweight,
@@
 
-In Quizius, experts create a prompt; in our case, we asked students to
+In Quizius, experts create a prompt; in our case, students were to
 create small but ``interesting'' programs using the SMoL
@@
 ignoring some small syntactic differences). Thus, for
-each cluster, we obtain a set of rationales.
+each cluster, I obtain a set of rationales.
 
 After running Quizius in the course (\Cref{s:populations}),
-we took over as experts. Determining which is the right
+I took over as experts. Determining which is the right
 answer is easy. Where expert knowledge is useful is in
@@
 
-Prior work using Quizius \citep{saarinenHarnessingWisdomClasses2019} finds that students do author programs
-that the experts did not imagine. In our case, we seeded Quizius with
-programs from prior papers (\Cref{s:background-misconceptions}), which gives the first
-few students programs to respond to. However, we found
-that Quizius significantly expanded the scope of our problems and
-misconceptions. In our final instrument, most programs were directly
-or indirectly inspired by the output of Quizius.
+Prior work using Quizius \citep{saarinenHarnessingWisdomClasses2019}
+finds that students do author programs that the experts did not
+imagine. In our case, I seeded Quizius with programs from prior papers
+(\Cref{s:background-misconceptions}), which gives the first few
+students programs to respond to. However, I found that Quizius
+significantly expanded the scope of our problems and misconceptions.
+In our final instrument, most programs were directly or indirectly
+inspired by the output of Quizius.
+
+The Quizius instrucment was set up by other people. I analyzed the
+collected data.
 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\linewidth]{images/smol-quizzes-example.png}
-  \caption{Screenshot of a SMoL Quiz question.}
-  \label{f:smol-quizzes-example}
-\end{figure}
+% \begin{figure}[t]
+%   \centering
+%   \includegraphics[width=\linewidth]{images/smol-quizzes-example.png}
+%   \caption{Screenshot of a SMoL Quiz question.}
+%   \label{f:smol-quizzes-example}
+% \end{figure}
 
@@
 
-Having curated the output, we had to confirm that these programs were
+Having curated the output, I had to confirm that these programs were
 still effective! That is, they needed to actually find student errors.
@@
 I therefore turned the programs into a set of quizzes (in the US
-sense: namely, a brief test of knowledge) that we call the SMoL
+sense: namely, a brief test of knowledge) that I call the SMoL
 Quizzes. There were three quizzes, ordered by linguistic
@@
 
-The goal of the SMoL Quizzes was to confirm that the aforementioned
-processes of cleansing and enriching the problems was successful. The
-quizzes were therefore administered in the third year of this project in the
-same course.
-\Cref{f:smol-quizzes-example} shows a
-sample question. Every question got an ``Other'' option. If chosen,
-the quiz gave the user a text box with the caption ``Please
-specify''. The goal here was to record any other answers, which in
-turn might lead to fresh misconceptions.
+Each quiz is a sequence of MCQs. The options are from the Quizius
+data. In addition, there is an ``Other'' option. If chosen, the quiz
+gave the user a text box with the caption ``Please specify''. The goal
+here was to record any other answers, which in turn might lead to
+fresh misconceptions.
+
+The quizzes were administered in the third year in University 1.
 
@@
 some easy, warm-up questions initially, so those were kept at the
-beginning. Similarly, we wanted programs that are syntactically
+beginning. Similarly, I wanted programs that are syntactically
 similar to stay close to each other in the quiz. This is so that, when
@@
 
-Students only received feedback after having completed a whole
-quiz. At the end of each quiz, they received both summative feedback
+Students only received feedback after having completed a whole quiz.
+At the end of each quiz, they received both summative feedback
 \emph{and} a refutation text that explained every program that
 appeared in the quiz. Students were also encouraged to run the
-programs, but we have little reason to believe that they did (and
+programs, but I have little reason to believe that they did (and
 certainly they asked few questions on the class forum about them).
 
-Due to space limitations, we present the entire instrument in \supplRef{SMoL Quizzes/instrument}.
-Here we focus on a few programs where student choices correspond to
-misconceptions identified earlier, thereby also showing
-that the curated programs are
-still effective.
-Syntactically, \lstinline|#t| and \lstinline|#f| are true and false,
-while \lstinline|#(...)| is a vector. I use a * to indicate the correct
-answer (which also matches the implementation's output).
+Due to space limitations, I present the entire instrument in
+\supplRef{SMoL Quizzes/instrument}. Here I focus on a few programs
+where student choices correspond to misconceptions identified earlier,
+thereby also showing that the curated programs are still effective. I
+use a * to indicate the correct answer (which also matches the
+implementation's output).
+
+% Syntactically, \lstinline|#t| and \lstinline|#f| are true and false,
+% while \lstinline|#(...)| is a vector.
 
@@
     student answers. Answers that can be considered correct are marked
-    with \texttt{*}. Wrong answers that we discuss are
+    with \texttt{*}. Wrong answers that I discuss are
     marked with $^\dag$.}
@@
 \end{table}
-\Cref{t:quizzes-data-fun} lists programs in \quizFun{} that we
+\Cref{t:quizzes-data-fun} lists programs in \quizFun{} that I
 consider the most interesting. These data confirm the presence of
@@
     }                                               \\\hline
-    \lstinputlisting{programs/mvec_in_mvec.txt} & {
-        \begin{tabularx}{\linewidth}{r X}
-          *62\%       & \lstinline|#(#(100 2 3 4) #(100 2 3 4))| \\
-          $^\dag$26\% & \lstinline|#(#(1 2 3 4) #(100 2 3 4))|   \\
-          6\%         & \texttt{Error}                      \\
-          4\%         & \lstinline|#(#(1 2 3 4) #(1 2 3 4))|     \\
-          2\%         & Other                               \\
-        \end{tabularx}
-    }                                               \\\hline
-    \lstinputlisting{programs/vset_in_let.txt}  & {
-        \begin{tabularx}{\linewidth}{r X}
-          *59\%       & \lstinline|#(10)|  \\
-          $^\dag$36\% & \lstinline|#(123)| \\
-          5\%         & Other         \\
-        \end{tabularx}
-    }                                               \\\hline
+    % \lstinputlisting{programs/mvec_in_mvec.txt} & {
+    %     \begin{tabularx}{\linewidth}{r X}
+    %       *62\%       & \lstinline|#(#(100 2 3 4) #(100 2 3 4))| \\
+    %       $^\dag$26\% & \lstinline|#(#(1 2 3 4) #(100 2 3 4))|   \\
+    %       6\%         & \texttt{Error}                      \\
+    %       4\%         & \lstinline|#(#(1 2 3 4) #(1 2 3 4))|     \\
+    %       2\%         & Other                               \\
+    %     \end{tabularx}
+    % }                                               \\\hline
+    % \lstinputlisting{programs/vset_in_let.txt}  & {
+    %     \begin{tabularx}{\linewidth}{r X}
+    %       *59\%       & \lstinline|#(10)|  \\
+    %       $^\dag$36\% & \lstinline|#(123)| \\
+    %       5\%         & Other         \\
+    %     \end{tabularx}
+    % }                                               \\\hline
     \lstinputlisting{programs/var_as_arg.txt}   & {
@@
 
-% This quiz includes two groups of questions. Each group contains
-% programs that are really similar to each other. The order of the two
-% groups and the other questions are randomized. Question order within
-% each group is also randomized. The first two programs in the
-% \Cref{t:quizzes-data-hof} belongs to one group, which is about
-% changing a function's behavior by mutating a variable defined outside
-% the function. The other group is about equality between functions. The
-% latter group is not included in the main text because we think it is
-% not very related to the theme of this paper.
-
 \begin{table}[t]
@@
     }                                                     \\\hline
-    \lstinputlisting{programs/fun_and_state_2.txt}    & {
-        \begin{tabularx}{\linewidth}{r X}
-          *88\%       & \lstinline|4|            \\
-          $^\dag$10\% & \lstinline|3|            \\
-          3\%         & Other: \lstinline|Error| \\
-        \end{tabularx}
-    }                                                     \\\hline
+    % \lstinputlisting{programs/fun_and_state_2.txt}    & {
+    %     \begin{tabularx}{\linewidth}{r X}
+    %       *88\%       & \lstinline|4|            \\
+    %       $^\dag$10\% & \lstinline|3|            \\
+    %       3\%         & Other: \lstinline|Error| \\
+    %     \end{tabularx}
+    % }                                                     \\\hline
     \lstinputlisting{programs/counter.txt}            & {
@@
         a lambda can't possibly change the behavior of the lambda.
-        Perhaps interestingly, this
-        misconception seems to depend on how the lambda is constructed
-        (compare the middle two programs).
   \item Student understanding of the
@@
 \end{itemize}
-
-\section{The SMoL Tutor}
-\label{s:system-tutor}
-
-So far, we have focused on identifying problems. As noted earlier
-(\Cref{s:smol-quizzes}), we provided students with refutation texts after the quizzes,
-but it is unclear to what extent students read, understood, or
-internalized these. I also wanted to determine whether other
-populations run into these issues, and it was unclear whether we would
-be able to administer passive quizzes to them.
-
-Furthermore, while the quizzes may be a useful diagnostic, our goal is
-not only to find faults but to improve the understanding of basic
-programming language behavior. I believe (and presumably so does
-anyone else who writes a formal semantics!)\ that an understanding of
-these basic program behaviors is important, and even more so when it
-comes to understanding concurrency, ownership, and other advanced
-features. Even more simply, misunderstanding these clearly impacts
-development and debugging time.
-
-In response, we decided to create a \emph{tutor}: the SMoL Tutor. It is built
-around our quiz instruments and the detected misconceptions. We
-describe the Tutor from a user's perspective in \Cref{s:tutor-uiux}, and
-explain how it was populated from the
-SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
-I then discuss what
-we learned from its data (\Cref{s:misconceptions-tutor}), and finally
-evaluate its effectiveness as a \emph{tutor} (\Cref{s:effect-tutor}).
 
@@
 previous one.
+
+\section{The SMoL Tutor}
+\label{s:system-tutor}
+
+So far, we have focused on identifying problems. As noted earlier
+(\Cref{s:smol-quizzes}), we provided students with refutation texts after the quizzes,
+but it is unclear to what extent students read, understood, or
+internalized these. I also wanted to determine whether other
+populations run into these issues, and it was unclear whether we would
+be able to administer passive quizzes to them.
+
+Furthermore, while the quizzes may be a useful diagnostic, our goal is
+not only to find faults but to improve the understanding of basic
+programming language behavior. I believe (and presumably so does
+anyone else who writes a formal semantics!)\ that an understanding of
+these basic program behaviors is important, and even more so when it
+comes to understanding concurrency, ownership, and other advanced
+features. Even more simply, misunderstanding these clearly impacts
+development and debugging time.
+
+In response, we decided to create a \emph{tutor}: the SMoL Tutor. It is built
+around our quiz instruments and the detected misconceptions. We
+describe the Tutor from a user's perspective in \Cref{s:tutor-uiux}, and
+explain how it was populated from the
+SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
+I then discuss what
+we learned from its data (\Cref{s:misconceptions-tutor}), and finally
+evaluate its effectiveness as a \emph{tutor} (\Cref{s:effect-tutor}).
 

### 2023-11-07 proposal.tex
sec 7

@@
 
-\subsection{Reduce Syntactic Bias}
+\subsection{Reduce Syntactic Bias}%
+\label{s:reduce-syntactic-bias}
 
@@
 \end{table}
-\Cref{t:quizzes-data-hof} lists interesting programs
-after the addition of closures and higher-order functions.
-It extends what we saw with loops in
-\Cref{s:intro}: that students have misconceptions about their interaction with mutable variables:
+\Cref{t:quizzes-data-hof} lists interesting programs after the
+addition of closures and higher-order functions. It extends what we
+saw with loops in \Cref{s:intro}: that students have misconceptions
+about their interaction with mutable variables:
 \begin{itemize}
@@
 
+\subsection{Collating Problems for the Tutor}%
+\label{s:tutor-refine-program-set}
+
+The SMoL Tutor's interpreting questions are the final instrument of
+this paper. I bootstrap it from the SMoL Quizzes, but made the
+following alterations:
+\begin{enumerate}
+
+  % fill gaps in goal sentences
+  \item I felt the existing programs were insufficient to cover the
+        ideas we wanted students to work through. For instance, SMoL
+        Quizzes did not alias vectors using \lstinline|defvar|, so we
+        added this program:
+
+\begin{lstlisting}
+  (defvar x (mvec 100))
+  (defvar y x)
+  (vec-set! x 0 200)
+  y
+\end{lstlisting}
+
+  \item I renamed vector operators to avoid confusion between
+        \lstinline|vset!|, which replaces a vector element, and \lstinline|set!|,
+        which mutates variables. I rename all vector operators from, for example,
+        \lstinline|vset!| to \lstinline|vec-set!|.
+
+  \item I deferred the local binding forms to the end, since we felt
+        they were less essential. I therefore rewrote programs that use
+        it to not depend on it.
+
+        % \item I removed some questions that we deemed ``Lispy'' (e.g.,
+        %   involving the difference between parenthetical and infix syntax).
+
+  \item I removed some programs that relied on underspecified aspects or
+        unimportant aspects of SMoL. For example, some depended on
+        whether operations on pairs could be applied to arbitrary vectors:
+
+\begin{lstlisting}
+  (pair? (mpair 1 2))
+  (pair? (mvec 1 2))
+  (pair? '#(1 2))
+  (pair? '(1 2))
+\end{lstlisting}
+
+        As another example, one hinged on
+        whether or not a function's formal parameter could have the same
+        name as the function itself:
+
+\begin{lstlisting}
+  (deffun (f f) f)
+  (f 5)
+\end{lstlisting}
+
+  \item I resolved ambiguities in some programs, either adding
+        answer choices or even adding other questions to tease
+        out different interpretations.
+
+  \item To reduce the number of concepts, I removed programs that
+        relied upon \emph{im}mutable vectors and lists, because they
+        did not seem to create problems. (For brevity, I leave these
+        out of the presentation of SMoL in \Cref{s:smol}, though they
+        are in the implementation.)
+
+  \item I removed questions related to function equality, which was
+        not a focus of the Tutor.
+
+  \item I removed programs of a ``Lispy'' nature (See
+        \Cref{s:reduce-syntactic-bias}).
+
+\end{enumerate}
+Most of these steps either modify or elide programs. Two cases,
+filling gaps in light of the goal sentences and resolving ambiguities,
+introduce programs. Starting with the 37 programs in the SMoL Quizzes,
+I added 52 more programs to arrive at a total of 89.
+
+In addition to generating this set of programs, I also modified the
+answer options. In addition to retaining the correct and incorrect
+answers from the SMoL Quizzes (and constructing our best guess of
+analogous incorrect answers for the new problems), I added more wrong
+answers to for the reasons given in \Cref{s:interpreting-tasks}. In
+general, I ended up increasing the number of choices substantially:
+only 8 out of the 89 questions have three or four choices; 70
+questions have 5--8 choices; and 11 questions have 9--14 choices.
+
 \section{What Pedagogic Techniques are Effective?}%
@@
 
-The SMoL Tutor is firmly grounded in one technique from
-cognitive and educational psychology. The fundamental problem is: how
-do you fix a misconception? One approach is to only ``present the
-right answer'', for fear that discussing the wrong conception might
-actually reinforce it. Instead, there is a body of literature starting
-from \cite{posnerTheoryConceptualChange1982} that presents a theory of conceptual
-change, at whose heart is the \emph{refutation text}. A refutation
-text tackles the misconception directly, discussing and providing a
-refutation for the incorrect idea. Several studies \citep{schroederRefutationTextFacilitates2022} have shown
-their effectiveness in multiple other domains.
-
-The SMoL Tutor's content structure is also influenced by work on \emph{case comparisons}
-(which draws analogies between examples).
+The SMoL Tutor is firmly grounded in one technique from cognitive and
+educational psychology. The fundamental problem is: how do you fix a
+misconception? One approach is to only ``present the right answer'',
+for fear that discussing the wrong conception might actually reinforce
+it. Instead, there is a body of literature starting from
+\cite{posnerTheoryConceptualChange1982} that presents a theory of
+conceptual change, at whose heart is the \emph{refutation text}. A
+refutation text tackles the misconception directly, discussing and
+providing a refutation for the incorrect idea. Several studies
+\citep{schroederRefutationTextFacilitates2022} have shown their
+effectiveness in multiple other domains.
+
+The SMoL Tutor's content structure is also influenced by work on
+\emph{case comparisons} (which draws analogies between examples).
 \cite{alfieriLearningCaseComparisons2013a} suggests that asking
@@
 and providing principles \emph{after} the comparisons (rather than
-before or not at all), are associated with better learning
-outcomes.
+before or not at all), are associated with better learning outcomes.
 
@@
 
-  [FILL Language Levels] Prior research argues for the benefits of \emph{language levels}
-[CITE]: a langauge has better be taught as a sequence of languages,
-where each later language includes more langauge constructs than the
-previous one.
+  [FILL Language Levels] Prior research argues for the benefits of
+\emph{language levels} [CITE]: a langauge has better be taught as a
+sequence of languages, where each later language includes more
+langauge constructs than the previous one.
 
@@
 So far, we have focused on identifying problems. As noted earlier
-(\Cref{s:smol-quizzes}), we provided students with refutation texts after the quizzes,
-but it is unclear to what extent students read, understood, or
-internalized these. I also wanted to determine whether other
-populations run into these issues, and it was unclear whether we would
-be able to administer passive quizzes to them.
+(\Cref{s:smol-quizzes}), we provided students with refutation texts
+after the quizzes, but it is unclear to what extent students read,
+understood, or internalized these. I also wanted to determine whether
+other populations run into these issues, and it was unclear whether we
+would be able to administer passive quizzes to them.
 
@@
 broken.
-
-\subsection{Collating Problems for the Tutor}%
-\label{s:tutor-refine-program-set}
-
-The SMoL Tutor's interpreting questions are the final instrument of
-this paper.
-I bootstrap it from the SMoL
-Quizzes, but made the following alterations:
-\begin{enumerate}
-
-  % fill gaps in goal sentences
-  \item In the process of developing the goal sentences, we felt the
-        existing programs were insufficient to cover the ideas we wanted
-        students to work through. For instance, SMoL Quizzes did not
-        alias vectors using \lstinline|defvar|, so we added this program:
-
-        \begin{lstlisting}
-(defvar x (mvec 100))
-(defvar y x)
-(vec-set! x 0 200)
-y
-\end{lstlisting}
-
-  \item I renamed vector operators to avoid confusion between
-        \lstinline|vset!|, which replaces a vector element, and \lstinline|set!|,
-        which mutates variables. I rename all vector operators from, for example,
-        \lstinline|vset!| to \lstinline|vec-set!|.
-
-  \item I deferred the local binding forms to the end, since we felt
-        they were less essential. I therefore rewrote programs that use
-        it to not depend on it.
-
-        % \item I removed some questions that we deemed ``Lispy'' (e.g.,
-        %   involving the difference between parenthetical and infix syntax).
-
-  \item I removed some programs that relied on underspecified aspects or
-        unimportant aspects of SMoL. For example, some depended on
-        whether operations on pairs could be applied to arbitrary vectors:
-
-        \begin{lstlisting}
-  (pair? (mpair 1 2))
-  (pair? (mvec 1 2))
-  (pair? '#(1 2))
-  (pair? '(1 2))
-\end{lstlisting}
-
-        As another example, one hinged on
-        whether or not a function's formal parameter could have the same
-        name as the function itself:
-
-        \begin{lstlisting}
-  (deffun (f f) f)
-  (f 5)
-\end{lstlisting}
-
-  \item I resolved ambiguities in some programs, either adding
-        answer choices or even adding other questions to tease
-        out different interpretations.
-
-  \item To reduce the number of concepts, we removed programs that relied
-        upon \emph{im}mutable vectors and lists, because they did not
-        seem to create problems. (For brevity, we leave these out of
-        the presentation of SMoL in \Cref{s:smol}, though they are in the implementation.)
-
-  \item I removed questions related to function equality, which was
-        not a focus of the Tutor.\footnote{However, we believe there is
-          a good deal of confusion about notions of equality. We
-          intend to
-          add that as a major tutorial component in future versions.}
-
-  \item I removed programs of a ``Lispy'' nature, such as one where
-        the answer depended on whether the reader correctly understood
-        this inequality:
-
-        \begin{lstlisting}
-      (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
-\end{lstlisting}
-
-        This checks whether $\verb|n| < \verb|3|$, but some presumably
-        vocalized it as ``greater than \lstinline|3|, \lstinline|n|?''.
-
-\end{enumerate}
-Most of these steps either modify or elide programs. Two cases,
-filling gaps in light of the goal sentences and resolving ambiguities,
-introduce
-programs. Starting with the 37 programs in the SMoL Quizzes, we added
-52 more programs to arrive at a total of 89.
-
-In addition to generating this set of programs, we also
-modified the answer options. In addition to retaining the correct and
-incorrect answers from the SMoL Quizzes (and constructing our best
-guess of analogous incorrect answers for the new problems), we added
-more wrong answers.
-
-The reason is as follows. The SMoL Quizzes often have very few wrong
-choices: of the 37 tasks, 26 have only three choices (including the
-correct answer and the ``Other'' option\footnote{In a few cases, we
-  removed the \emph{correct} answer from the choices. A student would
-  therefore have to click on ``Other'' and type it in. The idea was to
-  make the ``Other'' option more salient and plausible.}),
-seven have 4, and only four have 5 or 6.
-Thus, in most
-cases, students have a 25\% or even 33\% chance of just guessing the
-right answer or successfully using a process of elimination. By
-increasing the number of options, we hoped to
-greatly reduce the odds of getting the right answer by chance or
-by elimination.
-
-It was important to add wrong answers that are not utterly
-implausible, because those would become easy to eliminate. Therefore,
-we added numeric constants mentioned in the problem, permuted some of
-the values in case of multiple outputs, and so on. In general, we
-ended up increasing the number of choices substantially:
-only 8 out of the 89 questions have three or four choices; 70
-questions have 5--8 choices; and 11 questions have 9--14 choices. I hope
-this reduced both guessing and elimination, and forced students to
-actually think through the program. Of course, these new answers do
-not have a clear associated misconception, so their mistakes are given
-the generic explanation.
 

### 2023-11-07 proposal.tex
pushed through

@@
   the myriad ways in which students can misunderstand programs?
-  Furthermore, once identified, what can we do to correct them?
+  Furthermore, once identified, what can I do to correct them?
 
@@
 when transferring knowledge from old languages to new ones. In
-recognition of this deep commonality, in this paper we choose to call
+recognition of this deep commonality, in this paper I choose to call
 this the \emph{Standard Model of Languages} (SMoL).
@@
 programmers to understand. Consider the following scenario: to create a
-calculator, we have to construct a callback that can be attached to
+calculator, I have to construct a callback that can be attached to
 each button. The following Python program seems to achieve the
@@
 %   \item I implement these misconceptions as interpreters, which
-%         find weaknesses in our manual analysis and hold potential for
+%         find weaknesses in my manual analysis and hold potential for
 %         generative use in the future.
@@
 % \end{description}
-% I don't include \surmisedMiscon{CallerEnv} in our analysis because in our
+% I don't include \surmisedMiscon{CallerEnv} in my analysis because in our
 % data, all wrong answers that can be explained by \surmisedMiscon{CallerEnv}
@@
 
-% There are overlaps between our survey and
-% theirs. For instance, our \surmisedMiscon{CallerEnv} is their No.~47.
+% There are overlaps between my survey and
+% theirs. For instance, my \surmisedMiscon{CallerEnv} is their No.~47.
 % However, because their descriptions are brief, it is difficult to tell
-% whether a misconception in their survey matches our misconceptions.
+% whether a misconception in their survey matches my misconceptions.
 % Because
@@
 % At any rate,
-% we certainly find
+% I certainly find
 % no equivalent of \miscon{FlatEnv}, \miscon{DeepClosure}, and
@@
 are chosen with great care. Each one has been validated so that if a
-student picks it, we can quite unambiguously determine \emph{what
+student picks it, I can quite unambiguously determine \emph{what
 misconception the student has}. For instance, if the question is
@@
 inner loop iterates through UI events within a task. I follow the same
-structure in my tutor (\Cref{s:tutor-uiux}).
+structure in my tutor (\Cref{s:tutor}).
 
@@
 current takeaway is that refutation text, case comparison, notional
-machines, and language levels are helpful. I have created a tutor
-(\Cref{s:system-tutor}) that incorporates all these ideas
-(\Cref{s:tutor-uiux}) and have collected data from all populations. I
-plan to explore more literature and to improve the tutor if I find new
-helpful techniques.
+machines, and language levels are helpful. I have created a tutor that
+incorporates all these ideas (\Cref{s:tutor}) and have collected
+data from all populations. I plan to explore more literature and to
+improve the tutor if I find new helpful techniques.
 
@@
 
-Each tutorial might be delivered as multiple sections to avoid forcing
+Each tutorial might be delivered as multiple modules to avoid forcing
 students to continuesly work for more than 30 minutes.
@@
   However, grading by correctness is not suitable for a tutoring
-  system because we do not mind students doing badly on earlier tasks
+  system because I do not mind students doing badly on earlier tasks
   as long as they eventually master the concepts. Actually, if
@@
 \begin{quote}
-  When we see a variable reference, how do we find the corresponding
+  When I see a variable reference, how do I find the corresponding
   declaration, if any?
@@
 If the \textbf{TaskSpecific} model best explains the data, I deduce
-that some programs are more confusing and that we need further
+that some programs are more confusing and that I need further
 investigation into the relationship between the tasks and
@@
 requires me to carefully check the dependencies between current
-explanations and goal sentences and resolve those dependencies if any.
+explanations and other texts in the tutor and resolve those
+dependencies if any.
 
@@
 
-% Additionally, in this paper, we intentionally use the word
+% Additionally, in this paper, I intentionally use the word
 % mis\emph{conceptions} rather than mis\emph{take}. A mistake can happen
@@
 
-%   \item we can get past the expert blind spot,
-
-%   \item we have a sense of what misconceptions students have, and
-
-%   \item we can generally associate wrong answers with specific
+%   \item I can get past the expert blind spot,
+
+%   \item I have a sense of what misconceptions students have, and
+
+%   \item I can generally associate wrong answers with specific
 %         misconceptions, approaching a concept inventory.
@@
 
-In Quizius, experts create a prompt; in our case, students were to
+In Quizius, experts create a prompt; in my case, students were to
 create small but ``interesting'' programs using the SMoL
@@
 finds that students do author programs that the experts did not
-imagine. In our case, I seeded Quizius with programs from prior papers
+imagine. In my case, I seeded Quizius with programs from prior papers
 (\Cref{s:background-misconceptions}), which gives the first few
 students programs to respond to. However, I found that Quizius
-significantly expanded the scope of our problems and misconceptions.
-In our final instrument, most programs were directly or indirectly
+significantly expanded the scope of my problems and misconceptions.
+In my final instrument, most programs were directly or indirectly
 inspired by the output of Quizius.
@@
   \item I felt the existing programs were insufficient to cover the
-        ideas we wanted students to work through. For instance, SMoL
+        ideas I wanted students to work through. For instance, SMoL
         Quizzes did not alias vectors using \lstinline|defvar|, so we
@@
 
-  \item I deferred the local binding forms to the end, since we felt
+  \item I deferred the local binding forms to the end, since I felt
         they were less essential. I therefore rewrote programs that use
@@
 
-        % \item I removed some questions that we deemed ``Lispy'' (e.g.,
+        % \item I removed some questions that I deemed ``Lispy'' (e.g.,
         %   involving the difference between parenthetical and infix syntax).
@@
 answer options. In addition to retaining the correct and incorrect
-answers from the SMoL Quizzes (and constructing our best guess of
+answers from the SMoL Quizzes (and constructing my best guess of
 analogous incorrect answers for the new problems), I added more wrong
@@
 
-\section{The SMoL Tutor}
-\label{s:system-tutor}
-
-So far, we have focused on identifying problems. As noted earlier
-(\Cref{s:smol-quizzes}), we provided students with refutation texts
-after the quizzes, but it is unclear to what extent students read,
-understood, or internalized these. I also wanted to determine whether
-other populations run into these issues, and it was unclear whether we
-would be able to administer passive quizzes to them.
-
-Furthermore, while the quizzes may be a useful diagnostic, our goal is
-not only to find faults but to improve the understanding of basic
-programming language behavior. I believe (and presumably so does
-anyone else who writes a formal semantics!)\ that an understanding of
-these basic program behaviors is important, and even more so when it
-comes to understanding concurrency, ownership, and other advanced
-features. Even more simply, misunderstanding these clearly impacts
-development and debugging time.
-
-In response, we decided to create a \emph{tutor}: the SMoL Tutor. It is built
-around our quiz instruments and the detected misconceptions. We
-describe the Tutor from a user's perspective in \Cref{s:tutor-uiux}, and
-explain how it was populated from the
-SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
-I then discuss what
-we learned from its data (\Cref{s:misconceptions-tutor}), and finally
-evaluate its effectiveness as a \emph{tutor} (\Cref{s:effect-tutor}).
-
-\subsection{The User Experience}
-\label{s:tutor-uiux}
-
-The Tutor covers five major topics, shown on the left in
+\section{The SMoL Tutor}%
+\label{s:tutor}
+
+% So far, I have focused on identifying problems. As noted earlier
+% (\Cref{s:smol-quizzes}), I provided students with refutation texts
+% after the quizzes, but it is unclear to what extent students read,
+% understood, or internalized these. I also wanted to determine whether
+% other populations run into these issues, and it was unclear whether we
+% would be able to administer passive quizzes to them.
+
+% Furthermore, while the quizzes may be a useful diagnostic, my goal is
+% not only to find faults but to improve the understanding of basic
+% programming language behavior. I believe (and presumably so does
+% anyone else who writes a formal semantics!)\ that an understanding of
+% these basic program behaviors is important, and even more so when it
+% comes to understanding concurrency, ownership, and other advanced
+% features. Even more simply, misunderstanding these clearly impacts
+% development and debugging time.
+
+% In response, I decided to create a \emph{tutor}: the SMoL Tutor. It is built
+% around my quiz instruments and the detected misconceptions. We
+% describe the Tutor from a user's perspective in \Cref{s:tutor}, and
+% explain how it was populated from the
+% SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
+% I then discuss what
+% I learned from its data (\Cref{s:misconceptions-tutor}), and finally
+% evaluate its effectiveness as a \emph{tutor} (\Cref{s:effect-tutor}).
+
+% \subsection{The User Experience}
+
+This section describe the current status of the tutor.
+
+The Tutor includes five tutorials, shown on the left in
 \Cref{t:tutor-goal-sentences}. The larger topics are further broken
 down into 2--3 modules. The goal was for students to spend at most
-20--30 minutes per module. Our data show that in practice, students
+20--30 minutes per module. my data show that in practice, students
 spent about 9.8 (median) minutes.
 
-Each tutorial consists of a sequence of
-questions. Most questions in the Tutor are \newterm{interpreting
-  questions}.\footnote{The Tutor also asked students to perform some
-  other activities, like showing programs and asking for their heap
-  content, which we do not cover in this paper.}
-These questions (illustrated in
-\Cref{f:smol-tutor-example}) are versions of the SMoL
-Quizzes (obtained by the process described in \Cref{s:tutor-refine-program-set}).
-In each of these questions, students are shown a program
-and asked to predict the program's running result(s) by answering an
-MCQ (with an ``Other'' option). After making a choice, students receive
-feedback. If a student chooses incorrectly, they are (a) given an
-explanation that is \emph{based on the misconception associated with
-  that wrong answer} (or a generic one, if there is not a specific
-misconception),
-and then (b) asked to answer a
-second question:
-\begin{enumerate}
-
-  \item The second question is semantically the same as the first, but
-        with superficial changes (e.g., variable names, constants, and operators are
-        changed) so that the student cannot immediately
-        guess the answer.
-
-  \item Instead of multiple-choice, students must \emph{type} the answer
-        into a text box. (The Tutor normalizes text to accommodate variations.)
-        This is
-        intentional. First, we wanted to force reflection on the
-        explanation just given, whereas with an
-        MCQ, students could just guess. Second, we felt that students would
-        find typing more onerous than clicking. In case students
-        had just guessed on a question, our hope is that the penalty of having to
-        type means, on \emph{subsequent} tasks, they
-        would be more likely to pause and reflect before choosing.
-
-\end{enumerate}
+\Cref{f:smol-tutor-example} illustrates an interpreting task as
+descrbed in \Cref{s:interpreting-tasks}. The Tutor also asked students
+to perform some other activities, like showing programs and asking for
+their heap content, which I do not cover in this paper.
+
+% Each tutorial consists of a sequence of questions. Most questions in
+% the Tutor are \newterm{interpreting questions}.\footnote{The Tutor
+% also asked students to perform some other activities, like showing
+% programs and asking for their heap content, which I do not cover in
+% this paper.} These questions (illustrated in
+% \Cref{f:smol-tutor-example}) are versions of the SMoL Quizzes
+% (obtained by the process described in
+% \Cref{s:tutor-refine-program-set}). In each of these questions,
+% students are shown a program and asked to predict the program's
+% running result(s) by answering an MCQ (with an ``Other'' option).
+% After making a choice, students receive feedback. If a student chooses
+% incorrectly, they are (a) given an explanation that is \emph{based on
+% the misconception associated with that wrong answer} (or a generic
+% one, if there is not a specific misconception), and then (b) asked to
+% answer a second question:
+% \begin{enumerate}
+
+%   \item The second question is semantically the same as the first, but
+%         with superficial changes (e.g., variable names, constants, and operators are
+%         changed) so that the student cannot immediately
+%         guess the answer.
+
+%   \item Instead of multiple-choice, students must \emph{type} the answer
+%         into a text box. (The Tutor normalizes text to accommodate variations.)
+%         This is
+%         intentional. First, I wanted to force reflection on the
+%         explanation just given, whereas with an
+%         MCQ, students could just guess. Second, I felt that students would
+%         find typing more onerous than clicking. In case students
+%         had just guessed on a question, my hope is that the penalty of having to
+%         type means, on \emph{subsequent} tasks, they
+%         would be more likely to pause and reflect before choosing.
+
+% \end{enumerate}
 
@@
 \newterm{goal sentences}. \Cref{t:tutor-goal-sentences} lists
-the (abbreviated) goal sentences for each tutorial.
-Readers can find the full Tutor in \supplRef{SMoL Tutor/instrument}.
+the (abbreviated) goal sentences for each tutorial. I plan to keep
+goal sentences in the tutor.
 
@@
 Some later tutorials include questions about earlier tutorials so
-that we can check whether students remember the concepts across
+that I can check whether students remember the concepts across
 modules. In particular, the \textbf{mut-vars} tutorial starts with questions
 about \textbf{scope}, and \textbf{lambda} starts with questions about
-\textbf{mut-vars} and \textbf{vectors}.
-
-Of the modules, \textbf{local} is the least portable across
-languages. While local binding is of course present in other
-languages, this is mostly covered using nested \lstinline|defvar|s in
-earlier modules, starting with \textbf{scope}. The distinction central
-to this module (between three local-binding constructs with slightly
-different scopes) is primarily a focus of Lispy languages. Therefore,
-we exclude this module from our analysis and instruments, and indeed
-plan to deprecate the module in future versions.
-
-\Cref{s:smol-quizzes} discusses the partial randomization employed by
-the SMoL Quizzes. In contrast, the SMoL Tutor does \emph{not}
-randomize question order. This is because the Tutor consists of more
-than just questions: it also has explanatory text. This text is based
-on the preceding problems. Authoring it is therefore somewhat like
-writing a textbook, with complex dependencies that can easily be
-broken.
+\textbf{mut-vars} and \textbf{vectors}. I plan to keep this design to
+check retention.
+
+% Of the modules, \textbf{local} is the least portable across
+% languages. While local binding is of course present in other
+% languages, this is mostly covered using nested \lstinline|defvar|s in
+% earlier modules, starting with \textbf{scope}. The distinction central
+% to this module (between three local-binding constructs with slightly
+% different scopes) is primarily a focus of Lispy languages. Therefore,
+% I exclude this module from my analysis and instruments, and indeed
+% plan to deprecate the module in future versions.
+
+% \Cref{s:smol-quizzes} discusses the partial randomization employed by
+% the SMoL Quizzes. In contrast, the SMoL Tutor does \emph{not}
+% randomize question order. This is because the Tutor consists of more
+% than just questions: it also has explanatory text. This text is based
+% on the preceding problems. Authoring it is therefore somewhat like
+% writing a textbook, with complex dependencies that can easily be
+% broken.
 
@@
 
-I now examine what we learned from the interpreting
+I now examine what I learned from the interpreting
 questions in terms of program behavior
-misconceptions. But first, we explain how we associate incorrect
+misconceptions. But first, I explain how I associate incorrect
 answers with misconceptions.
 
-I do not present the results of \textbf{begin} and
-\textbf{local} (\Cref{t:tutor-goal-sentences}), because they focus on constructs not usually found in
-non-Lispy languages. They were included in the Tutor to help students
-write programs for the course, but are not interesting from a SMoL
-perspective.
+I do not present the results of and \textbf{local}
+(\Cref{t:tutor-goal-sentences}), because they focus on constructs not
+usually found in non-Lispy languages. They were included in the Tutor
+to help students write programs for the course, but are not
+interesting from a SMoL perspective.
 
@@
 In principle, an expert can identify what misconceptions a particular
-incorrect answer might correspond to. In practice, we found this
+incorrect answer might correspond to. In practice, I found this
 rather difficult for three reasons. First, with 89 programs, each with
-several answers, it's easy to make mistakes. Second, our own expert blind
-spot may prevent us from seeing an interpretation that would
-lead to an additional association. Finally, and specific to our case,
-since we had either curated or written all the programs and answers,
-we were likely to miss some associations we had not intended.
-
-To address this problem, we formalized misconceptions as
-interpreters. That is, for each misconception we created a
+several answers, it's easy to make mistakes. Second, my own expert blind
+spot may prevent me from seeing an interpretation that would
+lead to an additional association. Finally, and specific to my case,
+since I had either curated or written all the programs and answers,
+I was likely to miss some associations I had not intended.
+
+To address this problem, I formalized misconceptions as
+interpreters. That is, for each misconception I created a
 corresponding \emph{misinterpreter}. This is an interpreter for the
@@
 
-By running programs through the misinterpreters, we can more uniformly
+By running programs through the misinterpreters, I can more uniformly
 and rigorously identify \emph{all} the misconceptions associated with
-a wrong answer. Furthermore, if we identify a new misconception (or
+a wrong answer. Furthermore, if I identify a new misconception (or
 alter a misinterpreter), it is easy to automatically re-classify all
-the answers. By using misinterpreters, we indeed found new
+the answers. By using misinterpreters, I indeed found new
 interpretations for existing program-answer pairs.
-I provide all the misinterpreters in the artifact.
 
@@
 \Cref{s:misconceptions-quiz}. These are misconceptions for which we
-have reasonable validation (due to the prose in Quizius), so we call
+have reasonable validation (due to the prose in Quizius), so I call
 them \emph{grounded} misconceptions. I then looked at wrong answers
-not covered by these but chosen by students, and did our best to
+not covered by these but chosen by students, and did my best to
 distill these into misconceptions. These are \emph{surmised}
@@
   the time spent in each module reasonable, the SMoL Tutor does not
-  ask students for rationales for their programs. However, we could
+  ask students for rationales for their programs. However, I could
   easily modify it to do so a small number of times per user, to make
@@
 (b) difficult for
-us to attribute to a misconception, or (c) appeared to us to be
+us to attribute to a misconception, or (c) appeared to me to be
 ``Lispy'' and hence not of broad interest.
@@
 list the final catalog.
-For each, we also
+For each, I also
 present a Tutor question
@@
 misconception.\footnote{An astute reader may well imagine another
-  misinterpretation; but we presumably did not find evidence of it in our
+  misinterpretation; but I presumably did not find evidence of it in our
   data. I do welcome other misinterpretations, for which we
@@
 
-\paragraph{Confirmed Misconceptions}
-
-The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.
-
-\paragraph{Added Misconceptions}
-
-The misinterpreters helped us find misinterpretations that we had
-overlooked. For instance, consider this program from \quizState:
-\begin{lstlisting}
-(defvar x 12)
-(deffun (f x)
-  (set! x 0))
-(f x)
-x
-\end{lstlisting}
-I had assumed the wrong answer
-\texttt{0} is caused by \miscon{CallByRef}.
-However, our misinterpreters made us realize it can also be explained
-by \miscon{FlatEnv}. This could also explain why we see a difference in
-error rate when the formal and actual parameter have the same
-name (as mentioned in \Cref{s:misconceptions-quiz:state}).
-
-% Tutor's refined program sets and misinterpreters also helped us to
+% \paragraph{Confirmed Misconceptions}
+
+% The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.
+
+% \paragraph{Added Misconceptions}
+
+% The misinterpreters helped me find misinterpretations that I had
+% overlooked. For instance, consider this program from \quizState:
+% \begin{lstlisting}
+% (defvar x 12)
+% (deffun (f x)
+%   (set! x 0))
+% (f x)
+% x
+% \end{lstlisting}
+% I had assumed the wrong answer
+% \texttt{0} is caused by \miscon{CallByRef}.
+% However, my misinterpreters made me realize it can also be explained
+% by \miscon{FlatEnv}. This could also explain why I see a difference in
+% error rate when the formal and actual parameter have the same
+% name (as mentioned in \Cref{s:misconceptions-quiz:state}).
+
+% Tutor's refined program sets and misinterpreters also helped me to
 % identify programs that detect one of the misconceptions but not the
@@
 56\% of students asserted that it produces
-\texttt{1 3}. Based on this, we surmise that students might have
-another misconception, which we define as \surmisedMiscon{Lazy}. (Recall that
+\texttt{1 3}. Based on this, I surmise that students might have
+another misconception, which I define as \surmisedMiscon{Lazy}. (Recall that
 SMoL is eager, but even in many lazy languages, this would be an error.)
@@
 \lstinline|get-x| and the inner \lstinline|x| are in the same scope block. The
-answer \lstinline|1| cannot be explained by any of our existing
-misconceptions. Based on this, we surmise a new misconception,
+answer \lstinline|1| cannot be explained by any of my existing
+misconceptions. Based on this, I surmise a new misconception,
 \surmisedMiscon{NestedDef}. (Though its frequency falls below our
-threshold, our reading of answers suggests this may be more
-widespread, and we feel it needs to be investigated more.)
-
-Once we turned this into a misinterpreter, we found that it
+threshold, my reading of answers suggests this may be more
+widespread, and I feel it needs to be investigated more.)
+
+Once I turned this into a misinterpreter, I found that it
 unexpectedly captured the program-answer pair for the following
@@
 \end{lstlisting}
-Previously, we had interpreted this only as \miscon{DefOrSet}, because
+Previously, I had interpreted this only as \miscon{DefOrSet}, because
 students had stated that the second \lstinline|(defvar x ....)|
@@
 did not recognize that the original misconception is underspecified
-until we uncovered the new (surmised) misconception.
-
-\paragraph{Summary}
-
-To summarize, we used the SMoL Tutor, with an expanded set of
-programs, to investigate student misconceptions. I implemented the
-idea of misinterpreters to help us properly classify student
-performance. I were able to reconfirm all the previously identified
-misconceptions, refine some of them, and also identify potential new
-ones (that need further investigation).
+until I uncovered the new (surmised) misconception.
+
+\paragraph{Confirmed Misconceptions}
+
+The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.
+
+% \paragraph{Summary}
+
+% To summarize, I used the SMoL Tutor, with an expanded set of
+% programs, to investigate student misconceptions. I implemented the
+% idea of misinterpreters to help me properly classify student
+% performance. I were able to reconfirm all the previously identified
+% misconceptions, refine some of them, and also identify potential new
+% ones (that need further investigation).
 
@@
 Recall that the SMoL Tutor is not only a collection of MCQs: it is
-also a \emph{tutor}! So far, we have investigated the value of the
-MCQs. Now we examine its tutoring aspect. Concretely, we ask:
-\begin{description}
-
-  \item[RQ] How effective is the Tutor at correcting each misconception?
-
-\end{description}
-To study this, we perform the following analysis per
-misconception. Using misinterpreters, we identify those questions
+also a \emph{tutor}! So far, I have investigated the value of the
+MCQs. Now I examine its tutoring aspect. Concretely, I ask:
+\begin{center}
+
+  How effective is the Tutor at correcting each misconception?
+
+\end{center}
+To study this, I perform the following analysis per
+misconception. Using misinterpreters, I identify those questions
 whose wrong answers fit only one misconception (i.e., only one output
@@
 misinterpreter). I then examine student performance over time across
-those questions. This would let us examine how they do on just that
+those questions. This would let me examine how they do on just that
 topic, in isolation, over time.
@@
 
-Indeed, that is what we see in most of the graphs. The exceptions are
+Indeed, that is what I see in most of the graphs. The exceptions are
 \miscon{CallsCopyStructs}, \surmisedMiscon{NestedDef}, and
@@
 reassess all the problems manually. Once evaluated using
-misinterpreters, we found far fewer problems than we would have liked.
+misinterpreters, I found far fewer problems than I would have liked.
 Concretely, only 40 of the 71 eligible problems (after removing the
@@
 analysis. It is quite possible that some of the other problems would
-have found issues. Rather, what we really see is value in the
+have found issues. Rather, what I really see is value in the
 misinterpreter concept. It is not only useful for analysis, it is also
 valuable for \emph{problem design}: in the next iteration of the
-Tutor, we will use misinterpreters actively to shape the incorrect answers and,
+Tutor, I will use misinterpreters actively to shape the incorrect answers and,
 as necessary, update the programs as well. I therefore hope to have
@@
 There is, of course, a significant danger that the data above have
-been overfitted to only one institution (\Cref{s:populations},
-henceforth University 1),
-thereby actually reflecting the state of its curriculum rather than
-some greater truth about program behavior understanding.  I already
-have some reason to believe this is not the case: the related work
-discussed in \Cref{s:background-misconceptions} is drawn from many institutions in
-multiple countries with different educational preparations, levels,
-and demographics. Nevertheless, that gives us only
-limited information about the specific questions and misconceptions
-described above.
-
-Fortunately, we were able to deploy the Tutor on two other
-populations:
-\begin{itemize}
-
-  \item University 2 is a primarily public university in the US. It is
-        one of the largest Hispanic-serving institutions in the country. As
-        such, its demographic is extremely different from those whose data
-        were used above. The Tutor was used in one course in Spring 2023,
-        taken by 12 students.
-        The course is a third-year, programming
-        language course. The students are required to have taken two
-        introductory programming courses (C++ focused).
-
-  \item A separate instance of the Tutor was published on the
-        website of a programming languages textbook [ANON]. Over the course
-        of 8 months, 597 people started with the first module and 103
-        users made it to the last one. To protect privacy, we intentionally
-        do not record demographic information, but we conjecture that the
-        population is largely self-learners  (who are known to use the
-        accompanying book),
-        including some professional
-        programmers. It is extremely
-        unlikely to be the students from either university, because they
-        would not get credit for their work on the public instance; they
-        needed to use the university-specific instance. Furthermore, since
-        they were not penalized for wrong answers, it would make little
-        sense to do a ``test run'' on the public instance. Finally, we note
-        that there is no overlap between the dates of submission on the
-        public instance and the semester at University 1.
-
-\end{itemize}
-These two populations are therefore at least somewhat different from
-the original population, and help us assess whether the problems we
-identify are merely an artifact of the first institution.
-
-To evaluate, we computed a Spearman's rank correlation $\rho$, ranking
+been overfitted to University 1, thereby actually reflecting the state
+of its curriculum rather than some greater truth about program
+behavior understanding.
+
+% I already have some reason to believe this is
+% not the case: the related work discussed in
+% \Cref{s:background-misconceptions} is drawn from many institutions in
+% multiple countries with different educational preparations, levels,
+% and demographics. Nevertheless, that gives me only limited information
+% about the specific questions and misconceptions described above.
+
+Fortunately, I were able to deploy the Tutor on two other
+populations: University 2 and Texbook (See \Cref{s:populations} for their details).
+% \begin{itemize}
+
+%   \item University 2 is a primarily public university in the US. It is
+%         one of the largest Hispanic-serving institutions in the country. As
+%         such, its demographic is extremely different from those whose data
+%         were used above. The Tutor was used in one course in Spring 2023,
+%         taken by 12 students.
+%         The course is a third-year, programming
+%         language course. The students are required to have taken two
+%         introductory programming courses (C++ focused).
+
+%   \item A separate instance of the Tutor was published on the
+%         website of a programming languages textbook [ANON]. Over the course
+%         of 8 months, 597 people started with the first module and 103
+%         users made it to the last one. To protect privacy, I intentionally
+%         do not record demographic information, but I conjecture that the
+%         population is largely self-learners  (who are known to use the
+%         accompanying book),
+%         including some professional
+%         programmers. It is extremely
+%         unlikely to be the students from either university, because they
+%         would not get credit for their work on the public instance; they
+%         needed to use the university-specific instance. Furthermore, since
+%         they were not penalized for wrong answers, it would make little
+%         sense to do a ``test run'' on the public instance. Finally, I note
+%         that there is no overlap between the dates of submission on the
+%         public instance and the semester at University 1.
+
+% \end{itemize}
+% These two populations are therefore at least somewhat different from
+% the original population, and help me assess whether the problems we
+% identify are merely an artifact of the first institution.
+
+To evaluate, I computed a Spearman's rank correlation $\rho$, ranking
 questions by what percentage of students got the question
@@
 we obtain a \pValue~=~2.013e-07. Between the original university and
-the online population, we obtain a \pValue~<~2.2e-16. These show that
+the online population, I obtain a \pValue~<~2.2e-16. These show that
 the other two populations performed similarly to the original one.
@@
 
-\section{Threats to Validity}
-\label{s:threats}
-
-\subsection{Construct Validity}
-
-Did we measure the right thing?
-While our goal is to study student understanding of language
-behavior, what we actually measure is performance on MCQs on select
-programs. MCQs as a mechanism introduce various clear biases, though
-we add the ``Other'' options to somewhat alleviate them. The small
-size, syntactic details (such as intentionally meaningless variable
-names), and other aspects of the programs may also impact our ability
-to measure understanding. (This could go both ways: students may have
-no trouble understanding behavior in a small program but may struggle
-to do so in a larger one.)
-
-A particularly notable threat is the use of Lispy
-syntax. I chose it for two reasons: both because it helps clarify
-scope (\Cref{s:smol}) and because the rest of the course used
-Racket. However, students may well perform differently with more
-familiar syntaxes. It seems unlikely their performance would be \emph{too}
-different given the many languages that have produced similar
-misconceptions (\Cref{t:rel-work}). Nevertheless, we intend to use a
-variety of syntaxes to examine this issue further.
-
-Finally, we curated programs by hand, so they may well reflect our own
-biases about misconceptions. It may be possible to mitigate this
-problem by synthesizing MCQ programs using the misinterpreters.
-
-\subsection{Internal Validity}
-
-Is our reasoning valid? I have applied standard techniques and
-measurements for evaluating student responses. However, our
-instruments still lack the validity of a proper concept
-inventory. Their creation requires heavyweight processes (such as
-Delphi methods \citep{goldmanSettingScopeConcept2010}) that require many hours of expert attention as
-well as conversations with learners, and can hence be prohibitive in
-cost. The Quizius method (\Cref{s:quizius}) was created precisely to
-be an inexpensive method that provides a good proxy.
-
-Ultimately, our goal is to provide a reasonable instrument for
-widespread use. While the set of all misconceptions could be
-unbounded, we believe our tasks, especially as embedded in the Tutor,
-provide a good starting point for others. In particular, if students
-select a wrong answer, that is still of \emph{some} use to an
-educator, even if the precise misconception cannot be pinned down with
-the highest accuracy. I therefore believe our instruments, and our
-misinterpreter technique, are of general value.
-
-A failure in our Tutor's logging infrastructure led us to miss some
-responses. These are unlikely to be task-specific because the Tutor
-has a generic framework that should perform the same across
-tasks. Moreover, on average only $0.4\%\ (\textrm{sd} = 0.6\%)$ of values are missing. Therefore, we do not believe
-this had a noticeable impact. (In addition, every wrong answer is
-still wrong! I may just not have exactly the right proportions of
-them.)
-
-Because we don't randomize Tutor question order, it is hard to tell
-whether the general decreasing trend in \cref{f:tutor-trend-plots}
-shows misconceptions being corrected, or shows properties of the
-questions (e.g., maybe later questions in the plots are easier because
-their programs happen to be shorter).
-
-\subsection{External Validity}
-
-How well do our results generalize? Certainly there is reason to
-question whether our results would apply to other
-populations. \Cref{s:generality-tutor} provides preliminary evidence
-that the results are not specific to one institution, and
-\Cref{s:background-misconceptions} suggests these issues are widespread. Nevertheless,
-much more broad testing is needed to confirm our specific instruments.
-
-The other major concern is the tie to Lispy syntax. Learners in other
-settings may do worse or even better with it. Building a Tutor that
-supports multiple, and more traditional, syntaxes should help address
-this issue. I defer this to future work (\cref{s:future-work}).
+% \section{Threats to Validity}
+% \label{s:threats}
+
+% \subsection{Construct Validity}
+
+% Did I measure the right thing?
+% While my goal is to study student understanding of language
+% behavior, what I actually measure is performance on MCQs on select
+% programs. MCQs as a mechanism introduce various clear biases, though
+% I add the ``Other'' options to somewhat alleviate them. The small
+% size, syntactic details (such as intentionally meaningless variable
+% names), and other aspects of the programs may also impact my ability
+% to measure understanding. (This could go both ways: students may have
+% no trouble understanding behavior in a small program but may struggle
+% to do so in a larger one.)
+
+% A particularly notable threat is the use of Lispy
+% syntax. I chose it for two reasons: both because it helps clarify
+% scope (\Cref{s:smol}) and because the rest of the course used
+% Racket. However, students may well perform differently with more
+% familiar syntaxes. It seems unlikely their performance would be \emph{too}
+% different given the many languages that have produced similar
+% misconceptions (\Cref{t:rel-work}). Nevertheless, I intend to use a
+% variety of syntaxes to examine this issue further.
+
+% Finally, I curated programs by hand, so they may well reflect my own
+% biases about misconceptions. It may be possible to mitigate this
+% problem by synthesizing MCQ programs using the misinterpreters.
+
+% \subsection{Internal Validity}
+
+% Is my reasoning valid? I have applied standard techniques and
+% measurements for evaluating student responses. However, our
+% instruments still lack the validity of a proper concept
+% inventory. Their creation requires heavyweight processes (such as
+% Delphi methods \citep{goldmanSettingScopeConcept2010}) that require many hours of expert attention as
+% well as conversations with learners, and can hence be prohibitive in
+% cost. The Quizius method (\Cref{s:quizius}) was created precisely to
+% be an inexpensive method that provides a good proxy.
+
+% Ultimately, my goal is to provide a reasonable instrument for
+% widespread use. While the set of all misconceptions could be
+% unbounded, I believe my tasks, especially as embedded in the Tutor,
+% provide a good starting point for others. In particular, if students
+% select a wrong answer, that is still of \emph{some} use to an
+% educator, even if the precise misconception cannot be pinned down with
+% the highest accuracy. I therefore believe my instruments, and our
+% misinterpreter technique, are of general value.
+
+% A failure in my Tutor's logging infrastructure led me to miss some
+% responses. These are unlikely to be task-specific because the Tutor
+% has a generic framework that should perform the same across
+% tasks. Moreover, on average only $0.4\%\ (\textrm{sd} = 0.6\%)$ of values are missing. Therefore, I do not believe
+% this had a noticeable impact. (In addition, every wrong answer is
+% still wrong! I may just not have exactly the right proportions of
+% them.)
+
+% Because I don't randomize Tutor question order, it is hard to tell
+% whether the general decreasing trend in \cref{f:tutor-trend-plots}
+% shows misconceptions being corrected, or shows properties of the
+% questions (e.g., maybe later questions in the plots are easier because
+% their programs happen to be shorter).
+
+% \subsection{External Validity}
+
+% How well do my results generalize? Certainly there is reason to
+% question whether my results would apply to other
+% populations. \Cref{s:generality-tutor} provides preliminary evidence
+% that the results are not specific to one institution, and
+% \Cref{s:background-misconceptions} suggests these issues are widespread. Nevertheless,
+% much more broad testing is needed to confirm my specific instruments.
+
+% The other major concern is the tie to Lispy syntax. Learners in other
+% settings may do worse or even better with it. Building a Tutor that
+% supports multiple, and more traditional, syntaxes should help address
+% this issue. I defer this to future work (\cref{s:future-work}).
 
@@
 
-\section{References}
-
 \bibliography{bibfile}

### 2023-11-08 proposal.tex
all set

@@
 
-The SMoL Tutor is firmly grounded in one technique from cognitive and
-educational psychology. The fundamental problem is: how do you fix a
-misconception? One approach is to only ``present the right answer'',
-for fear that discussing the wrong conception might actually reinforce
-it. Instead, there is a body of literature starting from
+The fundamental problem about correcting misconceptions is: how do you
+fix a misconception? One approach is to only ``present the right
+answer'', for fear that discussing the wrong conception might actually
+reinforce it. Instead, there is a body of literature starting from
 \cite{posnerTheoryConceptualChange1982} that presents a theory of
@@
 
-The SMoL Tutor's content structure is also influenced by work on
-\emph{case comparisons} (which draws analogies between examples).
+\emph{Case comparisons} draws analogies between examples.
 \cite{alfieriLearningCaseComparisons2013a} suggests that asking
@@
 
-  [FILL Notional Machine]
-
-  [FILL Language Levels] Prior research argues for the benefits of
-\emph{language levels} [CITE]: a langauge has better be taught as a
-sequence of languages, where each later language includes more
-langauge constructs than the previous one.
+Several groups of authors have argued for the benefits of visualizing
+program executions
+\citep{dicksonExperiencesImplementingUtilizing2022,
+guoOnlinePythonTutor2013, karnalimUsePythonTutor2017,
+napsExploringRoleVisualization2002}. These visualization seems to
+increase engadgments and help students understand how programs works.
+These work often mention \emph{notional machines}
+\citep{duboulayBlackBoxGlass1981}, which refers to the underlying
+mental models rather than the visual representations.
+
+Prior research argues for the benefits of \emph{language levels}
+\citep{findlerDrSchemeProgrammingEnvironment2002,crestaniExperienceReportGrowing2010}:
+a langauge has better be taught as a sequence of languages, where each
+later language includes more langauge constructs than the previous
+one.
 

### 2023-11-08 proposal.tex
disable line numbers
grammar check

@@
 
-\usepackage[switch]{lineno}
+% \usepackage[switch]{lineno}
+% \renewcommand\thelinenumber{\textcolor{red}{\arabic{linenumber}}}
 \usepackage{xcolor}
-\renewcommand\thelinenumber{\textcolor{red}{\arabic{linenumber}}}
 \usepackage[utf8]{inputenc} % Use UTF-8 encoding
@@
 %% abbrev. for RQs
-\newcommand{\rqI}{What are the program-behavior misconceptions that apply to SMoL and present (even) in students with prior programming backgrounds?}
-\newcommand{\rqII}{What would be a tutoring system that aims at correcting the misconceptions, draws on cognitive, educational, or psychological concepts, and seems to be effective?}
+\newcommand{\rqI}{What program-behavior misconceptions apply to SMoL and present (even) in students with prior programming backgrounds?}
+\newcommand{\rqII}{What would be a tutoring system that aims to correct the misconceptions, draws on cognitive, educational, or psychological concepts, and seems to be effective?}
 
@@
 \maketitle
-\linenumbers
+% \linenumbers
 
@@
 In the rest of this document, I first define SMoL (\Cref{s:smol}),
-then presents literature review of misconceptions
+then present a literature review of misconceptions
 (\Cref{s:background-misconceptions}) and tutoring systems
-(\Cref{s:background-tutoring-systems}), then outline my research plans
-for the RQs (\Cref{s:plan-rq1,s:plan-rq2}). The remaining sections
-detail my current progress.
+(\Cref{s:background-tutoring-systems}), and then outline my research
+plans for the RQs (\Cref{s:plan-rq1,s:plan-rq2}). The remaining
+sections detail my current progress.
 
@@
 I am contacting professors from various other universities. Several of
-them have expressed interests in using my instruments. So the final
+them have expressed interest in using my instruments. So, the final
 results will likely include more populations.
@@
 I have another source of misconceptions. I have been creating a tutor
-to answer my RQ2, which I explain in details in \Cref{s:plan-rq2}. The
-tutor includes many tasks that asks students to predict the result(s)
+to answer my RQ2, which I explain in detail in \Cref{s:plan-rq2}. The
+tutor includes many tasks that ask students to predict the result(s)
 of running a program. Although these programs originate from
 student-written programs (\Cref{s:find-misconceptions-with-quizius}), they
-have been modified to fit tutoring purpose. Perhaps because of the
+have been modified to fit tutoring purposes. Perhaps because of the
 modification or because of the population variance, I have found
@@
 
-I plan to change the tutor such that it asks students' to explain
+I plan to change the tutor such that it asks students to explain
 their prediction. I plan to eyeball the explanations to find new
-misconceptions. This piece of data also helps me to confirm that
-anticipated wrong answers do represent the expected misconceptions.
+misconceptions. This piece of data also helps me confirm that
+anticipated wrong answers represent the expected misconceptions.
 
@@
 from my instruments, there is still a risk of having biase in student
-errors. I plan to adjust the process descrbed in
-\Cref{s:find-misconceptions-with-tutor} such that it presents programs
+errors. I plan to adjust the process described in
+\Cref{s:find-misconceptions-with-tutor} so that it presents programs
 in multiple languages: the SMoL version, a JavaScript translation, and
@@
 misconception comes with multiple programs. For each of these
-programs, the wrong result that correspond to the misconception should
+programs, the wrong result that corresponds to the misconception should
 be commonly predicted by students from various population.
@@
 \end{lstlisting}
-X\% of them responded \lstinline|12 0|; and when asked to predict
+X\% of them responded \lstinline|12 0|; when asked to predict
 \begin{lstlisting}
@@
 incorporates all these ideas (\Cref{s:tutor}) and have collected
-data from all populations. I plan to explore more literature and to
+data from all populations. I plan to explore more literature and
 improve the tutor if I find new helpful techniques.
@@
 incorporate ideas of refutation text (\Cref{s:refutationt-text}), case
-comparsion (\Cref{s:case-comparison}), and notional machines
+comparison (\Cref{s:case-comparison}), and notional machines
 (\Cref{s:notional-machine}). Finally, I present my plan to evaluate
@@
 There are many misconceptions to correct. To make the tutoring process
-managable, I have design the tutor as a sequence of tutorials.
+manageable, I have designed the tutor as a sequence of tutorials.
 Inspired by prior research on language levels
 (\Cref{s:pedagogic-techniques}), my sequence of tutorials is ordered,
-and each later tutorials cover one or more language constructs
+and each tutorial covers one or more new language constructs
 (\Cref{t:tutorials}).
@@
 Each tutorial might be delivered as multiple modules to avoid forcing
-students to continuesly work for more than 30 minutes.
+students to continuously work for more than 30 minutes.
 
@@
 
-The tutor is mostly consist of \emph{interpreting tasks}:
-multiple-choice questions (MCQs) that asks students to interprete
+The tutor is mostly consisted of \emph{interpreting tasks}:
+multiple-choice questions (MCQs) that asks students to interpret
 program.
@@
 wrong answers that I do not anticipate, the tutor has been providing
-an ``Other'' option that, once choosen, allows students to enter
-arbitrary answer.
+an ``Other'' option that, once chosen, allows students to enter
+arbitrary answers.
 
@@
 In this case, the options should include the correct answer
-\lstinline|12 0| and a wrong answer \lstinline|0 0|, which correspond
-to the \miscon{DefByRef} misconception. If an MCQ only has a small
-number of options like this example, students have a great chance (in
-this case, 50\%) of just guessing the correct answer or successfully
-using a process of elimination. So I have been adding additional
-options that are not obviously wrong. The tutor is adding options with
-an ad hoc rules. Eventually, I plan to generate options by moving
-and/or copying numbers in the current options. In this case, my plan
-would add \lstinline|12 12| and \lstinline|0 12|. In summary, each MCQ
-has the following options:
+\lstinline|12 0| and a wrong answer \lstinline|0 0|, which corresponds
+to the \miscon{DefByRef} misconception. If an MCQ only has few options
+like this example, students have a great chance (in this case, 50\%)
+of just guessing the correct answer or successfully using a process of
+elimination. So I have been adding additional options that are not
+obviously wrong. The tutor is adding options with an ad hoc rule.
+Eventually, I plan to generate options by moving and/or copying
+numbers in the current options. In this case, my plan would add
+\lstinline|12 12| and \lstinline|0 12|. In summary, each MCQ has the
+following options:
 \begin{itemize}
@@
         current options
-  \item the ``Other'' option (which, once choosen, allows students to
-        enter arbitrary answer)
+  \item the ``Other'' option (which, once chosen, allows students to
+        enter arbitrary answers)
 \end{itemize}
 
-\paragraph{Panelty to Wrong Choices} To discourage students from
-making a random choice, the tutor has been giving panelty. Designing
-the strategy of assigning panelty is non-trivial:
+\paragraph{Penalty to Wrong Choices} To discourage students from
+making a random choice, the tutor has been giving penalty. Designing
+the strategy of assigning penalty is non-trivial:
 \begin{itemize}
@@
   failing -- it does not teach anything to the students.
-  \item A alternative common strategy is to grade by completion.
-  However, grading by completion alone encourage students to rush
+  \item An alternative common strategy is to grade by completion.
+  However, grading by completion alone encourages students to rush
   through the tutorials and hence might encourage students to make
@@
 \begin{itemize}
-  \item They won't lose grade as long as they finish the tutorials.
+  \item They won't lose grades as long as they finish the tutorials.
   \item They should avoid giving wrong answers.
@@
 The tutor has been presenting a refutation text when a student chooses
-a wrong answer that correspond to known misconception(s). Let's
+a wrong answer that corresponds to known misconception(s). Let's
 reconsider the program in \Cref{s:interpreting-tasks}. If a student
@@
   \lstinline|x|) rather than to \lstinline|x|. So changing the value
-  of \lstinline|y| does not changes the value of \lstinline|x|.
+  of \lstinline|y| does not change the value of \lstinline|x|.
 \end{quote}
-If a students gave a wrong answer that does not correspond to a
+If a student gave a wrong answer that does not correspond to a
 misconception (either by choosing one of the generated option or by
@@
 
-The tutor provides opportunities for making case comparison: each
+The tutor provides opportunities for making case comparisons: each
 interpreting task presents a SMoL program and the program's result.
 These program-result pairs are concrete cases of the semantics of
-SMoL. To incorporate the idea of case comparsion, the tutor currently
+SMoL. To incorporate the idea of case comparison, the tutor currently
 prompts students to compare these cases and then to summarize the
@@
 to expand this to \emph{all} tutorials and use a consistent prompt.
-The prompt should asks about the behavior of new language constructs
+The prompt should ask about the behavior of new language constructs
 and their interaction with all previously introduced constructs.
 
-The current scope tutorial uses the following prompt
+The current scope tutorial uses the following prompt:
 \begin{quote}
@@
   \item It is unclear to adapt it to other tutorials.
-  \item Students tend to give brief response possibly because the
+  \item Students tend to give brief responses possibly because the
   questions is asking for a lot and because students are tired when
   they reach the end of a tutorial.
-\end{itemize}.
+\end{itemize}
 
@@
 Prior research argues (\Cref{s:pedagogic-techniques})that illustrating
-how programs runs with notional machines (NMs) is benefitial for
+how programs runs with notional machines (NMs) is beneficial for
 students. I have created a web-based notional machine
@@
 
-The notional machine can be configurated with URL parameters. For
+The notional machine can be configured with URL parameters. For
 example,
 \href{https://lukuangchen.github.io/stacker-2023/?program=\%28defvar+x+12\%29\%0A\%28defvar+y+x\%29\%0A\%28set\%21+y+0\%29\%0Ax\%0Ay}{this
-link} opens the NM with a program embeded.
+link} opens the NM with a program embedded.
 
@@
 To fully answer my research questions, I should also argue that my
-tutor is effective at correcting misconceptions. In this secton, I
+tutor is effective at correcting misconceptions. In this section, I
 describe how I want to refine the design of the tutor to help me make
@@
 
-I can infer whether students hold a misconception by analysing their
+I can infer whether students hold a misconception by analyzing their
 responses to interpreting tasks. There are three possible responses to
@@
 \emph{and} students perform better in later tasks, I deduce that the
-tutor seeems to be correcting misconceptions. If studnets perform
-worse in later taks, the tutor is potentially doing harm.
+tutor seems to be correcting misconceptions. If students perform
+worse in later tasks, the tutor is potentially doing harm.
 
@@
   \item Programs shouldn't be too long. Some misconceptions are really
-  similar and as a consequence have only a small number of short
-  programs to tell them apart.
+  similar and as a consequence have few short programs to tell them
+  apart.
 \end{enumerate}
@@
 The current tutor is not in the final shape yet. In particular, many
-of the wrong answers correspond to more than one misconceptions; some
+of the wrong answers correspond to more than one misconception; some
 misconceptions are represented by only one question; the order of
-interpreting tasks are not randomized. I plan to tackles the first two
+interpreting tasks are not randomized. I plan to tackle the first two
 differences by adding and removing programs. Adding randomization
@@
 \Cref{f:smol-tutor-example} illustrates an interpreting task as
-descrbed in \Cref{s:interpreting-tasks}. The Tutor also asked students
+described in \Cref{s:interpreting-tasks}. The Tutor also asked students
 to perform some other activities, like showing programs and asking for
@@
 \begin{enumerate}
-  \item Ask students to explained when they gave a wrong answer in a
-        interepreting question. With student explanations, I check whether
+  \item Ask students to explain when they gave a wrong answer in an
+        interpreting question. With student explanations, I check whether
         choosing the wrong answer indicates holding the corresponding misconception(s).
@@
 During Fall 2024, I plan to collect a new round of data from Brown. I
-expect this datasets and the datasets from Spring 2024 would give a
+expect these datasets and the datasets from Spring 2024 would give a
 more rigorously evaluation of the effectiveness of SMoL Tutor. The

### 2023-11-08 proposal.tex
grammar

@@
   \item Students tend to give brief responses possibly because the
-  questions is asking for a lot and because students are tired when
+  questions are asking for a lot and because students are tired when
   they reach the end of a tutorial.
@@
   \end{quote}
-  Otherwise, please \emph{response with} what you are uncertain about these language constructs.
+  Otherwise, please \emph{respond with} what you are uncertain about these language constructs.
 \end{quote}
@@
   \end{itemize}
-  If you believe you know the answers perfectly, please \emph{response with}
+  If you believe you know the answers perfectly, please \emph{respond with}
   \begin{quote}
@@
   \end{quote}
-  Otherwise, please \emph{response with} what you are uncertain about these language constructs.
+  Otherwise, please \emph{respond with} what you are uncertain about these language constructs.
 \end{quote}
-I am hoping these new prompt are
 
@@
 
-My plan is to change the tutor such that it provides a
+My plan is to change the tutor so that it provides a
 program-specific link after each interpreting task and instructs
@@
 the argument, what data I want to collect, and what (statistical)
-analysis do I want to perform on the data.
+analysis I want to perform on the data.
 
@@
   \item They give the correct answer.
-  \item They give a wrong answer that correspond to some
+  \item They give a wrong answer that corresponds to some
         misconceptions.
-  \item They give a wrong answer that do not correspond to any (known)
+  \item They give a wrong answer that does not correspond to any (known)
         misconceptions.
 \end{enumerate}
-If a student gives a wrong answer corresponds to exactly one
+If a student gives a wrong answer that corresponds to exactly one
 misconception \emph{and} an explanation that matches the
@@
 misconceptions are represented by only one question; the order of
-interpreting tasks are not randomized. I plan to tackle the first two
+interpreting tasks is not randomized. I plan to tackle the first two
 differences by adding and removing programs. Adding randomization
@@
 explanations and other texts in the tutor and resolve those
-dependencies if any.
+dependencies, if any.
 
@@
 
-\emph{Case comparisons} draws analogies between examples.
-\cite{alfieriLearningCaseComparisons2013a} suggests that asking
+\emph{Case comparisons} draw analogies between examples.
+\cite{alfieriLearningCaseComparisons2013a} suggest that asking
 (rather than not asking) students to find similarities between cases,
@@
 guoOnlinePythonTutor2013, karnalimUsePythonTutor2017,
-napsExploringRoleVisualization2002}. These visualization seems to
-increase engadgments and help students understand how programs works.
+napsExploringRoleVisualization2002}. Visualization seems to
+increase engagement and help students understand how programs work.
 These work often mention \emph{notional machines}
@@
 \citep{findlerDrSchemeProgrammingEnvironment2002,crestaniExperienceReportGrowing2010}:
-a langauge has better be taught as a sequence of languages, where each
-later language includes more langauge constructs than the previous
+a language has better be taught as a sequence of languages, where each
+later language includes more language constructs than the previous
 one.
@@
 as necessary, update the programs as well. I therefore hope to have
-much more thorough analyses in future work.
+much more thorough analyzes in future work.
 
@@
 
-Fortunately, I were able to deploy the Tutor on two other
+Fortunately, I was able to deploy the Tutor on two other
 populations: University 2 and Texbook (See \Cref{s:populations} for their details).
@@
   \item Adjust the collection of interpreting questions such that
-        every misconception has at least two questions that aims at
+        every misconception has at least two questions that aim at
         correcting the misconception.
@@
 expect these datasets and the datasets from Spring 2024 would give a
-more rigorously evaluation of the effectiveness of SMoL Tutor. The
+more rigorous evaluation of the effectiveness of SMoL Tutor. The
 improvements to be made in Fall 2023 should resolve many of the
@@
 
-I plan to defense in Spring 2025.
+I plan to defend in Spring 2025.
 

### 2023-11-09 proposal.tex
future plan

@@
 
-I plan to change the tutor such that it asks students to explain
-their prediction. I plan to eyeball the explanations to find new
-misconceptions. This piece of data also helps me confirm that
-anticipated wrong answers represent the expected misconceptions.
-
-\subsection{Reduce Syntactic Bias}%
-\label{s:reduce-syntactic-bias}
-
-The processes described in
-\Cref{s:find-misconceptions-with-quizius,s:find-misconceptions-with-tutor}
-have been conducted with SMoL's Lispy syntax (\Cref{s:smol}). This can
-potentially introduce biases to the found misconceptions. For example,
-current results find that some students think the following program
-produces \lstinline|'(4 5)| rather than \lstinline|'(1 2)|.
-\begin{lstlisting}
-  (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
-\end{lstlisting}
-This checks whether $\verb|3| > \verb|n|$, but those students
-presumably vocalized it as ``greater than \lstinline|3|,
-\lstinline|n|?''. Although I have tried to remove programs like this
-from my instruments, there is still a risk of having biase in student
-errors. I plan to adjust the process described in
-\Cref{s:find-misconceptions-with-tutor} so that it presents programs
-in multiple languages: the SMoL version, a JavaScript translation, and
-a Python version.
+I plan to change the tutor such that it collects data to uncover these
+unknown misconceptions and to confirm known ones.
 
@@
 definitional interpreter for each misconception that I have found. I
-plan to do the same things for any new misconceptions.
+plan to do the same things for any new misconceptions. I call these
+interpreters \emph{misinterpreters}.
 
@@
 To answer this RQ, I have done some literature review on what
-pedagogic techniques are helpful (\Cref{s:pedagogic-techniques}). My
-current takeaway is that refutation text, case comparison, notional
+pedagogic techniques are helpful (\Cref{s:pedagogic-techniques}).
+My current takeaway is that refutation text, case comparison, notional
 machines, and language levels are helpful. I have created a tutor that
-incorporates all these ideas (\Cref{s:tutor}) and have collected
-data from all populations. I plan to explore more literature and
-improve the tutor if I find new helpful techniques.
+incorporates all these ideas (\Cref{s:tutor}) and have collected data
+from all populations.
 
@@
 elimination. So I have been adding additional options that are not
-obviously wrong. The tutor is adding options with an ad hoc rule.
-Eventually, I plan to generate options by moving and/or copying
-numbers in the current options. In this case, my plan would add
-\lstinline|12 12| and \lstinline|0 12|. In summary, each MCQ has the
-following options:
-\begin{itemize}
-  \item the correct answer,
-  \item wrong answers that correspond to known misconceptions,
-  \item options generated by moving and/or copying numbers in the
-        current options
-  \item the ``Other'' option (which, once chosen, allows students to
-        enter arbitrary answers)
-\end{itemize}
-
-\paragraph{Penalty to Wrong Choices} To discourage students from
-making a random choice, the tutor has been giving penalty. Designing
-the strategy of assigning penalty is non-trivial:
+obviously wrong. The tutor is adding many options, but with an ad hoc
+rule. I plan to revise the rules of adding extra options.
+
+\paragraph{Penalty to Wrong Choices}\label{s:penalty} To discourage
+students from making a random choice, the tutor has been giving
+penalty. Designing the strategy of assigning penalty is non-trivial:
 \begin{itemize}
@@
 prompts students to compare these cases and then to summarize the
-rules of program behavior at the end of \emph{some} tutorials. I plan
-to expand this to \emph{all} tutorials and use a consistent prompt.
-The prompt should ask about the behavior of new language constructs
-and their interaction with all previously introduced constructs.
+rules of program behavior at the end of \emph{some} tutorials.
+
+\subsection{Notional Machine}%
+\label{s:notional-machine}
+
+Prior research argues (\Cref{s:pedagogic-techniques})that illustrating
+how programs runs with notional machines (NMs) is beneficial for
+students. I have been giving students opportunities and sometimes
+explicitly require them to run programs in a NM visualization tool.
+
+\subsection{Evaluate the Tutor}%
+\label{s:evaluation}
+
+To fully answer my research questions, I should also argue that my
+tutor is effective at correcting misconceptions. In this section, I
+describe how I want to refine the design of the tutor to help me make
+the argument, what data I want to collect, and what (statistical)
+analysis I want to perform on the data.
+
+I can infer whether students hold a misconception by analyzing their
+responses to interpreting tasks. There are three possible responses to
+an interpreting task:
+\begin{enumerate}
+  \item They give the correct answer.
+  \item They give a wrong answer that corresponds to some
+        misconceptions.
+  \item They give a wrong answer that does not correspond to any (known)
+        misconceptions.
+\end{enumerate}
+If a student gives a wrong answer that corresponds to exactly one
+misconception \emph{and} an explanation that matches the
+misconception, I deduce that the student holds the misconception when
+doing the task.
+
+Given this observation, I plan to set up the tutor such that
+\begin{itemize}
+  \item Every wrong answer corresponds to up to one misconception. In
+        this case, I say the wrong answer \emph{represents} the
+        misconception.
+  \item Every misconception is represented by some questions.
+  \item In each tutorial, a misconception is either not represented by
+        any interpreting tasks or represented by at least two tasks.
+  \item Within each tutorial, the order of interpreting tasks is
+        randomized.
+\end{itemize}
+For each tutorial and each misconception, I plan to do a model
+comparison (with the standard AIC and BIC criteria) on the following
+models/theories:
+\begin{description}
+  \item[TaskSpecific] Some students are more likely to choose the
+  representing wrong answers in certain tasks.
+  \item[PositionSpecific] Students are more likely to choose the
+  representing wrong answer in task that appears earlier (or later) in
+  the tutorial.
+\end{description}
+
+If the \textbf{TaskSpecific} model best explains the data, I deduce
+that some programs are more confusing and that I need further
+investigation into the relationship between the tasks and
+misconceptions. If the \textbf{PositionSpecific} model best explains
+\emph{and} students perform better in later tasks, I deduce that the
+tutor seems to be correcting misconceptions. If students perform
+worse in later tasks, the tutor is potentially doing harm.
+
+To know students well, I would like to have as many tasks per
+misconception as possible. But in practice, the number of tasks is
+subject to the following limitations:
+\begin{enumerate}
+  \item I don't want the length of any tutorial section to exceed 30
+  minutes.
+  \item Programs shouldn't be too long. Some misconceptions are really
+  similar and as a consequence have few short programs to tell them
+  apart.
+\end{enumerate}
+
+The current tutor is not prepared for this analysis yet. I describe how
+I plan to change the tutor in \Cref{s:future-work}.
+
+\section{Generating and Collating Problems}%
+\label{s:gen-prob-quizius}
+
+% Additionally, in this paper, I intentionally use the word
+% mis\emph{conceptions} rather than mis\emph{take}. A mistake can happen
+% for any reason (e.g., selecting the wrong answer from a menu). A
+% misconception, however, implies a conceptual problem: the student has
+% formed an incorrect concept in their head. For instance, they may
+% think that mutable structures are copied on function calls, or that
+% scope is resolved dynamically. This requires probing what they are
+% thinking.
+
+% All these, however, add up to a somewhat challenging demand. I want
+% to produce a list of questions (each one an MCQ) such that
+% \begin{enumerate}
+
+%   \item I can get past the expert blind spot,
+
+%   \item I have a sense of what misconceptions students have, and
+
+%   \item I can generally associate wrong answers with specific
+%         misconceptions, approaching a concept inventory.
+
+% \end{enumerate}
+
+This section describes preliminary results on finding misconceptions
+from student-written programs (\Cref{s:quizius}) and what I have done
+to distill misconception-detecting programs from student inputs
+(\Cref{s:smol-quizzes,s:tutor-refine-program-set}).
+
+\subsection{Generating Problems Using Quizius}
+\label{s:quizius}
+
+My main solution to the expert blind spot is to use the Quizius system
+\citep{saarinenHarnessingWisdomClasses2019}. In contrast to the very
+heavyweight process (involving a lot of expert time) that is generally
+used to create a concept inventory, Quizius uses a lightweight,
+interactive approach to obtain fairly comparable data, which an expert
+can then shape into a quality instrument.
+
+In Quizius, experts create a prompt; in my case, students were to
+create small but ``interesting'' programs using the SMoL
+language. Quizius shows this prompt to students and gathers their
+answers. Each student is then shown a set of programs created by other
+students and asked to predict (without running it) the value produced
+by the program.\footnote{In the course (\Cref{s:populations}),
+  students were given credit for using Quizius but not penalized for wrong
+  answers, reducing their incentive to ``cheat'' by running
+  programs. They were also told that doing so would diminish the value of
+  their answers. Some students seemed to do so anyway, but most
+  honored the directive.} Students are also asked to provide a rationale for why
+they think it will produce that output.
+
+Quizius runs interactively during an assignment period.
+At each point, it needs to determine which
+previously authored program to show a student.
+It can
+either ``exploit'' a given program that already has responses
+or ``explore'' a new
+one. Quizius thus treats this as a multi-armed bandit problem
+\citep{katehakisMultiArmedBanditProblem1987} and uses that to choose a program.
+
+The output from Quizius is (a) a collection of programs; (b) for
+each program, a collection of predicted answers; and (c) for each
+answer, a rationale. Clustering the answers is easy (after
+ignoring some small syntactic differences). Thus, for
+each cluster, I obtain a set of rationales.
+
+After running Quizius in the course (\Cref{s:populations}),
+I took over as experts. Determining which is the right
+answer is easy. Where expert knowledge is useful is in
+\emph{clustering the rationales}. If all the rationales for a wrong
+answer are fairly similar, this is strong evidence that there is a
+common misconception that generates it. If, however, there are
+multiple rationale clusters, that means the program
+is not discriminative enough to distinguish the misconceptions, and it
+needs to be further refined to tell them apart. Interestingly, even
+the correct answer needs to be analyzed, because sometimes correct
+answers do have incorrect rationales (again, suggesting the program
+needs refinement to discriminate correct conceptions
+from misconceptions).
+
+Prior work using Quizius \citep{saarinenHarnessingWisdomClasses2019}
+finds that students do author programs that the experts did not
+imagine. In my case, I seeded Quizius with programs from prior papers
+(\Cref{s:background-misconceptions}), which gives the first few
+students programs to respond to. However, I found that Quizius
+significantly expanded the scope of my problems and misconceptions.
+In my final instrument, most programs were directly or indirectly
+inspired by the output of Quizius.
+
+The Quizius instrucment was set up by other people. I analyzed the
+collected data.
+
+\subsection{Collating Problems}
+\label{s:smol-quizzes}
+
+While Quizius is very useful in principle, it also produced data that
+needed significant curation for the following reasons:
+\begin{itemize}
+
+  \item A problem may have produced diverse outputs simply because it
+        was written in a very confusing way. Such programs do not reveal
+        any useful \emph{behavior} misconceptions, and must therefore be
+        filtered out. For instance:
+
+        \begin{lstlisting}
+  (defvar x 1)
+  (defvar y 2)
+  (defvar z 3)
+  (deffun (sum a ...) (+ a ...))
+  (sum x y z)
+  \end{lstlisting}
+
+        A reader might think that \lstinline|sum| takes variable arguments (so
+        the program produces \lstinline|6|), but in fact \lstinline|...| is a single
+        variable, so this produces an arity error.
+
+  \item Some programs relied on (or stumbled upon) underspecified aspects
+        of unimportant (in particular, non-standard) parts of SMoL, such as floating-point versus
+        rational arithmetic.
+
+  \item A problem may have produced diverse outputs simply because it is
+        hard to parse or to (mentally) trace its execution. One example was a
+        17-line program with 6 similar-looking and -named functions. As
+        another example:
+
+        \begin{lstlisting}
+  (defvar a (or (/ 1 (- 0.25 0.25)) (/ 1 0.0)))
+  (defvar b (and (/ 1 (- 0.25 0.25)) (/ 1 0.0)))
+  (defvar c (and (/ 1 0.0) (/ 1 (- 0.25 -0.25))))
+  (defvar d (or (/ 1 0) (/ 1 (- 0.25 -0.25))))
+  (and (or a c) (or b d))
+\end{lstlisting}
+
+        This program is not only confusing, it \emph{also} tests interpretations of (a) exact
+        versus inexact numbers and (b) truthy/falsiness, leading to
+        significant (but not very useful) answer diversity.
+
+  \item As noted above, a program's wrong (or even correct) answers may
+        correspond to multiple (mis)conceptions. In these cases, the program
+        must be refined to be more discriminative.
+
+\end{itemize}
+and so on.
+I therefore manually curated the Quizius output
+to address these issues.
+
+% \begin{figure}[t]
+%   \centering
+%   \includegraphics[width=\linewidth]{images/smol-quizzes-example.png}
+%   \caption{Screenshot of a SMoL Quiz question.}
+%   \label{f:smol-quizzes-example}
+% \end{figure}
+
+\subsection{The SMoL Quizzes}%
+\label{s:misconceptions-quiz}
+
+Having curated the output, I had to confirm that these programs were
+still effective! That is, they needed to actually find student errors.
+
+I therefore turned the programs into a set of quizzes (in the US
+sense: namely, a brief test of knowledge) that I call the SMoL
+Quizzes.
+% There were three quizzes, ordered by linguistic
+% complexity. The first consisted of only basic operators and
+% first-order functions. The second added variable and structure
+% mutation. The third added \lstinline|lambda| and higher-order functions.
+Each quiz is a sequence of MCQs. The options are from the Quizius
+data. In addition, there is an ``Other'' option. If chosen, the quiz
+gave the user a text box with the caption ``Please specify''. The goal
+here was to record any other answers, which in turn might lead to
+fresh misconceptions.
+
+The quizzes were administered in the third year in University 1.
+
+% Question orders were partially randomized. I wanted students to get
+% some easy, warm-up questions initially, so those were kept at the
+% beginning. Similarly, I wanted programs that are syntactically
+% similar to stay close to each other in the quiz. This is so that, when
+% students got a second such program, they would not have to look far to
+% find the first one and confirm that they are indeed (slightly)
+% different, rather than wonder if they were seeing the same program
+% again.
+
+% Students only received feedback after having completed a whole quiz.
+% At the end of each quiz, they received both summative feedback
+% \emph{and} a refutation text that explained every program that
+% appeared in the quiz. Students were also encouraged to run the
+% programs, but I have little reason to believe that they did (and
+% certainly they asked few questions on the class forum about them).
+
+% Due to space limitations, I present the entire instrument in
+% \supplRef{SMoL Quizzes/instrument}. Here I focus on a few programs
+% where student choices correspond to misconceptions identified earlier,
+% thereby also showing that the curated programs are still effective. I
+% use a * to indicate the correct answer (which also matches the
+% implementation's output).
+
+% Syntactically, \lstinline|#t| and \lstinline|#f| are true and false,
+% while \lstinline|#(...)| is a vector.
+
+% This quiz starts with two warm-up and not very interesting questions
+% and then other questions. The order of the other questions are fully
+% randomized.
+
+\begin{table}[t]
+  \centering
+  \begin{tabularx}{\linewidth}{l X}
+    \textbf{Questions}                       &
+    \textbf{Table of Answers}                    \\\hline
+    \lstinputlisting{programs/def_twice.txt} & {
+        \begin{tabularx}{\linewidth}{r X}
+          $^\dag$59\% & \lstinline|2 0|             \\
+          *32\%       & \lstinline|Error|           \\
+          4\%         & \lstinline|Other: 2 2|      \\
+          3\%         & ``Nothing is printed'' \\
+          2\%         & \lstinline|0 0|             \\
+        \end{tabularx}
+    }                                            \\\hline
+    \lstinputlisting{programs/dyn_scope.txt} & {
+        \begin{tabularx}{\linewidth}{r X}
+          *65\%       & \lstinline|Error| \\
+          $^\dag$29\% & \lstinline|42 #t| \\
+          6\%         & Other        \\
+        \end{tabularx}
+    }                                            \\\hline
+    % \lstinputlisting{programs/reassign.txt}  & {
+    %     \begin{tabularx}{\linewidth}{r X}
+    %       *53\%       & \lstinline|#(6 5) 5|                 \\
+    %       *20\%       & \lstinline|Error|                    \\
+    %       $^\dag$11\% & \lstinline|#(6 6) 5|                 \\
+    %       $^\dag$11\% & \lstinline|#(6 6) 6|                 \\
+    %       5\%         & ``Nothing is printed'' or Other \\
+    %     \end{tabularx}
+    % }                                            \\\hline
+  \end{tabularx}
+  \caption{Some questions and student answers from \quizFun. Each
+    table row is a program and the (relative) frequencies of
+    student answers. Answers that can be considered correct are marked
+    with \texttt{*}.}
+  \label{t:quizzes-data-fun}
+\end{table}
+\Cref{t:quizzes-data-fun} lists programs in \quizFun{} that I
+consider the most interesting. These data confirm the presence of
+scope-related misconceptions.
+% :
+% \begin{enumerate}
+%   \item At least 59\% of students incorrectly believe that a
+%         variable can be defined twice in one block.
+%   \item 29\% of students believe that a variable defined in a function
+%         will be available in the top-level (or global) environment
+%         after the function is called. That is, these students may have a
+%         dynamic scope misconception.
+%   \item 22\% (11\% + 11\%) of students believe variables themselves
+%         can be passed as arguments and redefined inside the function.
+%         They disagree on whether the redefinition persists after the
+%         function call.
+% \end{enumerate}
+
+% This quiz randomizes the order of all questions.
+
+\begin{table}[t]
+  \centering
+  \begin{tabularx}{\linewidth}{l X}
+    \textbf{Question}                           &
+    \textbf{Table of Answers}                       \\\hline
+    \lstinputlisting{programs/circularity.txt}  & {
+        \begin{tabularx}{\linewidth}{r X}
+          $^\dag$50\% & \lstinline|#(#(2 #(2 3)) #(2 3))| \\
+          *29\%       & \lstinline|x=#(x x)| or
+          something similar. Both \lstinline|(left x)| and \lstinline|(right x)|
+          are \lstinline|x| itself.                       \\
+          $^\dag$16\% & \lstinline|Error|                 \\
+          5\%         & Other                        \\
+        \end{tabularx}
+    }                                               \\\hline
+    % \lstinputlisting{programs/mvec_in_mvec.txt} & {
+    %     \begin{tabularx}{\linewidth}{r X}
+    %       *62\%       & \lstinline|#(#(100 2 3 4) #(100 2 3 4))| \\
+    %       $^\dag$26\% & \lstinline|#(#(1 2 3 4) #(100 2 3 4))|   \\
+    %       6\%         & \texttt{Error}                      \\
+    %       4\%         & \lstinline|#(#(1 2 3 4) #(1 2 3 4))|     \\
+    %       2\%         & Other                               \\
+    %     \end{tabularx}
+    % }                                               \\\hline
+    % \lstinputlisting{programs/vset_in_let.txt}  & {
+    %     \begin{tabularx}{\linewidth}{r X}
+    %       *59\%       & \lstinline|#(10)|  \\
+    %       $^\dag$36\% & \lstinline|#(123)| \\
+    %       5\%         & Other         \\
+    %     \end{tabularx}
+    % }                                               \\\hline
+    \lstinputlisting{programs/var_as_arg.txt}   & {
+        \begin{tabularx}{\linewidth}{r X}
+          *65\%       & \texttt{12} \\
+          $^\dag$31\% & \texttt{0}  \\
+          4\%         & Other       \\
+        \end{tabularx}
+    }                                               \\\hline
+    % \lstinputlisting{programs/setset.txt}       & {
+    %     \begin{tabularx}{\linewidth}{r X}
+    %       *59\%       & Other: \texttt{5 7}   \\
+    %       $^\dag$27\% & \texttt{6 7}          \\
+    %       11\%        & \texttt{5 5}          \\
+    %       2\%         & Other: \texttt{Error} \\
+    %       1\%         & Other: \texttt{5 6}   \\
+    %     \end{tabularx}
+    % }                                               \\\hline
+  \end{tabularx}
+  \caption{Some questions and student answers from \quizState. Each
+    table row is a program and the (relative) frequencies of
+    student answers. Answers that can be considered correct are marked
+    with \texttt{*}.}
+  \label{t:quizzes-data-state}
+\end{table}
+\Cref{t:quizzes-data-state} lists interesting programs after the
+addition of state. These data suggest the students have aliasing-related misconceptions.
+% :
+% \begin{itemize}
+%   \item Up to 50\% of students think vectors are copied rather than
+%         aliased.
+%   \item 16\% of students think trying to construct and print a
+%         self-referring vector would error. (This program is ambiguous:
+%         \emph{constructing} works in most SMoL languages, but \emph{printing}
+%         may well cause a problem. These identified ambiguities are
+%         addressed in \Cref{s:tutor-refine-program-set}.)
+%   \item 31\% of students think a variable is aliased by a parameter if
+%         the two variables have the same name. Perhaps interestingly,
+%         fewer students (27\%) think the variable aliasing would
+%         happen if the two variables had different names.
+% \end{itemize}
+
+\begin{table}[t]
+  \centering
+  \begin{tabularx}{\linewidth}{l X}
+    \textbf{Question}                                 &
+    \textbf{Table of Answers}                             \\\hline
+    \lstinputlisting{programs/fun_returns_lambda.txt} & {
+        \begin{tabularx}{\linewidth}{r X}
+          *82\%       & \lstinline|3|     \\
+          $^\dag$17\% & \lstinline|Error| \\
+          1\%         & Other        \\
+        \end{tabularx}
+    }                                                     \\\hline
+    \lstinputlisting{programs/fun_and_state_1.txt}    & {
+        \begin{tabularx}{\linewidth}{r X}
+          *74\%       & \lstinline|4|            \\
+          $^\dag$21\% & \lstinline|3|            \\
+          6\%         & Other: \lstinline|Error| \\
+        \end{tabularx}
+    }                                                     \\\hline
+    % \lstinputlisting{programs/fun_and_state_2.txt}    & {
+    %     \begin{tabularx}{\linewidth}{r X}
+    %       *88\%       & \lstinline|4|            \\
+    %       $^\dag$10\% & \lstinline|3|            \\
+    %       3\%         & Other: \lstinline|Error| \\
+    %     \end{tabularx}
+    % }                                                     \\\hline
+    \lstinputlisting{programs/counter.txt}            & {
+        \begin{tabularx}{\linewidth}{r X}
+          *62\%       & \lstinline|1 1 2 3 2| \\
+          $^\dag$34\% & \lstinline|1 1 1 1 1| \\
+          4\%         & Other            \\
+          1\%         & \lstinline|1 1 2 3 4| \\
+        \end{tabularx}
+    }                                                     \\\hline
+  \end{tabularx}
+  \caption{Some questions and student answers from \quizHof. Each
+    table row is a program and the (relative) frequencies of
+    student answers. Answers that can be considered correct are marked
+    with \texttt{*}.}
+  \label{t:quizzes-data-hof}
+\end{table}
+\Cref{t:quizzes-data-hof} lists interesting programs after the
+addition of closures and higher-order functions. It extends what we
+saw with loops in \Cref{s:intro}: that students have misconceptions
+about their interaction with mutable variables.
+
+\subsection{Collating Problems for the Tutor}%
+\label{s:tutor-refine-program-set}
+
+The SMoL Tutor's interpreting questions are the final instrument of
+this paper. I bootstrap it from the SMoL Quizzes, but made the
+following alterations:
+\begin{enumerate}
+
+  % fill gaps in goal sentences
+  \item I felt the existing programs were insufficient to cover the
+        ideas I wanted students to work through. For instance, SMoL
+        Quizzes did not alias vectors using \lstinline|defvar|, so we
+        added this program:
+
+\begin{lstlisting}
+  (defvar x (mvec 100))
+  (defvar y x)
+  (vec-set! x 0 200)
+  y
+\end{lstlisting}
+
+  \item I renamed vector operators to avoid confusion between
+        \lstinline|vset!|, which replaces a vector element, and \lstinline|set!|,
+        which mutates variables. I rename all vector operators from, for example,
+        \lstinline|vset!| to \lstinline|vec-set!|.
+
+  \item I deferred the local binding forms to the end, since I felt
+        they were less essential. I therefore rewrote programs that use
+        it to not depend on it.
+
+        % \item I removed some questions that I deemed ``Lispy'' (e.g.,
+        %   involving the difference between parenthetical and infix syntax).
+
+  \item I removed some programs that relied on underspecified aspects or
+        unimportant aspects of SMoL. For example, some depended on
+        whether operations on pairs could be applied to arbitrary vectors:
+
+\begin{lstlisting}
+  (pair? (mpair 1 2))
+  (pair? (mvec 1 2))
+  (pair? '#(1 2))
+  (pair? '(1 2))
+\end{lstlisting}
+
+        As another example, one hinged on
+        whether or not a function's formal parameter could have the same
+        name as the function itself:
+
+\begin{lstlisting}
+  (deffun (f f) f)
+  (f 5)
+\end{lstlisting}
+
+  \item I resolved ambiguities in some programs, either adding
+        answer choices or even adding other questions to tease
+        out different interpretations.
+
+  \item To reduce the number of concepts, I removed programs that
+        relied upon \emph{im}mutable vectors and lists, because they
+        did not seem to create problems. (For brevity, I leave these
+        out of the presentation of SMoL in \Cref{s:smol}, though they
+        are in the implementation.)
+
+  \item I removed questions related to function equality, which was
+        not a focus of the Tutor.
+
+  \item I removed programs of a ``Lispy'' nature (See
+        \Cref{s:reduce-syntactic-bias}).
+
+\end{enumerate}
+Most of these steps either modify or elide programs. Two cases,
+filling gaps in light of the goal sentences and resolving ambiguities,
+introduce programs. Starting with the 37 programs in the SMoL Quizzes,
+I added 52 more programs to arrive at a total of 89.
+
+In addition to generating this set of programs, I also modified the
+answer options. In addition to retaining the correct and incorrect
+answers from the SMoL Quizzes (and constructing my best guess of
+analogous incorrect answers for the new problems), I added more wrong
+answers to for the reasons given in \Cref{s:interpreting-tasks}. In
+general, I ended up increasing the number of choices substantially:
+only 8 out of the 89 questions have three or four choices; 70
+questions have 5--8 choices; and 11 questions have 9--14 choices.
+
+\section{Preliminary List of Misconceptions}%
+\label{s:misconceptions-tutor}
+
+I now examine what I learned from the tutor's interpreting tasks in
+terms of program behavior misconceptions.
+
+I do not present the results of and \textbf{local}
+(\Cref{t:tutor-goal-sentences}), because they focus on constructs not
+usually found in non-Lispy languages. They were included in the Tutor
+to help students write programs for the course, but are not
+interesting from a SMoL perspective.
+
+% \subsection{Misinterpreters}
+% \label{s:misinterp-intro}
+
+% In principle, an expert can identify what misconceptions a particular
+% incorrect answer might correspond to. In practice, I found this
+% rather difficult for three reasons. First, with 89 programs, each with
+% several answers, it's easy to make mistakes. Second, my own expert blind
+% spot may prevent me from seeing an interpretation that would
+% lead to an additional association. Finally, and specific to my case,
+% since I had either curated or written all the programs and answers,
+% I was likely to miss some associations I had not intended.
+
+% To address this problem, I formalized misconceptions as
+% interpreters. That is, for each misconception I created a
+% corresponding \emph{misinterpreter}. This is an interpreter for the
+% SMoL syntax that intentionally has a semantic error corresponding to
+% that misconception. Put differently, a misinterpreter is a like a
+% ``definitional interpreter for a misconception''.
+
+% By running programs through the misinterpreters, I can more uniformly
+% and rigorously identify \emph{all} the misconceptions associated with
+% a wrong answer. Furthermore, if I identify a new misconception (or
+% alter a misinterpreter), it is easy to automatically re-classify all
+% the answers. By using misinterpreters, I indeed found new
+% interpretations for existing program-answer pairs.
+
+% \subsection{A Catalog of Misconceptions}
+
+I iteratively created our final catalog of misconceptions (from the
+perspective of the data in this paper). I started with misinterpreters
+representing the misconceptions described in
+\Cref{s:misconceptions-quiz}. These are misconceptions for which we
+have reasonable validation (due to the prose in Quizius), so I call
+them \emph{grounded} misconceptions. I then looked at wrong answers
+not covered by these but chosen by students, and did my best to
+distill these into misconceptions. These are \emph{surmised}
+misconceptions (which we identify with a $^\ddag$), which need to be
+validated in the future. I then re-ran the misinterpreters against the
+chosen answers. I terminated when all the remaining wrong answers were
+either (a) found in very few students (we found a gap between 23\% and
+13\%, and hence took 14\% as the threshold), (b) difficult for us to
+attribute to a misconception, or (c) appeared to me to be ``Lispy''
+and hence not of broad interest.
+
+\Cref{t:tutor-grounded-misconceptions-1,t:tutor-grounded-misconceptions-2,t:tutor-surmised-misconceptions}
+list the final catalog. For each, I also present a Tutor question for
+which the marked wrong answer can be explained by \emph{only} the
+named misconception. That is, that program-answer pair is a
+representative example of that misconception.
+
+\begin{table}[!ht]
+  \centering\begin{tabularx}{\linewidth}{X l X}
+    \textbf{Misconception}                                                     &
+    \textbf{Question}                                                          &
+    \textbf{Table of Answers}
+    \\\hline
+    \miscon{CallByRef} Function calls alias variables.                         &
+    % mut-vars.not_aliased_by_funarg_2 (35)
+    \lstinputlisting{programs/call-by-ref.txt}                                 & {
+        \begin{tabularx}{\linewidth}{r X}
+          78\%   & \lstinline|12|    \\
+          11\%   & \lstinline|error| \\
+          **10\% & \lstinline|0|     \\
+          1\%    & \lstinline|23|    \\
+        \end{tabularx}
+      }
+    \\\hline
+    \miscon{CallsCopyStructs} Function calls copy data structures.             &
+    % vectors2.alias_with_funcall (48)
+    \lstinputlisting{programs/calls-copy-structs.txt}
+                                                                               & {
+        \begin{tabularx}{\linewidth}{r X}
+          90\%   & \lstinline|#(173 0)| \\
+          **10\% & \lstinline|#(1 0)|   \\
+        \end{tabularx}
+    }                                                                              \\\hline
+    \miscon{DeepClosure} Closures copy the \textit{values} of free variables.  &
+    % lambda3.lambda_remembers_env (70)
+    \lstinputlisting{programs/deep-closure.txt}                                & {
+        \begin{tabularx}{\linewidth}{r X}
+          86\%   & \lstinline|4|      \\
+          **10\% & \lstinline|3|      \\
+          3\%    & \lstinline|error|  \\
+          1\%    & \lstinline|lambda| \\
+        \end{tabularx}
+    }                                                                              \\\hline
+    \miscon{DefByRef} Variable definitions alias variables.                    &
+    % mut-vars.not_aliased_by_defvar_1 (29)
+    \lstinputlisting{programs/def-by-ref.txt}                                  & {
+        \begin{tabularx}{\linewidth}{r X}
+          85\%   & \lstinline|0 12|                \\
+          **12\% & \lstinline|0 0|                 \\
+          2\%    & \lstinline|error|               \\
+          1\%    & depends on implementation. \\
+        \end{tabularx}
+      }
+    \\\hline
+    \miscon{DefOrSet}  Both definitions and variable assignments are
+    interpreted as follows: if a variable is not defined in the
+    current environment, it is defined. Otherwise, it is mutated to
+    the new value.
+                                                                               &
+    % program: mut-vars.update_undefined (28)
+    \lstinputlisting{programs/def-or-set.txt}                                  & {
+        \begin{tabularx}{\linewidth}{r X}
+          85\%   & \lstinline|error| \\
+          **15\% & \lstinline|2|     \\
+        \end{tabularx}
+      }
+    \\\hline
+    \miscon{DefsCopyStructs} Variable definitions copy structures recursively. &
+    % vectors2.alias_with_defvar (47)
+    \lstinputlisting{programs/defs-copy-structs.txt}                           & {
+        \begin{tabularx}{\linewidth}{r X}
+          **67\% & \lstinline|#(100)| \\
+          30\%   & \lstinline|#(200)| \\
+          1\%    & \lstinline|#(300)| \\
+          1\%    & \lstinline|error|  \\
+        \end{tabularx}
+    }                                                                              \\\hline
+  \end{tabularx}
+  \caption{Ground misconceptions identified by the SMoL~Tutor. Answers marked
+    with ``**'' represent the misconception. (Part I)}
+  \label{t:tutor-grounded-misconceptions-1}
+\end{table}
+
+\begin{table}[!ht]
+  \centering
+  \begin{tabularx}{\linewidth}{X l X}
+    \textbf{Misconception}                                                                 &
+    \textbf{Question}                                                                      &
+    \textbf{Table of Answers}
+    \\\hline
+    \miscon{FlatEnv} There is only one environment, the global
+    environment. (This misconception is a kind of dynamic scope.)                          &
+    % scope.error_when_refer_to_undefined (9)
+    \lstinputlisting{programs/flat-env.txt}                                                & {
+        \begin{tabularx}{\linewidth}{r X}
+          96\%  & \lstinline|error| \\
+          **4\% & \lstinline|402|   \\
+        \end{tabularx}
+      }
+    \\\hline
+    \miscon{FunNotVal} Functions are \textit{not} considered
+    first-class values. They can't be bound to other variables, passed as arguments, or referred to by data structures.
+                                                                                           &
+    % lambda1.fun_as_parameter (62)
+    \lstinputlisting{programs/fun-not-val.txt}                                             & {
+        \begin{tabularx}{\linewidth}{r X}
+          83\%   & \lstinline|4|     \\
+          **11\% & \lstinline|error| \\
+          4\%    & \lstinline|2|     \\
+          1\%    & \lstinline|8|     \\
+        \end{tabularx}
+    }                                                                                          \\\hline
+    \miscon{IsolatedFun} Functions can't refer to free variables
+    except for the built-in ones.
+                                                                                           &
+    % scope.refer_global_is_possible (5)
+    \lstinputlisting{programs/isolated-fun.txt}                                            & {
+        \begin{tabularx}{\linewidth}{r X}
+          77\%   & \lstinline|3|     \\
+          **23\% & \lstinline|error| \\
+        \end{tabularx}
+      }
+    \\\hline
+    \miscon{NoCircularity} Data structures can't (possibly indirectly)
+    refer to themselves.                                                                   &
+    % vector2.warmup_circularity (53)
+    \lstinputlisting{programs/no-circularity.txt}                                          & {
+        \begin{tabularx}{\linewidth}{r X}
+          76\%   & \lstinline|3|                   \\
+          **14\% & \lstinline|error|               \\
+          9\%    & Run out of memory or time. \\
+          1\%    & \lstinline|+inf|                \\
+        \end{tabularx}
+    }                                                                                          \\\hline
+    \miscon{StructByRef} Data structures might refer to variables by their references.     &
+    % vector2.alias_var_in_mvec (49)
+    \lstinputlisting{programs/struct-by-ref.txt}                                           & {
+        \begin{tabularx}{\linewidth}{r X}
+          67\%   & \lstinline|#(1 2 3)| \\
+          **24\% & \lstinline|#(1 2 4)| \\
+          9\%    & error           \\
+        \end{tabularx}
+    }                                                                                          \\\hline
+    \miscon{StructsCopyStructs} Storing data structures into data structures makes copies. &
+    % lambda1.smol_quiz_circularity (58)
+    \lstinputlisting{programs/structs-copy-struct.txt}                                     & {
+        \begin{tabularx}{\linewidth}{r X}
+          65\%  & \lstinline|#0=#(#0# #0#)| (Racket circular object notation) \\
+          24\%  & \lstinline|#(#(2 3) #(2 3))|                                \\
+          **6\% & \lstinline|#(#(2 #(2 3)) #(2 3))|                           \\
+          6\%   & error                                                  \\
+        \end{tabularx}
+    }                                                                                          \\\hline
+  \end{tabularx}
+  \caption{Ground misconceptions identified by the SMoL~Tutor. Answers marked
+    with ``**'' represent the misconception. (Part II)}
+  \label{t:tutor-grounded-misconceptions-2}
+\end{table}
+
+\begin{table}[!ht]
+  \centering
+  \begin{tabularx}{\linewidth}{X l X}
+    \textbf{Misconception}                                                             &
+    \textbf{Question}                                                                  &
+    \textbf{Table of Answers}                                                              \\\hline
+    \surmisedMiscon{NestedDef} Sequences of definitions are
+    interpreted as if they are written in nested blocks. A definition
+    is not in the scope of later definitions.
+                                                                                       &
+    % program: scope.what_is_x_4 (13)
+    \lstinputlisting{programs/nested-def.txt}                                          & {
+        \begin{tabularx}{\linewidth}{r X}
+          92\%  & \lstinline|2| \\
+          **8\% & \lstinline|1| \\
+        \end{tabularx}
+    }                                                                                      \\\hline
+    \surmisedMiscon{Lazy} Expressions are only evaluated when their values are needed. &
+    % order.bad_order (20)
+    \lstinputlisting{programs/lazy.txt}                                                & {
+        \begin{tabularx}{\linewidth}{r X}
+          **57\% & \lstinline|1 3|   \\
+          43\%   & \lstinline|error| \\
+        \end{tabularx}
+    }                                                                                      \\\hline
+  \end{tabularx}
+  \caption{Surmised misconceptions identified by the SMoL~Tutor. Answers marked
+    with ``**'' represent the misconception.}
+  \label{t:tutor-surmised-misconceptions}
+\end{table}
+
+% \paragraph{Confirmed Misconceptions}
+
+% The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.
+
+% \paragraph{Added Misconceptions}
+
+% The misinterpreters helped me find misinterpretations that I had
+% overlooked. For instance, consider this program from \quizState:
+% \begin{lstlisting}
+% (defvar x 12)
+% (deffun (f x)
+%   (set! x 0))
+% (f x)
+% x
+% \end{lstlisting}
+% I had assumed the wrong answer
+% \texttt{0} is caused by \miscon{CallByRef}.
+% However, my misinterpreters made me realize it can also be explained
+% by \miscon{FlatEnv}. This could also explain why I see a difference in
+% error rate when the formal and actual parameter have the same
+% name (as mentioned in \Cref{s:misconceptions-quiz:state}).
+
+% Tutor's refined program sets and misinterpreters also helped me to
+% identify programs that detect one of the misconceptions but not the
+% other. Consider this new program
+% % mut-vars.not_aliased_by_defvar_1 (29)
+% \begin{lstlisting}
+%   (defvar x 12)
+%   (defvar y x)
+%   (set! x 0)
+%   x
+%   y
+% \end{lstlisting}
+% 12\% of students asserted that the following program produces
+% \texttt{0 0}, which suggest that they have the \textbf{BindByRef}
+% misconception. Students with a \miscon{FlatEnv} misconception would
+% have chosen \texttt{0 12}, which happens to be the correct answer.
+% (1\% wrote \textit{depends on
+% implementation}; 2\% asserted \texttt{error}; the remaining 85\%
+% asserted \texttt{0 12}).
+
+% The second program in \Cref{t:quizzes-data-fun} is already a good
+% detector for \miscon{FlatEnv}. SMoL Tutor includes an alternative
+% program that does not depend on equality checks.
+% % scope.error_when_refer_to_undefined (9)
+% \begin{lstlisting}
+%   (deffun (addy x)
+%     (defvar y 200)
+%     (+ x y))
+%   (+ (addy 2) y)
+% \end{lstlisting}
+% 4\% of students asserted that the following program produces
+% \texttt{402}, which suggest that they have the \miscon{FlatEnv}
+% misconception. Students with a \textbf{BindByRef} misconception would
+% have chosen \texttt{error}, which happens to be the correct answer.
+% (the remaining 96\% asserted \texttt{error}).
+
+\paragraph{A New Potential Misconception}
+
+For the following program, added in the Tutor:
+\begin{lstlisting}
+  (defvar y (+ x 2))
+  (defvar x 1)
+  x
+  y
+\end{lstlisting}
+56\% of students asserted that it produces
+\texttt{1 3}. Based on this, I surmise that students might have
+another misconception, which I define as \surmisedMiscon{Lazy}. (Recall that
+SMoL is eager, but even in many lazy languages, this would be an error.)
+
+\paragraph{Another Potential Misconception, and Its Effect on
+  Interpreting Descriptions.}%
+\label{ss:misinterpreters:submisconception}
+
+Consider this program from the SMoL Tutor:
+\begin{lstlisting}
+(defvar x 1)
+(deffun (main)
+  (deffun (get-x) x)
+  (defvar x 2)
+  (get-x))
+
+(main)
+\end{lstlisting}
+This program, suitably translated, would produce \lstinline|2| in a wide
+variety of languages (Python, JavaScript, Racket, Java, etc.), because
+\lstinline|get-x| and the inner \lstinline|x| are in the same scope block. The
+answer \lstinline|1| cannot be explained by any of my existing
+misconceptions. Based on this, I surmise a new misconception,
+\surmisedMiscon{NestedDef}. (Though its frequency falls below our
+threshold, my reading of answers suggests this may be more
+widespread, and I feel it needs to be investigated more.)
+
+Once I turned this into a misinterpreter, I found that it
+unexpectedly captured the program-answer pair for the following
+program from \quizFun---which should be an error, due to the
+double-binding of \lstinline|x| in the same scope block---and the answer
+\lstinline|2 0|:
+\begin{lstlisting}
+  (defvar x 0)
+  (defvar y x)
+  (defvar x 2)
+  x
+  y
+\end{lstlisting}
+Previously, I had interpreted this only as \miscon{DefOrSet}, because
+students had stated that the second \lstinline|(defvar x ....)|
+``mutates'' or ``redefines'' \lstinline|x|. This is reminiscent of the
+behavior of languages like Python, which use the same syntax both for
+binding new variables and for mutating existing ones.
+
+The problem here is that the word ``redefine'' underspecifies how the
+second definition is interpreted. I had interpreted it as
+\textit{mutating} the binding established by the first definition,
+which fits \miscon{DefOrSet}. However, another possibility is that
+it \emph{shadows} the binding (i.e., establishes a new scope block). We
+did not recognize that the original misconception is underspecified
+until I uncovered the new (surmised) misconception.
+
+\paragraph{Confirmed Misconceptions}
+
+The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.
+
+% \paragraph{Summary}
+
+% To summarize, I used the SMoL Tutor, with an expanded set of
+% programs, to investigate student misconceptions. I implemented the
+% idea of misinterpreters to help me properly classify student
+% performance. I were able to reconfirm all the previously identified
+% misconceptions, refine some of them, and also identify potential new
+% ones (that need further investigation).
+
+\section{What Pedagogic Techniques are Effective?}%
+\label{s:pedagogic-techniques}
+
+The fundamental problem about correcting misconceptions is: how do you
+fix a misconception? One approach is to only ``present the right
+answer'', for fear that discussing the wrong conception might actually
+reinforce it. Instead, there is a body of literature starting from
+\cite{posnerTheoryConceptualChange1982} that presents a theory of
+conceptual change, at whose heart is the \emph{refutation text}. A
+refutation text tackles the misconception directly, discussing and
+providing a refutation for the incorrect idea. Several studies
+\citep{schroederRefutationTextFacilitates2022} have shown their
+effectiveness in multiple other domains.
+
+\emph{Case comparisons} draw analogies between examples.
+\cite{alfieriLearningCaseComparisons2013a} suggest that asking
+(rather than not asking) students to find similarities between cases,
+and providing principles \emph{after} the comparisons (rather than
+before or not at all), are associated with better learning outcomes.
+
+Several groups of authors have argued for the benefits of visualizing
+program executions
+\citep{dicksonExperiencesImplementingUtilizing2022,
+guoOnlinePythonTutor2013, karnalimUsePythonTutor2017,
+napsExploringRoleVisualization2002}. Visualization seems to
+increase engagement and help students understand how programs work.
+These work often mention \emph{notional machines}
+\citep{duboulayBlackBoxGlass1981}, which refers to the underlying
+mental models rather than the visual representations.
+
+Prior research argues for the benefits of \emph{language levels}
+\citep{findlerDrSchemeProgrammingEnvironment2002,crestaniExperienceReportGrowing2010}:
+a language has better be taught as a sequence of languages, where each
+later language includes more language constructs than the previous
+one.
+
+\section{The SMoL Tutor}%
+\label{s:tutor}
+
+% So far, I have focused on identifying problems. As noted earlier
+% (\Cref{s:smol-quizzes}), I provided students with refutation texts
+% after the quizzes, but it is unclear to what extent students read,
+% understood, or internalized these. I also wanted to determine whether
+% other populations run into these issues, and it was unclear whether we
+% would be able to administer passive quizzes to them.
+
+% Furthermore, while the quizzes may be a useful diagnostic, my goal is
+% not only to find faults but to improve the understanding of basic
+% programming language behavior. I believe (and presumably so does
+% anyone else who writes a formal semantics!)\ that an understanding of
+% these basic program behaviors is important, and even more so when it
+% comes to understanding concurrency, ownership, and other advanced
+% features. Even more simply, misunderstanding these clearly impacts
+% development and debugging time.
+
+% In response, I decided to create a \emph{tutor}: the SMoL Tutor. It is built
+% around my quiz instruments and the detected misconceptions. We
+% describe the Tutor from a user's perspective in \Cref{s:tutor}, and
+% explain how it was populated from the
+% SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
+% I then discuss what
+% I learned from its data (\Cref{s:misconceptions-tutor}), and finally
+% evaluate its effectiveness as a \emph{tutor} (\Cref{s:effect-tutor}).
+
+% \subsection{The User Experience}
+
+This section describe the current status of the tutor.
+
+The Tutor includes five tutorials, shown on the left in
+\Cref{t:tutor-goal-sentences}. The larger topics are further broken
+down into 2--3 modules. The goal was for students to spend at most
+20--30 minutes per module. my data show that in practice, students
+spent about 9.8 (median) minutes.
+
+\Cref{f:smol-tutor-example} illustrates an interpreting task as
+described in \Cref{s:interpreting-tasks}. The Tutor also asked students
+to perform some other activities, like showing programs and asking for
+their heap content, which I do not cover in this paper.
+
+% Each tutorial consists of a sequence of questions. Most questions in
+% the Tutor are \newterm{interpreting questions}.\footnote{The Tutor
+% also asked students to perform some other activities, like showing
+% programs and asking for their heap content, which I do not cover in
+% this paper.} These questions (illustrated in
+% \Cref{f:smol-tutor-example}) are versions of the SMoL Quizzes
+% (obtained by the process described in
+% \Cref{s:tutor-refine-program-set}). In each of these questions,
+% students are shown a program and asked to predict the program's
+% running result(s) by answering an MCQ (with an ``Other'' option).
+% After making a choice, students receive feedback. If a student chooses
+% incorrectly, they are (a) given an explanation that is \emph{based on
+% the misconception associated with that wrong answer} (or a generic
+% one, if there is not a specific misconception), and then (b) asked to
+% answer a second question:
+% \begin{enumerate}
+
+%   \item The second question is semantically the same as the first, but
+%         with superficial changes (e.g., variable names, constants, and operators are
+%         changed) so that the student cannot immediately
+%         guess the answer.
+
+%   \item Instead of multiple-choice, students must \emph{type} the answer
+%         into a text box. (The Tutor normalizes text to accommodate variations.)
+%         This is
+%         intentional. First, I wanted to force reflection on the
+%         explanation just given, whereas with an
+%         MCQ, students could just guess. Second, I felt that students would
+%         find typing more onerous than clicking. In case students
+%         had just guessed on a question, my hope is that the penalty of having to
+%         type means, on \emph{subsequent} tasks, they
+%         would be more likely to pause and reflect before choosing.
+
+% \end{enumerate}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{images/smol-tutor-example.png}
+  \caption{Screenshots of an interpreting question in SMoL Tutor. The
+    top-left shows the initial state, where the question is presented
+    as an MCQ. If a student chooses a wrong answer, they will receive
+    feedback (bottom-left) and will be asked a similar question
+    (right). The similar question must be answered by typing.}
+  \label{f:smol-tutor-example}
+\end{figure}
+
+In addition to asking students questions, the Tutor along the way
+introduces terminology and states the true conceptions. These are the
+teaching goals of the Tutor. I therefore refer to these as
+\newterm{goal sentences}. \Cref{t:tutor-goal-sentences} lists
+the (abbreviated) goal sentences for each tutorial. I plan to keep
+goal sentences in the tutor.
+
+\begin{table}[t]
+  \centering\begin{tabularx}{\linewidth}{p{0.2\linewidth} X}
+    \textbf{Tutorial}                                              & \textbf{Goal Sentences}
+    \\\hline
+    \newterm{scope}: Variable definitions and function definitions &
+    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
+      \item Referring to an unbound variable leads to an error.
+      \item Define "blocks".
+      \item SMoL is lexically scoped rather than dynamically scoped.
+      \item SMoL disallows defining a variable twice in one block.
+      \item SMoL is eager (and evaluates from left to right) rather
+            than lazy, reactive, or relational.
+    \end{enumerate}
+    \\\hline
+    \newterm{mut-vars}: Variable updates                           &
+    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
+      \item Variable assignment respects the hierarchical structure of blocks.
+      \item Variables do not alias.
+    \end{enumerate}
+    \\\hline
+    \newterm{begin}: Sequencing expressions                        &
+    A sequencing expression evaluates its sub-expressions from left to right and
+    returns the value of the last sub-expression.
+    \\\hline
+    \newterm{vectors}: Vectors and vector updates                  &
+    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
+      \item Define heap and memory addresses.
+      \item Vectors can be referred to by vectors, including themselves.
+      \item Vectors are not copied but aliased by bindings.
+      \item Constructing vectors doesn't alter environments.
+      \item Introducing bindings doesn't alter the (value part of) heap.
+    \end{enumerate}
+    \\\hline
+    \newterm{lambda}: Lambda expressions                           &
+    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
+      \item Functions are (first-class) values.
+      \item Functions remember the environment where they were created.
+      \item Functions can be created with lambda expressions.
+      \item A function definition can be viewed as a variable definition plus a lambda expression.
+    \end{enumerate}
+    \\\hline
+    \newterm{local}: Local binding forms                           &
+    Introduce \texttt{let}, \texttt{letrec}, and \texttt{let*}.
+    % \\\hline
+  \end{tabularx}
+  \caption{SMoL Tutor tutorials and their goal sentences.}
+  \label{t:tutor-goal-sentences}
+\end{table}
+
+Some later tutorials include questions about earlier tutorials so
+that I can check whether students remember the concepts across
+modules. In particular, the \textbf{mut-vars} tutorial starts with questions
+about \textbf{scope}, and \textbf{lambda} starts with questions about
+\textbf{mut-vars} and \textbf{vectors}. I plan to keep this design to
+check retention.
+
+% Of the modules, \textbf{local} is the least portable across
+% languages. While local binding is of course present in other
+% languages, this is mostly covered using nested \lstinline|defvar|s in
+% earlier modules, starting with \textbf{scope}. The distinction central
+% to this module (between three local-binding constructs with slightly
+% different scopes) is primarily a focus of Lispy languages. Therefore,
+% I exclude this module from my analysis and instruments, and indeed
+% plan to deprecate the module in future versions.
+
+% \Cref{s:smol-quizzes} discusses the partial randomization employed by
+% the SMoL Quizzes. In contrast, the SMoL Tutor does \emph{not}
+% randomize question order. This is because the Tutor consists of more
+% than just questions: it also has explanatory text. This text is based
+% on the preceding problems. Authoring it is therefore somewhat like
+% writing a textbook, with complex dependencies that can easily be
+% broken.
+
+\section{How Effective Is the Tutor?}
+\label{s:effect-tutor}
+
+\begin{figure}[p]
+
+  \includegraphics[width=\linewidth]{images/Misconceptions_Decay.png}
+
+  \caption{How many students chose a wrong answer that (uniquely)
+    represents a misconception? (Downward tendency suggests improvement
+    over time.)}
+  \label{f:tutor-trend-plots}
+
+\end{figure}
+
+% Recall that the SMoL Tutor is not only a collection of MCQs: it is
+% also a \emph{tutor}! So far, I have investigated the value of the
+% MCQs. Now I examine its tutoring aspect. Concretely, I ask:
+% \begin{center}
+
+%   How effective is the Tutor at correcting each misconception?
+
+% \end{center}
+% To study this, I perform the following analysis per
+% misconception. Using misinterpreters, I identify those questions
+% whose wrong answers fit only one misconception (i.e., only one output
+% matches that produced by that misinterpreter, and it matches only that
+% misinterpreter). I then examine student performance over time across
+% those questions. This would let me examine how they do on just that
+% topic, in isolation, over time.
+
+This section presents part of my preliminary answer to \textbf{RQ2}
+(\Cref{s:plan-rq2}): is the tutor effective?
+
+I present the result in two forms. Graphically, we show plots in
+\Cref{f:tutor-trend-plots}. Each figure shows the percentage of
+students who chose the answer corresponding to that misconception.
+Ideally, we would like to see these percentages diminish.
+
+Indeed, that is what I see in most of the graphs. The exceptions are
+\miscon{CallsCopyStructs}, \surmisedMiscon{NestedDef}, and
+\miscon{StructsCopyStructs}, which have only one problem (and hence no
+trend), and \miscon{DefOrSet} and \miscon{FunNotVal}, which show an
+increase. The lack of improvement for \miscon{FunNotVal} is
+unsurprising because the Tutor does not explicitly address this issue,
+focusing on closures created by \lstinline|lambda|s rather than named
+functions. (However, this does not explain the increase!)
+
+I also perform a logistic regression to see whether these
+improvements are significant (at a $p < 0.05$ threshold); details are
+in \supplRef{Paper.html}. Of the 11:
+\begin{itemize}
+
+  \item Of the nine seemingly improved (i.e., decreased)
+        misconceptions:
+        \begin{itemize}
+          \item Two are \emph{not} significant: \miscon{CallByRef}
+                (\pValue\ = .074); \miscon{NoCircularity} (\pValue\ = .075).
+          \item The other seven \emph{are} significant.
+        \end{itemize}
+  \item However, the two with an increasing trend (\miscon{DefOrSet}
+        and \miscon{FunNotVal}) are \emph{also} significant.
+\end{itemize}
+
+Overall, the data suggest that the Tutor \emph{most likely did not do
+harm} and perhaps even \emph{may have done some good}. Revisions on
+the tutor (\Cref{s:future-work}), in light of misinterpreters, is
+necessarily to give a more accurate evaluation. For example I found
+far fewer problems than I would have liked. Concretely, only 40 of the
+71 eligible problems (after removing the non-SMoL modules) were useful
+in the above analysis.
+
+% These data broadly suggest that the Tutor is a net positive.
+% Ultimately, however, what these data really show is a need for
+% improvement in the Tutor. When designing the Tutor questions, many
+% more were intended to be representative of single misconceptions.
+% However, as noted in \Cref{s:misinterp-intro}, it is easy to be
+% incomplete (or incorrect) in ascribing misconceptions. Furthermore, as
+% their set grows, it is difficult to reassess all the problems
+% manually. Once evaluated using misinterpreters,
+
+% I therefore view the above data as
+% purely formative: they . However, it would be improper to read too much into the
+% analysis. It is quite possible that some of the other problems would
+% have found issues. Rather, what I really see is value in the
+% misinterpreter concept. It is not only useful for analysis, it is also
+% valuable for \emph{problem design}: in the next iteration of the
+% Tutor, I will use misinterpreters actively to shape the incorrect answers and,
+% as necessary, update the programs as well. I therefore hope to have
+% much more thorough analyzes in future work.
+
+\section{How Generalizable is My Observations?}
+\label{s:generality-tutor}
+
+% There is, of course, a significant danger that the data above have
+% been overfitted to University 1, thereby actually reflecting the state
+% of its curriculum rather than some greater truth about program
+% behavior understanding.
+
+% I already have some reason to believe this is
+% not the case: the related work discussed in
+% \Cref{s:background-misconceptions} is drawn from many institutions in
+% multiple countries with different educational preparations, levels,
+% and demographics. Nevertheless, that gives me only limited information
+% about the specific questions and misconceptions described above.
+
+% Fortunately, I was able to deploy the Tutor on two other
+% populations: University 2 and Texbook (See \Cref{s:populations} for their details).
+% \begin{itemize}
+
+%   \item University 2 is a primarily public university in the US. It is
+%         one of the largest Hispanic-serving institutions in the country. As
+%         such, its demographic is extremely different from those whose data
+%         were used above. The Tutor was used in one course in Spring 2023,
+%         taken by 12 students.
+%         The course is a third-year, programming
+%         language course. The students are required to have taken two
+%         introductory programming courses (C++ focused).
+
+%   \item A separate instance of the Tutor was published on the
+%         website of a programming languages textbook [ANON]. Over the course
+%         of 8 months, 597 people started with the first module and 103
+%         users made it to the last one. To protect privacy, I intentionally
+%         do not record demographic information, but I conjecture that the
+%         population is largely self-learners  (who are known to use the
+%         accompanying book),
+%         including some professional
+%         programmers. It is extremely
+%         unlikely to be the students from either university, because they
+%         would not get credit for their work on the public instance; they
+%         needed to use the university-specific instance. Furthermore, since
+%         they were not penalized for wrong answers, it would make little
+%         sense to do a ``test run'' on the public instance. Finally, I note
+%         that there is no overlap between the dates of submission on the
+%         public instance and the semester at University 1.
+
+% \end{itemize}
+% These two populations are therefore at least somewhat different from
+% the original population, and help me assess whether the problems we
+% identify are merely an artifact of the first institution.
+
+% To evaluate,
+I computed a Spearman's rank correlation $\rho$, ranking questions by
+what percentage of students got the question right. Between the
+original university and University 2, we obtain a \pValue~=~2.013e-07.
+Between the original university and the online population, I obtain a
+\pValue~<~2.2e-16. These show that the other two populations performed
+similarly to the original one. While further validation on other
+populations remains essential, these suggest that the questions are
+\emph{finding misconceptions that may be universal}.
+
+Of course, other threats to generalizability remains: both University
+1 and University 2 use the same textbook; SMoL's Lispy syntax might
+have bias in probing errors. I plan to address these issues in the
+future (\Cref{s:future-work}).
+
+% \section{Threats to Validity}
+% \label{s:threats}
+
+% \subsection{Construct Validity}
+
+% Did I measure the right thing?
+% While my goal is to study student understanding of language
+% behavior, what I actually measure is performance on MCQs on select
+% programs. MCQs as a mechanism introduce various clear biases, though
+% I add the ``Other'' options to somewhat alleviate them. The small
+% size, syntactic details (such as intentionally meaningless variable
+% names), and other aspects of the programs may also impact my ability
+% to measure understanding. (This could go both ways: students may have
+% no trouble understanding behavior in a small program but may struggle
+% to do so in a larger one.)
+
+% A particularly notable threat is the use of Lispy
+% syntax. I chose it for two reasons: both because it helps clarify
+% scope (\Cref{s:smol}) and because the rest of the course used
+% Racket. However, students may well perform differently with more
+% familiar syntaxes. It seems unlikely their performance would be \emph{too}
+% different given the many languages that have produced similar
+% misconceptions (\Cref{t:rel-work}). Nevertheless, I intend to use a
+% variety of syntaxes to examine this issue further.
+
+% Finally, I curated programs by hand, so they may well reflect my own
+% biases about misconceptions. It may be possible to mitigate this
+% problem by synthesizing MCQ programs using the misinterpreters.
+
+% \subsection{Internal Validity}
+
+% Is my reasoning valid? I have applied standard techniques and
+% measurements for evaluating student responses. However, our
+% instruments still lack the validity of a proper concept
+% inventory. Their creation requires heavyweight processes (such as
+% Delphi methods \citep{goldmanSettingScopeConcept2010}) that require many hours of expert attention as
+% well as conversations with learners, and can hence be prohibitive in
+% cost. The Quizius method (\Cref{s:quizius}) was created precisely to
+% be an inexpensive method that provides a good proxy.
+
+% Ultimately, my goal is to provide a reasonable instrument for
+% widespread use. While the set of all misconceptions could be
+% unbounded, I believe my tasks, especially as embedded in the Tutor,
+% provide a good starting point for others. In particular, if students
+% select a wrong answer, that is still of \emph{some} use to an
+% educator, even if the precise misconception cannot be pinned down with
+% the highest accuracy. I therefore believe my instruments, and our
+% misinterpreter technique, are of general value.
+
+% A failure in my Tutor's logging infrastructure led me to miss some
+% responses. These are unlikely to be task-specific because the Tutor
+% has a generic framework that should perform the same across
+% tasks. Moreover, on average only $0.4\%\ (\textrm{sd} = 0.6\%)$ of values are missing. Therefore, I do not believe
+% this had a noticeable impact. (In addition, every wrong answer is
+% still wrong! I may just not have exactly the right proportions of
+% them.)
+
+% Because I don't randomize Tutor question order, it is hard to tell
+% whether the general decreasing trend in \cref{f:tutor-trend-plots}
+% shows misconceptions being corrected, or shows properties of the
+% questions (e.g., maybe later questions in the plots are easier because
+% their programs happen to be shorter).
+
+% \subsection{External Validity}
+
+% How well do my results generalize? Certainly there is reason to
+% question whether my results would apply to other
+% populations. \Cref{s:generality-tutor} provides preliminary evidence
+% that the results are not specific to one institution, and
+% \Cref{s:background-misconceptions} suggests these issues are widespread. Nevertheless,
+% much more broad testing is needed to confirm my specific instruments.
+
+% The other major concern is the tie to Lispy syntax. Learners in other
+% settings may do worse or even better with it. Building a Tutor that
+% supports multiple, and more traditional, syntaxes should help address
+% this issue. I defer this to future work (\cref{s:future-work}).
+
+\section{Related Work}%
+\label{s:related-work}
+
+\subsection{Program Behavior}
+
+Our idea of ``goal sentences'' is not substantially different from the
+\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
+I only used these sentences (or rules) to guide the
+design of the Tutor and as text in the Tutor itself.
+\cite{duranRulesProgramBehavior2021} argue for other uses of such
+sentences.
+
+\subsection{Misinterpreters}
+
+Our idea of misinterpreters is related to mystery languages
+\citep{diwanPLdetectiveSystemTeaching2004,
+  pombrioTeachingProgrammingLanguages2017}. Both approaches use
+evaluators that represent alternative semantics to the same syntax.
+However, the two are complementary.
+In mystery languages, instructors design the space of semantics with
+pedagogic intent, and
+students must create programs to explore that space. Misinterpreters,
+in contrast, are driven by student input, while the programs are provided
+by instructors. The two approaches also have different goals:
+mystery languages focus on encouraging students to experiment with
+languages; misinterpreters aim at capturing students' misconceptions.
+
+\section{Future Work \& Timeline}%
+\label{s:future-work}
+
+\subsection{Notional Machine}
+
+The preliminary tutor provides a ``Copy This Program'' button that
+help students to run the program in DrRacket, where a SMoL
+implementation is provided.
+
+This approach is inconvenient both for the students and for me. For
+students, they have to click, pastes and run, which is less convenient
+than a single click. For me, because students run program in DrRacket
+rather than ``within'' the tutor, I do not know whether they have run
+the programs. This limit my ability to claim that the tutor is using
+NM to help teaching.
+
+Given these shortcomings, I have created a web-based notional machine
+\begin{center}
+  \href{https://lukuangchen.github.io/stacker-2023}{https://lukuangchen.github.io/stacker-2023}
+\end{center}
+The notional machine can be configured with URL parameters. For
+example,
+\href{https://lukuangchen.github.io/stacker-2023/?program=\%28defvar+x+12\%29\%0A\%28defvar+y+x\%29\%0A\%28set\%21+y+0\%29\%0Ax\%0Ay}{this
+link} opens the NM with a program embedded.
+
+The latest tutor provides a button on the top-right of each program.
+If students click the button, they will open a new browser tab that
+shows the NM populated with the program. I have deployed the latest
+tutor to the latest population at Brown (i.e., University 1), but I
+have not analyzed the data, which I plan to do in Spring 2024.
+
+I have not changed the logging systems to log when students click the
+buttons. I plan to make the change before Spring 2024.
+
+Another issue with the current design is that students might run a
+program before giving an answer. I don't know yet how often this
+happens. But given that students receive penalty (\Cref{s:penalty})
+for giving wrong answers, they might do so when the program is overly
+complicated. This problem is more of an issue after the switch to a
+web-based notional machine because now the ``cost'' of running a
+program is cheaper. My plan is to change the tutor so that it provides
+a program-specific link \emph{only} after each interpreting task.
+
+\subsection{Reduce Syntactic Bias}
+
+The processes described in
+\Cref{s:find-misconceptions-with-quizius,s:find-misconceptions-with-tutor}
+have been conducted with SMoL's Lispy syntax (\Cref{s:smol}). This can
+potentially introduce biases to the found misconceptions. For example,
+current results find that some students think the following program
+produces \lstinline|'(4 5)| rather than \lstinline|'(1 2)|.
+\begin{lstlisting}
+  (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
+\end{lstlisting}
+This checks whether $\verb|3| > \verb|n|$, but those students
+presumably vocalized it as ``greater than \lstinline|3|,
+\lstinline|n|?''. Although I have tried to remove programs like this
+from my instruments, there is still a risk of having biases in student
+errors. This semester (Fall 2023) I am already deploying a newer SMoL
+Tutor at Brown. This new tutor is able to display programs in
+JavaScript and Python, in addition to the Lispy syntax. I plan to
+compare in Spring 2024 the new tutor data with the current results,
+i.e., a comparison between uni-syntax tutor and multi-syntax tutor.
+
+\subsection{Use Case Comparisons More Consistently}
+
+The tutor currently prompts students to compare program-output pairs
+and then to summarize the rules of program behavior at the end of
+\emph{some} tutorials.
+
+I plan to expand this to \emph{all} tutorials and use a consistent
+prompt. The prompt should ask about the behavior of new language
+constructs and their interaction with all previously introduced
+constructs.
 
@@
 
-\subsection{Notional Machine}%
-\label{s:notional-machine}
-
-Prior research argues (\Cref{s:pedagogic-techniques})that illustrating
-how programs runs with notional machines (NMs) is beneficial for
-students. I have created a web-based notional machine
-\begin{center}
-  \href{https://lukuangchen.github.io/stacker-2023}{https://lukuangchen.github.io/stacker-2023}
-\end{center}
-
-The notional machine can be configured with URL parameters. For
-example,
-\href{https://lukuangchen.github.io/stacker-2023/?program=\%28defvar+x+12\%29\%0A\%28defvar+y+x\%29\%0A\%28set\%21+y+0\%29\%0Ax\%0Ay}{this
-link} opens the NM with a program embedded.
-
-My plan is to change the tutor so that it provides a
-program-specific link after each interpreting task and instructs
-students to click the link for more explanation.
-
-\subsection{Evaluate the Tutor}%
-\label{s:evaluation}
-
-To fully answer my research questions, I should also argue that my
-tutor is effective at correcting misconceptions. In this section, I
-describe how I want to refine the design of the tutor to help me make
-the argument, what data I want to collect, and what (statistical)
-analysis I want to perform on the data.
-
-I can infer whether students hold a misconception by analyzing their
-responses to interpreting tasks. There are three possible responses to
-an interpreting task:
+\subsection{Options in Interpreting Tasks}
+
+The tutor is adding extra options in interpreting tasks in an ad hoc
+way. Eventually, I plan to generate options by moving and/or copying
+numbers in the current options. In the case of the example question in
+\Cref{s:interpreting-tasks}, my plan would add \lstinline|12 12| and
+\lstinline|0 12|. In summary, each MCQ has the following options:
+\begin{itemize}
+  \item the correct answer,
+  \item wrong answers that correspond to known misconceptions,
+  \item options generated by moving and/or copying numbers in the
+        current options
+  \item the ``Other'' option (which, once chosen, allows students to
+        enter arbitrary answers)
+\end{itemize}
+
+\subsection{Other Changes in the Tutor}
+
+By the end of Fall 2023, I also want to improve the SMoL Tutor at
+least on the following aspects:
 \begin{enumerate}
-  \item They give the correct answer.
-  \item They give a wrong answer that corresponds to some
-        misconceptions.
-  \item They give a wrong answer that does not correspond to any (known)
-        misconceptions.
+  \item The tutor currently does not ask students to explain when they
+        gave a wrong answer in an interpreting question. This limit my
+        ability to find new misconceptions and to confirm that the
+        tasks do detect the anticipated misconceptions. I plan to
+        change the tutor so that it asks for an explanation in each
+        interpreting task.
+        % With student explanations, I check whether
+        % choosing the wrong answer indicates holding the corresponding
+        % misconception(s). I plan to eyeball the explanations to find
+        % new misconceptions. This piece of data also helps me confirm
+        % that anticipated wrong answers represent the expected
+        % misconceptions.
+  \item As I said in \Cref{s:effect-tutor}, for many of the
+        misconceptions, the tutor has too few questions. On the other
+        hand, there are many questions that do not produce useful data
+        because their wrong answers do not uniquely point to a
+        misconception. I plan to adjust the collection of interpreting
+        questions such that every misconception has at least two
+        questions that aim at correcting the misconception.
+  \item The tutor is not randomizing the order of questions. This make
+        it difficult to compare questions for one misconception --
+        earlier questions might be difficult simply because the
+        programs are somehow more difficult. I plan to randomize the
+        order of interpreting tasks within each tutorial module.
+  \item After randomizing question orders, I would need to check
+        carefully the dependencies between explanations and to
+        resolve these dependencies, if any.
+  \item The tutor is using many variable names, some are separated by
+        dash (e.g., \lstinline|get-x|). Students might misread these
+        variables. For example, some students might misread
+        \lstinline|(get-x)| (a function named \lstinline|get-x| is
+        called with no argument) as \lstinline|(get x)| (a function
+        named \lstinline|get| is called with one argument
+        \lstinline|x|). These misreading add noise to the data. So I plan
+        to use consistent and simple variable names in all programs.
 \end{enumerate}
-If a student gives a wrong answer that corresponds to exactly one
-misconception \emph{and} an explanation that matches the
-misconception, I deduce that the student holds the misconception when
-doing the task.
-
-Given this observation, I plan to set up the tutor such that
-\begin{itemize}
-  \item Every wrong answer corresponds to up to one misconception. In
-        this case, I say the wrong answer \emph{represents} the
-        misconception.
-  \item Every misconception is represented by some questions.
-  \item In each tutorial, a misconception is either not represented by
-        any interpreting tasks or represented by at least two tasks.
-  \item Within each tutorial, the order of interpreting tasks is
-        randomized.
-\end{itemize}
-For each tutorial and each misconception, I plan to do a model
-comparison (with the standard AIC and BIC criteria) on the following
-models/theories:
-\begin{description}
-  \item[TaskSpecific] Some students are more likely to choose the
-  representing wrong answers in certain tasks.
-  \item[PositionSpecific] Students are more likely to choose the
-  representing wrong answer in task that appears earlier (or later) in
-  the tutorial.
-\end{description}
-
-If the \textbf{TaskSpecific} model best explains the data, I deduce
-that some programs are more confusing and that I need further
-investigation into the relationship between the tasks and
-misconceptions. If the \textbf{PositionSpecific} model best explains
-\emph{and} students perform better in later tasks, I deduce that the
-tutor seems to be correcting misconceptions. If students perform
-worse in later tasks, the tutor is potentially doing harm.
-
-To know students well, I would like to have as many tasks per
-misconception as possible. But in practice, the number of tasks is
-subject to the following limitations:
-\begin{enumerate}
-  \item I don't want the length of any tutorial section to exceed 30
-  minutes.
-  \item Programs shouldn't be too long. Some misconceptions are really
-  similar and as a consequence have few short programs to tell them
-  apart.
-\end{enumerate}
-
-The current tutor is not in the final shape yet. In particular, many
-of the wrong answers correspond to more than one misconception; some
-misconceptions are represented by only one question; the order of
-interpreting tasks is not randomized. I plan to tackle the first two
-differences by adding and removing programs. Adding randomization
-requires me to carefully check the dependencies between current
-explanations and other texts in the tutor and resolve those
-dependencies, if any.
-
-\section{Generating and Collating Problems}%
-\label{s:gen-prob-quizius}
-
-% Additionally, in this paper, I intentionally use the word
-% mis\emph{conceptions} rather than mis\emph{take}. A mistake can happen
-% for any reason (e.g., selecting the wrong answer from a menu). A
-% misconception, however, implies a conceptual problem: the student has
-% formed an incorrect concept in their head. For instance, they may
-% think that mutable structures are copied on function calls, or that
-% scope is resolved dynamically. This requires probing what they are
-% thinking.
-
-% All these, however, add up to a somewhat challenging demand. I want
-% to produce a list of questions (each one an MCQ) such that
-% \begin{enumerate}
-
-%   \item I can get past the expert blind spot,
-
-%   \item I have a sense of what misconceptions students have, and
-
-%   \item I can generally associate wrong answers with specific
-%         misconceptions, approaching a concept inventory.
-
-% \end{enumerate}
-
-This section describes preliminary results on finding misconceptions
-from student-written programs (\Cref{s:quizius}) and what I have done
-to distill misconception-detecting programs from student inputs
-(\Cref{s:smol-quizzes,s:tutor-refine-program-set}).
-
-\subsection{Generating Problems Using Quizius}
-\label{s:quizius}
-
-My main solution to the expert blind spot is to use the Quizius system
-\citep{saarinenHarnessingWisdomClasses2019}. In contrast to the very
-heavyweight process (involving a lot of expert time) that is generally
-used to create a concept inventory, Quizius uses a lightweight,
-interactive approach to obtain fairly comparable data, which an expert
-can then shape into a quality instrument.
-
-In Quizius, experts create a prompt; in my case, students were to
-create small but ``interesting'' programs using the SMoL
-language. Quizius shows this prompt to students and gathers their
-answers. Each student is then shown a set of programs created by other
-students and asked to predict (without running it) the value produced
-by the program.\footnote{In the course (\Cref{s:populations}),
-  students were given credit for using Quizius but not penalized for wrong
-  answers, reducing their incentive to ``cheat'' by running
-  programs. They were also told that doing so would diminish the value of
-  their answers. Some students seemed to do so anyway, but most
-  honored the directive.} Students are also asked to provide a rationale for why
-they think it will produce that output.
-
-Quizius runs interactively during an assignment period.
-At each point, it needs to determine which
-previously authored program to show a student.
-It can
-either ``exploit'' a given program that already has responses
-or ``explore'' a new
-one. Quizius thus treats this as a multi-armed bandit problem
-\citep{katehakisMultiArmedBanditProblem1987} and uses that to choose a program.
-
-The output from Quizius is (a) a collection of programs; (b) for
-each program, a collection of predicted answers; and (c) for each
-answer, a rationale. Clustering the answers is easy (after
-ignoring some small syntactic differences). Thus, for
-each cluster, I obtain a set of rationales.
-
-After running Quizius in the course (\Cref{s:populations}),
-I took over as experts. Determining which is the right
-answer is easy. Where expert knowledge is useful is in
-\emph{clustering the rationales}. If all the rationales for a wrong
-answer are fairly similar, this is strong evidence that there is a
-common misconception that generates it. If, however, there are
-multiple rationale clusters, that means the program
-is not discriminative enough to distinguish the misconceptions, and it
-needs to be further refined to tell them apart. Interestingly, even
-the correct answer needs to be analyzed, because sometimes correct
-answers do have incorrect rationales (again, suggesting the program
-needs refinement to discriminate correct conceptions
-from misconceptions).
-
-Prior work using Quizius \citep{saarinenHarnessingWisdomClasses2019}
-finds that students do author programs that the experts did not
-imagine. In my case, I seeded Quizius with programs from prior papers
-(\Cref{s:background-misconceptions}), which gives the first few
-students programs to respond to. However, I found that Quizius
-significantly expanded the scope of my problems and misconceptions.
-In my final instrument, most programs were directly or indirectly
-inspired by the output of Quizius.
-
-The Quizius instrucment was set up by other people. I analyzed the
-collected data.
-
-\subsection{Collating Problems}
-\label{s:smol-quizzes}
-
-While Quizius is very useful in principle, it also produced data that
-needed significant curation for the following reasons:
-\begin{itemize}
-
-  \item A problem may have produced diverse outputs simply because it
-        was written in a very confusing way. Such programs do not reveal
-        any useful \emph{behavior} misconceptions, and must therefore be
-        filtered out. For instance:
-
-        \begin{lstlisting}
-  (defvar x 1)
-  (defvar y 2)
-  (defvar z 3)
-  (deffun (sum a ...) (+ a ...))
-  (sum x y z)
-  \end{lstlisting}
-
-        A reader might think that \lstinline|sum| takes variable arguments (so
-        the program produces \lstinline|6|), but in fact \lstinline|...| is a single
-        variable, so this produces an arity error.
-
-  \item Some programs relied on (or stumbled upon) underspecified aspects
-        of unimportant (in particular, non-standard) parts of SMoL, such as floating-point versus
-        rational arithmetic.
-
-  \item A problem may have produced diverse outputs simply because it is
-        hard to parse or to (mentally) trace its execution. One example was a
-        17-line program with 6 similar-looking and -named functions. As
-        another example:
-
-        \begin{lstlisting}
-  (defvar a (or (/ 1 (- 0.25 0.25)) (/ 1 0.0)))
-  (defvar b (and (/ 1 (- 0.25 0.25)) (/ 1 0.0)))
-  (defvar c (and (/ 1 0.0) (/ 1 (- 0.25 -0.25))))
-  (defvar d (or (/ 1 0) (/ 1 (- 0.25 -0.25))))
-  (and (or a c) (or b d))
-\end{lstlisting}
-
-        This program is not only confusing, it \emph{also} tests interpretations of (a) exact
-        versus inexact numbers and (b) truthy/falsiness, leading to
-        significant (but not very useful) answer diversity.
-
-  \item As noted above, a program's wrong (or even correct) answers may
-        correspond to multiple (mis)conceptions. In these cases, the program
-        must be refined to be more discriminative.
-
-\end{itemize}
-and so on.
-I therefore manually curated the Quizius output
-to address these issues.
-
-% \begin{figure}[t]
-%   \centering
-%   \includegraphics[width=\linewidth]{images/smol-quizzes-example.png}
-%   \caption{Screenshot of a SMoL Quiz question.}
-%   \label{f:smol-quizzes-example}
-% \end{figure}
-
-\subsection{The SMoL Quizzes}%
-\label{s:misconceptions-quiz}
-
-Having curated the output, I had to confirm that these programs were
-still effective! That is, they needed to actually find student errors.
-
-I therefore turned the programs into a set of quizzes (in the US
-sense: namely, a brief test of knowledge) that I call the SMoL
-Quizzes. There were three quizzes, ordered by linguistic
-complexity. The first consisted of only basic operators and
-first-order functions. The second added variable and structure
-mutation. The third added \lstinline|lambda| and higher-order functions.
-
-Each quiz is a sequence of MCQs. The options are from the Quizius
-data. In addition, there is an ``Other'' option. If chosen, the quiz
-gave the user a text box with the caption ``Please specify''. The goal
-here was to record any other answers, which in turn might lead to
-fresh misconceptions.
-
-The quizzes were administered in the third year in University 1.
-
-Question orders were partially randomized. I wanted students to get
-some easy, warm-up questions initially, so those were kept at the
-beginning. Similarly, I wanted programs that are syntactically
-similar to stay close to each other in the quiz. This is so that, when
-students got a second such program, they would not have to look far to
-find the first one and confirm that they are indeed (slightly)
-different, rather than wonder if they were seeing the same program
-again.
-
-Students only received feedback after having completed a whole quiz.
-At the end of each quiz, they received both summative feedback
-\emph{and} a refutation text that explained every program that
-appeared in the quiz. Students were also encouraged to run the
-programs, but I have little reason to believe that they did (and
-certainly they asked few questions on the class forum about them).
-
-Due to space limitations, I present the entire instrument in
-\supplRef{SMoL Quizzes/instrument}. Here I focus on a few programs
-where student choices correspond to misconceptions identified earlier,
-thereby also showing that the curated programs are still effective. I
-use a * to indicate the correct answer (which also matches the
-implementation's output).
-
-% Syntactically, \lstinline|#t| and \lstinline|#f| are true and false,
-% while \lstinline|#(...)| is a vector.
-
-\subsubsection{Quiz 1: Basic Operators and First-Order Functions}%
-\label{s:misconceptions-quiz:fun}
-
-% This quiz starts with two warm-up and not very interesting questions
-% and then other questions. The order of the other questions are fully
-% randomized.
-
-\begin{table}[t]
-  \centering
-  \begin{tabularx}{\linewidth}{l X}
-    \textbf{Questions}                       &
-    \textbf{Table of Answers}                    \\\hline
-    \lstinputlisting{programs/def_twice.txt} & {
-        \begin{tabularx}{\linewidth}{r X}
-          $^\dag$59\% & \lstinline|2 0|             \\
-          *32\%       & \lstinline|Error|           \\
-          4\%         & \lstinline|Other: 2 2|      \\
-          3\%         & ``Nothing is printed'' \\
-          2\%         & \lstinline|0 0|             \\
-        \end{tabularx}
-    }                                            \\\hline
-    \lstinputlisting{programs/dyn_scope.txt} & {
-        \begin{tabularx}{\linewidth}{r X}
-          *65\%       & \lstinline|Error| \\
-          $^\dag$29\% & \lstinline|42 #t| \\
-          6\%         & Other        \\
-        \end{tabularx}
-    }                                            \\\hline
-    \lstinputlisting{programs/reassign.txt}  & {
-        \begin{tabularx}{\linewidth}{r X}
-          *53\%       & \lstinline|#(6 5) 5|                 \\
-          *20\%       & \lstinline|Error|                    \\
-          $^\dag$11\% & \lstinline|#(6 6) 5|                 \\
-          $^\dag$11\% & \lstinline|#(6 6) 6|                 \\
-          5\%         & ``Nothing is printed'' or Other \\
-        \end{tabularx}
-    }                                            \\\hline
-  \end{tabularx}
-  \caption{Some questions and student answers from \quizFun. Each
-    table row is a program and the (relative) frequencies of
-    student answers. Answers that can be considered correct are marked
-    with \texttt{*}. Wrong answers that I discuss are
-    marked with $^\dag$.}
-  \label{t:quizzes-data-fun}
-\end{table}
-\Cref{t:quizzes-data-fun} lists programs in \quizFun{} that I
-consider the most interesting. These data confirm the presence of
-scope-related misconceptions:
-\begin{enumerate}
-  \item At least 59\% of students incorrectly believe that a
-        variable can be defined twice in one block.
-  \item 29\% of students believe that a variable defined in a function
-        will be available in the top-level (or global) environment
-        after the function is called. That is, these students may have a
-        dynamic scope misconception.
-  \item 22\% (11\% + 11\%) of students believe variables themselves
-        can be passed as arguments and redefined inside the function.
-        They disagree on whether the redefinition persists after the
-        function call.
-\end{enumerate}
-
-\subsubsection{Quiz 2: Adding Variable and Structure Mutation}%
-\label{s:misconceptions-quiz:state}
-
-% This quiz randomizes the order of all questions.
-
-\begin{table}[t]
-  \centering
-  \begin{tabularx}{\linewidth}{l X}
-    \textbf{Question}                           &
-    \textbf{Table of Answers}                       \\\hline
-    \lstinputlisting{programs/circularity.txt}  & {
-        \begin{tabularx}{\linewidth}{r X}
-          $^\dag$50\% & \lstinline|#(#(2 #(2 3)) #(2 3))| \\
-          *29\%       & \lstinline|x=#(x x)| or
-          something similar. Both \lstinline|(left x)| and \lstinline|(right x)|
-          are \lstinline|x| itself.                       \\
-          $^\dag$16\% & \lstinline|Error|                 \\
-          5\%         & Other                        \\
-        \end{tabularx}
-    }                                               \\\hline
-    % \lstinputlisting{programs/mvec_in_mvec.txt} & {
-    %     \begin{tabularx}{\linewidth}{r X}
-    %       *62\%       & \lstinline|#(#(100 2 3 4) #(100 2 3 4))| \\
-    %       $^\dag$26\% & \lstinline|#(#(1 2 3 4) #(100 2 3 4))|   \\
-    %       6\%         & \texttt{Error}                      \\
-    %       4\%         & \lstinline|#(#(1 2 3 4) #(1 2 3 4))|     \\
-    %       2\%         & Other                               \\
-    %     \end{tabularx}
-    % }                                               \\\hline
-    % \lstinputlisting{programs/vset_in_let.txt}  & {
-    %     \begin{tabularx}{\linewidth}{r X}
-    %       *59\%       & \lstinline|#(10)|  \\
-    %       $^\dag$36\% & \lstinline|#(123)| \\
-    %       5\%         & Other         \\
-    %     \end{tabularx}
-    % }                                               \\\hline
-    \lstinputlisting{programs/var_as_arg.txt}   & {
-        \begin{tabularx}{\linewidth}{r X}
-          *65\%       & \texttt{12} \\
-          $^\dag$31\% & \texttt{0}  \\
-          4\%         & Other       \\
-        \end{tabularx}
-    }                                               \\\hline
-    \lstinputlisting{programs/setset.txt}       & {
-        \begin{tabularx}{\linewidth}{r X}
-          *59\%       & Other: \texttt{5 7}   \\
-          $^\dag$27\% & \texttt{6 7}          \\
-          11\%        & \texttt{5 5}          \\
-          2\%         & Other: \texttt{Error} \\
-          1\%         & Other: \texttt{5 6}   \\
-        \end{tabularx}
-    }                                               \\\hline
-  \end{tabularx}
-  \caption{Some questions and student answers from \quizState. Each
-    table row is a program and the (relative) frequencies of
-    student answers. Answers that can be considered correct are marked
-    with \texttt{*}. Wrong answers that I discuss are
-    marked with $^\dag$.}
-  \label{t:quizzes-data-state}
-\end{table}
-\Cref{t:quizzes-data-state} lists interesting programs after the
-addition of state. These data suggest the students have aliasing-related misconceptions:
-\begin{itemize}
-  \item Up to 50\% of students think vectors are copied rather than
-        aliased.
-  \item 16\% of students think trying to construct and print a
-        self-referring vector would error. (This program is ambiguous:
-        \emph{constructing} works in most SMoL languages, but \emph{printing}
-        may well cause a problem. These identified ambiguities are
-        addressed in \Cref{s:tutor-refine-program-set}.)
-  \item 31\% of students think a variable is aliased by a parameter if
-        the two variables have the same name. Perhaps interestingly,
-        fewer students (27\%) think the variable aliasing would
-        happen if the two variables had different names.
-\end{itemize}
-
-\subsubsection{Quiz 3: Adding Closures and Higher-Order Functions}%
-\label{s:misconceptions-quiz:hof}
-
-\begin{table}[t]
-  \centering
-  \begin{tabularx}{\linewidth}{l X}
-    \textbf{Question}                                 &
-    \textbf{Table of Answers}                             \\\hline
-    \lstinputlisting{programs/fun_returns_lambda.txt} & {
-        \begin{tabularx}{\linewidth}{r X}
-          *82\%       & \lstinline|3|     \\
-          $^\dag$17\% & \lstinline|Error| \\
-          1\%         & Other        \\
-        \end{tabularx}
-    }                                                     \\\hline
-    \lstinputlisting{programs/fun_and_state_1.txt}    & {
-        \begin{tabularx}{\linewidth}{r X}
-          *74\%       & \lstinline|4|            \\
-          $^\dag$21\% & \lstinline|3|            \\
-          6\%         & Other: \lstinline|Error| \\
-        \end{tabularx}
-    }                                                     \\\hline
-    % \lstinputlisting{programs/fun_and_state_2.txt}    & {
-    %     \begin{tabularx}{\linewidth}{r X}
-    %       *88\%       & \lstinline|4|            \\
-    %       $^\dag$10\% & \lstinline|3|            \\
-    %       3\%         & Other: \lstinline|Error| \\
-    %     \end{tabularx}
-    % }                                                     \\\hline
-    \lstinputlisting{programs/counter.txt}            & {
-        \begin{tabularx}{\linewidth}{r X}
-          *62\%       & \lstinline|1 1 2 3 2| \\
-          $^\dag$34\% & \lstinline|1 1 1 1 1| \\
-          4\%         & Other            \\
-          1\%         & \lstinline|1 1 2 3 4| \\
-        \end{tabularx}
-    }                                                     \\\hline
-  \end{tabularx}
-  \caption{Some questions and student answers from \quizHof. Each
-    table row is a program and the (relative) frequencies of
-    student answers. Answers that can be considered correct are marked
-    with \texttt{*}. Wrong answers that I discuss are
-    marked with $^\dag$.}
-  \label{t:quizzes-data-hof}
-\end{table}
-\Cref{t:quizzes-data-hof} lists interesting programs after the
-addition of closures and higher-order functions. It extends what we
-saw with loops in \Cref{s:intro}: that students have misconceptions
-about their interaction with mutable variables:
-\begin{itemize}
-  \item 17\% of students think \texttt{lambda} functions can't refer
-        to free variables.
-  \item 21\% of the students think mutating a variable defined outside
-        a lambda can't possibly change the behavior of the lambda.
-  \item Student understanding of the
-        \lstinline|let|-over-\lstinline|lambda| pattern is weak.
-\end{itemize}
-
-\subsection{Collating Problems for the Tutor}%
-\label{s:tutor-refine-program-set}
-
-The SMoL Tutor's interpreting questions are the final instrument of
-this paper. I bootstrap it from the SMoL Quizzes, but made the
-following alterations:
-\begin{enumerate}
-
-  % fill gaps in goal sentences
-  \item I felt the existing programs were insufficient to cover the
-        ideas I wanted students to work through. For instance, SMoL
-        Quizzes did not alias vectors using \lstinline|defvar|, so we
-        added this program:
-
-\begin{lstlisting}
-  (defvar x (mvec 100))
-  (defvar y x)
-  (vec-set! x 0 200)
-  y
-\end{lstlisting}
-
-  \item I renamed vector operators to avoid confusion between
-        \lstinline|vset!|, which replaces a vector element, and \lstinline|set!|,
-        which mutates variables. I rename all vector operators from, for example,
-        \lstinline|vset!| to \lstinline|vec-set!|.
-
-  \item I deferred the local binding forms to the end, since I felt
-        they were less essential. I therefore rewrote programs that use
-        it to not depend on it.
-
-        % \item I removed some questions that I deemed ``Lispy'' (e.g.,
-        %   involving the difference between parenthetical and infix syntax).
-
-  \item I removed some programs that relied on underspecified aspects or
-        unimportant aspects of SMoL. For example, some depended on
-        whether operations on pairs could be applied to arbitrary vectors:
-
-\begin{lstlisting}
-  (pair? (mpair 1 2))
-  (pair? (mvec 1 2))
-  (pair? '#(1 2))
-  (pair? '(1 2))
-\end{lstlisting}
-
-        As another example, one hinged on
-        whether or not a function's formal parameter could have the same
-        name as the function itself:
-
-\begin{lstlisting}
-  (deffun (f f) f)
-  (f 5)
-\end{lstlisting}
-
-  \item I resolved ambiguities in some programs, either adding
-        answer choices or even adding other questions to tease
-        out different interpretations.
-
-  \item To reduce the number of concepts, I removed programs that
-        relied upon \emph{im}mutable vectors and lists, because they
-        did not seem to create problems. (For brevity, I leave these
-        out of the presentation of SMoL in \Cref{s:smol}, though they
-        are in the implementation.)
-
-  \item I removed questions related to function equality, which was
-        not a focus of the Tutor.
-
-  \item I removed programs of a ``Lispy'' nature (See
-        \Cref{s:reduce-syntactic-bias}).
-
-\end{enumerate}
-Most of these steps either modify or elide programs. Two cases,
-filling gaps in light of the goal sentences and resolving ambiguities,
-introduce programs. Starting with the 37 programs in the SMoL Quizzes,
-I added 52 more programs to arrive at a total of 89.
-
-In addition to generating this set of programs, I also modified the
-answer options. In addition to retaining the correct and incorrect
-answers from the SMoL Quizzes (and constructing my best guess of
-analogous incorrect answers for the new problems), I added more wrong
-answers to for the reasons given in \Cref{s:interpreting-tasks}. In
-general, I ended up increasing the number of choices substantially:
-only 8 out of the 89 questions have three or four choices; 70
-questions have 5--8 choices; and 11 questions have 9--14 choices.
-
-\section{What Pedagogic Techniques are Effective?}%
-\label{s:pedagogic-techniques}
-
-The fundamental problem about correcting misconceptions is: how do you
-fix a misconception? One approach is to only ``present the right
-answer'', for fear that discussing the wrong conception might actually
-reinforce it. Instead, there is a body of literature starting from
-\cite{posnerTheoryConceptualChange1982} that presents a theory of
-conceptual change, at whose heart is the \emph{refutation text}. A
-refutation text tackles the misconception directly, discussing and
-providing a refutation for the incorrect idea. Several studies
-\citep{schroederRefutationTextFacilitates2022} have shown their
-effectiveness in multiple other domains.
-
-\emph{Case comparisons} draw analogies between examples.
-\cite{alfieriLearningCaseComparisons2013a} suggest that asking
-(rather than not asking) students to find similarities between cases,
-and providing principles \emph{after} the comparisons (rather than
-before or not at all), are associated with better learning outcomes.
-
-Several groups of authors have argued for the benefits of visualizing
-program executions
-\citep{dicksonExperiencesImplementingUtilizing2022,
-guoOnlinePythonTutor2013, karnalimUsePythonTutor2017,
-napsExploringRoleVisualization2002}. Visualization seems to
-increase engagement and help students understand how programs work.
-These work often mention \emph{notional machines}
-\citep{duboulayBlackBoxGlass1981}, which refers to the underlying
-mental models rather than the visual representations.
-
-Prior research argues for the benefits of \emph{language levels}
-\citep{findlerDrSchemeProgrammingEnvironment2002,crestaniExperienceReportGrowing2010}:
-a language has better be taught as a sequence of languages, where each
-later language includes more language constructs than the previous
-one.
-
-\section{The SMoL Tutor}%
-\label{s:tutor}
-
-% So far, I have focused on identifying problems. As noted earlier
-% (\Cref{s:smol-quizzes}), I provided students with refutation texts
-% after the quizzes, but it is unclear to what extent students read,
-% understood, or internalized these. I also wanted to determine whether
-% other populations run into these issues, and it was unclear whether we
-% would be able to administer passive quizzes to them.
-
-% Furthermore, while the quizzes may be a useful diagnostic, my goal is
-% not only to find faults but to improve the understanding of basic
-% programming language behavior. I believe (and presumably so does
-% anyone else who writes a formal semantics!)\ that an understanding of
-% these basic program behaviors is important, and even more so when it
-% comes to understanding concurrency, ownership, and other advanced
-% features. Even more simply, misunderstanding these clearly impacts
-% development and debugging time.
-
-% In response, I decided to create a \emph{tutor}: the SMoL Tutor. It is built
-% around my quiz instruments and the detected misconceptions. We
-% describe the Tutor from a user's perspective in \Cref{s:tutor}, and
-% explain how it was populated from the
-% SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
-% I then discuss what
-% I learned from its data (\Cref{s:misconceptions-tutor}), and finally
-% evaluate its effectiveness as a \emph{tutor} (\Cref{s:effect-tutor}).
-
-% \subsection{The User Experience}
-
-This section describe the current status of the tutor.
-
-The Tutor includes five tutorials, shown on the left in
-\Cref{t:tutor-goal-sentences}. The larger topics are further broken
-down into 2--3 modules. The goal was for students to spend at most
-20--30 minutes per module. my data show that in practice, students
-spent about 9.8 (median) minutes.
-
-\Cref{f:smol-tutor-example} illustrates an interpreting task as
-described in \Cref{s:interpreting-tasks}. The Tutor also asked students
-to perform some other activities, like showing programs and asking for
-their heap content, which I do not cover in this paper.
-
-% Each tutorial consists of a sequence of questions. Most questions in
-% the Tutor are \newterm{interpreting questions}.\footnote{The Tutor
-% also asked students to perform some other activities, like showing
-% programs and asking for their heap content, which I do not cover in
-% this paper.} These questions (illustrated in
-% \Cref{f:smol-tutor-example}) are versions of the SMoL Quizzes
-% (obtained by the process described in
-% \Cref{s:tutor-refine-program-set}). In each of these questions,
-% students are shown a program and asked to predict the program's
-% running result(s) by answering an MCQ (with an ``Other'' option).
-% After making a choice, students receive feedback. If a student chooses
-% incorrectly, they are (a) given an explanation that is \emph{based on
-% the misconception associated with that wrong answer} (or a generic
-% one, if there is not a specific misconception), and then (b) asked to
-% answer a second question:
-% \begin{enumerate}
-
-%   \item The second question is semantically the same as the first, but
-%         with superficial changes (e.g., variable names, constants, and operators are
-%         changed) so that the student cannot immediately
-%         guess the answer.
-
-%   \item Instead of multiple-choice, students must \emph{type} the answer
-%         into a text box. (The Tutor normalizes text to accommodate variations.)
-%         This is
-%         intentional. First, I wanted to force reflection on the
-%         explanation just given, whereas with an
-%         MCQ, students could just guess. Second, I felt that students would
-%         find typing more onerous than clicking. In case students
-%         had just guessed on a question, my hope is that the penalty of having to
-%         type means, on \emph{subsequent} tasks, they
-%         would be more likely to pause and reflect before choosing.
-
-% \end{enumerate}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=\textwidth]{images/smol-tutor-example.png}
-  \caption{Screenshots of an interpreting question in SMoL Tutor. The
-    top-left shows the initial state, where the question is presented
-    as an MCQ. If a student chooses a wrong answer, they will receive
-    feedback (bottom-left) and will be asked a similar question
-    (right). The similar question must be answered by typing.}
-  \label{f:smol-tutor-example}
-\end{figure}
-
-In addition to asking students questions, the Tutor along the way
-introduces terminology and states the true conceptions. These are the
-teaching goals of the Tutor. I therefore refer to these as
-\newterm{goal sentences}. \Cref{t:tutor-goal-sentences} lists
-the (abbreviated) goal sentences for each tutorial. I plan to keep
-goal sentences in the tutor.
-
-\begin{table}[t]
-  \centering\begin{tabularx}{\linewidth}{p{0.2\linewidth} X}
-    \textbf{Tutorial}                                              & \textbf{Goal Sentences}
-    \\\hline
-    \newterm{scope}: Variable definitions and function definitions &
-    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
-      \item Referring to an unbound variable leads to an error.
-      \item Define "blocks".
-      \item SMoL is lexically scoped rather than dynamically scoped.
-      \item SMoL disallows defining a variable twice in one block.
-      \item SMoL is eager (and evaluates from left to right) rather
-            than lazy, reactive, or relational.
-    \end{enumerate}
-    \\\hline
-    \newterm{mut-vars}: Variable updates                           &
-    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
-      \item Variable assignment respects the hierarchical structure of blocks.
-      \item Variables do not alias.
-    \end{enumerate}
-    \\\hline
-    \newterm{begin}: Sequencing expressions                        &
-    A sequencing expression evaluates its sub-expressions from left to right and
-    returns the value of the last sub-expression.
-    \\\hline
-    \newterm{vectors}: Vectors and vector updates                  &
-    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
-      \item Define heap and memory addresses.
-      \item Vectors can be referred to by vectors, including themselves.
-      \item Vectors are not copied but aliased by bindings.
-      \item Constructing vectors doesn't alter environments.
-      \item Introducing bindings doesn't alter the (value part of) heap.
-    \end{enumerate}
-    \\\hline
-    \newterm{lambda}: Lambda expressions                           &
-    \begin{enumerate}[topsep=0pt, partopsep=0pt, itemsep=0pt, parsep=0pt]
-      \item Functions are (first-class) values.
-      \item Functions remember the environment where they were created.
-      \item Functions can be created with lambda expressions.
-      \item A function definition can be viewed as a variable definition plus a lambda expression.
-    \end{enumerate}
-    \\\hline
-    \newterm{local}: Local binding forms                           &
-    Introduce \texttt{let}, \texttt{letrec}, and \texttt{let*}.
-    % \\\hline
-  \end{tabularx}
-  \caption{SMoL Tutor tutorials and their goal sentences.}
-  \label{t:tutor-goal-sentences}
-\end{table}
-
-Some later tutorials include questions about earlier tutorials so
-that I can check whether students remember the concepts across
-modules. In particular, the \textbf{mut-vars} tutorial starts with questions
-about \textbf{scope}, and \textbf{lambda} starts with questions about
-\textbf{mut-vars} and \textbf{vectors}. I plan to keep this design to
-check retention.
-
-% Of the modules, \textbf{local} is the least portable across
-% languages. While local binding is of course present in other
-% languages, this is mostly covered using nested \lstinline|defvar|s in
-% earlier modules, starting with \textbf{scope}. The distinction central
-% to this module (between three local-binding constructs with slightly
-% different scopes) is primarily a focus of Lispy languages. Therefore,
-% I exclude this module from my analysis and instruments, and indeed
-% plan to deprecate the module in future versions.
-
-% \Cref{s:smol-quizzes} discusses the partial randomization employed by
-% the SMoL Quizzes. In contrast, the SMoL Tutor does \emph{not}
-% randomize question order. This is because the Tutor consists of more
-% than just questions: it also has explanatory text. This text is based
-% on the preceding problems. Authoring it is therefore somewhat like
-% writing a textbook, with complex dependencies that can easily be
-% broken.
-
-\section{Misconceptions Detected by the SMoL Tutor}%
-\label{s:misconceptions-tutor}
-
-I now examine what I learned from the interpreting
-questions in terms of program behavior
-misconceptions. But first, I explain how I associate incorrect
-answers with misconceptions.
-
-I do not present the results of and \textbf{local}
-(\Cref{t:tutor-goal-sentences}), because they focus on constructs not
-usually found in non-Lispy languages. They were included in the Tutor
-to help students write programs for the course, but are not
-interesting from a SMoL perspective.
-
-\subsection{Misinterpreters}
-\label{s:misinterp-intro}
-
-In principle, an expert can identify what misconceptions a particular
-incorrect answer might correspond to. In practice, I found this
-rather difficult for three reasons. First, with 89 programs, each with
-several answers, it's easy to make mistakes. Second, my own expert blind
-spot may prevent me from seeing an interpretation that would
-lead to an additional association. Finally, and specific to my case,
-since I had either curated or written all the programs and answers,
-I was likely to miss some associations I had not intended.
-
-To address this problem, I formalized misconceptions as
-interpreters. That is, for each misconception I created a
-corresponding \emph{misinterpreter}. This is an interpreter for the
-SMoL syntax that intentionally has a semantic error corresponding to
-that misconception. Put differently, a misinterpreter is a like a
-``definitional interpreter for a misconception''.
-
-By running programs through the misinterpreters, I can more uniformly
-and rigorously identify \emph{all} the misconceptions associated with
-a wrong answer. Furthermore, if I identify a new misconception (or
-alter a misinterpreter), it is easy to automatically re-classify all
-the answers. By using misinterpreters, I indeed found new
-interpretations for existing program-answer pairs.
-
-\subsection{A Catalog of Misconceptions}
-
-I iteratively created
-our final catalog of misconceptions
-(from the perspective of the data in this paper).
-I started
-with misinterpreters representing the misconceptions described in
-\Cref{s:misconceptions-quiz}. These are misconceptions for which we
-have reasonable validation (due to the prose in Quizius), so I call
-them \emph{grounded} misconceptions. I then looked at wrong answers
-not covered by these but chosen by students, and did my best to
-distill these into misconceptions. These are \emph{surmised}
-misconceptions (which we
-identify with a $^\ddag$), which need to be validated in the future.\footnote{To keep
-  the time spent in each module reasonable, the SMoL Tutor does not
-  ask students for rationales for their programs. However, I could
-  easily modify it to do so a small number of times per user, to make
-  progress towards this need.} I then re-ran the misinterpreters
-against the chosen answers. I terminated when all the remaining wrong
-answers were either (a) found in very few students (we found a gap
-between 23\% and 13\%, and hence took 14\% as the threshold),
-(b) difficult for
-us to attribute to a misconception, or (c) appeared to me to be
-``Lispy'' and hence not of broad interest.
-
-\Cref{t:tutor-grounded-misconceptions-1,t:tutor-grounded-misconceptions-2,t:tutor-surmised-misconceptions}
-list the final catalog.
-For each, I also
-present a Tutor question
-for which the marked wrong answer can be explained by \emph{only} the
-named misconception.
-That is, that program-answer pair is a representative example of that
-misconception.\footnote{An astute reader may well imagine another
-  misinterpretation; but I presumably did not find evidence of it in our
-  data. I do welcome other misinterpretations, for which we
-  can add more misinterpreters!}
-
-\begin{table}[!ht]
-  \centering\begin{tabularx}{\linewidth}{X l X}
-    \textbf{Misconception}                                                     &
-    \textbf{Question}                                                          &
-    \textbf{Table of Answers}
-    \\\hline
-    \miscon{CallByRef} Function calls alias variables.                         &
-    % mut-vars.not_aliased_by_funarg_2 (35)
-    \lstinputlisting{programs/call-by-ref.txt}                                 & {
-        \begin{tabularx}{\linewidth}{r X}
-          78\%   & \lstinline|12|    \\
-          11\%   & \lstinline|error| \\
-          **10\% & \lstinline|0|     \\
-          1\%    & \lstinline|23|    \\
-        \end{tabularx}
-      }
-    \\\hline
-    \miscon{CallsCopyStructs} Function calls copy data structures.             &
-    % vectors2.alias_with_funcall (48)
-    \lstinputlisting{programs/calls-copy-structs.txt}
-                                                                               & {
-        \begin{tabularx}{\linewidth}{r X}
-          90\%   & \lstinline|#(173 0)| \\
-          **10\% & \lstinline|#(1 0)|   \\
-        \end{tabularx}
-    }                                                                              \\\hline
-    \miscon{DeepClosure} Closures copy the \textit{values} of free variables.  &
-    % lambda3.lambda_remembers_env (70)
-    \lstinputlisting{programs/deep-closure.txt}                                & {
-        \begin{tabularx}{\linewidth}{r X}
-          86\%   & \lstinline|4|      \\
-          **10\% & \lstinline|3|      \\
-          3\%    & \lstinline|error|  \\
-          1\%    & \lstinline|lambda| \\
-        \end{tabularx}
-    }                                                                              \\\hline
-    \miscon{DefByRef} Variable definitions alias variables.                    &
-    % mut-vars.not_aliased_by_defvar_1 (29)
-    \lstinputlisting{programs/def-by-ref.txt}                                  & {
-        \begin{tabularx}{\linewidth}{r X}
-          85\%   & \lstinline|0 12|                \\
-          **12\% & \lstinline|0 0|                 \\
-          2\%    & \lstinline|error|               \\
-          1\%    & depends on implementation. \\
-        \end{tabularx}
-      }
-    \\\hline
-    \miscon{DefOrSet}  Both definitions and variable assignments are
-    interpreted as follows: if a variable is not defined in the
-    current environment, it is defined. Otherwise, it is mutated to
-    the new value.
-                                                                               &
-    % program: mut-vars.update_undefined (28)
-    \lstinputlisting{programs/def-or-set.txt}                                  & {
-        \begin{tabularx}{\linewidth}{r X}
-          85\%   & \lstinline|error| \\
-          **15\% & \lstinline|2|     \\
-        \end{tabularx}
-      }
-    \\\hline
-    \miscon{DefsCopyStructs} Variable definitions copy structures recursively. &
-    % vectors2.alias_with_defvar (47)
-    \lstinputlisting{programs/defs-copy-structs.txt}                           & {
-        \begin{tabularx}{\linewidth}{r X}
-          **67\% & \lstinline|#(100)| \\
-          30\%   & \lstinline|#(200)| \\
-          1\%    & \lstinline|#(300)| \\
-          1\%    & \lstinline|error|  \\
-        \end{tabularx}
-    }                                                                              \\\hline
-  \end{tabularx}
-  \caption{Ground misconceptions identified by the SMoL~Tutor. Answers marked
-    with ``**'' represent the misconception. (Part I)}
-  \label{t:tutor-grounded-misconceptions-1}
-\end{table}
-
-\begin{table}[!ht]
-  \centering
-  \begin{tabularx}{\linewidth}{X l X}
-    \textbf{Misconception}                                                                 &
-    \textbf{Question}                                                                      &
-    \textbf{Table of Answers}
-    \\\hline
-    \miscon{FlatEnv} There is only one environment, the global
-    environment. (This misconception is a kind of dynamic scope.)                          &
-    % scope.error_when_refer_to_undefined (9)
-    \lstinputlisting{programs/flat-env.txt}                                                & {
-        \begin{tabularx}{\linewidth}{r X}
-          96\%  & \lstinline|error| \\
-          **4\% & \lstinline|402|   \\
-        \end{tabularx}
-      }
-    \\\hline
-    \miscon{FunNotVal} Functions are \textit{not} considered
-    first-class values. They can't be bound to other variables, passed as arguments, or referred to by data structures.
-                                                                                           &
-    % lambda1.fun_as_parameter (62)
-    \lstinputlisting{programs/fun-not-val.txt}                                             & {
-        \begin{tabularx}{\linewidth}{r X}
-          83\%   & \lstinline|4|     \\
-          **11\% & \lstinline|error| \\
-          4\%    & \lstinline|2|     \\
-          1\%    & \lstinline|8|     \\
-        \end{tabularx}
-    }                                                                                          \\\hline
-    \miscon{IsolatedFun} Functions can't refer to free variables
-    except for the built-in ones.
-                                                                                           &
-    % scope.refer_global_is_possible (5)
-    \lstinputlisting{programs/isolated-fun.txt}                                            & {
-        \begin{tabularx}{\linewidth}{r X}
-          77\%   & \lstinline|3|     \\
-          **23\% & \lstinline|error| \\
-        \end{tabularx}
-      }
-    \\\hline
-    \miscon{NoCircularity} Data structures can't (possibly indirectly)
-    refer to themselves.                                                                   &
-    % vector2.warmup_circularity (53)
-    \lstinputlisting{programs/no-circularity.txt}                                          & {
-        \begin{tabularx}{\linewidth}{r X}
-          76\%   & \lstinline|3|                   \\
-          **14\% & \lstinline|error|               \\
-          9\%    & Run out of memory or time. \\
-          1\%    & \lstinline|+inf|                \\
-        \end{tabularx}
-    }                                                                                          \\\hline
-    \miscon{StructByRef} Data structures might refer to variables by their references.     &
-    % vector2.alias_var_in_mvec (49)
-    \lstinputlisting{programs/struct-by-ref.txt}                                           & {
-        \begin{tabularx}{\linewidth}{r X}
-          67\%   & \lstinline|#(1 2 3)| \\
-          **24\% & \lstinline|#(1 2 4)| \\
-          9\%    & error           \\
-        \end{tabularx}
-    }                                                                                          \\\hline
-    \miscon{StructsCopyStructs} Storing data structures into data structures makes copies. &
-    % lambda1.smol_quiz_circularity (58)
-    \lstinputlisting{programs/structs-copy-struct.txt}                                     & {
-        \begin{tabularx}{\linewidth}{r X}
-          65\%  & \lstinline|#0=#(#0# #0#)| (Racket circular object notation) \\
-          24\%  & \lstinline|#(#(2 3) #(2 3))|                                \\
-          **6\% & \lstinline|#(#(2 #(2 3)) #(2 3))|                           \\
-          6\%   & error                                                  \\
-        \end{tabularx}
-    }                                                                                          \\\hline
-  \end{tabularx}
-  \caption{Ground misconceptions identified by the SMoL~Tutor. Answers marked
-    with ``**'' represent the misconception. (Part II)}
-  \label{t:tutor-grounded-misconceptions-2}
-\end{table}
-
-\begin{table}[!ht]
-  \centering
-  \begin{tabularx}{\linewidth}{X l X}
-    \textbf{Misconception}                                                             &
-    \textbf{Question}                                                                  &
-    \textbf{Table of Answers}                                                              \\\hline
-    \surmisedMiscon{NestedDef} Sequences of definitions are
-    interpreted as if they are written in nested blocks. A definition
-    is not in the scope of later definitions.
-                                                                                       &
-    % program: scope.what_is_x_4 (13)
-    \lstinputlisting{programs/nested-def.txt}                                          & {
-        \begin{tabularx}{\linewidth}{r X}
-          92\%  & \lstinline|2| \\
-          **8\% & \lstinline|1| \\
-        \end{tabularx}
-    }                                                                                      \\\hline
-    \surmisedMiscon{Lazy} Expressions are only evaluated when their values are needed. &
-    % order.bad_order (20)
-    \lstinputlisting{programs/lazy.txt}                                                & {
-        \begin{tabularx}{\linewidth}{r X}
-          **57\% & \lstinline|1 3|   \\
-          43\%   & \lstinline|error| \\
-        \end{tabularx}
-    }                                                                                      \\\hline
-  \end{tabularx}
-  \caption{Surmised misconceptions identified by the SMoL~Tutor. Answers marked
-    with ``**'' represent the misconception.}
-  \label{t:tutor-surmised-misconceptions}
-\end{table}
-
-% \paragraph{Confirmed Misconceptions}
-
-% The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.
-
-% \paragraph{Added Misconceptions}
-
-% The misinterpreters helped me find misinterpretations that I had
-% overlooked. For instance, consider this program from \quizState:
-% \begin{lstlisting}
-% (defvar x 12)
-% (deffun (f x)
-%   (set! x 0))
-% (f x)
-% x
-% \end{lstlisting}
-% I had assumed the wrong answer
-% \texttt{0} is caused by \miscon{CallByRef}.
-% However, my misinterpreters made me realize it can also be explained
-% by \miscon{FlatEnv}. This could also explain why I see a difference in
-% error rate when the formal and actual parameter have the same
-% name (as mentioned in \Cref{s:misconceptions-quiz:state}).
-
-% Tutor's refined program sets and misinterpreters also helped me to
-% identify programs that detect one of the misconceptions but not the
-% other. Consider this new program
-% % mut-vars.not_aliased_by_defvar_1 (29)
-% \begin{lstlisting}
-%   (defvar x 12)
-%   (defvar y x)
-%   (set! x 0)
-%   x
-%   y
-% \end{lstlisting}
-% 12\% of students asserted that the following program produces
-% \texttt{0 0}, which suggest that they have the \textbf{BindByRef}
-% misconception. Students with a \miscon{FlatEnv} misconception would
-% have chosen \texttt{0 12}, which happens to be the correct answer.
-% (1\% wrote \textit{depends on
-% implementation}; 2\% asserted \texttt{error}; the remaining 85\%
-% asserted \texttt{0 12}).
-
-% The second program in \Cref{t:quizzes-data-fun} is already a good
-% detector for \miscon{FlatEnv}. SMoL Tutor includes an alternative
-% program that does not depend on equality checks.
-% % scope.error_when_refer_to_undefined (9)
-% \begin{lstlisting}
-%   (deffun (addy x)
-%     (defvar y 200)
-%     (+ x y))
-%   (+ (addy 2) y)
-% \end{lstlisting}
-% 4\% of students asserted that the following program produces
-% \texttt{402}, which suggest that they have the \miscon{FlatEnv}
-% misconception. Students with a \textbf{BindByRef} misconception would
-% have chosen \texttt{error}, which happens to be the correct answer.
-% (the remaining 96\% asserted \texttt{error}).
-
-\paragraph{A New Potential Misconception}
-
-For the following program, added in the Tutor:
-\begin{lstlisting}
-  (defvar y (+ x 2))
-  (defvar x 1)
-  x
-  y
-\end{lstlisting}
-56\% of students asserted that it produces
-\texttt{1 3}. Based on this, I surmise that students might have
-another misconception, which I define as \surmisedMiscon{Lazy}. (Recall that
-SMoL is eager, but even in many lazy languages, this would be an error.)
-
-\paragraph{Another Potential Misconception, and Its Effect on
-  Interpreting Descriptions.}%
-\label{ss:misinterpreters:submisconception}
-
-Consider this program from the SMoL Tutor:
-\begin{lstlisting}
-(defvar x 1)
-(deffun (main)
-  (deffun (get-x) x)
-  (defvar x 2)
-  (get-x))
-
-(main)
-\end{lstlisting}
-This program, suitably translated, would produce \lstinline|2| in a wide
-variety of languages (Python, JavaScript, Racket, Java, etc.), because
-\lstinline|get-x| and the inner \lstinline|x| are in the same scope block. The
-answer \lstinline|1| cannot be explained by any of my existing
-misconceptions. Based on this, I surmise a new misconception,
-\surmisedMiscon{NestedDef}. (Though its frequency falls below our
-threshold, my reading of answers suggests this may be more
-widespread, and I feel it needs to be investigated more.)
-
-Once I turned this into a misinterpreter, I found that it
-unexpectedly captured the program-answer pair for the following
-program from \quizFun---which should be an error, due to the
-double-binding of \lstinline|x| in the same scope block---and the answer
-\lstinline|2 0|:
-\begin{lstlisting}
-  (defvar x 0)
-  (defvar y x)
-  (defvar x 2)
-  x
-  y
-\end{lstlisting}
-Previously, I had interpreted this only as \miscon{DefOrSet}, because
-students had stated that the second \lstinline|(defvar x ....)|
-``mutates'' or ``redefines'' \lstinline|x|. This is reminiscent of the
-behavior of languages like Python, which use the same syntax both for
-binding new variables and for mutating existing ones.
-
-The problem here is that the word ``redefine'' underspecifies how the
-second definition is interpreted. I had interpreted it as
-\textit{mutating} the binding established by the first definition,
-which fits \miscon{DefOrSet}. However, another possibility is that
-it \emph{shadows} the binding (i.e., establishes a new scope block). We
-did not recognize that the original misconception is underspecified
-until I uncovered the new (surmised) misconception.
-
-\paragraph{Confirmed Misconceptions}
-
-The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.
-
-% \paragraph{Summary}
-
-% To summarize, I used the SMoL Tutor, with an expanded set of
-% programs, to investigate student misconceptions. I implemented the
-% idea of misinterpreters to help me properly classify student
-% performance. I were able to reconfirm all the previously identified
-% misconceptions, refine some of them, and also identify potential new
-% ones (that need further investigation).
-
-\section{Is The Tutor Effective?}
-\label{s:effect-tutor}
-
-\begin{figure}[p]
-
-  \includegraphics[width=\linewidth]{images/Misconceptions_Decay.png}
-
-  \caption{How many students chose a wrong answer that (uniquely)
-    represents a misconception? (Downward tendency suggests improvement
-    over time.)}
-  \label{f:tutor-trend-plots}
-
-\end{figure}
-
-Recall that the SMoL Tutor is not only a collection of MCQs: it is
-also a \emph{tutor}! So far, I have investigated the value of the
-MCQs. Now I examine its tutoring aspect. Concretely, I ask:
-\begin{center}
-
-  How effective is the Tutor at correcting each misconception?
-
-\end{center}
-To study this, I perform the following analysis per
-misconception. Using misinterpreters, I identify those questions
-whose wrong answers fit only one misconception (i.e., only one output
-matches that produced by that misinterpreter, and it matches only that
-misinterpreter). I then examine student performance over time across
-those questions. This would let me examine how they do on just that
-topic, in isolation, over time.
-
-I present the result of this analysis in two forms. Graphically, we
-show plots in \Cref{f:tutor-trend-plots}. Each figure shows the percentage of students who
-chose the answer corresponding to that misconception. Ideally, we
-would like to see these percentages diminish.
-
-Indeed, that is what I see in most of the graphs. The exceptions are
-\miscon{CallsCopyStructs}, \surmisedMiscon{NestedDef}, and
-\miscon{StructsCopyStructs}, which have only one problem (and hence no
-trend), and \miscon{DefOrSet} and \miscon{FunNotVal}, which show an
-increase. The lack of improvement for \miscon{FunNotVal} is
-unsurprising because the Tutor does not explicitly address this issue,
-focusing on closures created by \lstinline|lambda|s rather than named
-functions. (However, this does not explain the increase!)
-
-I also perform a logistic regression to see whether these
-improvements are significant (at a $p < 0.05$ threshold); details are
-in \supplRef{Paper.html}. Of the 11:
-\begin{itemize}
-
-  \item Of the nine seemingly improved (i.e., decreased)
-        misconceptions:
-        \begin{itemize}
-          \item Two are \emph{not} significant: \miscon{CallByRef}
-                (\pValue\ = .074); \miscon{NoCircularity} (\pValue\ = .075).
-          \item The other seven \emph{are} significant.
-        \end{itemize}
-  \item However, the two with an increasing trend (\miscon{DefOrSet}
-        and \miscon{FunNotVal}) are \emph{also} significant.
-\end{itemize}
-
-These data broadly suggest that the Tutor is a net positive.
-Ultimately, however, what these data really show is a need for
-improvement in the Tutor. When designing the Tutor questions, many
-more were intended to be representative of single
-misconceptions. However, as noted in \Cref{s:misinterp-intro}, it is
-easy to be incomplete (or incorrect) in ascribing
-misconceptions. Furthermore, as their set grows, it is difficult to
-reassess all the problems manually. Once evaluated using
-misinterpreters, I found far fewer problems than I would have liked.
-Concretely, only 40 of the 71 eligible problems (after removing the
-non-SMoL modules) were useful in the above
-analysis.
-
-I therefore view the above data as
-purely formative: they suggest that the Tutor \emph{most likely did
-  not do harm} and perhaps even \emph{may have done some
-  good}. However, it would be improper to read too much into the
-analysis. It is quite possible that some of the other problems would
-have found issues. Rather, what I really see is value in the
-misinterpreter concept. It is not only useful for analysis, it is also
-valuable for \emph{problem design}: in the next iteration of the
-Tutor, I will use misinterpreters actively to shape the incorrect answers and,
-as necessary, update the programs as well. I therefore hope to have
-much more thorough analyzes in future work.
-
-\section{Performance on Other Populations}
-\label{s:generality-tutor}
-
-There is, of course, a significant danger that the data above have
-been overfitted to University 1, thereby actually reflecting the state
-of its curriculum rather than some greater truth about program
-behavior understanding.
-
-% I already have some reason to believe this is
-% not the case: the related work discussed in
-% \Cref{s:background-misconceptions} is drawn from many institutions in
-% multiple countries with different educational preparations, levels,
-% and demographics. Nevertheless, that gives me only limited information
-% about the specific questions and misconceptions described above.
-
-Fortunately, I was able to deploy the Tutor on two other
-populations: University 2 and Texbook (See \Cref{s:populations} for their details).
-% \begin{itemize}
-
-%   \item University 2 is a primarily public university in the US. It is
-%         one of the largest Hispanic-serving institutions in the country. As
-%         such, its demographic is extremely different from those whose data
-%         were used above. The Tutor was used in one course in Spring 2023,
-%         taken by 12 students.
-%         The course is a third-year, programming
-%         language course. The students are required to have taken two
-%         introductory programming courses (C++ focused).
-
-%   \item A separate instance of the Tutor was published on the
-%         website of a programming languages textbook [ANON]. Over the course
-%         of 8 months, 597 people started with the first module and 103
-%         users made it to the last one. To protect privacy, I intentionally
-%         do not record demographic information, but I conjecture that the
-%         population is largely self-learners  (who are known to use the
-%         accompanying book),
-%         including some professional
-%         programmers. It is extremely
-%         unlikely to be the students from either university, because they
-%         would not get credit for their work on the public instance; they
-%         needed to use the university-specific instance. Furthermore, since
-%         they were not penalized for wrong answers, it would make little
-%         sense to do a ``test run'' on the public instance. Finally, I note
-%         that there is no overlap between the dates of submission on the
-%         public instance and the semester at University 1.
-
-% \end{itemize}
-% These two populations are therefore at least somewhat different from
-% the original population, and help me assess whether the problems we
-% identify are merely an artifact of the first institution.
-
-To evaluate, I computed a Spearman's rank correlation $\rho$, ranking
-questions by what percentage of students got the question
-right. Between the original university and University 2,
-we obtain a \pValue~=~2.013e-07. Between the original university and
-the online population, I obtain a \pValue~<~2.2e-16. These show that
-the other two populations performed similarly to the original one.
-While
-further validation on other populations remains essential, these
-suggest that the questions are \emph{finding misconceptions that may
-  be universal}.
-
-% \section{Threats to Validity}
-% \label{s:threats}
-
-% \subsection{Construct Validity}
-
-% Did I measure the right thing?
-% While my goal is to study student understanding of language
-% behavior, what I actually measure is performance on MCQs on select
-% programs. MCQs as a mechanism introduce various clear biases, though
-% I add the ``Other'' options to somewhat alleviate them. The small
-% size, syntactic details (such as intentionally meaningless variable
-% names), and other aspects of the programs may also impact my ability
-% to measure understanding. (This could go both ways: students may have
-% no trouble understanding behavior in a small program but may struggle
-% to do so in a larger one.)
-
-% A particularly notable threat is the use of Lispy
-% syntax. I chose it for two reasons: both because it helps clarify
-% scope (\Cref{s:smol}) and because the rest of the course used
-% Racket. However, students may well perform differently with more
-% familiar syntaxes. It seems unlikely their performance would be \emph{too}
-% different given the many languages that have produced similar
-% misconceptions (\Cref{t:rel-work}). Nevertheless, I intend to use a
-% variety of syntaxes to examine this issue further.
-
-% Finally, I curated programs by hand, so they may well reflect my own
-% biases about misconceptions. It may be possible to mitigate this
-% problem by synthesizing MCQ programs using the misinterpreters.
-
-% \subsection{Internal Validity}
-
-% Is my reasoning valid? I have applied standard techniques and
-% measurements for evaluating student responses. However, our
-% instruments still lack the validity of a proper concept
-% inventory. Their creation requires heavyweight processes (such as
-% Delphi methods \citep{goldmanSettingScopeConcept2010}) that require many hours of expert attention as
-% well as conversations with learners, and can hence be prohibitive in
-% cost. The Quizius method (\Cref{s:quizius}) was created precisely to
-% be an inexpensive method that provides a good proxy.
-
-% Ultimately, my goal is to provide a reasonable instrument for
-% widespread use. While the set of all misconceptions could be
-% unbounded, I believe my tasks, especially as embedded in the Tutor,
-% provide a good starting point for others. In particular, if students
-% select a wrong answer, that is still of \emph{some} use to an
-% educator, even if the precise misconception cannot be pinned down with
-% the highest accuracy. I therefore believe my instruments, and our
-% misinterpreter technique, are of general value.
-
-% A failure in my Tutor's logging infrastructure led me to miss some
-% responses. These are unlikely to be task-specific because the Tutor
-% has a generic framework that should perform the same across
-% tasks. Moreover, on average only $0.4\%\ (\textrm{sd} = 0.6\%)$ of values are missing. Therefore, I do not believe
-% this had a noticeable impact. (In addition, every wrong answer is
-% still wrong! I may just not have exactly the right proportions of
-% them.)
-
-% Because I don't randomize Tutor question order, it is hard to tell
-% whether the general decreasing trend in \cref{f:tutor-trend-plots}
-% shows misconceptions being corrected, or shows properties of the
-% questions (e.g., maybe later questions in the plots are easier because
-% their programs happen to be shorter).
-
-% \subsection{External Validity}
-
-% How well do my results generalize? Certainly there is reason to
-% question whether my results would apply to other
-% populations. \Cref{s:generality-tutor} provides preliminary evidence
-% that the results are not specific to one institution, and
-% \Cref{s:background-misconceptions} suggests these issues are widespread. Nevertheless,
-% much more broad testing is needed to confirm my specific instruments.
-
-% The other major concern is the tie to Lispy syntax. Learners in other
-% settings may do worse or even better with it. Building a Tutor that
-% supports multiple, and more traditional, syntaxes should help address
-% this issue. I defer this to future work (\cref{s:future-work}).
-
-\section{Related Work}%
-\label{s:related-work}
-
-\subsection{Program Behavior}
-
-Our idea of ``goal sentences'' is not substantially different from the
-\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
-I only used these sentences (or rules) to guide the
-design of the Tutor and as text in the Tutor itself.
-\cite{duranRulesProgramBehavior2021} argue for other uses of such
-sentences.
-
-\subsection{Misinterpreters}
-
-Our idea of misinterpreters is related to mystery languages
-\citep{diwanPLdetectiveSystemTeaching2004,
-  pombrioTeachingProgrammingLanguages2017}. Both approaches use
-evaluators that represent alternative semantics to the same syntax.
-However, the two are complementary.
-In mystery languages, instructors design the space of semantics with
-pedagogic intent, and
-students must create programs to explore that space. Misinterpreters,
-in contrast, are driven by student input, while the programs are provided
-by instructors. The two approaches also have different goals:
-mystery languages focus on encouraging students to experiment with
-languages; misinterpreters aim at capturing students' misconceptions.
-
-\section{Future Work \& Timeline}%
-\label{s:future-work}
-
-This semester (Fall 2023) I am already deploying a newer SMoL Tutor at
-Brown. This new tutor is able to display programs in JavaScript and
-Python, in addition to the Lispy syntax. The goal of this semester is
-to compare the new data with the current results, i.e., a comparison
-between uni-syntax tutor and multi-syntax tutor.
-
-By the end of Fall 2023, I would like to improve the SMoL Tutor:
-\begin{enumerate}
-  \item Ask students to explain when they gave a wrong answer in an
-        interpreting question. With student explanations, I check whether
-        choosing the wrong answer indicates holding the corresponding misconception(s).
-  \item Adjust the collection of interpreting questions such that
-        every misconception has at least two questions that aim at
-        correcting the misconception.
-  \item Partially randomize the order of questions.
-\end{enumerate}
+
+\subsection{Explore More Pedagogic Techniques}
+
+I have learned (\Cref{s:pedagogic-techniques}) that refutation text,
+case comparison, notional machines, and language levels are helpful. I
+plan to explore more literature and improve the tutor if I find new
+helpful techniques.
+
+\subsection{Expand Populations}
 
@@
 using the Tutor in Spring 2024. My goal is to collect data from at
-least two other institutions. I am officially taking a personal leave
-in Spring 2024, so I do not plan to make more progress other than
-data collection.
+least two other institutions.
+
+\subsection{Further Plans}
+
+I am officially taking a personal leave to do military service in my
+country dring Spring 2024, so I do not plan to make more progress
+other than data collection.
 

### 2023-11-09 proposal.tex
grammar check

@@
 I plan to change the tutor such that it collects data to uncover these
-unknown misconceptions and to confirm known ones.
+unknown misconceptions and confirm known ones.
 
@@
 The preliminary tutor provides a ``Copy This Program'' button that
-help students to run the program in DrRacket, where a SMoL
+helps students to run the program in DrRacket, where a SMoL
 implementation is provided.
@@
 This approach is inconvenient both for the students and for me. For
-students, they have to click, pastes and run, which is less convenient
+students, they have to click, paste, and run, which is less convenient
 than a single click. For me, because students run program in DrRacket
 rather than ``within'' the tutor, I do not know whether they have run
-the programs. This limit my ability to claim that the tutor is using
-NM to help teaching.
+the programs. This limits my ability to claim that the tutor is using
+NM to help teach.
 
@@
 program before giving an answer. I don't know yet how often this
-happens. But given that students receive penalty (\Cref{s:penalty})
+happens. But, given that students receive a penalty (\Cref{s:penalty})
 for giving wrong answers, they might do so when the program is overly
@@
 
-The tutor is adding extra options in interpreting tasks in an ad hoc
+The tutor is adding extra options to interpreting tasks in an ad hoc
 way. Eventually, I plan to generate options by moving and/or copying
@@
   \item The tutor currently does not ask students to explain when they
-        gave a wrong answer in an interpreting question. This limit my
+        gave a wrong answer in an interpreting question. This limits my
         ability to find new misconceptions and to confirm that the
@@
         misconception. I plan to adjust the collection of interpreting
-        questions such that every misconception has at least two
-        questions that aim at correcting the misconception.
-  \item The tutor is not randomizing the order of questions. This make
+        questions so that every misconception has at least two
+        questions that aim to correct the misconception.
+  \item The tutor is not randomizing the order of questions. This makes
         it difficult to compare questions for one misconception --
@@
   \item After randomizing question orders, I would need to check
-        carefully the dependencies between explanations and to
+        carefully the dependencies between explanations and
         resolve these dependencies, if any.
-  \item The tutor is using many variable names, some are separated by
-        dash (e.g., \lstinline|get-x|). Students might misread these
+  \item The tutor uses many variable names, some are separated by dashes
+        (e.g., \lstinline|get-x|). Students might misread these
         variables. For example, some students might misread
@@
         named \lstinline|get| is called with one argument
-        \lstinline|x|). These misreading add noise to the data. So I plan
-        to use consistent and simple variable names in all programs.
+        \lstinline|x|). These misreadings add noise to the data. So I
+        plan to use consistent and simple variable names in all
+        programs.
 \end{enumerate}
@@
 I am officially taking a personal leave to do military service in my
-country dring Spring 2024, so I do not plan to make more progress
+country in Spring 2024, so I do not plan to make more progress
 other than data collection.
 
-During Fall 2024, I plan to collect a new round of data from Brown. I
-expect these datasets and the datasets from Spring 2024 would give a
-more rigorous evaluation of the effectiveness of SMoL Tutor. The
-improvements to be made in Fall 2023 should resolve many of the
-threats to validity. In this semester I also want to write a new paper
-about the SMoL Tutor.
+During Fall 2024, I plan to collect a new round of data from Brown, analyze
+data that I collected in Spring 2024, and hopefully write a paper about the
+updates.
 

### 2023-11-09 proposal.tex
clean up high-level structures

@@
 
+\clearpage
+
+\tableofcontents
+
+\clearpage
+
 \section{Introduction}
@@
 
-\section{Plan for \textbf{RQ1}}%
+\section{Plan for \textbf{RQ1}: Find Misconceptions}%
 \label{s:plan-rq1}
@@
 
-\subsection{Populations}%
+\subsection{In What Populations to Find?}%
 \label{s:populations}
@@
 
-\subsection{Find Misconceptions in Student-written Programs}%
+\subsection{Find Misconceptions by Analyzing Student-written Programs}%
 \label{s:find-misconceptions-with-quizius}
@@
 
-\subsection{Find Misconceptions with Expert-written Programs}%
+\subsection{Find Misconceptions by Testing Students with Expert-written Programs}%
 \label{s:find-misconceptions-with-tutor}
@@
 plan to do the same things for any new misconceptions. I call these
-interpreters \emph{misinterpreters}.
-
-\section{Plan for \textbf{RQ2}}%
+interpreters \emph{misinterpreters}. Having definitional interpreters
+makes it convenient (and less error-prone) to associate wrong answers 
+with misconceptions (\Cref{s:misconceptions-tutor}).
+
+\section{Plan for \textbf{RQ2}: Correct Misconceptions}%
 \label{s:plan-rq2}
@@
 \end{description}
+
+To answer this RQ, I plan to create a tutor (\Cref{s:design-tutor}) and
+argue that the tutor is effective (\Cref{s:evaluation}).
+
+\subsection{Design a Tutor for Correcting Misconceptions}%
+\label{s:design-tutor}
 
@@
 
-\subsection{Tutorials and Language Levels}%
+\subsubsection{Tutorials and Language Levels}%
 \label{s:tutorials}
@@
 and each tutorial covers one or more new language constructs
-(\Cref{t:tutorials}).
-
-Each tutorial might be delivered as multiple modules to avoid forcing
+(\Cref{t:tutorials}). Each tutorial might be delivered as multiple modules to avoid forcing
 students to continuously work for more than 30 minutes.
@@
 
-\subsection{Interpreting Tasks}%
+\subsubsection{Interpreting Tasks}%
 \label{s:interpreting-tasks}
@@
 multiple-choice questions (MCQs) that asks students to interpret
-program.
-
-\paragraph{The MCQ Options} The options include (among other options
+program. The options include (among other options
 that I will explain shortly) the correct answer and wrong answers that
@@
 
-\subsection{Refutation Text and Other Explanations}%
+\subsubsection{Design with Pedagogic Techniques Known to Work}
+
+\paragraph{Refutation Text and Other Explanations}%
 \label{s:refutationt-text}
-
 The tutor has been presenting a refutation text when a student chooses
@@
 
-\subsection{Case Comparison}%
+\paragraph{Case Comparison}%
 \label{s:case-comparison}
@@
 
-\subsection{Notional Machine}%
+\paragraph{Notional Machine}%
 \label{s:notional-machine}
@@
 
-\subsection{Collating Problems for the Tutor}%
+\subsection{Further Collating Problems, for the Tutor}%
 \label{s:tutor-refine-program-set}
@@
 
-  \item I removed programs of a ``Lispy'' nature (See
-        \Cref{s:reduce-syntactic-bias}).
+  \item I removed programs of a ``Lispy'' nature.
 
@@
 
-\subsection{Notional Machine}
+\subsection{Keep Exploring More Pedagogic Techniques}
+
+I have learned (\Cref{s:pedagogic-techniques}) that refutation text,
+case comparison, notional machines, and language levels are helpful. I
+plan to explore more literature and improve the tutor if I find new
+helpful techniques.
+
+\subsection{Implemented But Not Evaluated Improvements in the Tutor}
+
+\paragraph{Ease Access to the Notional Machine}
 
@@
 students, they have to click, paste, and run, which is less convenient
-than a single click. For me, because students run program in DrRacket
-rather than ``within'' the tutor, I do not know whether they have run
-the programs. This limits my ability to claim that the tutor is using
-NM to help teach.
+than a single click.
 
@@
 tutor to the latest population at Brown (i.e., University 1), but I
-have not analyzed the data, which I plan to do in Spring 2024.
-
-I have not changed the logging systems to log when students click the
-buttons. I plan to make the change before Spring 2024.
-
-Another issue with the current design is that students might run a
-program before giving an answer. I don't know yet how often this
-happens. But, given that students receive a penalty (\Cref{s:penalty})
-for giving wrong answers, they might do so when the program is overly
-complicated. This problem is more of an issue after the switch to a
-web-based notional machine because now the ``cost'' of running a
-program is cheaper. My plan is to change the tutor so that it provides
-a program-specific link \emph{only} after each interpreting task.
-
-\subsection{Reduce Syntactic Bias}
+have not analyzed the data, so the new tutor is not presented in
+the preliminary results.
+
+\paragraph{Reduce Syntactic Bias with Multiple Syntax}
 
@@
 potentially introduce biases to the found misconceptions. For example,
-current results find that some students think the following program
+when using SMoL Quizzes, I found that some students think the following program
 produces \lstinline|'(4 5)| rather than \lstinline|'(1 2)|.
@@
 \lstinline|n|?''. Although I have tried to remove programs like this
-from my instruments, there is still a risk of having biases in student
+from the tutor, there is still a risk of having biases in student
 errors. This semester (Fall 2023) I am already deploying a newer SMoL
@@
 
-\subsection{Use Case Comparisons More Consistently}
+\subsection{The Rest of Fall 2023: Improve the Tutor}
+
+\paragraph{Log Access to the Notional Machine}
+
+I have not changed the logging systems to log when students click the
+buttons to access the NM. I plan to make the change before Spring 2024.
+
+\paragraph{Limit Access to the Notional Machine}
+
+Another issue with the current design is that students might run a
+program before giving an answer. I don't know how often this
+happens. But, given that students receive a penalty (\Cref{s:penalty})
+for giving wrong answers, they might do so when the program is overly
+complicated. This problem is more of an issue after the switch to a
+web-based notional machine because now the ``cost'' of running a
+program is cheaper. My plan is to change the tutor so that it provides
+a program-specific link \emph{only} after each interpreting task.
+
+\paragraph{Use Case Comparisons More Consistently}
 
@@
 
-\subsection{Options in Interpreting Tasks}
+\paragraph{Use More Sensible Options in Interpreting Tasks}
 
@@
 
-\subsection{Other Changes in the Tutor}
-
-By the end of Fall 2023, I also want to improve the SMoL Tutor at
-least on the following aspects:
-\begin{enumerate}
-  \item The tutor currently does not ask students to explain when they
+\paragraph{Ask For Explanation on Choices}
+
+The tutor currently does not ask students to explain when they
         gave a wrong answer in an interpreting question. This limits my
@@
         interpreting task.
+
+\paragraph{Clean Up the Program Set}
         % With student explanations, I check whether
@@
         % misconceptions.
-  \item As I said in \Cref{s:effect-tutor}, for many of the
+As I said in \Cref{s:effect-tutor}, for many of the
         misconceptions, the tutor has too few questions. On the other
@@
         questions that aim to correct the misconception.
-  \item The tutor is not randomizing the order of questions. This makes
+
+\paragraph{Randomize the Order of Interpreting Tasks}
+
+The tutor is not randomizing the order of questions. This makes
         it difficult to compare questions for one misconception --
@@
         order of interpreting tasks within each tutorial module.
-  \item After randomizing question orders, I would need to check
+After randomizing question orders, I would need to check
         carefully the dependencies between explanations and
         resolve these dependencies, if any.
-  \item The tutor uses many variable names, some are separated by dashes
+
+\paragraph{Clean Up Variable Names}
+
+The tutor uses many variable names, some are separated by dashes
         (e.g., \lstinline|get-x|). Students might misread these
@@
         programs.
-\end{enumerate}
-
-\subsection{Explore More Pedagogic Techniques}
-
-I have learned (\Cref{s:pedagogic-techniques}) that refutation text,
-case comparison, notional machines, and language levels are helpful. I
-plan to explore more literature and improve the tutor if I find new
-helpful techniques.
-
-\subsection{Expand Populations}
+
+\subsection{Spring 2024: Collect More Data}
 
@@
 
-\subsection{Further Plans}
-
 I am officially taking a personal leave to do military service in my
@@
 
+\subsection{Summer 2024: Data Analysis and Write a Paper}
+
+During this time, I plan to analyze the data I collected in Fall 2023 
+and Spring 2024 and start working on paper about the changes in the tutor.
+
+\subsection{Fall 2024: Collect More Data and Write a Paper}
+
 During Fall 2024, I plan to collect a new round of data from Brown, analyze
-data that I collected in Spring 2024, and hopefully write a paper about the
-updates.
+data that I collected in Spring 2024, and hopefully finish writing the new paper.
+
+\subsection{Spring 2025: Defense}
 

### 2023-11-10 proposal.tex
group sections

@@
 
-%% abbrev. for quizzes
-\newcommand{\QuizFun}{Quiz 1}
-\newcommand{\QuizState}{Quiz 2}
-\newcommand{\QuizHof}{Quiz 3}
-\newcommand{\quizFun}{Quiz 1}
-\newcommand{\quizState}{Quiz 2}
-\newcommand{\quizHof}{Quiz 3}
-
 %% abbrev. for tutorials
@@
 two-year process to find (a) student-written programs that tend to
-trip up students and (b) incorrect responses to those programs
-(\Cref{s:gen-prob-quizius}), using a tool for the purpose
+trip up students and (b) incorrect responses to those programs using a tool for the purpose
 (\Cref{s:quizius}). I have been cleaning up those programs
-(\Cref{s:smol-quizzes,s:tutor-refine-program-set}) and confirming that
-these programs are still tricky in the same way
-(\Cref{s:misconceptions-quiz,s:misconceptions-tutor}).
+(\Cref{s:collating-problems}) and confirming that
+these programs are still tricky in the same way.
 \Cref{s:misconceptions-tutor} presents the final list of
@@
 
-\section{Generating and Collating Problems}%
-\label{s:gen-prob-quizius}
-
-% Additionally, in this paper, I intentionally use the word
-% mis\emph{conceptions} rather than mis\emph{take}. A mistake can happen
-% for any reason (e.g., selecting the wrong answer from a menu). A
-% misconception, however, implies a conceptual problem: the student has
-% formed an incorrect concept in their head. For instance, they may
-% think that mutable structures are copied on function calls, or that
-% scope is resolved dynamically. This requires probing what they are
-% thinking.
-
-% All these, however, add up to a somewhat challenging demand. I want
-% to produce a list of questions (each one an MCQ) such that
-% \begin{enumerate}
-
-%   \item I can get past the expert blind spot,
-
-%   \item I have a sense of what misconceptions students have, and
-
-%   \item I can generally associate wrong answers with specific
-%         misconceptions, approaching a concept inventory.
-
-% \end{enumerate}
-
-This section describes preliminary results on finding misconceptions
-from student-written programs (\Cref{s:quizius}) and what I have done
-to distill misconception-detecting programs from student inputs
-(\Cref{s:smol-quizzes,s:tutor-refine-program-set}).
+\section{Preliminary Results}
 
@@
 
-My main solution to the expert blind spot is to use the Quizius system
+As said in \Cref{s:plan-rq1}, my main solution to the expert blind spot is to use the Quizius system
 \citep{saarinenHarnessingWisdomClasses2019}. In contrast to the very
@@
 \subsection{Collating Problems}
-\label{s:smol-quizzes}
+\label{s:collating-problems}
 
@@
 
-\end{itemize}
-and so on.
-I therefore manually curated the Quizius output
-to address these issues.
-
-% \begin{figure}[t]
-%   \centering
-%   \includegraphics[width=\linewidth]{images/smol-quizzes-example.png}
-%   \caption{Screenshot of a SMoL Quiz question.}
-%   \label{f:smol-quizzes-example}
-% \end{figure}
-
-\subsection{The SMoL Quizzes}%
-\label{s:misconceptions-quiz}
-
-Having curated the output, I had to confirm that these programs were
-still effective! That is, they needed to actually find student errors.
-
-I therefore turned the programs into a set of quizzes (in the US
-sense: namely, a brief test of knowledge) that I call the SMoL
-Quizzes.
-% There were three quizzes, ordered by linguistic
-% complexity. The first consisted of only basic operators and
-% first-order functions. The second added variable and structure
-% mutation. The third added \lstinline|lambda| and higher-order functions.
-Each quiz is a sequence of MCQs. The options are from the Quizius
-data. In addition, there is an ``Other'' option. If chosen, the quiz
-gave the user a text box with the caption ``Please specify''. The goal
-here was to record any other answers, which in turn might lead to
-fresh misconceptions.
-
-The quizzes were administered in the third year in University 1.
-
-% Question orders were partially randomized. I wanted students to get
-% some easy, warm-up questions initially, so those were kept at the
-% beginning. Similarly, I wanted programs that are syntactically
-% similar to stay close to each other in the quiz. This is so that, when
-% students got a second such program, they would not have to look far to
-% find the first one and confirm that they are indeed (slightly)
-% different, rather than wonder if they were seeing the same program
-% again.
-
-% Students only received feedback after having completed a whole quiz.
-% At the end of each quiz, they received both summative feedback
-% \emph{and} a refutation text that explained every program that
-% appeared in the quiz. Students were also encouraged to run the
-% programs, but I have little reason to believe that they did (and
-% certainly they asked few questions on the class forum about them).
-
-% Due to space limitations, I present the entire instrument in
-% \supplRef{SMoL Quizzes/instrument}. Here I focus on a few programs
-% where student choices correspond to misconceptions identified earlier,
-% thereby also showing that the curated programs are still effective. I
-% use a * to indicate the correct answer (which also matches the
-% implementation's output).
-
-% Syntactically, \lstinline|#t| and \lstinline|#f| are true and false,
-% while \lstinline|#(...)| is a vector.
-
-% This quiz starts with two warm-up and not very interesting questions
-% and then other questions. The order of the other questions are fully
-% randomized.
-
-\begin{table}[t]
-  \centering
-  \begin{tabularx}{\linewidth}{l X}
-    \textbf{Questions}                       &
-    \textbf{Table of Answers}                    \\\hline
-    \lstinputlisting{programs/def_twice.txt} & {
-        \begin{tabularx}{\linewidth}{r X}
-          $^\dag$59\% & \lstinline|2 0|             \\
-          *32\%       & \lstinline|Error|           \\
-          4\%         & \lstinline|Other: 2 2|      \\
-          3\%         & ``Nothing is printed'' \\
-          2\%         & \lstinline|0 0|             \\
-        \end{tabularx}
-    }                                            \\\hline
-    \lstinputlisting{programs/dyn_scope.txt} & {
-        \begin{tabularx}{\linewidth}{r X}
-          *65\%       & \lstinline|Error| \\
-          $^\dag$29\% & \lstinline|42 #t| \\
-          6\%         & Other        \\
-        \end{tabularx}
-    }                                            \\\hline
-    % \lstinputlisting{programs/reassign.txt}  & {
-    %     \begin{tabularx}{\linewidth}{r X}
-    %       *53\%       & \lstinline|#(6 5) 5|                 \\
-    %       *20\%       & \lstinline|Error|                    \\
-    %       $^\dag$11\% & \lstinline|#(6 6) 5|                 \\
-    %       $^\dag$11\% & \lstinline|#(6 6) 6|                 \\
-    %       5\%         & ``Nothing is printed'' or Other \\
-    %     \end{tabularx}
-    % }                                            \\\hline
-  \end{tabularx}
-  \caption{Some questions and student answers from \quizFun. Each
-    table row is a program and the (relative) frequencies of
-    student answers. Answers that can be considered correct are marked
-    with \texttt{*}.}
-  \label{t:quizzes-data-fun}
-\end{table}
-\Cref{t:quizzes-data-fun} lists programs in \quizFun{} that I
-consider the most interesting. These data confirm the presence of
-scope-related misconceptions.
-% :
-% \begin{enumerate}
-%   \item At least 59\% of students incorrectly believe that a
-%         variable can be defined twice in one block.
-%   \item 29\% of students believe that a variable defined in a function
-%         will be available in the top-level (or global) environment
-%         after the function is called. That is, these students may have a
-%         dynamic scope misconception.
-%   \item 22\% (11\% + 11\%) of students believe variables themselves
-%         can be passed as arguments and redefined inside the function.
-%         They disagree on whether the redefinition persists after the
-%         function call.
-% \end{enumerate}
-
-% This quiz randomizes the order of all questions.
-
-\begin{table}[t]
-  \centering
-  \begin{tabularx}{\linewidth}{l X}
-    \textbf{Question}                           &
-    \textbf{Table of Answers}                       \\\hline
-    \lstinputlisting{programs/circularity.txt}  & {
-        \begin{tabularx}{\linewidth}{r X}
-          $^\dag$50\% & \lstinline|#(#(2 #(2 3)) #(2 3))| \\
-          *29\%       & \lstinline|x=#(x x)| or
-          something similar. Both \lstinline|(left x)| and \lstinline|(right x)|
-          are \lstinline|x| itself.                       \\
-          $^\dag$16\% & \lstinline|Error|                 \\
-          5\%         & Other                        \\
-        \end{tabularx}
-    }                                               \\\hline
-    % \lstinputlisting{programs/mvec_in_mvec.txt} & {
-    %     \begin{tabularx}{\linewidth}{r X}
-    %       *62\%       & \lstinline|#(#(100 2 3 4) #(100 2 3 4))| \\
-    %       $^\dag$26\% & \lstinline|#(#(1 2 3 4) #(100 2 3 4))|   \\
-    %       6\%         & \texttt{Error}                      \\
-    %       4\%         & \lstinline|#(#(1 2 3 4) #(1 2 3 4))|     \\
-    %       2\%         & Other                               \\
-    %     \end{tabularx}
-    % }                                               \\\hline
-    % \lstinputlisting{programs/vset_in_let.txt}  & {
-    %     \begin{tabularx}{\linewidth}{r X}
-    %       *59\%       & \lstinline|#(10)|  \\
-    %       $^\dag$36\% & \lstinline|#(123)| \\
-    %       5\%         & Other         \\
-    %     \end{tabularx}
-    % }                                               \\\hline
-    \lstinputlisting{programs/var_as_arg.txt}   & {
-        \begin{tabularx}{\linewidth}{r X}
-          *65\%       & \texttt{12} \\
-          $^\dag$31\% & \texttt{0}  \\
-          4\%         & Other       \\
-        \end{tabularx}
-    }                                               \\\hline
-    % \lstinputlisting{programs/setset.txt}       & {
-    %     \begin{tabularx}{\linewidth}{r X}
-    %       *59\%       & Other: \texttt{5 7}   \\
-    %       $^\dag$27\% & \texttt{6 7}          \\
-    %       11\%        & \texttt{5 5}          \\
-    %       2\%         & Other: \texttt{Error} \\
-    %       1\%         & Other: \texttt{5 6}   \\
-    %     \end{tabularx}
-    % }                                               \\\hline
-  \end{tabularx}
-  \caption{Some questions and student answers from \quizState. Each
-    table row is a program and the (relative) frequencies of
-    student answers. Answers that can be considered correct are marked
-    with \texttt{*}.}
-  \label{t:quizzes-data-state}
-\end{table}
-\Cref{t:quizzes-data-state} lists interesting programs after the
-addition of state. These data suggest the students have aliasing-related misconceptions.
-% :
-% \begin{itemize}
-%   \item Up to 50\% of students think vectors are copied rather than
-%         aliased.
-%   \item 16\% of students think trying to construct and print a
-%         self-referring vector would error. (This program is ambiguous:
-%         \emph{constructing} works in most SMoL languages, but \emph{printing}
-%         may well cause a problem. These identified ambiguities are
-%         addressed in \Cref{s:tutor-refine-program-set}.)
-%   \item 31\% of students think a variable is aliased by a parameter if
-%         the two variables have the same name. Perhaps interestingly,
-%         fewer students (27\%) think the variable aliasing would
-%         happen if the two variables had different names.
-% \end{itemize}
-
-\begin{table}[t]
-  \centering
-  \begin{tabularx}{\linewidth}{l X}
-    \textbf{Question}                                 &
-    \textbf{Table of Answers}                             \\\hline
-    \lstinputlisting{programs/fun_returns_lambda.txt} & {
-        \begin{tabularx}{\linewidth}{r X}
-          *82\%       & \lstinline|3|     \\
-          $^\dag$17\% & \lstinline|Error| \\
-          1\%         & Other        \\
-        \end{tabularx}
-    }                                                     \\\hline
-    \lstinputlisting{programs/fun_and_state_1.txt}    & {
-        \begin{tabularx}{\linewidth}{r X}
-          *74\%       & \lstinline|4|            \\
-          $^\dag$21\% & \lstinline|3|            \\
-          6\%         & Other: \lstinline|Error| \\
-        \end{tabularx}
-    }                                                     \\\hline
-    % \lstinputlisting{programs/fun_and_state_2.txt}    & {
-    %     \begin{tabularx}{\linewidth}{r X}
-    %       *88\%       & \lstinline|4|            \\
-    %       $^\dag$10\% & \lstinline|3|            \\
-    %       3\%         & Other: \lstinline|Error| \\
-    %     \end{tabularx}
-    % }                                                     \\\hline
-    \lstinputlisting{programs/counter.txt}            & {
-        \begin{tabularx}{\linewidth}{r X}
-          *62\%       & \lstinline|1 1 2 3 2| \\
-          $^\dag$34\% & \lstinline|1 1 1 1 1| \\
-          4\%         & Other            \\
-          1\%         & \lstinline|1 1 2 3 4| \\
-        \end{tabularx}
-    }                                                     \\\hline
-  \end{tabularx}
-  \caption{Some questions and student answers from \quizHof. Each
-    table row is a program and the (relative) frequencies of
-    student answers. Answers that can be considered correct are marked
-    with \texttt{*}.}
-  \label{t:quizzes-data-hof}
-\end{table}
-\Cref{t:quizzes-data-hof} lists interesting programs after the
-addition of closures and higher-order functions. It extends what we
-saw with loops in \Cref{s:intro}: that students have misconceptions
-about their interaction with mutable variables.
-
-\subsection{Further Collating Problems, for the Tutor}%
-\label{s:tutor-refine-program-set}
-
-The SMoL Tutor's interpreting questions are the final instrument of
-this paper. I bootstrap it from the SMoL Quizzes, but made the
-following alterations:
-\begin{enumerate}
-
-  % fill gaps in goal sentences
   \item I felt the existing programs were insufficient to cover the
-        ideas I wanted students to work through. For instance, SMoL
-        Quizzes did not alias vectors using \lstinline|defvar|, so we
+        ideas I wanted students to work through. For instance, I found no program that aliases vectors using \lstinline|defvar|, so we
         added this program:
@@
 
-\end{enumerate}
-Most of these steps either modify or elide programs. Two cases,
-filling gaps in light of the goal sentences and resolving ambiguities,
-introduce programs. Starting with the 37 programs in the SMoL Quizzes,
-I added 52 more programs to arrive at a total of 89.
-
-In addition to generating this set of programs, I also modified the
-answer options. In addition to retaining the correct and incorrect
-answers from the SMoL Quizzes (and constructing my best guess of
-analogous incorrect answers for the new problems), I added more wrong
-answers to for the reasons given in \Cref{s:interpreting-tasks}. In
-general, I ended up increasing the number of choices substantially:
-only 8 out of the 89 questions have three or four choices; 70
-questions have 5--8 choices; and 11 questions have 9--14 choices.
-
-\section{Preliminary List of Misconceptions}%
+\end{itemize}
+
+\subsection{Preliminary List of Misconceptions}%
 \label{s:misconceptions-tutor}
@@
 
-I do not present the results of and \textbf{local}
-(\Cref{t:tutor-goal-sentences}), because they focus on constructs not
-usually found in non-Lispy languages. They were included in the Tutor
-to help students write programs for the course, but are not
-interesting from a SMoL perspective.
-
-% \subsection{Misinterpreters}
-% \label{s:misinterp-intro}
-
-% In principle, an expert can identify what misconceptions a particular
-% incorrect answer might correspond to. In practice, I found this
-% rather difficult for three reasons. First, with 89 programs, each with
-% several answers, it's easy to make mistakes. Second, my own expert blind
-% spot may prevent me from seeing an interpretation that would
-% lead to an additional association. Finally, and specific to my case,
-% since I had either curated or written all the programs and answers,
-% I was likely to miss some associations I had not intended.
-
-% To address this problem, I formalized misconceptions as
-% interpreters. That is, for each misconception I created a
-% corresponding \emph{misinterpreter}. This is an interpreter for the
-% SMoL syntax that intentionally has a semantic error corresponding to
-% that misconception. Put differently, a misinterpreter is a like a
-% ``definitional interpreter for a misconception''.
-
-% By running programs through the misinterpreters, I can more uniformly
-% and rigorously identify \emph{all} the misconceptions associated with
-% a wrong answer. Furthermore, if I identify a new misconception (or
-% alter a misinterpreter), it is easy to automatically re-classify all
-% the answers. By using misinterpreters, I indeed found new
-% interpretations for existing program-answer pairs.
-
-% \subsection{A Catalog of Misconceptions}
-
 I iteratively created our final catalog of misconceptions (from the
 perspective of the data in this paper). I started with misinterpreters
-representing the misconceptions described in
-\Cref{s:misconceptions-quiz}. These are misconceptions for which we
+representing the misconceptions for which we
 have reasonable validation (due to the prose in Quizius), so I call
@@
 
-% \paragraph{Confirmed Misconceptions}
-
-% The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.
-
-% \paragraph{Added Misconceptions}
-
-% The misinterpreters helped me find misinterpretations that I had
-% overlooked. For instance, consider this program from \quizState:
-% \begin{lstlisting}
-% (defvar x 12)
-% (deffun (f x)
-%   (set! x 0))
-% (f x)
-% x
-% \end{lstlisting}
-% I had assumed the wrong answer
-% \texttt{0} is caused by \miscon{CallByRef}.
-% However, my misinterpreters made me realize it can also be explained
-% by \miscon{FlatEnv}. This could also explain why I see a difference in
-% error rate when the formal and actual parameter have the same
-% name (as mentioned in \Cref{s:misconceptions-quiz:state}).
-
-% Tutor's refined program sets and misinterpreters also helped me to
-% identify programs that detect one of the misconceptions but not the
-% other. Consider this new program
-% % mut-vars.not_aliased_by_defvar_1 (29)
-% \begin{lstlisting}
-%   (defvar x 12)
-%   (defvar y x)
-%   (set! x 0)
-%   x
-%   y
-% \end{lstlisting}
-% 12\% of students asserted that the following program produces
-% \texttt{0 0}, which suggest that they have the \textbf{BindByRef}
-% misconception. Students with a \miscon{FlatEnv} misconception would
-% have chosen \texttt{0 12}, which happens to be the correct answer.
-% (1\% wrote \textit{depends on
-% implementation}; 2\% asserted \texttt{error}; the remaining 85\%
-% asserted \texttt{0 12}).
-
-% The second program in \Cref{t:quizzes-data-fun} is already a good
-% detector for \miscon{FlatEnv}. SMoL Tutor includes an alternative
-% program that does not depend on equality checks.
-% % scope.error_when_refer_to_undefined (9)
-% \begin{lstlisting}
-%   (deffun (addy x)
-%     (defvar y 200)
-%     (+ x y))
-%   (+ (addy 2) y)
-% \end{lstlisting}
-% 4\% of students asserted that the following program produces
-% \texttt{402}, which suggest that they have the \miscon{FlatEnv}
-% misconception. Students with a \textbf{BindByRef} misconception would
-% have chosen \texttt{error}, which happens to be the correct answer.
-% (the remaining 96\% asserted \texttt{error}).
-
 \paragraph{A New Potential Misconception}
@@
 unexpectedly captured the program-answer pair for the following
-program from \quizFun---which should be an error, due to the
+program --- which should be an error, due to the
 double-binding of \lstinline|x| in the same scope block---and the answer
@@
 
-\paragraph{Confirmed Misconceptions}
-
-The Tutor confirmed all the misconceptions from Quizius via the SMoL Quizzes.
-
-% \paragraph{Summary}
-
-% To summarize, I used the SMoL Tutor, with an expanded set of
-% programs, to investigate student misconceptions. I implemented the
-% idea of misinterpreters to help me properly classify student
-% performance. I were able to reconfirm all the previously identified
-% misconceptions, refine some of them, and also identify potential new
-% ones (that need further investigation).
-
-\section{What Pedagogic Techniques are Effective?}%
+\subsection{What Pedagogic Techniques are Effective?}%
 \label{s:pedagogic-techniques}
@@
 
-\section{The SMoL Tutor}%
+\subsection{The SMoL Tutor}%
 \label{s:tutor}
-
-% So far, I have focused on identifying problems. As noted earlier
-% (\Cref{s:smol-quizzes}), I provided students with refutation texts
-% after the quizzes, but it is unclear to what extent students read,
-% understood, or internalized these. I also wanted to determine whether
-% other populations run into these issues, and it was unclear whether we
-% would be able to administer passive quizzes to them.
-
-% Furthermore, while the quizzes may be a useful diagnostic, my goal is
-% not only to find faults but to improve the understanding of basic
-% programming language behavior. I believe (and presumably so does
-% anyone else who writes a formal semantics!)\ that an understanding of
-% these basic program behaviors is important, and even more so when it
-% comes to understanding concurrency, ownership, and other advanced
-% features. Even more simply, misunderstanding these clearly impacts
-% development and debugging time.
-
-% In response, I decided to create a \emph{tutor}: the SMoL Tutor. It is built
-% around my quiz instruments and the detected misconceptions. We
-% describe the Tutor from a user's perspective in \Cref{s:tutor}, and
-% explain how it was populated from the
-% SMoL Quizzes in \Cref{s:tutor-refine-program-set}.
-% I then discuss what
-% I learned from its data (\Cref{s:misconceptions-tutor}), and finally
-% evaluate its effectiveness as a \emph{tutor} (\Cref{s:effect-tutor}).
-
-% \subsection{The User Experience}
 
@@
 their heap content, which I do not cover in this paper.
-
-% Each tutorial consists of a sequence of questions. Most questions in
-% the Tutor are \newterm{interpreting questions}.\footnote{The Tutor
-% also asked students to perform some other activities, like showing
-% programs and asking for their heap content, which I do not cover in
-% this paper.} These questions (illustrated in
-% \Cref{f:smol-tutor-example}) are versions of the SMoL Quizzes
-% (obtained by the process described in
-% \Cref{s:tutor-refine-program-set}). In each of these questions,
-% students are shown a program and asked to predict the program's
-% running result(s) by answering an MCQ (with an ``Other'' option).
-% After making a choice, students receive feedback. If a student chooses
-% incorrectly, they are (a) given an explanation that is \emph{based on
-% the misconception associated with that wrong answer} (or a generic
-% one, if there is not a specific misconception), and then (b) asked to
-% answer a second question:
-% \begin{enumerate}
-
-%   \item The second question is semantically the same as the first, but
-%         with superficial changes (e.g., variable names, constants, and operators are
-%         changed) so that the student cannot immediately
-%         guess the answer.
-
-%   \item Instead of multiple-choice, students must \emph{type} the answer
-%         into a text box. (The Tutor normalizes text to accommodate variations.)
-%         This is
-%         intentional. First, I wanted to force reflection on the
-%         explanation just given, whereas with an
-%         MCQ, students could just guess. Second, I felt that students would
-%         find typing more onerous than clicking. In case students
-%         had just guessed on a question, my hope is that the penalty of having to
-%         type means, on \emph{subsequent} tasks, they
-%         would be more likely to pause and reflect before choosing.
-
-% \end{enumerate}
 
@@
 
-% Of the modules, \textbf{local} is the least portable across
-% languages. While local binding is of course present in other
-% languages, this is mostly covered using nested \lstinline|defvar|s in
-% earlier modules, starting with \textbf{scope}. The distinction central
-% to this module (between three local-binding constructs with slightly
-% different scopes) is primarily a focus of Lispy languages. Therefore,
-% I exclude this module from my analysis and instruments, and indeed
-% plan to deprecate the module in future versions.
-
-% \Cref{s:smol-quizzes} discusses the partial randomization employed by
-% the SMoL Quizzes. In contrast, the SMoL Tutor does \emph{not}
-% randomize question order. This is because the Tutor consists of more
-% than just questions: it also has explanatory text. This text is based
-% on the preceding problems. Authoring it is therefore somewhat like
-% writing a textbook, with complex dependencies that can easily be
-% broken.
-
-\section{How Effective Is the Tutor?}
+\subsection{How Effective Is the Tutor?}
 \label{s:effect-tutor}
@@
 
-\section{How Generalizable is My Observations?}
+\subsection{How Generalizable are My Results?}
 \label{s:generality-tutor}
@@
 potentially introduce biases to the found misconceptions. For example,
-when using SMoL Quizzes, I found that some students think the following program
+I found that some students think the following program
 produces \lstinline|'(4 5)| rather than \lstinline|'(1 2)|.

### 2023-11-10 proposal.tex
grammar check

@@
 % TODO
-% - clean up the background sections
+% - clean-up the background sections
 
@@
 
-%% abbrev. for referring to supplementary materials
-\newcommand{\supplRef}[1]{\texttt{#1} in the supplementary materials}
-
 \title{Dissertation Proposal}
@@
   by experts, and the design of an automated, self-guided tutoring
-  system. The tutor builds on strategies in the cognitive and
+  system. The Tutor builds on strategies in the cognitive and
   educational literature and is explicitly designed around identifying
   and correcting misconceptions. Preliminary results suggest (a) the
-  misconceptions I found are widespread, and (b) the tutor appears to
+  misconceptions I found are widespread, and (b) the Tutor appears to
   improve understanding.
@@
 when transferring knowledge from old languages to new ones. In
-recognition of this deep commonality, in this paper I choose to call
+recognition of this deep commonality, in this proposal I choose to call
 this the \emph{Standard Model of Languages} (SMoL).
@@
 
-This lead to my \textbf{research questions}:
+This led to my \textbf{research questions}:
 
@@
 sections detail my current progress.
-
-% To answer my research questions, I have (a) run a multi-year,
-% multi-phase study to identify these misunderstandings while trying to
-% work around the blind spots of experts (and also drawing on prior
-% work), and (b) constructed the SMoL Tutor, an interactive tutor to
-% address these misunderstandings. The SMoL Tutor is not a
-% \emph{programming} tutor; rather, it assumes the user has some basic
-% facility with programming (i.e., that they are already familiar with
-% concepts like vectors, mutation, and higher-order functions). It is
-% built around SMoL's features and their interactions, and draws on
-% cognitive and educational psychology concepts to expressly identify
-% and correct misconceptions.
-
-% Concretely, I have made the following contributions:
-% \begin{enumerate}
-
-%   \item I provide a collection of brief programs for detecting
-%         misconceptions about the features of SMoL and their
-%         interaction. These programs are short and most are readily
-%         translatable to a wide variety of programming languages that
-%         share the SMoL characteristics.
-
-%   \item I show that multiple populations have problems with these
-%         programs.
-
-%   \item I present a list of curated misconceptions generated
-%         after multiple rounds of analysis.
-
-%   \item I implement these misconceptions as interpreters, which
-%         find weaknesses in my manual analysis and hold potential for
-%         generative use in the future.
-
-%   \item I present a tutoring system, the SMoL Tutor, that measurably
-%         corrects these misconceptions.
-
-%   \item I present a notional machine, the Stacker, that is used by the
-%         SMoL Tutor to explain programs and might be useful in itsown.
-
-% \end{enumerate}
-
-% This document first presents the current results and ends with future
-% work \& timeline (\cref{s:future-work}).
-
-% \paragraph{A Note on Terminology} I use the term
-% ``behavior'' to refer to the meaning of programs in terms of the
-% answers they produce. A more standard term for this would, of course, be
-% ``semantics''. However, the term ``semantics tutor'' might
-% mislead some readers into thinking it
-% teaches people to read or write a formal semantics, e.g., an
-% introduction to ``Greek'' notation. Because that is not the kind of
-% tutor I are describing, to avoid confusion, I use the term
-% ``behavior'' instead.
 
@@
 
-SMoL defines a semantic but not the syntax. However, to repsent SMoL
+SMoL defines a semantic but not the syntax. However, to represent SMoL
 to students, I have to choose a syntax. I chose a Lispy syntax, which
@@
 Nevertheless, most SMoL programs are easy to translate to other
-languages. I present machine-translated Python and JavaScript versions
-of my programs in \supplRef{Translated\_Programs.html}. Furthermore, I
-intend to make a multi-lingual tutor in the future
-(\Cref{s:future-work}).
+languages. Furthermore, I intend to make a multi-lingual Tutor in the
+future (\Cref{s:future-work}).
 
@@
 inner loop iterates through UI events within a task. I follow the same
-structure in my tutor (\Cref{s:tutor}).
+structure in my Tutor (\Cref{s:tutor}).
 
@@
 \citep{andersonLISPTutor1985}), and in the process undoubtedly address
-some program behavior misconceptions. My tutor differs in a notable
+some program behavior misconceptions. My Tutor differs in a notable
 way: it does not try to teach programming per se. Instead, it assumes
 a basic programming background and focuses entirely on program
-behavior misconceptions and correcting them. I are not aware of a
+behavior misconceptions and correcting them. I am not aware of a
 tutoring system (in computer science) that has this specific design.
@@
 
-I have another source of misconceptions. I have been creating a tutor
+I have another source of misconceptions. I have been creating a Tutor
 to answer my RQ2, which I explain in detail in \Cref{s:plan-rq2}. The
-tutor includes many tasks that ask students to predict the result(s)
+Tutor includes many tasks that ask students to predict the result(s)
 of running a program. Although these programs originate from
@@
 
-I plan to change the tutor such that it collects data to uncover these
+I plan to change the Tutor so that it collects data to uncover these
 unknown misconceptions and confirm known ones.
@@
 programs, the wrong result that corresponds to the misconception should
-be commonly predicted by students from various population.
+be commonly predicted by students from various populations.
 Furthermore, these students should give similar explanations for all
@@
 
-To answer this RQ, I plan to create a tutor (\Cref{s:design-tutor}) and
-argue that the tutor is effective (\Cref{s:evaluation}).
+To answer this RQ, I plan to create a Tutor (\Cref{s:design-tutor}) and
+argue that the Tutor is effective (\Cref{s:evaluation}).
 
@@
 My current takeaway is that refutation text, case comparison, notional
-machines, and language levels are helpful. I have created a tutor that
+machines, and language levels are helpful. I have created a Tutor that
 incorporates all these ideas (\Cref{s:tutor}) and have collected data
@@
 
-I would like to first explain key aspects of the tutor
+I would like to first explain key aspects of the Tutor
 (\Cref{s:tutorials,s:interpreting-tasks}) and then my plan to
@@
 (\Cref{s:notional-machine}). Finally, I present my plan to evaluate
-the effectiveness of the tutor and how evaluation impacts the tutor
+the effectiveness of the Tutor and how evaluation impacts the Tutor
 design (\Cref{s:evaluation}).
@@
 There are many misconceptions to correct. To make the tutoring process
-manageable, I have designed the tutor as a sequence of tutorials.
+manageable, I have designed the Tutor as a sequence of tutorials.
 Inspired by prior research on language levels
@@
 
-The tutor is mostly consisted of \emph{interpreting tasks}:
+The Tutor mostly consists of \emph{interpreting tasks}:
 multiple-choice questions (MCQs) that asks students to interpret
@@
 correspond to known misconceptions. Because students might conceive
-wrong answers that I do not anticipate, the tutor has been providing
+wrong answers that I do not anticipate, the Tutor has been providing
 an ``Other'' option that, once chosen, allows students to enter
@@
 elimination. So I have been adding additional options that are not
-obviously wrong. The tutor is adding many options, but with an ad hoc
+obviously wrong. The Tutor is adding many options, but with an ad hoc
 rule. I plan to revise the rules of adding extra options.
@@
 \paragraph{Penalty to Wrong Choices}\label{s:penalty} To discourage
-students from making a random choice, the tutor has been giving
-penalty. Designing the strategy of assigning penalty is non-trivial:
+students from making a random choice, the Tutor has been giving
+penalties. Designing the strategy of assigning penalties is non-trivial:
 \begin{itemize}
@@
   However, grading by completion alone encourages students to rush
-  through the tutorials and hence might encourage students to make
+  through the tutorials and, hence, might encourage students to make
   random choices.
@@
 To encourage this feeling, I have been doing the following: when I
-deployed the tutor in my institution, University 1, I graded by
+deployed the Tutor in my institution, University 1, I graded by
 completion\footnote{I encourage collaborators to grade by
 completeness, but I have no real control over how they actually use
-the tutor.}; when a student fails a task, the tutor gives an
+the Tutor.}; when a student fails a task, the Tutor gives an
 \emph{extra} task:
@@
   \item Instead of multiple-choice, students must \emph{type} the
-        answer into a text box. (The tutor normalizes text to accommodate
+        answer into a text box. (The Tutor normalizes text to accommodate
         variations.) This is intentional. First, I want to force
@@
 \label{s:refutationt-text}
-The tutor has been presenting a refutation text when a student chooses
+The Tutor has been presenting a refutation text when a student chooses
 a wrong answer that corresponds to known misconception(s). Let's
@@
 answers \lstinline|0 0|, which suggests they hold the
-\miscon{DefByRef} misconception, the tutor responds
+\miscon{DefByRef} misconception, the Tutor responds
 \begin{quote}
@@
 \end{quote}
-If a student gave a wrong answer that does not correspond to a
-misconception (either by choosing one of the generated option or by
-choosing the ``Other'' option), the tutor presents a generic
+If a student gives a wrong answer that does not correspond to a
+misconception (either by choosing one of the generated options or by
+choosing the ``Other'' option), the Tutor presents a generic
 explanation. For example, if a student answers \lstinline|12 12|, the
-tutor responds
+Tutor responds
 \begin{quote}
@@
 \end{quote}
-If a student gave the correct answer, the tutor also presents the
+If a student gives the correct answer, the Tutor also presents the
 generic explanation.
@@
 
-The tutor provides opportunities for making case comparisons: each
+The Tutor provides opportunities for making case comparisons: each
 interpreting task presents a SMoL program and the program's result.
 These program-result pairs are concrete cases of the semantics of
-SMoL. To incorporate the idea of case comparison, the tutor currently
+SMoL. To incorporate the idea of case comparison, the Tutor currently
 prompts students to compare these cases and then to summarize the
@@
 Prior research argues (\Cref{s:pedagogic-techniques})that illustrating
-how programs runs with notional machines (NMs) is beneficial for
+how programs run with notional machines (NMs) is beneficial for
 students. I have been giving students opportunities and sometimes
-explicitly require them to run programs in a NM visualization tool.
+explicitly requiring them to run programs in the NM visualization tool.
 
@@
 
-To fully answer my research questions, I should also argue that my
-tutor is effective at correcting misconceptions. In this section, I
-describe how I want to refine the design of the tutor to help me make
+To fully answer my research questions, I should also argue that the
+Tutor is effective at correcting misconceptions. In this section, I
+describe how I want to refine the design of the Tutor to help me make
 the argument, what data I want to collect, and what (statistical)
@@
 
-Given this observation, I plan to set up the tutor such that
+Given this observation, I plan to set up the Tutor such that
 \begin{itemize}
@@
 \emph{and} students perform better in later tasks, I deduce that the
-tutor seems to be correcting misconceptions. If students perform
-worse in later tasks, the tutor is potentially doing harm.
+Tutor seems to be correcting misconceptions. If students perform
+worse in later tasks, the Tutor is potentially doing harm.
 
@@
 
-The current tutor is not prepared for this analysis yet. I describe how
-I plan to change the tutor in \Cref{s:future-work}.
+The current Tutor is not prepared for this analysis yet. I describe how
+I plan to change the Tutor in \Cref{s:future-work}.
 
@@
 I took over as experts. Determining which is the right
-answer is easy. Where expert knowledge is useful is in
+answer is easy, whereas expert knowledge is useful is in
 \emph{clustering the rationales}. If all the rationales for a wrong
@@
 
-The Quizius instrucment was set up by other people. I analyzed the
+The Quizius instrument was set up by other people. I analyzed the
 collected data.
@@
 
-  \item I resolved ambiguities in some programs, either adding
-        answer choices or even adding other questions to tease
+  \item I resolved ambiguities in some programs by adding
+        answer choices or even other questions to tease
         out different interpretations.
@@
 
-I now examine what I learned from the tutor's interpreting tasks in
+I now examine what I learned from the Tutor's interpreting tasks in
 terms of program behavior misconceptions.
@@
 I iteratively created our final catalog of misconceptions (from the
-perspective of the data in this paper). I started with misinterpreters
+perspective of the data in this document). I started with misinterpreters
 representing the misconceptions for which we
@@
 
-This section describe the current status of the tutor.
+This section describe the current status of the Tutor.
 
@@
 to perform some other activities, like showing programs and asking for
-their heap content, which I do not cover in this paper.
+their heap content, which I do not cover in this document.
 
@@
 the (abbreviated) goal sentences for each tutorial. I plan to keep
-goal sentences in the tutor.
+goal sentences in the Tutor.
 
@@
 % Recall that the SMoL Tutor is not only a collection of MCQs: it is
-% also a \emph{tutor}! So far, I have investigated the value of the
+% also a \emph{Tutor}! So far, I have investigated the value of the
 % MCQs. Now I examine its tutoring aspect. Concretely, I ask:
@@
 This section presents part of my preliminary answer to \textbf{RQ2}
-(\Cref{s:plan-rq2}): is the tutor effective?
+(\Cref{s:plan-rq2}): is the Tutor effective?
 
@@
 I also perform a logistic regression to see whether these
-improvements are significant (at a $p < 0.05$ threshold); details are
-in \supplRef{Paper.html}. Of the 11:
+improvements are significant (at a $p < 0.05$ threshold). Of the 11:
 \begin{itemize}
@@
 harm} and perhaps even \emph{may have done some good}. Revisions on
-the tutor (\Cref{s:future-work}), in light of misinterpreters, is
+the Tutor (\Cref{s:future-work}), in light of misinterpreters, is
 necessarily to give a more accurate evaluation. For example I found
@@
 case comparison, notional machines, and language levels are helpful. I
-plan to explore more literature and improve the tutor if I find new
+plan to explore more literature and improve the Tutor if I find new
 helpful techniques.
@@
 
-The preliminary tutor provides a ``Copy This Program'' button that
+The preliminary Tutor provides a ``Copy This Program'' button that
 helps students to run the program in DrRacket, where a SMoL
@@
 
-The latest tutor provides a button on the top-right of each program.
+The latest Tutor provides a button on the top-right of each program.
 If students click the button, they will open a new browser tab that
 shows the NM populated with the program. I have deployed the latest
-tutor to the latest population at Brown (i.e., University 1), but I
-have not analyzed the data, so the new tutor is not presented in
+Tutor to the latest population at Brown (i.e., University 1), but I
+have not analyzed the data, so the new Tutor is not presented in
 the preliminary results.
@@
 \lstinline|n|?''. Although I have tried to remove programs like this
-from the tutor, there is still a risk of having biases in student
+from the Tutor, there is still a risk of having biases in student
 errors. This semester (Fall 2023) I am already deploying a newer SMoL
-Tutor at Brown. This new tutor is able to display programs in
+Tutor at Brown. This new Tutor is able to display programs in
 JavaScript and Python, in addition to the Lispy syntax. I plan to
-compare in Spring 2024 the new tutor data with the current results,
-i.e., a comparison between uni-syntax tutor and multi-syntax tutor.
+compare in Spring 2024 the new Tutor data with the current results,
+i.e., a comparison between uni-syntax Tutor and multi-syntax Tutor.
 
@@
 web-based notional machine because now the ``cost'' of running a
-program is cheaper. My plan is to change the tutor so that it provides
+program is cheaper. My plan is to change the Tutor so that it provides
 a program-specific link \emph{only} after each interpreting task.
@@
 
-The tutor currently prompts students to compare program-output pairs
+The Tutor currently prompts students to compare program-output pairs
 and then to summarize the rules of program behavior at the end of
@@
 I plan to use a similar prompt in later tutorials. For example, in \textbf{mut-vars},
-I would want the tutor to ask
+I would want the Tutor to ask
 \begin{quote}
@@
 
-The tutor is adding extra options to interpreting tasks in an ad hoc
+The Tutor is adding extra options to interpreting tasks in an ad hoc
 way. Eventually, I plan to generate options by moving and/or copying
@@
 
-The tutor currently does not ask students to explain when they
-        gave a wrong answer in an interpreting question. This limits my
+The Tutor currently does not ask students to explain when they
+        give a wrong answer in an interpreting question. This limits my
         ability to find new misconceptions and to confirm that the
         tasks do detect the anticipated misconceptions. I plan to
-        change the tutor so that it asks for an explanation in each
+        change the Tutor so that it asks for an explanation in each
         interpreting task.
 
-\paragraph{Clean Up the Program Set}
+\paragraph{Clean-Up the Program Set}
         % With student explanations, I check whether
@@
 As I said in \Cref{s:effect-tutor}, for many of the
-        misconceptions, the tutor has too few questions. On the other
+        misconceptions, the Tutor has too few questions. On the other
         hand, there are many questions that do not produce useful data
@@
 
-The tutor is not randomizing the order of questions. This makes
+The Tutor is not randomizing the order of questions. This makes
         it difficult to compare questions for one misconception --
@@
 
-\paragraph{Clean Up Variable Names}
-
-The tutor uses many variable names, some are separated by dashes
+\paragraph{Clean-Up Variable Names}
+
+The Tutor uses many variable names; some are separated by dashes
         (e.g., \lstinline|get-x|). Students might misread these
@@
 During this time, I plan to analyze the data I collected in Fall 2023
-and Spring 2024 and start working on paper about the changes in the tutor.
+and Spring 2024 and start working on a paper about the changes in the Tutor.
 

### 2023-11-14 proposal.tex
Updates from Overleaf
title
Clean up

@@
 %% abbrev. for RQs
-\newcommand{\rqI}{What program-behavior misconceptions apply to SMoL and present (even) in students with prior programming backgrounds?}
-\newcommand{\rqII}{What would be a tutoring system that aims to correct the misconceptions, draws on cognitive, educational, or psychological concepts, and seems to be effective?}
+\newcommand{\rqI}{What program-behavior misconceptions apply to SMoL and present even in students with prior programming backgrounds?}
+\newcommand{\rqII}{How effective is a tutoring system driven by cognitive and educational principles at correcting these misconceptions?}
 
@@
 
-\title{Dissertation Proposal}
+\title{Identifying and Correcting Programming Language Behavior Misconceptions}
 \author{Kuang-Chen Lu}
@@
   the myriad ways in which students can misunderstand programs?
-  Furthermore, once identified, what can I do to correct them?
+  Furthermore, once identified, what can we do to correct them?
 
@@
 programmers to understand. Consider the following scenario: to create a
-calculator, I have to construct a callback that can be attached to
+calculator, we have to construct a callback that can be attached to
 each button. The following Python program seems to achieve the
@@
 (\Cref{s:background-tutoring-systems}), and then outline my research
-plans for the RQs (\Cref{s:plan-rq1,s:plan-rq2}). The remaining
-sections detail my current progress.
+plans for the RQs (\Cref{s:plan-rq1,s:plan-rq2}). After that, I describe preliminary results (\Cref{s:preliminary-results} and propose my next steps (\Cref{s:future-work}).
 
@@
 are chosen with great care. Each one has been validated so that if a
-student picks it, I can quite unambiguously determine \emph{what
+student picks it, we can quite unambiguously determine \emph{what
 misconception the student has}. For instance, if the question is
@@
 
-\section{Plan for \textbf{RQ1}: Find Misconceptions}%
+\section{\textbf{RQ1}: Find Misconceptions}%
 \label{s:plan-rq1}
@@
 \begin{description}
-  \item[University 1] students in a ``Principles of Programming
+  \item[University 1] includes students in a ``Principles of Programming
     Languages'' class at Brown University. The class has about 70--75
@@
     required, but students were graded on effort, not correctness.
-  \item[University 2] a primarily public university in the US. It is
+  \item[University 2] includes students at a primarily public university in the US. It is
     one of the largest Hispanic-serving institutions in the country. As
@@
     introductory programming courses (C++ focused).
-  \item[Textbook] an instrument was published on the website of a
+  \item[Textbook] includes visitors to the website of a
     programming languages textbook. Over the course of 8 months,
@@
 
-\section{Plan for \textbf{RQ2}: Correct Misconceptions}%
+\section{\textbf{RQ2}: Correct Misconceptions}%
 \label{s:plan-rq2}
@@
 
-\section{Preliminary Results}
+\section{Preliminary Results}%
+\label{s:preliminary-results}
 
@@
 % this issue. I defer this to future work (\cref{s:future-work}).
-
-\section{Related Work}%
-\label{s:related-work}
-
-\subsection{Program Behavior}
-
-Our idea of ``goal sentences'' is not substantially different from the
-\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
-I only used these sentences (or rules) to guide the
-design of the Tutor and as text in the Tutor itself.
-\cite{duranRulesProgramBehavior2021} argue for other uses of such
-sentences.
-
-\subsection{Misinterpreters}
-
-Our idea of misinterpreters is related to mystery languages
-\citep{diwanPLdetectiveSystemTeaching2004,
-  pombrioTeachingProgrammingLanguages2017}. Both approaches use
-evaluators that represent alternative semantics to the same syntax.
-However, the two are complementary.
-In mystery languages, instructors design the space of semantics with
-pedagogic intent, and
-students must create programs to explore that space. Misinterpreters,
-in contrast, are driven by student input, while the programs are provided
-by instructors. The two approaches also have different goals:
-mystery languages focus on encouraging students to experiment with
-languages; misinterpreters aim at capturing students' misconceptions.
 
@@
 \end{quote}
-This prompt has a few problems:
-\begin{itemize}
-  \item It is unclear to adapt it to other tutorials.
-  \item Students tend to give brief responses possibly because the
-  questions are asking for a lot and because students are tired when
-  they reach the end of a tutorial.
-\end{itemize}
-
-So I plan to change the question format into a notice-and-wonder style
-of questions, which is known to be less demanding
-(\Cref{s:pedagogic-techniques}). So I plan to replace the
+It is unclear to adapt it to other tutorials.
+
+So I plan to change the question format into a more consistent style. So I plan to replace the
 \textbf{scope} prompt with
 \begin{quote}
-  \emph{Think about} the following questions:
-  \begin{itemize}
-    \item How do \lstinline|defvar|s work?
-    \item How do \lstinline|deffun|s work?
-    \item How do variable referencess (e.g., \lstinline|x| in \lstinline|(+ x 1)|) work?
-    \item How do \lstinline|defvar|s, \lstinline|deffun|s, and variable references interact with each other?
-  \end{itemize}
-  If you believe you know the answers perfectly, please \emph{response with}
-  \begin{quote}
-    I know the answers with 100\% certainty.
-  \end{quote}
-  Otherwise, please \emph{respond with} what you are uncertain about these language constructs.
+  How do variable referencess (e.g., the \lstinline|x| in \lstinline|(+ x 1)|) , \lstinline|defvar|s, and \lstinline|deffun|s interact with each other?
 \end{quote}
@@
 \begin{quote}
-  \emph{Think about} the following questions:
-  \begin{itemize}
-    \item How do \lstinline|set!|s work?
-    \item How do \lstinline|set!|s interact with other language constructs (namely, \lstinline|defvar|s, \lstinline|deffun|s, and variable references)?
-  \end{itemize}
-  If you believe you know the answers perfectly, please \emph{respond with}
-  \begin{quote}
-    I know the answers with 100\% certainty.
-  \end{quote}
-  Otherwise, please \emph{respond with} what you are uncertain about these language constructs.
+  How do \lstinline|set!|s interact with other language constructs (namely, \lstinline|defvar|s, \lstinline|deffun|s, and variable references)?
 \end{quote}
@@
 
+\section{Related Work}%
+\label{s:related-work}
+
+\subsection{Program Behavior}
+
+Our idea of ``goal sentences'' is not substantially different from the
+\textit{rules of program behavior} from \cite{duranRulesProgramBehavior2021}.
+I only used these sentences (or rules) to guide the
+design of the Tutor and as text in the Tutor itself.
+\cite{duranRulesProgramBehavior2021} argue for other uses of such
+sentences.
+
+\subsection{Misinterpreters}
+
+Our idea of misinterpreters is related to mystery languages
+\citep{diwanPLdetectiveSystemTeaching2004,
+  pombrioTeachingProgrammingLanguages2017}. Both approaches use
+evaluators that represent alternative semantics to the same syntax.
+However, the two are complementary.
+In mystery languages, instructors design the space of semantics with
+pedagogic intent, and
+students must create programs to explore that space. Misinterpreters,
+in contrast, are driven by student input, while the programs are provided
+by instructors. The two approaches also have different goals:
+mystery languages focus on encouraging students to experiment with
+languages; misinterpreters aim at capturing students' misconceptions.
+
 \bibliography{bibfile}

### 2023-11-14 proposal.tex
grammar check

@@
 (\Cref{s:background-tutoring-systems}), and then outline my research
-plans for the RQs (\Cref{s:plan-rq1,s:plan-rq2}). After that, I describe preliminary results (\Cref{s:preliminary-results} and propose my next steps (\Cref{s:future-work}).
+plans for the RQs (\Cref{s:plan-rq1,s:plan-rq2}). After that, I describe the preliminary results (\Cref{s:preliminary-results} and propose my next steps (\Cref{s:future-work}).
 
@@
 The Tutor mostly consists of \emph{interpreting tasks}:
-multiple-choice questions (MCQs) that asks students to interpret
+multiple-choice questions (MCQs) that ask students to interpret
 program. The options include (among other options
@@
   \item Programs shouldn't be too long. Some misconceptions are really
-  similar and as a consequence have few short programs to tell them
+  similar and, as a consequence, have few short programs to tell them
   apart.
@@
 ignoring some small syntactic differences). Thus, for
-each cluster, I obtain a set of rationales.
+each cluster, I obtained a set of rationales.
 
 After running Quizius in the course (\Cref{s:populations}),
-I took over as experts. Determining which is the right
-answer is easy, whereas expert knowledge is useful is in
+I took over as an expert. Determining which is the right
+answer is easy, whereas expert knowledge is useful in
 \emph{clustering the rationales}. If all the rationales for a wrong
@@
   \item I deferred the local binding forms to the end, since I felt
-        they were less essential. I therefore rewrote programs that use
+        they were less important. I therefore rewrote programs that use
         it to not depend on it.
@@
 increase engagement and help students understand how programs work.
-These work often mention \emph{notional machines}
+These works often mention \emph{notional machines}
 \citep{duboulayBlackBoxGlass1981}, which refers to the underlying
@@
 
-This section describe the current status of the Tutor.
+This section describes the current status of the Tutor.
 
@@
 down into 2--3 modules. The goal was for students to spend at most
-20--30 minutes per module. my data show that in practice, students
-spent about 9.8 (median) minutes.
+20--30 minutes per module. My data show that in practice, students
+spent about 9.8 minutes (median).
 
@@
 
-In addition to asking students questions, the Tutor along the way
+In addition to asking students questions, the Tutor, along the way,
 introduces terminology and states the true conceptions. These are the
-teaching goals of the Tutor. I therefore refer to these as
+teaching goals of the Tutor. I, therefore, refer to these as
 \newterm{goal sentences}. \Cref{t:tutor-goal-sentences} lists
@@
 the Tutor (\Cref{s:future-work}), in light of misinterpreters, is
-necessarily to give a more accurate evaluation. For example I found
+necessarily to give a more accurate evaluation. For example, I found
 far fewer problems than I would have liked. Concretely, only 40 of the
@@
 what percentage of students got the question right. Between the
-original university and University 2, we obtain a \pValue~=~2.013e-07.
-Between the original university and the online population, I obtain a
+original university and University 2, I obtained a \pValue~=~2.013e-07.
+Between the original university and the online population, I obtained a
 \pValue~<~2.2e-16. These show that the other two populations performed
@@
 
-Of course, other threats to generalizability remains: both University
+Of course, other threats to generalizability remain: both University
 1 and University 2 use the same textbook; SMoL's Lispy syntax might
-have bias in probing errors. I plan to address these issues in the
+have biases in probing errors. I plan to address these issues in the
 future (\Cref{s:future-work}).
@@
 from the Tutor, there is still a risk of having biases in student
-errors. This semester (Fall 2023) I am already deploying a newer SMoL
+errors. This semester (Fall 2023), I am already deploying a newer SMoL
 Tutor at Brown. This new Tutor is able to display programs in
